### 2022-07-29 ms/ms_02_suppl.tex
ENH Added MS to this repository

%  CONFIGURE NEW SINGLE-PAGE FORMAT

\onecolumn % go back to one column
\fancyhead{} % make sure we get no headers
\renewcommand{\floatpagefraction}{0.1}
\lfoot[\bSupInf]{\dAuthor}
\rfoot[\dAuthor]{\cSupInf}
\newpage

\captionsetup*{format=largeformat} % make figure legend slightly larger than in the paper
\setcounter{figure}{0} % reset figure counter for Supp. Figures
\setcounter{equation}{0} % reset equation counter for Supp. Equations
%\setcounter{page}{1} % reset page count
\makeatletter
\renewcommand{\thefigure}{S\@arabic\c@figure} % make Figure legend start with Figure S
\makeatother
\def\theequation{S\arabic{equation}}

%  MAIN TEXT

\newpage
\section*{Supplementary Information}

TODO

### 2022-07-29 ms/ms02_floats.tex
RFR reorganize float ms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Tab 1: Winner variants
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% full size table is table*
\begin{table*}[]
  \caption{Variants involved in 5 or more GWAS categories.}\label{tab:pleitropic_variants}
\centering
\scriptsize
\hline
\csvreader[separator=tab,
tabular=rrrcp{0.6\textwidth},
head,
table head=\bfseries Chrom. & \bfseries Pos. (hg38) & \bfseries Variant & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
]{fig/tab_rsid_most_pleiotropic.tsv}{}% use head of csv as column names
{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
\hline
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 1: Histograms variants vs GWAS, egene and etissues
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[]
\centering
%
\begin{subfigure}[]{.32\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
\end{subfigure}
%
\begin{subfigure}[]{.32\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
\end{subfigure}
%
\begin{subfigure}[]{.32\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
\end{subfigure}
%
\caption{\textbf{Probability density of GWAS phenotypes, eGene or eQTL samples.} (\textbf{a}) TODO.} \label{fig:hist_gwas_egene_etissue}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Scatter plots of winner regions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[!ht]

\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
\end{subfigure}

\begin{subfigure}[]{.33\textwidth}
\textbf{d}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{e}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{f}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
\end{subfigure}

\begin{subfigure}[]{.33\textwidth}
\textbf{g}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{h}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{i}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
\end{subfigure}

\caption{\textbf{Count of GWAS categories, eGenes and eTissues in pleiotropic genomic regions.} (\textbf{a,d,g}) Region 5:132,239,645-132,497,907, \textbf{b,e,h}) region 6:31,034,839-32,478,149,  and \textbf{c,g,i}) region 12:111,395,984-111,645,358} \label{fig:region_gwas_egenes_tissues}
%
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/missense_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/splice_region_variant.png}
%
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/3_prime_UTR_variant.png}
%
\end{subfigure}
%
\caption{\textbf{Location of of variants computed as VEP consequence.} (\textbf{a}) TODO.} \label{fig:gwas_cat_vs_egene_and_sample}
%
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 4: Violin plots. eGenes, etissues and gwas per variants
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_egene_per_variant.py/vlnplt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_etissue_per_variant_egene.py/vlnplt.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_gwas_per_variant_egene_etissue.py/vlnplt.png}
\end{subfigure}
%
\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:distrib_variant_egene_etissue_gwas}
%
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Tab 2: Winner regions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% full size table is table*
\begin{table*}[]
\caption{Pleiotropic regions involving more than 4 GWAS categories.}\label{tab:pleiotropic_regions}
\centering
\scriptsize
\hline
\csvreader[
separator=tab,
tabular=rrrcp{0.5\textwidth},
head,
table head=\bfseries Chrom. & \bfseries Start (hg38) & \bfseries End (hg38) & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
]{fig/tab_region_window_100000_pleio_highest.tsv}{}% use head of csv as column names
{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
\hline
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 5: Length histogram of regions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[]
\centering
%
\includegraphics[width=0.33\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
%
\caption{\textbf{Length distribution of pleiotropic regions.} (\textbf{a}) TODO.} \label{fig:pleiotropy_region_distribution}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: TF count per GWAS category count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}[!]
\centering
%
\begin{subfigure}[]{.33\textwidth}
%
\includegraphics[width=\textwidth]{\floatRelativePath/plt_bxplt_remaptf_per_rsid.py/bxplt_remaptf_per_rsid_flank_50.png}
\end{subfigure}
%
\caption{\textbf{Binding of transcription factors in the region (100 kb) around pleiotropic variants.} (\textbf{a}) TODO.} \label{fig:gwas_cat_vs_egene_and_sample}
%
\end{figure*}

### 2022-08-02 ms/ms02_floats.tex
DOC ms

@@
 %
-\caption{\textbf{Location of of variants computed as VEP consequence.} (\textbf{a}) TODO.} \label{fig:gwas_cat_vs_egene_and_sample}
+\caption{\textbf{Location of of variants computed as VEP consequence.} (\textbf{a}) TODO.} \label{fig:vep_consequence}
 %
@@
 %
-\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:distrib_variant_egene_etissue_gwas}
+\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:freq_gwas_egene_etisue_per_variant}
 %
@@
 %
-\caption{\textbf{Binding of transcription factors in the region (100 kb) around pleiotropic variants.} (\textbf{a}) TODO.} \label{fig:gwas_cat_vs_egene_and_sample}
+\caption{\textbf{Binding of transcription factors in the region (100 kb) around pleiotropic variants.} (\textbf{a}) TODO.} \label{fig:freq_tf_per_variant}
 %
 \end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 7: beta
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[!]
+\centering
+%
+\begin{subfigure}[]{.33\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_beta_pval.py/eqtl_beta.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.33\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_beta_pval.py/gwas_beta.png}
+\end{subfigure}
+%
+\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:beta}
+%
+\end{figure*}

### 2022-08-03 ms/ms02_floats.tex
ENH improved boxplots instead of violin plots

@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_gwas.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_egene.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_scttr_count_per_rsid_etissue.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_egene_per_variant.py/vlnplt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_y_egene.py/plt.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_etissue_per_variant_egene.py/vlnplt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_y_etissue.py/plt.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_gwas_per_variant_egene_etissue.py/vlnplt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_etissue_y_gwas.py/plt.png}
 \end{subfigure}
 %
-\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:freq_gwas_egene_etisue_per_variant}
+\caption{\textbf{Frequency of eGenes per variant, eTissues per variant-eGene and phenotypes per variant-eGene-eTissue.} (\textbf{a}) TODO.} \label{fig:freq_gwas_egene_etisue_per_variant}
 %
@@
 %
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_bxplt_remaptf_per_rsid.py/bxplt_remaptf_per_rsid_flank_50.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_0.png}
 \end{subfigure}
@@
 %
-% Fig 7: beta
+% Fig 7: beta and log pval
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_beta_pval.py/eqtl_beta.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_violin_beta_pval.py/gwas_beta.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_beta.py/gwas_beta.png}
+\end{subfigure}
+
+\begin{subfigure}[]{.33\textwidth}
+\textbf{c}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_logpval.py/eqtl.png}
 \end{subfigure}
 %
+\begin{subfigure}[]{.33\textwidth}
+\textbf{d}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_logpval.py/gwas.png}
+\end{subfigure}
+
 \caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:beta}

### 2022-08-03 scripts/plt_bar_nonredundant_variant_category_percentage.py

@@ -7,17 +7,16 @@ import pandas
 import pathlib
 import shlex
 import subprocess
-import seaborn as sns
+import seaborn
 
 from eqtl2gwas_pleiotropy.Logger import Logger
 from eqtl2gwas_pleiotropy.PathManager import PathManager
 from matplotlib import pyplot as plt
+from eqtl2gwas_pleiotropy.constants import seaborn_theme_dic
+from eqtl2gwas_pleiotropy.constants import dpi, tick_fontsize, label_fontsize
 
 plt.rcParams["figure.figsize"] = (16, 6)
-
-
-#%%
-from eqtl2gwas_pleiotropy.constants import dpi, tick_fontsize, label_fontsize
+seaborn.set_theme(**seaborn_theme_dic)
 
 #%% outdir path
 if not '__file__' in locals():
@@ -74,7 +73,7 @@ for pleiotropy in range(1, 6):
 
 variant_cat_perc_df = pandas.DataFrame(out_lst_2d_lst, columns=['Pleiotropy', 'Category', 'variant_cat_perc'])
 
-ax = sns.barplot(x="Category", y="variant_cat_perc", hue="Pleiotropy", data=variant_cat_perc_df)
+ax = seaborn.barplot(x="Category", y="variant_cat_perc", hue="Pleiotropy", data=variant_cat_perc_df)
 
 plt.grid(axis='y')
 plt.legend(fontsize=16) # using a size in points
@@ -86,8 +85,6 @@ plt.ylabel("Variant category [%]", fontsize=label_fontsize)
 plt.yticks(fontsize=tick_fontsize)
 
 plt.tight_layout()
-fig = ax.get_figure()
 png_path = os.path.join(outdir_path, "non_redundant_variant_category_pleio2.png")
-fig.savefig(png_path, dpi=dpi)
-plt.clf()
+plt.savefig(png_path, dpi=dpi)
 plt.close()

### 2022-08-03 scripts/plthst_gwas_egene_etissue.py

@@ -1,9 +1,5 @@
 #%%
-import pdb
 import sys
-
-from eqtl2gwas_pleiotropy.PathManager import PathManager
-
 import matplotlib.pyplot as plt
 import numpy
 import os
@@ -44,6 +40,7 @@ edgecolor='k'
 linewidth = 2
 grid_axis = 'y'
 hist_kwargs = {'density': 1, 'edgecolor': edgecolor, 'linewidth': linewidth}
+
 
 #%% gwas
 count_df = pandas.read_csv(gwas_count_tsv_path, sep="\t", header=0)

### 2022-08-03 ms/ms02_floats.tex
ENH

@@
 \begin{table*}[]
-\caption{Pleiotropic regions involving more than 4 GWAS categories.}\label{tab:pleiotropic_regions}
+\caption{Pleiotropic regions involving 6 or more GWAS categories.}\label{tab:pleiotropic_regions}
 \centering
@@
 \end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 7: graphical conclusions
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[!]
+\centering
+%
+\begin{subfigure}[]{\textwidth}
+
+\includegraphics[width=\textwidth]{fig/graphical_summary.png}
+\end{subfigure}
+
+\caption{\textbf{Molecular mechanism of pleiotropy.} (We have observed three possible mechanisms resulting in pleiotropy.
+(Left) Pleiotropic variants have more eGenes.
+One explanation is that these variants are present in splicing regions.
+(Center) eGenes of pleiotropic variants are active in more eTissues, which might be explained from variants being bound by more transcription factors.
+(Right) eGenes/eTissues are associated with more GWAS phenotypes.
+This might be explained because the variants have more frequently a missense consequence.} \label{fig:beta}
+%
+\end{figure*}

### 2022-08-10 ms/ms02_floats.tex
ENH cytoband

@@
 \csvreader[separator=tab,
-tabular=rrrcp{0.6\textwidth},
+tabular=rrrrcp{0.6\textwidth},
 head,
-table head=\bfseries Chrom. & \bfseries Pos. (hg38) & \bfseries Variant & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos. (hg38) & \bfseries Variant & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
 ]{fig/tab_rsid_most_pleiotropic.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv & \csvcolvi}% specify your coloumns here
 \hline
 \end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Tab 2: Winner regions
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% full size table is table*
+\begin{table*}[]
+\caption{Pleiotropic regions involving 6 or more GWAS categories.}\label{tab:pleiotropic_regions}
+\centering
+\scriptsize
+\hline
+\csvreader[
+separator=tab,
+tabular=rrrrcp{0.5\textwidth},
+head,
+table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start (hg38) & \bfseries End (hg38) & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
+]{fig/tab_region_window_100000_pleio_highest.tsv}{}% use head of csv as column names
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv & \csvcolvi}% specify your coloumns here
+\hline
+\end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 5: Length histogram of regions
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[]
+\centering
+%
+\includegraphics[width=0.33\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
+%
+\caption{\textbf{Length distribution of pleiotropic regions.} (\textbf{a}) TODO.} \label{fig:pleiotropy_region_distribution}
+\end{figure*}
 
@@
 %
-% Tab 2: Winner regions
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% full size table is table*
-\begin{table*}[]
-\caption{Pleiotropic regions involving 6 or more GWAS categories.}\label{tab:pleiotropic_regions}
-\centering
-\scriptsize
-\hline
-\csvreader[
-separator=tab,
-tabular=rrrcp{0.5\textwidth},
-head,
-table head=\bfseries Chrom. & \bfseries Start (hg38) & \bfseries End (hg38) & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
-]{fig/tab_region_window_100000_pleio_highest.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
-\hline
-\end{table*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 5: Length histogram of regions
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}[]
-\centering
-%
-\includegraphics[width=0.33\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
-%
-\caption{\textbf{Length distribution of pleiotropic regions.} (\textbf{a}) TODO.} \label{fig:pleiotropy_region_distribution}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 % Fig 6: TF count per GWAS category count

### 2022-08-10 scripts/cmpt_count_per_rsid.py

@@ -1,12 +1,3 @@
-import shlex
-import subprocess
-
-from eqtl2gwas_pleiotropy.EBIeQTLinfo import EBIeQTLinfo
-from eqtl2gwas_pleiotropy.OpenGWASinfo import OpenGWASinfo
-from eqtl2gwas_pleiotropy.PathManager import PathManager
-from eqtl2gwas_pleiotropy.URL import URL
-from eqtl2gwas_pleiotropy.constants import coloc_raw_tsv_path
-
 import os
 import pandas
 import pathlib
@@ -33,26 +24,28 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 #%%
 coloc_df = pandas.read_csv(annotated_tsv_path, sep="\t")
+coloc_df['chrom'] = coloc_df['chrom'].astype(int)
 
 #%% etissue pleiotropy
-pleio_etissue_df = coloc_df[['chrom', 'pos', 'rsid', 'etissue_category']].drop_duplicates().groupby(['chrom', 'pos', 'rsid']).agg({'etissue_category': ['size', (lambda x: (",".join(sorted(x))))]}).reset_index()
-pleio_etissue_df.columns = ['chrom', 'pos', 'rsid', 'etissue_label_count', 'etissue_subcategory_lst']
-pleio_etissue_df = pleio_etissue_df.sort_values(by=['etissue_label_count', 'etissue_subcategory_lst', 'rsid'], ascending=[False, True, True])
+pleio_etissue_df = coloc_df[['chrom', 'cytoband', 'pos', 'rsid', 'etissue_category']].drop_duplicates().groupby(['chrom', 'cytoband', 'pos', 'rsid']).agg({'etissue_category': ['size', (lambda x: (",".join(sorted(x))))]}).reset_index()
+pleio_etissue_df.columns = ['chrom', 'cytoband', 'pos', 'rsid', 'etissue_label_count', 'etissue_subcategory_lst']
+pleio_etissue_df = pleio_etissue_df.sort_values(by=['etissue_label_count', 'chrom', 'pos', 'etissue_subcategory_lst', 'rsid'], ascending=[False, True, True, True, True])
 tsv_path = os.path.join(outdir_path, "count_per_rsid_etissue.tsv")
 pleio_etissue_df.to_csv(tsv_path, sep="\t", index=False)
 
 #%% egene pleiotropy
[…]

### 2022-08-12 scripts/create_supplementary_table.py
ENH MS ongoing

@@ -38,12 +38,15 @@ sheet_name_lst.append("ST2")
 description_lst.append("Metadata and classification of GWAS")
 #
 sheet_name_lst.append("ST3")
-description_lst.append("Count of GWAS phenotypes of eQTL/GWAS variants")
+description_lst.append("Count and list of GWAS phenotypes, eGenes and eTissues for each eQTL/GWAS variant")
+#
+sheet_name_lst.append("ST4")
+description_lst.append("Count and list of GWAS phenotypes for each pleiotropic region")
 
 capt_df = pandas.DataFrame({'Supp. Tab.': sheet_name_lst, 'Description': description_lst})
 capt_df.to_excel(writer, sheet_name='Table descrip.', index=False, header=True)
 
-#%%
+#%% ST1
 sheet_name = 'ST{}'.format(sheet_counter)
 sheet_counter += 1
 etissue_category_ods_path = os.path.join(PathManager.get_project_path(), "config", "etissue_category.ods")
@@ -51,7 +54,7 @@ st_df = pandas.read_excel(etissue_category_ods_path, index_col=None, header=0)
 st_df.drop(['Unnamed: 7', 'etissue_category.1', 'count'], axis=1, inplace=True)
 st_df.to_excel(writer, sheet_name=sheet_name, index=False, header=True)
 
-#%%
+#%% ST2
 sheet_name = 'ST{}'.format(sheet_counter)
 sheet_counter += 1
 tsv_path = os.path.join(wdir_path, "annot_gwas_metadata.py/gwas413_metadata.tsv")
@@ -59,17 +62,35 @@ st_df = pandas.read_csv(tsv_path, sep="\t", header=0)
 st_df.sort_values(by=st_df.columns.tolist(), inplace=True)
 st_df.to_excel(writer, sheet_name=sheet_name, index=False, header=True)
 
-#%%
+#%% ST3
 sheet_name = 'ST{}'.format(sheet_counter)
 sheet_counter += 1
 tsv_path = os.path.join(wdir_path, "cmpt_count_per_rsid.py/count_per_rsid_gwas.tsv")
-st_df = pandas.read_csv(tsv_path, sep="\t", header=0)
+gwas_df = pandas.read_csv(tsv_path, sep="\t", header=0)
+
+tsv_path = os.path.join(wdir_path, "cmpt_count_per_rsid.py/count_per_rsid_egene.tsv")
[…]

### 2022-08-13 ms/ms02_floats.tex
ENH went through figures

@@
 \begin{table*}[]
-  \caption{Variants involved in 5 or more GWAS categories.}\label{tab:pleitropic_variants}
+  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories.}\label{tab:pleitropic_variants}
 \centering
@@
 \begin{table*}[]
-\caption{Pleiotropic regions involving 6 or more GWAS categories.}\label{tab:pleiotropic_regions}
+\caption{Pleiotropic regions involving 5 or more GWAS categories.}\label{tab:pleiotropic_regions}
 \centering
@@
 %
-\caption{\textbf{Length distribution of pleiotropic regions.} (\textbf{a}) TODO.} \label{fig:pleiotropy_region_distribution}
+\caption{Length distribution of pleiotropic regions.} \label{fig:pleiotropy_region_distribution}
 \end{figure*}
@@
 %
-\caption{\textbf{Probability density of GWAS phenotypes, eGene or eQTL samples.} (\textbf{a}) TODO.} \label{fig:hist_gwas_egene_etissue}
+\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS categories (\textbf{a}), egenes (\textbf{b}) and etissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
 \end{figure*}
@@
 
-\caption{\textbf{Count of GWAS categories, eGenes and eTissues in pleiotropic genomic regions.} (\textbf{a,d,g}) Region 5:132,239,645-132,497,907, \textbf{b,e,h}) region 6:31,034,839-32,478,149,  and \textbf{c,g,i}) region 12:111,395,984-111,645,358} \label{fig:region_gwas_egenes_tissues}
+\caption{Count of GWAS categories (\textbf{a-c}), egenes (\textbf{d-f}) and etissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1)} \label{fig:region_gwas_egenes_tissues}
 %
@@
 %
-\caption{\textbf{Location of of variants computed as VEP consequence.} (\textbf{a}) TODO.} \label{fig:vep_consequence}
+\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
 %
@@
 %
-\caption{\textbf{Frequency of eGenes per variant, eTissues per variant-eGene and phenotypes per variant-eGene-eTissue.} (\textbf{a}) TODO.} \label{fig:freq_gwas_egene_etisue_per_variant}
+\caption{Count of egenes per variant (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
 %
@@
 %
-\caption{\textbf{Binding of transcription factors in the region (100 kb) around pleiotropic variants.} (\textbf{a}) TODO.} \label{fig:freq_tf_per_variant}
+\caption{Binding of transcription factors in the region around pleiotropic variants (Radius 50 bp) as a function of the GWAS category count.} \label{fig:freq_tf_per_variant}
 %
@@
 
-\caption{\textbf{TODO.} (\textbf{a}) TODO.} \label{fig:beta}
+\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS category count.} \label{fig:beta_pval}
 %
@@
 
-\caption{\textbf{Molecular mechanism of pleiotropy.} (We have observed three possible mechanisms resulting in pleiotropy.
-(Left) Pleiotropic variants have more eGenes.
-One explanation is that these variants are present in splicing regions.
-(Center) eGenes of pleiotropic variants are active in more eTissues, which might be explained from variants being bound by more transcription factors.
-(Right) eGenes/eTissues are associated with more GWAS phenotypes.
-This might be explained because the variants have more frequently a missense consequence.} \label{fig:beta}
+\caption{\textbf{Model of regulatory variant pleiotropy.} I have investigated three possible mechanisms of pleiotropy.
+\textbf{Left}, Pleiotropic variants have more egenes that result in more functions and more phenotypes.
+This might arise from an enrichment of pleiotropic variants in splicing or 3' UTR regions.
+\textbf{Center}, I have also found that egenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
+This might be explained from variants being bound by more transcription factors.
+\textbf{Left} Triplets of variant-egene-etissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
+I have found that this might be explained by an enrichment of missense alleles.} \label{fig:beta}
 %

### 2022-08-13 eqtl2gwas_pleiotropy/constants.py

@@ -31,4 +31,4 @@ alpha = 0.5
 boxplot_kwargs = {'linewidth': 3, 'notch': True, 'palette': "rocket_r", 'showfliers': False, 'boxprops': dict(alpha=alpha)}
 
 # statannotations
-annotator_config_dic = {'fontsize': 20}
+annotator_config_dic = {'fontsize': 16}

### 2022-08-13 scripts/cmpt_pleiotropic_regions.py

@@ -161,14 +161,14 @@ hist_kwargs = {'density': False, 'edgecolor': edgecolor, 'linewidth': linewidth}
 # import pdb; pdb.set_trace()
 # plt.hist(region_lenght_ser/1000000, **hist_kwargs)
 data_ser = region_lenght_ser/100000
-seaborn.histplot(data_ser, stat='probability', discrete=True)
+seaborn.histplot(data_ser, stat='percent', discrete=True)
 
 plt.grid(visible=True, axis='y')
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel("Region length [1e5 bp]", fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize)
 plt.ylabel(ylabel, fontsize=label_fontsize)
-plt.yscale('log')
+# plt.yscale('log')
 plt.yticks(fontsize=tick_fontsize)
 
 plt.tight_layout()

### 2022-08-13 scripts/plt_vep_consequence.py

@@ -42,7 +42,7 @@ in_df.loc[in_df['p'] <= 1.00e-03, 'signif'] = '***'
 in_df.loc[in_df['p'] <= 1.00e-04, 'signif'] = '****'
 
 #%%
-consequence = 'missense_variant'
+# consequence = 'missense_variant'
 
 #%%
 for consequence in in_df['consequence'].unique():
@@ -53,15 +53,16 @@ for consequence in in_df['consequence'].unique():
     #%%
     order = sorted(plt_df['gwas_category_count'].unique())
     xticklabels = order.copy()
+    xticklabels[-1] = '≥{}'.format(order[-1])
     title = consequence
     xlabel = "GWAS category count"
     ylabel = "Odds ratio"
-    ylim = [0, 9]
+    # ylim = [0, 9]
     y = "oddsr"
     x = "gwas_category_count"
 
     #%%
-    ax = seaborn.barplot(x=x, y=y, data=plt_df, order=order, palette="rocket_r", ci=None)
+    ax = seaborn.barplot(x=x, y=y, data=plt_df, order=order, palette="rocket_r")
 
     #%%
     formatted_pvalues = plt_df['signif'].tolist()
@@ -73,12 +74,11 @@ for consequence in in_df['consequence'].unique():
     annotator.configure(**annotator_config_dic)
     annotator.annotate()
 
-
     plt.title(title, fontsize=label_fontsize)
     plt.xlabel(xlabel, fontsize=label_fontsize)
     plt.xticks(fontsize=tick_fontsize, rotation=0)
     plt.ylabel(ylabel, fontsize=label_fontsize)
-    plt.ylim(ylim)
+    # plt.ylim(ylim)
     plt.yticks(fontsize=tick_fontsize)
     ax.set_xticklabels(xticklabels)
 

### 2022-08-13 scripts/plt_violin_x_per_egene_y_interactors.py

@@ -101,7 +101,7 @@ m_df.groupby('gwas_category_count')['interactor_count'].apply(lambda x: x.descri
 order = [*range(1, upper_var_gwas_cat_count+1)]
 xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
-title = "Interactors per eGene"
+title = "Interactors per egene"
 xlabel = "GWAS category count"
 ylabel = "Phys. interactor count"
 y = "interactor_count"

### 2022-08-13 scripts/pltbox_x_per_variant_egene_etissue_y_gwas.py

@@ -75,7 +75,7 @@ m_df.groupby('gwas_category_count')['gwas_trait_count'].apply(lambda x: x.descri
 order = [str(x) for x in range(1, upper_var_gwas_cat_count+1)]
 xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
-title = "GWAS per var.-gene-tissue"
+title = "GWAS per variant-egene-etissue"
 xlabel = "GWAS category count"
 ylabel = "GWAS trait count"
 y = "gwas_trait_count"

### 2022-08-13 scripts/pltbox_x_per_variant_egene_y_etissue.py

@@ -75,7 +75,7 @@ m_df.groupby('gwas_category_count')['etissue_category_count'].apply(lambda x: x.
 order = [str(x) for x in range(1, upper_var_gwas_cat_count+1)]
 xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
-title = "eTissues per variant-eGene "
+title = "eTissues per variant-egene "
 xlabel = "GWAS category count"
 ylabel = "eTissue count"
 y = "etissue_category_count"

### 2022-08-13 scripts/pltsctr_x_per_rsid_y_egene.py

@@ -45,10 +45,11 @@ region_window_100000_df = pandas.read_csv(region_window_100000_tsv_path, sep="\t
 # outdir_path = os.path.join(PathManager.get_project_path(), "out", os.path.basename(__file__))
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
-title = "eGenes per variant"
+title = "Coloc. eQTL/GWAS variant"
 count_col_name = "egene_count"
-ylim = [0, 35]
+ylim = [0, 40]
 c = 'blue'
+ylabel = "eGene count"
 
 count_per_rsid_df['pos'] = count_per_rsid_df['pos'].astype('int')
 
@@ -72,7 +73,7 @@ for rowi, row in pleiotropic_regions_df.iterrows():
     plt.title(title, fontsize=label_fontsize)
     plt.xlabel("Chr{} position [Mbp]".format(chrom), fontsize=label_fontsize)
     plt.xticks(fontsize=tick_fontsize, rotation=45)
-    plt.ylabel("eGene count", fontsize=label_fontsize)
+    plt.ylabel(ylabel, fontsize=label_fontsize)
     plt.ylim(ylim)
     plt.yticks(fontsize=tick_fontsize)
 

### 2022-08-13 ms/ms02_floats.tex
ENH david go

@@
 \end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 1: GO of pleiotropic variants
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[]
+\centering
+%
+\begin{subfigure}[]{.66\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
+\end{subfigure}
+
+\begin{subfigure}[]{.66\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
+\end{subfigure}
+%
+\caption{Gene ontology analysis of egenes of variants associated to two (\textbf{a}) and three (\textbf{b}) GWAS categories compared to egenes of variants associate to one GWAS category.} \label{fig:hist_gwas_egene_etissue}
+\end{figure*}
 
@@
 %
-% Fig 5: Length histogram of regions
+% Fig 2: Length histogram of regions
 %

### 2022-08-17 ms/ms02_floats.tex
ENH updated pipeline

@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
 \end{subfigure}
 %
-\caption{Gene ontology analysis of egenes of variants associated to two (\textbf{a}) and three (\textbf{b}) GWAS categories compared to egenes of variants associate to one GWAS category.} \label{fig:hist_gwas_egene_etissue}
+\caption{Gene ontology analysis of egenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to egenes of variants associated to one GWAS category.} \label{fig:hist_gwas_egene_etissue}
 \end{figure*}

### 2022-08-17 scripts/annotate.py

@@ -77,3 +77,5 @@ coloc_df = coloc_df[columns]
 coloc_df.sort_values(by=coloc_df.columns.tolist(), inplace=True)
 Logger.info("Writing {}".format(annotated_tsv_gz_path))
 coloc_df.to_csv(annotated_tsv_gz_path, sep="\t", index=True, index_label='id', na_rep='na')
+
+#%% DBSQLITE

### 2022-08-18 ms/ms02_floats.tex
ENH suppl

--- 2022-08-13 ms/ms02_floats.tex
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
 \end{subfigure}
 %
-\caption{Gene ontology analysis of egenes of variants associated to two (\textbf{a}) and three (\textbf{b}) GWAS categories compared to egenes of variants associate to one GWAS category.} \label{fig:hist_gwas_egene_etissue}
+\caption{Gene ontology analysis of egenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to egenes of variants associated to one GWAS category.} \label{fig:gene_ontology}
 \end{figure*}

### 2022-08-18 ms/ms01_main.tex
ENH ms

\leadauthor{Scientist}

\title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
\shorttitle{Pleiotropy and gene regulatory variants}

\author[1,\Letter]{Aitor González \orcidlink{0000-0002-8402-1655}}
\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
\date{}

\maketitle

\begin{abstract}

Pleiotropic genetic variants affecting several phenotypes are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
%
Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
%
In this work, I investigate the link between gene regulation and phenotype pleiotropy for frequent variants.

% Verify numbers
First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
%
I have categorized the GWAS phenotypes into 96 phenotypes and categories to define pleiotropic as belonging at least to two of these categories.
%
Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more phenotypes.
%
These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
%
These variants with the GWAS phenotypes, egenes and etissues can be explored in a public website.

In conclusion, our work suggest that pleiotropy of gene regulatory variants arise from an increase of bound transcription factors, target egenes and etissues and severity of molecular effect consequences.

\end{abstract}

\begin{keywords}
keyword1 | keyword2 | keyword3
\end{keywords}

\begin{corrauthor}
aitor.gonzalez\at univ-amu.fr
\end{corrauthor}

\section*{Introduction}\label{sec:introduction}

Genome-wide association studies (GWAS) evaluate the impact of frequent variants on a trait in the population.
%
The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
%
For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71673 variant-trait associations from 3567 publications \citep{2018.Parkinson.Buniello}.

However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
%
First, identification of the causaul variant is difficult because of linkage disequilibrium.
%
Second, GWAS do not provide hints regarding the tissues involved in the phenotypes.
%
Third, GWAS variants affect most often non-coding regions.

Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
%
eQTLs have some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
%
On the other hand, eQTLs provide information about the target genes and tissues.

eQTLs

A large number of GWAS variants are found in non-conding regions.
To priori

Sentence describing EBI eQTL

In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.

\section*{Results}\label{s:results}

\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}

To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).

125 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.

Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
%
The first criterion was to exclude molecular phenotypes such as proteome or methylome.
%	
The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
%
The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental phenotypes such "employment status" or "self-reported" medical conditions.
%
The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
%
These filters resulted in 413 GWAS (Supplementary table 2).

I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
%
This analysis explored 30 261 unique variants and 2,479,065 potential variant colocalizations (OSF URL).
%
Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).

\subsection*{Which regulatory variants are the most pleiotropic?}

I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
%
In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
%
Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).

These category annotations were used to investigate the variant pleiotropy for GWAS phenotypes.
%
Colocalized eQTL/GWAS variants were classified according to the number of GWAS categories they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
%
Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).

% 12q24.12
For instance, the most pleiotropic variants in cytoband 12q24.12 are involved in allergies, cancer, cardiovascular and autoimmune diseases (Table \ref{tab:pleitropic_variants}).
%
These variants control egenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
%
These eQTLs are active in adipose tissue, arteries, blood, colon, immune cells, skin and induced pluripotent stem cells (iPSC).

\subsection*{What is the function of pleiotropic variants?}

Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
%
I analyzed the ontology of egenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
%
I found that egenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
For instance, egenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
%
IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
%
Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
%
These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.

%Variants of the 3q23 cytoband are involved in allergy, cancer and height phenotypes (Supplementary table 3).
%%
%Interestingly all these variants target the egene ZBTB3.
%%
%ZBTB3 is a transcription factor that controls pro-inflammatory factors such as IRF5 (10.1016/j.jaut.2016.08.001).
%%
%These variants are active in tissues such as blood, esophagus, immune cell, LCLs and muscle (Supplementary table 3).

%%11q13.5
%The variants in cytoband 11q13.5 are involved in allergy and autoimmune disease (Supplementary table 3).
%%
%All these variants target the egene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
%%
%These variants are active in LCL, Skin and iPSC (Supplementary table 3).
%
%In summary, I have identified pleiotropic variants with gene regulatory function.
%%
%Some of these loci are strongly involved in the immune system.

\subsection*{Analysis of pleiotropic regions}

Pleiotropic variants are aggregated in few cytobands.
%
Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions}).
%
I found 453 regions with 441 regions with 2 or more categories, 114 regions with 3 or more categories, 37 regions with 4 or more categories and 18 regions with 5 or more categories (Supplementary table 5).
%
75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}).

The most pleiotropic region is 12:111,395,984-111,645,358 in cytoband 12q24.12 that we discussed above (Table \ref{tab:pleiotropic_regions}).
%
The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension phenotypes (Supplementary table 5).
%
The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).

\subsection*{How specific are variants to GWAS phenotypes, egenes and etissues?}

Next I explored the phenotype, egene and etissue specificity of eQTL/GWAS variants.
%
83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more phenotype categories make 3\% or less (Fig. \ref{fig:hist_gwas_egene_etissue}(a).
%
Regarding egenes, half of variants modulate one egene, and the other half modulate two or more egenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
%
In the case of etissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
%
In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS phenotype. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{How do phenotype frequencies relates to eQTL and etissue frequency at the loci level?}

In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, egenes and eTissues for regulatory variants in three pleiotropic regions.
%
All three regions are very pleiotropic with very different associated phenotype categories such as cancer, cardiovascular and autoimmune diseases.
%
However these plots show that the relationship between the number of phenotypes, egenes and etissues can be very different.
%
The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of egenes but a high number of etissues.
%
By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both egenes and etissues.
%
Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest phenotype pleiotropy but the number of egenes and etissues is very low.
%
In summary, phenotype pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.

\subsection*{What are the mechanisms of pleiotropy}

Then, I wanted to understand the molecular mechanism of pleiotropic regions.

\subsubsection*{More severe variant effect consequences?}

I first analysed whether there are significant differences of effect consequences between more and less pleiotropic variants.
%
I separated variants according to the count of GWAS categories and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
%
I found a significant increase of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
%
I also found a significant increase of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
%
Finally, there is also a significant increase of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
%
These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.

\subsubsection*{More egenes per variant?}

Then I hypothesized that pleiotropic variants modulate more egenes.
%
I found that pleiotropic variants have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
%
This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).

\subsubsection*{More etissues per variant-egene?}

My next hypothesis was that pleiotropic variant-egene pairs are active in more tissues compared with variant-egene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
%
Indeed, I found that pleiotropic variant-egenes pairs are associated with a higher number of etissues (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).

\subsubsection*{How to explain that pleiotropic variants are active in more tissues?}

I hypothesized that pleiotropic variants are active in more etissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
%
I looked at the number of unique transcription factors bound in a window of 100 bp around each variant.
%
I found a significant increase of unique transcription factors around pleagonzaleziotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).

%TODO CRMs and variants

\subsubsection*{More GWAS per variant-egene-etissue?}

Another molecular mechanism of pleiotropy is that egenes correlate with more GWAS categories even after taking into account variants and etissues.
%
I computed the number of GWAS categories per variant-egene-etissue triplet.
%
Indeed I found a significant increase of GWAS categories for pleiotropic variant-egene-etissue triplets (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
%
This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).

\subsection*{What is the effect size of pleiotropic eQTLs?}

Then I wondered, how does the effect size (beta) and significance (p-value) relate to pleiotropy.
%
I found a significant decrease of effect size of eQTLs and GWAS variants with the exception of the GWAS effect size of variants with 3 categories (Fig. \ref{fig:beta_pval}a,b).

Then I looked at the significance of pleiotropic eQTLs and GWAS variants.
%
eQTL significance significantly decreased for pleiotropic variants (Fig. \ref{fig:beta_pval}c).
%
By contrast, the significance of GWAS associations increased for GWAS variants (Fig. \ref{fig:beta_pval}d).

In summary, pleiotropic eQTLs are weaker with a weaker beta and less significant p-values..
%
On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.

\section*{Discussion}

% Confirm eQTL studies
In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
%
This pipeline and the results is useful for three aspects.
%
First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
%
Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
%
% TODO how many
Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.

In this work, I have explore variants with a role in gene regulation that are highly pleiotropic.
%
For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.

For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.

%\subsection*{What is the function of the variants in the most pleiotropic regions?}
%
%% 12q24.12
%The cytoband 12q24.12 has five variants rs3184504, rs653178, rs7310615, rs597808, rs11065979, that are known in the literature to be associated to a wide palette of diseases.
%This locus has been used to explain links between inflammaiton and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
%%
%Our analysis add up that the variants in this cytoband control important genes such as ALDH2, ATXN2, MAPKAPK5, SH2B3 and TMEM116 in several tissues such as adipose, artery, blood, brain, digestive, skin, heart and skin tissue as well as immune cells.
%
%% 6p21.3
%% TODO
%
%%15q26.1
%The next pleiotropic locus is in cytoband 15q26.1 involved in cardiovascular, hypertension and psyciatric diseases with variants rs2071382, rs8039305 and rs6224.
%%
%eQTLs in cytoband 15q26.1 control genes like Furin, oncogene Fes and UNC45A involved in Osteootohepatoenteric syndrome (OMIM DB).
%%
%eQTLs are active in a very large number of cells including blood, artery, brain and immune cells (See supplementary table).
%
%% 1p13.3
%The locus in cytoband 1p13.3 is mainly related to cardiovascular diseases.
%Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
%
%
%
%The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular. Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
%
%The variant rs11236797 and the two others in locus 11q13.5 is mainly autoimmune.
%All three regulate the gene EMSY-DT and are active induced pluripotent cells, LCDs and skin.
%
%Another pleiotropic variants rs301806
%rs301807
%rs301802
%rs301817
%rs159963
%rs301816
%in the cytoband 1p36.23 that are involved in allergy, cardiovascular and hypertension.
%This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\

\section*{Methods}\label{sec:methods}

\subsection*{Data sets and data exploration}

I used these data sets here: eQTLs from the eQTL Catalogue \citep{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \citep{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \citep{2021.Ballester.Hammal}, and UCSC annotation data.
%
I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \citep{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
%
Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
%
The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/eqtl2gwas} and \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.

\subsection*{Colocalization}

I developed a colocalization pipeline based on a pipeline that is public at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
%
First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants in a radius of 500 000 nt are retrieved.
%
Colocalization between eQTL and GWAS variants is tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.

\subsection*{Definition of pleiotropic regions}

To define pleiotropic regions, I included "pleiotropic" variants defining as having more than 2 categories as long as the distance between the two "pleiotropic" variants is less than 100 000 bp.
%
Then the number of GWAS categories in Supplementary table TODO is given by the number of different variant categories in the region.

\subsection*{Characterization of regulatory QTLs}

\section*{Code availability}

The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/eqtl2gwas}.
%
The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.

\section*{Data availability}

The raw colocalization data before filtering in a single file is available at this URL: \url{https://osf.io/hvmje}.
%
A website to access colocalization variants with PP.H4.abf $\geq$ 0.8 is available at this URL \url{https://agonzalez.pythonanywhere.com}.

\section*{Acknowledgements}

Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
%
I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.

\section*{References}
\bibliographystyle{unsrt}
\bibliography{ms_pleiotropy.bib}

### 2022-08-18 ms/ms02_floats.tex
ENH ms

@@
 \begin{table*}[]
-  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories.}\label{tab:pleitropic_variants}
+  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
 \centering
@@
 %
-\caption{Gene ontology analysis of egenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to egenes of variants associated to one GWAS category.} \label{fig:gene_ontology}
+\caption{Gene ontology analysis of egenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to egenes of variants associated to one GWAS category.} \label{fig:geneontology}
 \end{figure*}
@@
 \begin{table*}[]
-\caption{Pleiotropic regions involving 5 or more GWAS categories.}\label{tab:pleiotropic_regions}
+\caption{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
 \centering
@@
 
-\caption{Count of GWAS categories (\textbf{a-c}), egenes (\textbf{d-f}) and etissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1)} \label{fig:region_gwas_egenes_tissues}
+\caption{Count of GWAS categories (\textbf{a-c}), egenes (\textbf{d-f}) and etissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
+Genomic coordinates are given for the hg38 assembly.} \label{fig:region_gwas_egenes_tissues}
 %

### 2022-08-18 scripts/pltbox_x_per_variant_egene_etissue_y_gwas.py

@@ -77,7 +77,7 @@ xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
 title = "GWAS per variant-egene-etissue"
 xlabel = "GWAS category count"
-ylabel = "GWAS trait count"
+ylabel = "GWAS pheno. count"
 y = "gwas_trait_count"
 x = "gwas_category_count"
 

### 2022-08-19 ms/ms01_main.tex
ENH added crm analysis

@@
 
-eQTLs
-
 A large number of GWAS variants are found in non-conding regions.
@@
 
-\subsection*{Which regulatory variants are the most pleiotropic?}
+\subsection*{Pleiotropic variants colocalized with eQTLs}
+
+\subsubsection*{Which variants are the most pleiotropic?}
 
@@
 
-\subsection*{What is the function of pleiotropic variants?}
+\subsubsection*{What is the function of pleiotropic variants?}
 
@@
 
-\subsection*{Analysis of pleiotropic regions}
+\subsection*{Pleiotropic regions}
 
@@
 The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).
+
+%TODO cumulated covering of genome by pleitopic
 
@@
 
+%TODO egene per variant-etissue
 Then I hypothesized that pleiotropic variants modulate more egenes.
@@
 
+%TODO idea that gene expression is causal to disease
+
 % Confirm eQTL studies
@@
 %
-This pipeline and the results is useful for three aspects.
+For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
+%
+The results of this pipeline provides three types of insights.
 %
@@
 
-In this work, I have explore variants with a role in gene regulation that are highly pleiotropic.
-%
-For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
+% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
+%
+Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
+
+These variants are under control of more transcription factors
 

### 2022-08-19 ms/ms02_floats.tex
ENH added crm analysis

@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/missense_variant.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/missense_variant.png}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/splice_region_variant.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/splice_region_variant.png}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/plt_vep_consequence.py/3_prime_UTR_variant.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_y_egene.py/plt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_etissue_y_egene.py/plt.png}
 \end{subfigure}
@@
 %
-\caption{Count of egenes per variant (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
+\caption{Count of egenes per variant-tissue (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
 %
@@
 \begin{subfigure}[]{.33\textwidth}
-%
+\textbf{a}
+\\
 \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_0.png}
@@
 %
-\caption{Binding of transcription factors in the region around pleiotropic variants (Radius 50 bp) as a function of the GWAS category count.} \label{fig:freq_tf_per_variant}
+\begin{subfigure}[]{.33\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_remap_crm_gwas_categories.py/remap_crm_fisher.png}
+\end{subfigure}
+%
+\caption{Analysis of transcription factor binding.
+\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS category count.
+\textbf{b}, Odds ratio of cis regulatory module-annotated vs non-annotated variants.} \label{fig:freq_tf_per_variant}
 %

### 2022-08-19 scripts/plt_overlap_crm_remap.py

@@ -1,92 +1,0 @@
-"""This script will compute a tissue enrichment in a given GWAS"""
-import os
-import pathlib
-import pandas
-import seaborn
-from statannot import add_stat_annotation
-
-from eqtl2gwas_pleiotropy.Logger import Logger
-from eqtl2gwas_pleiotropy.PathManager import PathManager
-from eqtl2gwas_pleiotropy.URL import URL
-from matplotlib import pyplot as plt
-from eqtl2gwas_pleiotropy.constants import label_fontsize, tick_fontsize, dpi
-from eqtl2gwas_pleiotropy.constants import seaborn_theme_dic
-
-seaborn.set_theme(**seaborn_theme_dic)
-
-#%% download gencode annotation
-url = "https://remap.univ-amu.fr/storage/remap2022/hg38/MACS2/remap2022_crm_macs2_hg38_v1_0.bed.gz"
-remap_crm_path = URL(url).download()
-
-#%% outdir path
-if not '__file__' in locals():
-    __file__ = "plt_overlap_crm_remap.py"
-outdir_path = os.path.join(PathManager.get_project_path(), "out", os.path.basename(__file__))
-pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
-
-#%% input dir cmpt_count_per_rsid
-cmpt_count_per_rsid_indir_path = os.path.join(PathManager.get_outdir_path(), "cmpt_count_per_rsid.py")
-
-#%% input dir cmpt_overlap_crm_remap
-cmpt_overlap_crm_remap_indir_path = os.path.join(PathManager.get_outdir_path(), "intrsct_remap_nr_gwas_categories.py")
-
-#%% barplot with proportion of variants in CRMs
-x_lst = []
-y_lst = []
-for count_pleio in range(1, 6):
-    bed_path = os.path.join(cmpt_count_per_rsid_indir_path, "variant_pleio_{}_flank_0.bed".format(count_pleio))
-    intersect_bed_path = os.path.join(cmpt_overlap_crm_remap_indir_path, "remap_crm_" + os.path.basename(bed_path))
-    count_variants = sum(1 for line in open(bed_path))
-    count_intersection = sum(1 for line in open(intersect_bed_path))
-    x_lst.append(count_pleio)
-    y_lst.append(count_intersection/count_variants)
-
[…]

### 2022-08-19 scripts/pltbox_x_per_variant_egene_y_etissue.py

@@ -1,16 +1,15 @@
-import seaborn
-from statannot import add_stat_annotation
 from statannotations.Annotator import Annotator
+from eqtl2gwas_pleiotropy.constants import label_fontsize, dpi, tick_fontsize, boxplot_kwargs, annotator_config_dic
 
-from eqtl2gwas_pleiotropy.PathManager import PathManager
 
 import os
 import pandas
 import pathlib
+import seaborn
 import sys
 import matplotlib.pyplot as plt
-from eqtl2gwas_pleiotropy.constants import label_fontsize, dpi, tick_fontsize, boxplot_kwargs, annotator_config_dic
 
+#%%
 plt.rcParams["figure.figsize"] = (8, 6)
 from eqtl2gwas_pleiotropy.constants import seaborn_theme_dic
 seaborn.set_theme(**seaborn_theme_dic)
@@ -55,7 +54,7 @@ gwas_category_count_max_int = count_per_rsid_gwas_df['gwas_category_count'].max(
 m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos', 'rsid'])
 
 # %%
-sel_cols = ['rsid', 'egene', 'etissue_category']  # tissue per egene
+sel_cols = ['rsid', 'egene', 'etissue_category']  # tissue per variant-egene
 
 #%% set upper_var_gwas_cat_count
 m_df = m_df[sel_cols + ['gwas_category_count']]

### 2022-08-19 scripts/pltbox_x_per_variant_etissue_y_egene.py

@@ -1,5 +1,5 @@
 from statannotations.Annotator import Annotator
-
+from eqtl2gwas_pleiotropy.constants import seaborn_theme_dic
 from eqtl2gwas_pleiotropy.constants import label_fontsize, tick_fontsize, dpi, boxplot_kwargs, annotator_config_dic
 
 import matplotlib.pyplot as plt
@@ -10,9 +10,9 @@ import seaborn
 import sys
 
 
+#%%
 # Plot parameters
 plt.rcParams["figure.figsize"] = (8, 6)
-from eqtl2gwas_pleiotropy.constants import seaborn_theme_dic
 seaborn.set_theme(**seaborn_theme_dic)
 
 #%%
@@ -55,7 +55,7 @@ gwas_category_count_max_int = count_per_rsid_gwas_df['gwas_category_count'].max(
 m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos', 'rsid'])
 
 # %%
-sel_cols = ['rsid', 'egene']  # egene per variant
+sel_cols = ['rsid', 'egene', 'etissue_category']  # egene per variant-etissuecategory
 
 #%%
 m2_df = m_df[['chrom', 'pos'] + sel_cols + ['egene_symbol', 'gwas_category_count']].drop_duplicates()
@@ -69,9 +69,9 @@ m_df.loc[m_df['gwas_category_count'] >= upper_var_gwas_cat_count, "gwas_category
 
 #%%
 m_df = m_df.drop_duplicates()
-m_df = m_df.groupby(['rsid', 'gwas_category_count']).count()
+m_df = m_df.groupby(['rsid', 'etissue_category', 'gwas_category_count']).count()
 m_df = m_df.reset_index()
-m_df.columns = ['rsid', 'gwas_category_count', 'egene_count']
+m_df.columns = ['rsid', 'etissue_category', 'gwas_category_count', 'egene_count']
 
 #%%
 describe_tsv_path = os.path.join(outdir_path, "describe.tsv")
@@ -81,7 +81,7 @@ m_df.groupby('gwas_category_count')['egene_count'].apply(lambda x: x.describe())
 order = [str(x) for x in range(1, upper_var_gwas_cat_count+1)]
 xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
-title = "eGenes per variant"
+title = "eGenes per variant-etissue"
 xlabel = "GWAS category count"
 ylabel = "eGene count"
 y = "egene_count"

### 2022-08-19 ms/ms02_floats.tex
ENH added region cumsum

@@
 %
-\includegraphics[width=0.33\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
+\begin{subfigure}[]{.33\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
+\end{subfigure}
 %
-\caption{Length distribution of pleiotropic regions.} \label{fig:pleiotropy_region_distribution}
+\begin{subfigure}[]{.33\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
+\end{subfigure}
+%
+\caption{Analysis of pleiotropic regions.
+\textbf{a}, Length distribution of pleiotropic regions.
+\textbf{b}, Cumulative sum ordered inversely with the number of GWAS categories.} \label{fig:pleiotropy_region_distribution}
 \end{figure*}

### 2022-08-19 ms/ms01_main.tex
ENH added region cumsum
ENH minor

@@
 %
-75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}).
+75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
 
@@
 
-%TODO cumulated covering of genome by pleitopic
+Very pleiotropic regions with 6 GWAS categories or more make less than 1Mb of the genome, while moderate pleiotropic regions with 3 GWAS categories or more make 2 Mb of the genome (Fig.
+\ref{fig:pleiotropy_region_distribution}b.
 
@@
 
-\subsubsection*{More egenes per variant?}
+\subsubsection*{More egenes per variant-etissue?}
 
-%TODO egene per variant-etissue
-Then I hypothesized that pleiotropic variants modulate more egenes.
+Then I hypothesized that pleiotropic variant-etissue pairs modulate more egenes.
 %
-I found that pleiotropic variants have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+I found that pleiotropic variant-etissues pairs have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
 %

### 2022-08-19 ms/ms01_main.tex
ENH added tissue enrichment in gwas

@@
 
-125 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
+127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
 
@@
 
-%TODO CRMs and variants
-
 \subsubsection*{More GWAS per variant-egene-etissue?}
@@
 
-%TODO idea that gene expression is causal to disease
-
-% Confirm eQTL studies
-In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
+In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
 %
@@
 %
-% TODO how many
 Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.
@@
 Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
+
+% Second insight: causal variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Third insight: mechanistic information for GWAS variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2022-08-19 scripts/cmpt_tissue_enrich_in_gwas.py

@@ -1,0 +1,89 @@
+"""This script will compute a tissue enrichment in a given GWAS with the following algorithm.
+Count all loci of each gwas and of each eqtl
+Then count the proportion of loci of a given eqtl in a given gwas
+Binomial test to evaluate significance
+FDR 0.05"""
+import os
+
+import numpy
+import pandas
+import pathlib
+import sys
+
+from scipy.stats import binomtest
+from statsmodels.stats.multitest import fdrcorrection
+from eqtl2gwas_pleiotropy.PathManager import PathManager
+
+
+#%%
+help_cmd_str = "todo"
+try:
+    h4_tsv_path = sys.argv[1]
+    tsv_path = sys.argv[2]
+    if len(sys.argv) > 3:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+
+# #%% Parameters
+# if not '__file__' in locals():
+#     __file__ = "cmpt_tissue_enrich_in_gwas.py"
+outdir_path = os.path.dirname(tsv_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
+
+#%%
+# h4_tsv_path = "/home/gonzalez/Repositories/eqtl2gwas_pleiotropy/out/gwas413/genome/5e-08/1000000/filter_h4.py/h4.tsv"
+df_raw = pandas.read_csv(h4_tsv_path, sep="\t")
+
+#%%
+columns = ['chrom', 'pos', 'rsid', 'eqtl_identifier', 'etissue_category', 'gwas_identifier', 'gwas_trait', 'gwas_category']
+df = df_raw[columns].drop_duplicates()
+
+#%% nb loci for each gwas
+ngwas_df = df[['chrom', 'pos', 'rsid', 'gwas_identifier', 'gwas_trait']].drop_duplicates().groupby(['gwas_identifier', 'gwas_trait']).size().reset_index()
+ngwas_df.rename({0: 'count_gwas'}, axis=1, inplace=1)
[…]

### 2022-08-21 ms/ms01_main.tex
ONGOING

@@
 
-For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
+For a posterior probability PP.H4.abf$\ge$0.8 of common causal eQTL and GWAS variants, we found 100,526 loci.
+%
+If I focus in causal variants with PP.H4.abf$\ge$0.8 and SNP.PP.H4$\ge$0.5, then I find 4684 variants.
 %

### 2022-08-21 scripts/filter_h4.py

@@ -19,7 +19,8 @@ except IndexError:
     {}""".format(help_cmd_str))
     sys.exit(1)
 
-pathlib.Path(os.path.dirname((h4_tsv_gz_path))).mkdir(parents=True, exist_ok=True)
+outdir_path = os.path.dirname(h4_tsv_gz_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 #%% input
 if not os.path.isfile(annotated_tsv_gz_path):
@@ -30,4 +31,19 @@ if not os.path.isfile(annotated_tsv_gz_path):
 coloc_df = pandas.read_csv(annotated_tsv_gz_path, sep="\t")
 coloc_df.drop(['id'], inplace=True, axis=1)
 coloc_df = coloc_df.loc[coloc_df['PP.H4.abf'] >= h4_cutoff, ]
+
+#%% how many coloc loci?
+loci_h4_df = coloc_df.sort_values(by=['PP.H4.abf', 'SNP.PP.H4'], ascending=False)
+loci_h4_df = loci_h4_df.drop_duplicates(subset=['PP.H4.abf', 'coloc_window', 'nsnps'], keep='first')
+loci_coloc_h4_path = os.path.join(outdir_path, "loci_h4.tsv")
+loci_h4_df.to_csv(loci_coloc_h4_path, sep="\t", index=False)
+
+#%% how many causal variants, SNP.PP.H4>=0.5?
+variants_h4_df = coloc_df.sort_values(by=['SNP.PP.H4'], ascending=False)
+variants_h4_df = variants_h4_df.drop_duplicates(subset=['chrom', 'pos', 'cytoband', 'rsid'], keep='first')
+variants_h4_df = variants_h4_df.loc[variants_h4_df['SNP.PP.H4'] >= 0.5, ]
+variants_h4_path = os.path.join(outdir_path, "variants_h4.tsv")
+variants_h4_df.to_csv(variants_h4_path, sep="\t", index=False)
+
+#%%
 coloc_df.to_csv(h4_tsv_gz_path, sep="\t", index=False)

### 2022-08-22 scripts/cmpt_tissue_enrich_in_pleio.py
ENH added tissue enrichment

@@ -1,80 +1,0 @@
-"""This script will compute a tissue enrichment in a given GWAS with the following algorithm.
-Count all loci of each gwas and of each eqtl
-Then count the proportion of loci of a given eqtl in a given gwas
-Binomial test to evaluate significance
-FDR 0.05"""
-import os
-
-import numpy
-import pandas
-import pathlib
-import sys
-
-from scipy.stats import fisher_exact
-from statsmodels.stats.multitest import fdrcorrection
-
-
-#%%
-help_cmd_str = "todo"
-try:
-    h4_tsv_path = sys.argv[1]
-    count_per_rsid_gwas_tsv_path = sys.argv[2]
-    upper_var_gwas_cat_count = int(sys.argv[3])
-    tsv_path = sys.argv[4]
-    if len(sys.argv) > 5:
-        print("""Two many arguments!
-        {}""".format(help_cmd_str))
-        sys.exit(1)
-except IndexError:
-    print("""Argument missing!
-    {}""".format(help_cmd_str))
-    sys.exit(1)
-
-
-# #%% Parameters
-outdir_path = os.path.dirname(tsv_path)
-pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
-
-#%%
-h4_df = pandas.read_csv(h4_tsv_path, sep="\t")
-
-#%%
-count_per_rsid_gwas_df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
-
-#%%
-m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos', 'cytoband', 'rsid'])
-
-#%%
-m2_df = m_df[['rsid', 'gwas_category_count', 'etissue_category']].drop_duplicates().sort_values('gwas_category_count', ascending=False)
-m2_df.loc[m2_df['gwas_category_count'] >= upper_var_gwas_cat_count, 'gwas_category_count'] = upper_var_gwas_cat_count
-
-#%%
-g_target_df = m2_df.groupby(['gwas_category_count', 'etissue_category']).count().reset_index()
-
-#%%
-g_backg_df = m2_df.groupby(['gwas_category_count']).count().reset_index()[['gwas_category_count', 'rsid']]
-
-#%%
-g_target_df = g_target_df.merge(g_backg_df, on='gwas_category_count')
[…]

### 2022-08-27 ms/ms01_titlepage.tex
ENH ms

--- 2022-08-19 ms/ms01_main.tex
@@
-\leadauthor{Scientist}
+%\leadauthor{Scientist}
 
 \title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
-\shorttitle{Pleiotropy and gene regulatory variants}
+%\shorttitle{Pleiotropy and gene regulatory variants}
 
-\author[1,\Letter]{Aitor González \orcidlink{0000-0002-8402-1655}}
+\author[1,*]{Aitor González\,\orcidlink{0000-0002-8402-1655}}
+
+%\author[1,\Letter]{Aitor González}
 \affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
+\affil[*]{Corresponding author: aitor.gonzalez@univ-amu.fr}
+% \orcidlink{0000-0002-8402-1655}
+%\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
 \date{}
@@
 \end{corrauthor}
-
-\section*{Introduction}\label{sec:introduction}
-
-Genome-wide association studies (GWAS) evaluate the impact of frequent variants on a trait in the population.
-%
-The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
-%
-For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71673 variant-trait associations from 3567 publications \citep{2018.Parkinson.Buniello}.
-
-However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
-%
-First, identification of the causaul variant is difficult because of linkage disequilibrium.
-%
-Second, GWAS do not provide hints regarding the tissues involved in the phenotypes.
-%
-Third, GWAS variants affect most often non-coding regions.
-
-Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
-%
-eQTLs have some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
-%
-On the other hand, eQTLs provide information about the target genes and tissues.
-
-A large number of GWAS variants are found in non-conding regions.
-To priori
-
-Sentence describing EBI eQTL
-
-In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.
-
-\section*{Results}\label{s:results}
-
-\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}
-
-To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
-
-125 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
-
-Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
-%
-The first criterion was to exclude molecular phenotypes such as proteome or methylome.
-%	
-The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
-%
-The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental phenotypes such "employment status" or "self-reported" medical conditions.
-%
-The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
-%
-These filters resulted in 413 GWAS (Supplementary table 2).
-
-I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
-%
-This analysis explored 30 261 unique variants and 2,479,065 potential variant colocalizations (OSF URL).
-%
-Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
-
-\subsection*{Pleiotropic variants colocalized with eQTLs}
-
-\subsubsection*{Which variants are the most pleiotropic?}
-
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
-%
-In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
-%
-Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).
-
-These category annotations were used to investigate the variant pleiotropy for GWAS phenotypes.
-%
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS categories they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-%
-Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-
-% 12q24.12
-For instance, the most pleiotropic variants in cytoband 12q24.12 are involved in allergies, cancer, cardiovascular and autoimmune diseases (Table \ref{tab:pleitropic_variants}).
-%
-These variants control egenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
-%
-These eQTLs are active in adipose tissue, arteries, blood, colon, immune cells, skin and induced pluripotent stem cells (iPSC).
-
-\subsubsection*{What is the function of pleiotropic variants?}
-
-Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
-%
-I analyzed the ontology of egenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
-%
-I found that egenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-For instance, egenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
-%
-IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
-%
-Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
-%
-These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
-
-%Variants of the 3q23 cytoband are involved in allergy, cancer and height phenotypes (Supplementary table 3).
-%%
-%Interestingly all these variants target the egene ZBTB3.
-%%
-%ZBTB3 is a transcription factor that controls pro-inflammatory factors such as IRF5 (10.1016/j.jaut.2016.08.001).
-%%
-%These variants are active in tissues such as blood, esophagus, immune cell, LCLs and muscle (Supplementary table 3).
-
-%%11q13.5
-%The variants in cytoband 11q13.5 are involved in allergy and autoimmune disease (Supplementary table 3).
-%%
-%All these variants target the egene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
-%%
-%These variants are active in LCL, Skin and iPSC (Supplementary table 3).
-%
-%In summary, I have identified pleiotropic variants with gene regulatory function.
-%%
-%Some of these loci are strongly involved in the immune system.
-
-\subsection*{Pleiotropic regions}
-
-Pleiotropic variants are aggregated in few cytobands.
-%
-Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions}).
-%
-I found 453 regions with 441 regions with 2 or more categories, 114 regions with 3 or more categories, 37 regions with 4 or more categories and 18 regions with 5 or more categories (Supplementary table 5).
-%
-75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
-
-The most pleiotropic region is 12:111,395,984-111,645,358 in cytoband 12q24.12 that we discussed above (Table \ref{tab:pleiotropic_regions}).
-%
-The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension phenotypes (Supplementary table 5).
-%
-The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).
-
-Very pleiotropic regions with 6 GWAS categories or more make less than 1Mb of the genome, while moderate pleiotropic regions with 3 GWAS categories or more make 2 Mb of the genome (Fig.
-\ref{fig:pleiotropy_region_distribution}b.
-
-\subsection*{How specific are variants to GWAS phenotypes, egenes and etissues?}
-
-Next I explored the phenotype, egene and etissue specificity of eQTL/GWAS variants.
-%
-83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more phenotype categories make 3\% or less (Fig.
-\ref{fig:hist_gwas_egene_etissue}(a).
-%
-Regarding egenes, half of variants modulate one egene, and the other half modulate two or more egenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
-%
-In the case of etissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
-%
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS phenotype. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do phenotype frequencies relates to eQTL and etissue frequency at the loci level?}
-
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, egenes and eTissues for regulatory variants in three pleiotropic regions.
-%
-All three regions are very pleiotropic with very different associated phenotype categories such as cancer, cardiovascular and autoimmune diseases.
-%
-However these plots show that the relationship between the number of phenotypes, egenes and etissues can be very different.
-%
-The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of egenes but a high number of etissues.
-%
-By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both egenes and etissues.
-%
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest phenotype pleiotropy but the number of egenes and etissues is very low.
-%
-In summary, phenotype pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
-
-\subsection*{What are the mechanisms of pleiotropy}
-
-Then, I wanted to understand the molecular mechanism of pleiotropic regions.
-
-\subsubsection*{More severe variant effect consequences?}
-
-I first analysed whether there are significant differences of effect consequences between more and less pleiotropic variants.
-%
-I separated variants according to the count of GWAS categories and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
-%
-I found a significant increase of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
-%
-I also found a significant increase of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
-%
-Finally, there is also a significant increase of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
-%
-These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
-
-\subsubsection*{More egenes per variant-etissue?}
-
-Then I hypothesized that pleiotropic variant-etissue pairs modulate more egenes.
-%
-I found that pleiotropic variant-etissues pairs have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).
-
-\subsubsection*{More etissues per variant-egene?}
-
-My next hypothesis was that pleiotropic variant-egene pairs are active in more tissues compared with variant-egene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
-%
-Indeed, I found that pleiotropic variant-egenes pairs are associated with a higher number of etissues (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
-
-\subsubsection*{How to explain that pleiotropic variants are active in more tissues?}
-
-I hypothesized that pleiotropic variants are active in more etissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
-%
-I looked at the number of unique transcription factors bound in a window of 100 bp around each variant.
-%
-I found a significant increase of unique transcription factors around pleagonzaleziotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
-
-%TODO CRMs and variants
-
-\subsubsection*{More GWAS per variant-egene-etissue?}
-
-Another molecular mechanism of pleiotropy is that egenes correlate with more GWAS categories even after taking into account variants and etissues.
-%
-I computed the number of GWAS categories per variant-egene-etissue triplet.
-%
-Indeed I found a significant increase of GWAS categories for pleiotropic variant-egene-etissue triplets (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
-
-\subsection*{What is the effect size of pleiotropic eQTLs?}
-
-Then I wondered, how does the effect size (beta) and significance (p-value) relate to pleiotropy.
-%
-I found a significant decrease of effect size of eQTLs and GWAS variants with the exception of the GWAS effect size of variants with 3 categories (Fig. \ref{fig:beta_pval}a,b).
-
-Then I looked at the significance of pleiotropic eQTLs and GWAS variants.
-%
-eQTL significance significantly decreased for pleiotropic variants (Fig. \ref{fig:beta_pval}c).
-%
-By contrast, the significance of GWAS associations increased for GWAS variants (Fig. \ref{fig:beta_pval}d).
-
-In summary, pleiotropic eQTLs are weaker with a weaker beta and less significant p-values..
-%
-On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.
-
-\section*{Discussion}
-
-%TODO idea that gene expression is causal to disease
-
-% Confirm eQTL studies
-In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
-%
-For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
-%
-The results of this pipeline provides three types of insights.
-%
-First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
-%
-Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
-%
-% TODO how many
-Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.
-
-% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
-%
-Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
-
-These variants are under control of more transcription factors
-
-%\subsection*{What is the function of the variants in the most pleiotropic regions?}
-%
-%% 12q24.12
-%The cytoband 12q24.12 has five variants rs3184504, rs653178, rs7310615, rs597808, rs11065979, that are known in the literature to be associated to a wide palette of diseases.
-%This locus has been used to explain links between inflammaiton and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
-%%
-%Our analysis add up that the variants in this cytoband control important genes such as ALDH2, ATXN2, MAPKAPK5, SH2B3 and TMEM116 in several tissues such as adipose, artery, blood, brain, digestive, skin, heart and skin tissue as well as immune cells.
-%
-%% 6p21.3
-%% TODO
-%
-%%15q26.1
-%The next pleiotropic locus is in cytoband 15q26.1 involved in cardiovascular, hypertension and psyciatric diseases with variants rs2071382, rs8039305 and rs6224.
-%%
-%eQTLs in cytoband 15q26.1 control genes like Furin, oncogene Fes and UNC45A involved in Osteootohepatoenteric syndrome (OMIM DB).
-%%
-%eQTLs are active in a very large number of cells including blood, artery, brain and immune cells (See supplementary table).
-%
-%% 1p13.3
-%The locus in cytoband 1p13.3 is mainly related to cardiovascular diseases.
-%Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
-%
-%
-%
-%The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
-%
-%The variant rs11236797 and the two others in locus 11q13.5 is mainly autoimmune.
-%All three regulate the gene EMSY-DT and are active induced pluripotent cells, LCDs and skin.
-%
-%Another pleiotropic variants rs301806
-%rs301807
-%rs301802
-%rs301817
-%rs159963
-%rs301816
-%in the cytoband 1p36.23 that are involved in allergy, cardiovascular and hypertension.
-%This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\
-
-\section*{Methods}\label{sec:methods}
-
-\subsection*{Data sets and data exploration}
-
-I used these data sets here: eQTLs from the eQTL Catalogue \citep{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \citep{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \citep{2021.Ballester.Hammal}, and UCSC annotation data.
-%
-I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \citep{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
-%
-Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
-%
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/eqtl2gwas} and \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\subsection*{Colocalization}
-
-I developed a colocalization pipeline based on a pipeline that is public at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
-%
-First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants in a radius of 500 000 nt are retrieved.
-%
-Colocalization between eQTL and GWAS variants is tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
-
-\subsection*{Definition of pleiotropic regions}
-
-To define pleiotropic regions, I included "pleiotropic" variants defining as having more than 2 categories as long as the distance between the two "pleiotropic" variants is less than 100 000 bp.
-%
-Then the number of GWAS categories in Supplementary table TODO is given by the number of different variant categories in the region.
-
-\subsection*{Characterization of regulatory QTLs}
-
-\section*{Code availability}
-
-The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/eqtl2gwas}.
-%
-The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\section*{Data availability}
-
-The raw colocalization data before filtering in a single file is available at this URL: \url{https://osf.io/hvmje}.
-%
-A website to access colocalization variants with PP.H4.abf $\geq$ 0.8 is available at this URL \url{https://agonzalez.pythonanywhere.com}.
-
-\section*{Acknowledgements}
-
-Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
-%
-I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.
-
-\section*{References}
-\bibliographystyle{unsrt}
-\bibliography{ms_pleiotropy.bib}

### 2022-08-27 ms/ms02_intro.tex
ENH ms

--- 2022-08-19 ms/ms01_main.tex
@@
-\leadauthor{Scientist}
-
-\title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
-\shorttitle{Pleiotropy and gene regulatory variants}
-
-\author[1,\Letter]{Aitor González \orcidlink{0000-0002-8402-1655}}
-\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
-\date{}
-
-\maketitle
-
-\begin{abstract}
-
-Pleiotropic genetic variants affecting several phenotypes are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
-%
-Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
-%
-In this work, I investigate the link between gene regulation and phenotype pleiotropy for frequent variants.
-
-% Verify numbers
-First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
-%
-I have categorized the GWAS phenotypes into 96 phenotypes and categories to define pleiotropic as belonging at least to two of these categories.
-%
-Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more phenotypes.
-%
-These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
-%
-These variants with the GWAS phenotypes, egenes and etissues can be explored in a public website.
-
-In conclusion, our work suggest that pleiotropy of gene regulatory variants arise from an increase of bound transcription factors, target egenes and etissues and severity of molecular effect consequences.
-
-\end{abstract}
-
-\begin{keywords}
-keyword1 | keyword2 | keyword3
-\end{keywords}
-
-\begin{corrauthor}
-aitor.gonzalez\at univ-amu.fr
-\end{corrauthor}
-
 \section*{Introduction}\label{sec:introduction}
@@
 In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.
-
-\section*{Results}\label{s:results}
-
-\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}
-
-To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
-
-125 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
-
-Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
-%
-The first criterion was to exclude molecular phenotypes such as proteome or methylome.
-%	
-The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
-%
-The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental phenotypes such "employment status" or "self-reported" medical conditions.
-%
-The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
-%
-These filters resulted in 413 GWAS (Supplementary table 2).
-
-I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
-%
-This analysis explored 30 261 unique variants and 2,479,065 potential variant colocalizations (OSF URL).
-%
-Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
-
-\subsection*{Pleiotropic variants colocalized with eQTLs}
-
-\subsubsection*{Which variants are the most pleiotropic?}
-
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
-%
-In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
-%
-Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).
-
-These category annotations were used to investigate the variant pleiotropy for GWAS phenotypes.
-%
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS categories they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-%
-Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-
-% 12q24.12
-For instance, the most pleiotropic variants in cytoband 12q24.12 are involved in allergies, cancer, cardiovascular and autoimmune diseases (Table \ref{tab:pleitropic_variants}).
-%
-These variants control egenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
-%
-These eQTLs are active in adipose tissue, arteries, blood, colon, immune cells, skin and induced pluripotent stem cells (iPSC).
-
-\subsubsection*{What is the function of pleiotropic variants?}
-
-Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
-%
-I analyzed the ontology of egenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
-%
-I found that egenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-For instance, egenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
-%
-IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
-%
-Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
-%
-These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
-
-%Variants of the 3q23 cytoband are involved in allergy, cancer and height phenotypes (Supplementary table 3).
-%%
-%Interestingly all these variants target the egene ZBTB3.
-%%
-%ZBTB3 is a transcription factor that controls pro-inflammatory factors such as IRF5 (10.1016/j.jaut.2016.08.001).
-%%
-%These variants are active in tissues such as blood, esophagus, immune cell, LCLs and muscle (Supplementary table 3).
-
-%%11q13.5
-%The variants in cytoband 11q13.5 are involved in allergy and autoimmune disease (Supplementary table 3).
-%%
-%All these variants target the egene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
-%%
-%These variants are active in LCL, Skin and iPSC (Supplementary table 3).
-%
-%In summary, I have identified pleiotropic variants with gene regulatory function.
-%%
-%Some of these loci are strongly involved in the immune system.
-
-\subsection*{Pleiotropic regions}
-
-Pleiotropic variants are aggregated in few cytobands.
-%
-Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions}).
-%
-I found 453 regions with 441 regions with 2 or more categories, 114 regions with 3 or more categories, 37 regions with 4 or more categories and 18 regions with 5 or more categories (Supplementary table 5).
-%
-75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
-
-The most pleiotropic region is 12:111,395,984-111,645,358 in cytoband 12q24.12 that we discussed above (Table \ref{tab:pleiotropic_regions}).
-%
-The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension phenotypes (Supplementary table 5).
-%
-The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).
-
-Very pleiotropic regions with 6 GWAS categories or more make less than 1Mb of the genome, while moderate pleiotropic regions with 3 GWAS categories or more make 2 Mb of the genome (Fig.
-\ref{fig:pleiotropy_region_distribution}b.
-
-\subsection*{How specific are variants to GWAS phenotypes, egenes and etissues?}
-
-Next I explored the phenotype, egene and etissue specificity of eQTL/GWAS variants.
-%
-83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more phenotype categories make 3\% or less (Fig.
-\ref{fig:hist_gwas_egene_etissue}(a).
-%
-Regarding egenes, half of variants modulate one egene, and the other half modulate two or more egenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
-%
-In the case of etissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
-%
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS phenotype. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do phenotype frequencies relates to eQTL and etissue frequency at the loci level?}
-
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, egenes and eTissues for regulatory variants in three pleiotropic regions.
-%
-All three regions are very pleiotropic with very different associated phenotype categories such as cancer, cardiovascular and autoimmune diseases.
-%
-However these plots show that the relationship between the number of phenotypes, egenes and etissues can be very different.
-%
-The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of egenes but a high number of etissues.
-%
-By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both egenes and etissues.
-%
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest phenotype pleiotropy but the number of egenes and etissues is very low.
-%
-In summary, phenotype pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
-
-\subsection*{What are the mechanisms of pleiotropy}
-
-Then, I wanted to understand the molecular mechanism of pleiotropic regions.
-
-\subsubsection*{More severe variant effect consequences?}
-
-I first analysed whether there are significant differences of effect consequences between more and less pleiotropic variants.
-%
-I separated variants according to the count of GWAS categories and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
-%
-I found a significant increase of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
-%
-I also found a significant increase of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
-%
-Finally, there is also a significant increase of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
-%
-These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
-
-\subsubsection*{More egenes per variant-etissue?}
-
-Then I hypothesized that pleiotropic variant-etissue pairs modulate more egenes.
-%
-I found that pleiotropic variant-etissues pairs have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).
-
-\subsubsection*{More etissues per variant-egene?}
-
-My next hypothesis was that pleiotropic variant-egene pairs are active in more tissues compared with variant-egene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
-%
-Indeed, I found that pleiotropic variant-egenes pairs are associated with a higher number of etissues (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
-
-\subsubsection*{How to explain that pleiotropic variants are active in more tissues?}
-
-I hypothesized that pleiotropic variants are active in more etissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
-%
-I looked at the number of unique transcription factors bound in a window of 100 bp around each variant.
-%
-I found a significant increase of unique transcription factors around pleagonzaleziotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
-
-%TODO CRMs and variants
-
-\subsubsection*{More GWAS per variant-egene-etissue?}
-
-Another molecular mechanism of pleiotropy is that egenes correlate with more GWAS categories even after taking into account variants and etissues.
-%
-I computed the number of GWAS categories per variant-egene-etissue triplet.
-%
-Indeed I found a significant increase of GWAS categories for pleiotropic variant-egene-etissue triplets (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
-
-\subsection*{What is the effect size of pleiotropic eQTLs?}
-
-Then I wondered, how does the effect size (beta) and significance (p-value) relate to pleiotropy.
-%
-I found a significant decrease of effect size of eQTLs and GWAS variants with the exception of the GWAS effect size of variants with 3 categories (Fig. \ref{fig:beta_pval}a,b).
-
-Then I looked at the significance of pleiotropic eQTLs and GWAS variants.
-%
-eQTL significance significantly decreased for pleiotropic variants (Fig. \ref{fig:beta_pval}c).
-%
-By contrast, the significance of GWAS associations increased for GWAS variants (Fig. \ref{fig:beta_pval}d).
-
-In summary, pleiotropic eQTLs are weaker with a weaker beta and less significant p-values..
-%
-On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.
-
-\section*{Discussion}
-
-%TODO idea that gene expression is causal to disease
-
-% Confirm eQTL studies
-In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 125 eQTL studies.
-%
-For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
-%
-The results of this pipeline provides three types of insights.
-%
-First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
-%
-Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
-%
-% TODO how many
-Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.
-
-% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
-%
-Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
-
-These variants are under control of more transcription factors
-
-%\subsection*{What is the function of the variants in the most pleiotropic regions?}
-%
-%% 12q24.12
-%The cytoband 12q24.12 has five variants rs3184504, rs653178, rs7310615, rs597808, rs11065979, that are known in the literature to be associated to a wide palette of diseases.
-%This locus has been used to explain links between inflammaiton and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
-%%
-%Our analysis add up that the variants in this cytoband control important genes such as ALDH2, ATXN2, MAPKAPK5, SH2B3 and TMEM116 in several tissues such as adipose, artery, blood, brain, digestive, skin, heart and skin tissue as well as immune cells.
-%
-%% 6p21.3
-%% TODO
-%
-%%15q26.1
-%The next pleiotropic locus is in cytoband 15q26.1 involved in cardiovascular, hypertension and psyciatric diseases with variants rs2071382, rs8039305 and rs6224.
-%%
-%eQTLs in cytoband 15q26.1 control genes like Furin, oncogene Fes and UNC45A involved in Osteootohepatoenteric syndrome (OMIM DB).
-%%
-%eQTLs are active in a very large number of cells including blood, artery, brain and immune cells (See supplementary table).
-%
-%% 1p13.3
-%The locus in cytoband 1p13.3 is mainly related to cardiovascular diseases.
-%Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
-%
-%
-%
-%The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
-%
-%The variant rs11236797 and the two others in locus 11q13.5 is mainly autoimmune.
-%All three regulate the gene EMSY-DT and are active induced pluripotent cells, LCDs and skin.
-%
-%Another pleiotropic variants rs301806
-%rs301807
-%rs301802
-%rs301817
-%rs159963
-%rs301816
-%in the cytoband 1p36.23 that are involved in allergy, cardiovascular and hypertension.
-%This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\
-
-\section*{Methods}\label{sec:methods}
-
-\subsection*{Data sets and data exploration}
-
-I used these data sets here: eQTLs from the eQTL Catalogue \citep{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \citep{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \citep{2021.Ballester.Hammal}, and UCSC annotation data.
-%
-I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \citep{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
-%
-Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
-%
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/eqtl2gwas} and \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\subsection*{Colocalization}
-
-I developed a colocalization pipeline based on a pipeline that is public at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
-%
-First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants in a radius of 500 000 nt are retrieved.
-%
-Colocalization between eQTL and GWAS variants is tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
-
-\subsection*{Definition of pleiotropic regions}
-
-To define pleiotropic regions, I included "pleiotropic" variants defining as having more than 2 categories as long as the distance between the two "pleiotropic" variants is less than 100 000 bp.
-%
-Then the number of GWAS categories in Supplementary table TODO is given by the number of different variant categories in the region.
-
-\subsection*{Characterization of regulatory QTLs}
-
-\section*{Code availability}
-
-The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/eqtl2gwas}.
-%
-The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\section*{Data availability}
-
-The raw colocalization data before filtering in a single file is available at this URL: \url{https://osf.io/hvmje}.
-%
-A website to access colocalization variants with PP.H4.abf $\geq$ 0.8 is available at this URL \url{https://agonzalez.pythonanywhere.com}.
-
-\section*{Acknowledgements}
-
-Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
-%
-I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.
-
-\section*{References}
-\bibliographystyle{unsrt}
-\bibliography{ms_pleiotropy.bib}

### 2022-08-27 ms/ms02_results.tex
ENH ms

--- 2022-08-19 ms/ms01_main.tex
@@
 
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
+I manually assigned the 413 GWAS to 9
+6 categories to aggregate identical or similar phenotypes (Supplementary table 2).
 %
@@
 On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.
-
-\section*{Discussion}
-
-In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
-%
-For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
-%
-The results of this pipeline provides three types of insights.
-%
-First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
-%
-Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
-%
-Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.
-
-% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For a posterior probability $\ge$ 0.8 that the GWAS variants and the eQTLs are common, we found TODO variants.
-%
-Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
-
-% Second insight: causal variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Third insight: mechanistic information for GWAS variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-These variants are under control of more transcription factors
-
-%\subsection*{What is the function of the variants in the most pleiotropic regions?}
-%
-%% 12q24.12
-%The cytoband 12q24.12 has five variants rs3184504, rs653178, rs7310615, rs597808, rs11065979, that are known in the literature to be associated to a wide palette of diseases.
-%This locus has been used to explain links between inflammaiton and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
-%%
-%Our analysis add up that the variants in this cytoband control important genes such as ALDH2, ATXN2, MAPKAPK5, SH2B3 and TMEM116 in several tissues such as adipose, artery, blood, brain, digestive, skin, heart and skin tissue as well as immune cells.
-%
-%% 6p21.3
-%% TODO
-%
-%%15q26.1
-%The next pleiotropic locus is in cytoband 15q26.1 involved in cardiovascular, hypertension and psyciatric diseases with variants rs2071382, rs8039305 and rs6224.
-%%
-%eQTLs in cytoband 15q26.1 control genes like Furin, oncogene Fes and UNC45A involved in Osteootohepatoenteric syndrome (OMIM DB).
-%%
-%eQTLs are active in a very large number of cells including blood, artery, brain and immune cells (See supplementary table).
-%
-%% 1p13.3
-%The locus in cytoband 1p13.3 is mainly related to cardiovascular diseases.
-%Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
-%
-%
-%
-%The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
-%
-%The variant rs11236797 and the two others in locus 11q13.5 is mainly autoimmune.
-%All three regulate the gene EMSY-DT and are active induced pluripotent cells, LCDs and skin.
-%
-%Another pleiotropic variants rs301806
-%rs301807
-%rs301802
-%rs301817
-%rs159963
-%rs301816
-%in the cytoband 1p36.23 that are involved in allergy, cardiovascular and hypertension.
-%This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\
-
-\section*{Methods}\label{sec:methods}
-
-\subsection*{Data sets and data exploration}
-
-I used these data sets here: eQTLs from the eQTL Catalogue \citep{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \citep{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \citep{2021.Ballester.Hammal}, and UCSC annotation data.
-%
-I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \citep{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
-%
-Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
-%
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/eqtl2gwas} and \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\subsection*{Colocalization}
-
-I developed a colocalization pipeline based on a pipeline that is public at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
-%
-First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants in a radius of 500 000 nt are retrieved.
-%
-Colocalization between eQTL and GWAS variants is tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
-
-\subsection*{Definition of pleiotropic regions}
-
-To define pleiotropic regions, I included "pleiotropic" variants defining as having more than 2 categories as long as the distance between the two "pleiotropic" variants is less than 100 000 bp.
-%
-Then the number of GWAS categories in Supplementary table TODO is given by the number of different variant categories in the region.
-
-\subsection*{Characterization of regulatory QTLs}
-
-\section*{Code availability}
-
-The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/eqtl2gwas}.
-%
-The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
-
-\section*{Data availability}
-
-The raw colocalization data before filtering in a single file is available at this URL: \url{https://osf.io/hvmje}.
-%
-A website to access colocalization variants with PP.H4.abf $\geq$ 0.8 is available at this URL \url{https://agonzalez.pythonanywhere.com}.
-
-\section*{Acknowledgements}
-
-Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
-%
-I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.
-
-\section*{References}
-\bibliographystyle{unsrt}
-\bibliography{ms_pleiotropy.bib}

### 2022-08-27 ms/ms03_results.tex
ENH ms

--- 2022-08-19 ms/ms01_main.tex
@@
-\leadauthor{Scientist}
-
-\title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
-\shorttitle{Pleiotropy and gene regulatory variants}
-
-\author[1,\Letter]{Aitor González \orcidlink{0000-0002-8402-1655}}
-\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
-\date{}
-
-\maketitle
-
-\begin{abstract}
-
-Pleiotropic genetic variants affecting several phenotypes are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
-%
-Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
-%
-In this work, I investigate the link between gene regulation and phenotype pleiotropy for frequent variants.
-
-% Verify numbers
-First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
-%
-I have categorized the GWAS phenotypes into 96 phenotypes and categories to define pleiotropic as belonging at least to two of these categories.
-%
-Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more phenotypes.
-%
-These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
-%
-These variants with the GWAS phenotypes, egenes and etissues can be explored in a public website.
-
-In conclusion, our work suggest that pleiotropy of gene regulatory variants arise from an increase of bound transcription factors, target egenes and etissues and severity of molecular effect consequences.
-
-\end{abstract}
-
-\begin{keywords}
-keyword1 | keyword2 | keyword3
-\end{keywords}
-
-\begin{corrauthor}
-aitor.gonzalez\at univ-amu.fr
-\end{corrauthor}
-
-\section*{Introduction}\label{sec:introduction}
-
-Genome-wide association studies (GWAS) evaluate the impact of frequent variants on a trait in the population.
-%
-The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
-%
-For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71673 variant-trait associations from 3567 publications \citep{2018.Parkinson.Buniello}.
-
-However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
-%
-First, identification of the causaul variant is difficult because of linkage disequilibrium.
-%
-Second, GWAS do not provide hints regarding the tissues involved in the phenotypes.
-%
-Third, GWAS variants affect most often non-coding regions.
-
-Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
-%
-eQTLs have some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
-%
-On the other hand, eQTLs provide information about the target genes and tissues.
-
-A large number of GWAS variants are found in non-conding regions.
-To priori
-
-Sentence describing EBI eQTL
-
-In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.
-
 \section*{Results}\label{s:results}

### 2022-08-27 ms/ms04_discussion.tex
ENH ms

--- 2022-08-21 ms/ms01_main.tex
@@
-\leadauthor{Scientist}
-
-\title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
-\shorttitle{Pleiotropy and gene regulatory variants}
-
-\author[1,\Letter]{Aitor González \orcidlink{0000-0002-8402-1655}}
-\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
-\date{}
-
-\maketitle
-
-\begin{abstract}
-
-Pleiotropic genetic variants affecting several phenotypes are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
-%
-Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
-%
-In this work, I investigate the link between gene regulation and phenotype pleiotropy for frequent variants.
-
-% Verify numbers
-First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
-%
-I have categorized the GWAS phenotypes into 96 phenotypes and categories to define pleiotropic as belonging at least to two of these categories.
-%
-Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more phenotypes.
-%
-These observations could be partially explained, because pleitropic variants are enriched in splicing and missense variants and bind more transcription factors.
-%
-These variants with the GWAS phenotypes, egenes and etissues can be explored in a public website.
-
-In conclusion, our work suggest that pleiotropy of gene regulatory variants arise from an increase of bound transcription factors, target egenes and etissues and severity of molecular effect consequences.
-
-\end{abstract}
-
-\begin{keywords}
-keyword1 | keyword2 | keyword3
-\end{keywords}
-
-\begin{corrauthor}
-aitor.gonzalez\at univ-amu.fr
-\end{corrauthor}
-
-\section*{Introduction}\label{sec:introduction}
-
-Genome-wide association studies (GWAS) evaluate the impact of frequent variants on a trait in the population.
-%
-The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
-%
-For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71673 variant-trait associations from 3567 publications \citep{2018.Parkinson.Buniello}.
-
-However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
-%
-First, identification of the causaul variant is difficult because of linkage disequilibrium.
-%
-Second, GWAS do not provide hints regarding the tissues involved in the phenotypes.
-%
-Third, GWAS variants affect most often non-coding regions.
-
-Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
-%
-eQTLs have some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
-%
-On the other hand, eQTLs provide information about the target genes and tissues.
-
-A large number of GWAS variants are found in non-conding regions.
-To priori
-
-Sentence describing EBI eQTL
-
-In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.
-
-\section*{Results}\label{s:results}
-
-\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}
-
-To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
-
-127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
-
-Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
-%
-The first criterion was to exclude molecular phenotypes such as proteome or methylome.
-%	
-The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
-%
-The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental phenotypes such "employment status" or "self-reported" medical conditions.
-%
-The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
-%
-These filters resulted in 413 GWAS (Supplementary table 2).
-
-I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
-%
-This analysis explored 30 261 unique variants and 2,479,065 potential variant colocalizations (OSF URL).
-%
-Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
-
-\subsection*{Pleiotropic variants colocalized with eQTLs}
-
-\subsubsection*{Which variants are the most pleiotropic?}
-
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
-%
-In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
-%
-Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).
-
-These category annotations were used to investigate the variant pleiotropy for GWAS phenotypes.
-%
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS categories they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-%
-Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-
-% 12q24.12
-For instance, the most pleiotropic variants in cytoband 12q24.12 are involved in allergies, cancer, cardiovascular and autoimmune diseases (Table \ref{tab:pleitropic_variants}).
-%
-These variants control egenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
-%
-These eQTLs are active in adipose tissue, arteries, blood, colon, immune cells, skin and induced pluripotent stem cells (iPSC).
-
-\subsubsection*{What is the function of pleiotropic variants?}
-
-Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
-%
-I analyzed the ontology of egenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
-%
-I found that egenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-For instance, egenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
-%
-IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
-%
-Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
-%
-These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
-
-%Variants of the 3q23 cytoband are involved in allergy, cancer and height phenotypes (Supplementary table 3).
-%%
-%Interestingly all these variants target the egene ZBTB3.
-%%
-%ZBTB3 is a transcription factor that controls pro-inflammatory factors such as IRF5 (10.1016/j.jaut.2016.08.001).
-%%
-%These variants are active in tissues such as blood, esophagus, immune cell, LCLs and muscle (Supplementary table 3).
-
-%%11q13.5
-%The variants in cytoband 11q13.5 are involved in allergy and autoimmune disease (Supplementary table 3).
-%%
-%All these variants target the egene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
-%%
-%These variants are active in LCL, Skin and iPSC (Supplementary table 3).
-%
-%In summary, I have identified pleiotropic variants with gene regulatory function.
-%%
-%Some of these loci are strongly involved in the immune system.
-
-\subsection*{Pleiotropic regions}
-
-Pleiotropic variants are aggregated in few cytobands.
-%
-Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions}).
-%
-I found 453 regions with 441 regions with 2 or more categories, 114 regions with 3 or more categories, 37 regions with 4 or more categories and 18 regions with 5 or more categories (Supplementary table 5).
-%
-75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
-
-The most pleiotropic region is 12:111,395,984-111,645,358 in cytoband 12q24.12 that we discussed above (Table \ref{tab:pleiotropic_regions}).
-%
-The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension phenotypes (Supplementary table 5).
-%
-The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).
-
-Very pleiotropic regions with 6 GWAS categories or more make less than 1Mb of the genome, while moderate pleiotropic regions with 3 GWAS categories or more make 2 Mb of the genome (Fig.
-\ref{fig:pleiotropy_region_distribution}b.
-
-\subsection*{How specific are variants to GWAS phenotypes, egenes and etissues?}
-
-Next I explored the phenotype, egene and etissue specificity of eQTL/GWAS variants.
-%
-83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more phenotype categories make 3\% or less (Fig.
-\ref{fig:hist_gwas_egene_etissue}(a).
-%
-Regarding egenes, half of variants modulate one egene, and the other half modulate two or more egenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
-%
-In the case of etissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
-%
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS phenotype. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do phenotype frequencies relates to eQTL and etissue frequency at the loci level?}
-
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, egenes and eTissues for regulatory variants in three pleiotropic regions.
-%
-All three regions are very pleiotropic with very different associated phenotype categories such as cancer, cardiovascular and autoimmune diseases.
-%
-However these plots show that the relationship between the number of phenotypes, egenes and etissues can be very different.
-%
-The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of egenes but a high number of etissues.
-%
-By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both egenes and etissues.
-%
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest phenotype pleiotropy but the number of egenes and etissues is very low.
-%
-In summary, phenotype pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
-
-\subsection*{What are the mechanisms of pleiotropy}
-
-Then, I wanted to understand the molecular mechanism of pleiotropic regions.
-
-\subsubsection*{More severe variant effect consequences?}
-
-I first analysed whether there are significant differences of effect consequences between more and less pleiotropic variants.
-%
-I separated variants according to the count of GWAS categories and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
-%
-I found a significant increase of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
-%
-I also found a significant increase of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
-%
-Finally, there is also a significant increase of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
-%
-These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
-
-\subsubsection*{More egenes per variant-etissue?}
-
-Then I hypothesized that pleiotropic variant-etissue pairs modulate more egenes.
-%
-I found that pleiotropic variant-etissues pairs have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).
-
-\subsubsection*{More etissues per variant-egene?}
-
-My next hypothesis was that pleiotropic variant-egene pairs are active in more tissues compared with variant-egene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
-%
-Indeed, I found that pleiotropic variant-egenes pairs are associated with a higher number of etissues (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
-
-\subsubsection*{How to explain that pleiotropic variants are active in more tissues?}
-
-I hypothesized that pleiotropic variants are active in more etissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
-%
-I looked at the number of unique transcription factors bound in a window of 100 bp around each variant.
-%
-I found a significant increase of unique transcription factors around pleagonzaleziotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
-
-\subsubsection*{More GWAS per variant-egene-etissue?}
-
-Another molecular mechanism of pleiotropy is that egenes correlate with more GWAS categories even after taking into account variants and etissues.
-%
-I computed the number of GWAS categories per variant-egene-etissue triplet.
-%
-Indeed I found a significant increase of GWAS categories for pleiotropic variant-egene-etissue triplets (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
-
-\subsection*{What is the effect size of pleiotropic eQTLs?}
-
-Then I wondered, how does the effect size (beta) and significance (p-value) relate to pleiotropy.
-%
-I found a significant decrease of effect size of eQTLs and GWAS variants with the exception of the GWAS effect size of variants with 3 categories (Fig. \ref{fig:beta_pval}a,b).
-
-Then I looked at the significance of pleiotropic eQTLs and GWAS variants.
-%
-eQTL significance significantly decreased for pleiotropic variants (Fig. \ref{fig:beta_pval}c).
-%
-By contrast, the significance of GWAS associations increased for GWAS variants (Fig. \ref{fig:beta_pval}d).
-
-In summary, pleiotropic eQTLs are weaker with a weaker beta and less significant p-values..
-%
-On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.
-
 \section*{Discussion}
@@
 I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.
-
-\section*{References}
-\bibliographystyle{unsrt}
-\bibliography{ms_pleiotropy.bib}

### 2022-08-27 ms/ms05_floats.tex
ENH ms

--- 2022-08-19 ms/ms02_floats.tex
@@
 % full size table is table*
-\begin{table*}[]
-  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+\begin{table*}[!tbp]
 \centering
@@
 \csvreader[separator=tab,
-tabular=rrrrcp{0.6\textwidth},
+tabular=ccrrp{0.4\textwidth},
 head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos. (hg38) & \bfseries Variant & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
 ]{fig/tab_rsid_most_pleiotropic.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv & \csvcolvi}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolvi}% specify your coloumns here
 \hline
+%
+\vspace{15pt}
+%
+  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
 \end{table*}
@@
 
-\begin{figure*}[]
+\begin{figure*}[!tbp]
 \centering
@@
 %
+% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[!tbp]
+\centering
+%
+\begin{subfigure}[]{.66\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_tissue_enrich_in_pleio.py/plt.png}
+\end{subfigure}
+%
+\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:geneontology}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % Tab 2: Winner regions
@@
 % full size table is table*
-\begin{table*}[]
-\caption{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
+\begin{table*}[!tbp]
 \centering
@@
 separator=tab,
-tabular=rrrrcp{0.5\textwidth},
+tabular=ccrrp{0.45\textwidth},
 head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start (hg38) & \bfseries End (hg38) & \bfseries Cat. ct. & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
 ]{fig/tab_region_window_100000_pleio_highest.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv & \csvcolvi}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolvi}% specify your coloumns here
 \hline
+%
+\vspace{15pt}
+\caption{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
 \end{table*}
@@
 
-\begin{figure*}[!ht]
+\begin{figure*}[!tbp]
 
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{a}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{b}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{c}
@@
 
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{d}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{e}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{f}
@@
 
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{g}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{h}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{i}
@@
 
-\begin{figure*}[!]
+\begin{figure*}[!tbp]
 \centering

### 2022-08-27 ms/ms01_titlepage.tex
ENH intro and discussion

@@
 
-\title{Phenotype pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
+\title{Trait pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
 %\shorttitle{Pleiotropy and gene regulatory variants}
@@
 
-Pleiotropic genetic variants affecting several phenotypes are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
+Pleiotropic genetic variants affecting several traits are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
 %
@@
 %
-In this work, I investigate the link between gene regulation and phenotype pleiotropy for frequent variants.
+In this work, I investigate the link between gene regulation and trait pleiotropy for frequent variants.
 
@@
 %
-I have categorized the GWAS phenotypes into 96 phenotypes and categories to define pleiotropic as belonging at least to two of these categories.
+I have categorized the GWAS traits into 96 traits and categories to define pleiotropic as belonging at least to two of these categories.
 %
-Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more phenotypes.
+Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
 %
@@
 %
-These variants with the GWAS phenotypes, egenes and etissues can be explored in a public website.
+These variants with the GWAS traits, egenes and etissues can be explored in a public website.
 

### 2022-08-27 ms/ms02_intro.tex
ENH intro and discussion

@@
 
-Genome-wide association studies (GWAS) evaluate the impact of frequent variants on a trait in the population.
+Genome-wide association studies (GWAS) allow to define the impact of genetic variants on a trait in the population.
 %
@@
 %
-eQTLs have some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
+eQTLs share some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
 %
-On the other hand, eQTLs provide information about the target genes and tissues.
+Also the distribution of eQTLs and GWAS variants in the genome have been suggested to be different because of negative selection.
 
-A large number of GWAS variants are found in non-conding regions.
-To priori
+Despite these previous disadvantages eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (egenes) and tissues (etissues) (REF).
+%
+In addition colocalization anlaysis can pinpoint more precisely the causal variants that either data set independently (REF).
 
-Sentence describing EBI eQTL
+The large number of GWAS has allowed to discover that many genetic variants are involved in different traits.
+%
+TODO explain Watanabe.
+%
+Because GWAS variants affect mostly non-coding regions, it makes sense to investigate trait pleiotropy in terms of non-coding gene regulatory functions.
 
-In this work, I investigate the molecular basis of pleiotropy of gene regulatory variants.
+For these reasons, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
+%
+I have used these colocalization results to investigate the gene regulatory properties of pleiotropic variants.
+%
+This global colocalization analysis is also useful to help interpret GWAS variants.

### 2022-08-27 ms/ms03_results.tex
ENH intro and discussion

@@
 %
-The first criterion was to exclude molecular phenotypes such as proteome or methylome.
+The first criterion was to exclude molecular traits such as proteome or methylome.
 %	
@@
 %
-The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental phenotypes such "employment status" or "self-reported" medical conditions.
+The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental traits such "employment status" or "self-reported" medical conditions.
 %
@@
 
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar phenotypes (Supplementary table 2).
+I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).
 %
@@
 
-These category annotations were used to investigate the variant pleiotropy for GWAS phenotypes.
+These category annotations were used to investigate the variant pleiotropy for GWAS traits.
 %
@@
 
-%Variants of the 3q23 cytoband are involved in allergy, cancer and height phenotypes (Supplementary table 3).
+%Variants of the 3q23 cytoband are involved in allergy, cancer and height traits (Supplementary table 3).
 %%
@@
 %
-The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension phenotypes (Supplementary table 5).
+The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension traits (Supplementary table 5).
 %
-The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune phenotypes  (Supplementary table 5).
+The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune traits  (Supplementary table 5).
 
@@
 
-\subsection*{How specific are variants to GWAS phenotypes, egenes and etissues?}
+\subsection*{How specific are variants to GWAS traits, egenes and etissues?}
 
-Next I explored the phenotype, egene and etissue specificity of eQTL/GWAS variants.
+Next I explored the trait, egene and etissue specificity of eQTL/GWAS variants.
 %
-83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more phenotype categories make 3\% or less (Fig.
+83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more trait categories make 3\% or less (Fig.
 \ref{fig:hist_gwas_egene_etissue}(a).
@@
 %
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS phenotype. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
+In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do phenotype frequencies relates to eQTL and etissue frequency at the loci level?}
+\subsection*{How do trait frequencies relates to eQTL and etissue frequency at the loci level?}
 
@@
 %
-All three regions are very pleiotropic with very different associated phenotype categories such as cancer, cardiovascular and autoimmune diseases.
+All three regions are very pleiotropic with very different associated trait categories such as cancer, cardiovascular and autoimmune diseases.
 %
-However these plots show that the relationship between the number of phenotypes, egenes and etissues can be very different.
+However these plots show that the relationship between the number of traits, egenes and etissues can be very different.
 %
@@
 %
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest phenotype pleiotropy but the number of egenes and etissues is very low.
+Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest trait pleiotropy but the number of egenes and etissues is very low.
 %
-In summary, phenotype pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
+In summary, trait pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
 

### 2022-08-27 ms/ms04_discussion.tex
ENH intro and discussion

@@
 
-In this work, I present a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
+In this article, I present the results of a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
 %
-For this purpose, I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
+I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
 %
-The results of this pipeline provides three types of insights.
+These results are provided as a downloadable and an interactive table that can be used to explore variants with the GWAS traits, etissues and egenes.
 %
-First these results allow to investigate the molecular characteristics of pleiotropic variants with a gene regulatory effect.
-%
-Second, the colocalized loci provide predictions of causal variants in given loci of GWAS.
-%
-Third, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of many GWAs traits.
-
-% First insight: pleiotropy %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 For a posterior probability PP.H4.abf$\ge$0.8 of common causal eQTL and GWAS variants, we found 100,526 loci.
 %
-If I focus in causal variants with PP.H4.abf$\ge$0.8 and SNP.PP.H4$\ge$0.5, then I find 4684 variants.
+If I focus on causal variants with PP.H4.abf$\ge$0.8 and SNP.PP.H4$\ge$0.5, then I find 4684 variants.
+
+This interactive table can be used for different applications.
 %
-Our analysis suggests that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
+For instance, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of GWAs traits.
+%
+Colocalized loci provide predictions of potential causal variants in a given loci of GWAS.
+%
+The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
+Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
 
-% Second insight: causal variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+I have extensively compared molecular properties of pleiotropic variants involved in several GWAS traits.
+%
+Our analysis provides a number of hints about pleiotropic variants and regions.
+%
+I have found that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
+%
+By contrast, the eQTLs of pleiotropic variants show lower beta and less significant p-values (Fig. TODO).
+%
+This can make sense, because these variants target a larger number of egenes and etissues, and larger effects are maybe negatively selected.
+%
+It has been recently observed that eQTLs are negatively selected from essential genes (Ref bioarxiv TODO).
+%
+On the other hand, the effect on GWAS traits is more significant suggesting that the additive effect of several egenes and etissues results in stronger effects at the level of the GWAS trait (Fig TODO).
 
-% Third insight: mechanistic information for GWAS variants %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+At the molecular level, I have found that the immune system might be partly involved in pleiotropy.
+%
+Immune related gene ontology is significantly found in pleiotropic egenes (Fig 1) and etissues related to immune cells are overrepresented in pleiotropic loci (Fig 2).
+%
+The most pleiotropic locus 12q24.12 includes the SH2B3 gene that regulates cytokine signaling.
+%
+This locus has been used to explain links between inflammation and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
+%
+There is also the 5q31.1 locus that includes important genes of the immune system such as IRF1 and IL4 and the MHC locus that was previously discussed (REF watanaba discussion TODO).
 
-These variants are under control of more transcription factors
-
-%\subsection*{What is the function of the variants in the most pleiotropic regions?}
-%
-%% 12q24.12
-%The cytoband 12q24.12 has five variants rs3184504, rs653178, rs7310615, rs597808, rs11065979, that are known in the literature to be associated to a wide palette of diseases.
-%This locus has been used to explain links between inflammaiton and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
-%%
-%Our analysis add up that the variants in this cytoband control important genes such as ALDH2, ATXN2, MAPKAPK5, SH2B3 and TMEM116 in several tissues such as adipose, artery, blood, brain, digestive, skin, heart and skin tissue as well as immune cells.
-%
 %% 6p21.3
@@
 %Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
-%
-%
-%
-%The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
 %

### 2022-08-28 ms/ms01_titlepage.tex
ENH added supp table to snakemake

@@
 %
-Pleiotropic variants are associated to more eQTL genes (egenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
+Pleiotropic variants are associated to more eQTL genes (eGenes) and more eQTL tissues (etissues) but egenes do also correlate with more traits.
 %
@@
 
-\begin{keywords}
-keyword1 | keyword2 | keyword3
-\end{keywords}
-
-\begin{corrauthor}
-aitor.gonzalez\at univ-amu.fr
-\end{corrauthor}
+\begin{keywords2}
+eGenes; eQTLs
+\end{keywords2}

### 2022-08-28 ms/ms02_intro.tex
ENH added supp table to snakemake

@@
 
-Despite these previous disadvantages eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (egenes) and tissues (etissues) (REF).
+Nevertheless eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (eGenes) and tissues (etissues) (REF).
 %
@@
 
-The large number of GWAS has allowed to discover that many genetic variants are involved in different traits.
+The large number of GWAS has allowed to discover that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
 %
-TODO explain Watanabe.
+In this work, the analysed loci covered 60\% of the genome and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
 %
-Because GWAS variants affect mostly non-coding regions, it makes sense to investigate trait pleiotropy in terms of non-coding gene regulatory functions.
+They also found that pleiotropic genes and variants where less tissue specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
+%
+The question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature.Geneticeffectsgene}.
 
-For these reasons, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
+In this work, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
 %
-I have used these colocalization results to investigate the gene regulatory properties of pleiotropic variants.
+I have used these colocalization results to investigate molecular differences between pleiotropic variants in terms of gene regulatory properties.
+%
+We provide a model to mechanistically understand our results.
 %

### 2022-08-28 ms/ms02_intro.tex
ENH beta and pval improvements

@@
 %
-The question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature.Geneticeffectsgene}.
+The question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature}.
 

### 2022-08-28 ms/ms03_results.tex
ENH beta and pval improvements

@@
 %
-These variants control egenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
+These variants control eGenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
 %
@@
 %
-I analyzed the ontology of egenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
+I analyzed the ontology of eGenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
 %
-I found that egenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-For instance, egenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
+I found that eGenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
+For instance, eGenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
 %
@@
 %%
-%Interestingly all these variants target the egene ZBTB3.
+%Interestingly all these variants target the eGene ZBTB3.
 %%
@@
 %%
-%All these variants target the egene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
+%All these variants target the eGene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
 %%
@@
 
-\subsection*{How specific are variants to GWAS traits, egenes and etissues?}
+\subsection*{How specific are variants to GWAS traits, eGenes and eTissues?}
 
-Next I explored the trait, egene and etissue specificity of eQTL/GWAS variants.
+Next I explored the trait, eGene and eTissue specificity of eQTL/GWAS variants.
 %
@@
 %
-Regarding egenes, half of variants modulate one egene, and the other half modulate two or more egenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
+Regarding eGenes, half of variants modulate one eGene, and the other half modulate two or more eGenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
 %
-In the case of etissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
+In the case of eTissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
 %
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait. By contrast, half of them only module one specific egene, and less than half are active in a single tissue.
+In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait. By contrast, half of them only module one specific eGene, and less than half are active in a single tissue.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do trait frequencies relates to eQTL and etissue frequency at the loci level?}
+\subsection*{How do trait frequencies relates to eQTL and eTissue frequency at the loci level?}
 
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, egenes and eTissues for regulatory variants in three pleiotropic regions.
+In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, eGenes and eTissues for regulatory variants in three pleiotropic regions.
 %
@@
 %
-However these plots show that the relationship between the number of traits, egenes and etissues can be very different.
+However these plots show that the relationship between the number of traits, eGenes and eTissues can be very different.
 %
-The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of egenes but a high number of etissues.
+The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of eGenes but a high number of eTissues.
 %
-By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both egenes and etissues.
+By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both eGenes and eTissues.
 %
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest trait pleiotropy but the number of egenes and etissues is very low.
+Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest trait pleiotropy but the number of eGenes and eTissues is very low.
 %
-In summary, trait pleiotropy can arise from different situations: few egenes and etissues, many egenes and etissues or a mix of both.
+In summary, trait pleiotropy can arise from different situations: few eGenes and eTissues, many eGenes and eTissues or a mix of both.
 
@@
 
-Then, I wanted to understand the molecular mechanism of pleiotropic regions.
+Then, I wanted to understand the molecular mechanism of pleiotropic variants and regions.
+%
+I hypothesized that pleiotropy arises from bias in regulatory effects of variants.
+%
+For instance, pleiotropic variants might significantly affect some molecular functions more often.
+%
+Another possibility is that pleiotropic variants affect more eGenes in more eTissues, which affect more GWAS traits.
 
@@
 
-I first analysed whether there are significant differences of effect consequences between more and less pleiotropic variants.
+I first analysed whether there are significant differences of some effect consequences between more and less pleiotropic variants.
 %
@@
 %
-I found a significant increase of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
+I found a significant larger number of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
 %
-I also found a significant increase of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
+I also found a significant larger number of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
 %
-Finally, there is also a significant increase of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
+Finally, there is also a larger of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
 %
+% TODO Discuss sQTLs in 2021.Li.Mu.GenomeBiology.impactcelltype paper
 These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
 
-\subsubsection*{More egenes per variant-etissue?}
+\subsubsection*{More eGenes per variant-eTissue?}
 
-Then I hypothesized that pleiotropic variant-etissue pairs modulate more egenes.
+Then I hypothesized that pleiotropic variants regulate more eGene even when we keep fixed the eTissue.
 %
-I found that pleiotropic variant-etissues pairs have significantly more egenes compared to variants with one GWAS category (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+To test this hypothesis, the eGenes per variant-eTissues pairs were counted.
+%
+Then these eGene counts were classified according to the GWAS category count.
+%
+If the GWAS category count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal one.
+%
+I found means of eGene counts of 1.4, 1.7, 1.7 and 1.6 for GWAS categories counts 5, 4, 3 and 2 compared to eGene count mean of 1.5 for GWAS category count 1 (Fig.
+\ref{fig:gwas_egene_etisue_per_variant}a).
+%
+This suggests that pleiotropic variant-eTissue pairs have slightly but significantly (Mann–Whitney U test) more eGenes compared to variants with one GWAS category (Fig.
+\ref{fig:gwas_egene_etisue_per_variant}a).
 %
@@
 
-\subsubsection*{More etissues per variant-egene?}
+\subsubsection*{More eTissues per variant-eGene?}
 
-My next hypothesis was that pleiotropic variant-egene pairs are active in more tissues compared with variant-egene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
+My next hypothesis was that pleiotropic variant-eGene pairs are active in more eTissues compared with variant-eGene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
 %
-Indeed, I found that pleiotropic variant-egenes pairs are associated with a higher number of etissues (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
+To test this hypothesis, the eTissues per variant-eGene pairs were counted.
+%
+Then these eTissue counts were classified according to the GWAS category count.
+%
+If the GWAS category count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS category count.
+%
+I found means of eTissue counts of 2.2, 3, 3 and 2.6 for GWAS categories counts 5, 4, 3 and 2 compared to eTissue count mean of 2.5 for GWAS category count 1 (Fig.
+\ref{fig:gwas_egene_etisue_per_variant}b).
+%
+This suggests that pleiotropic variant-eGene pairs are active slightly but significantly (Mann–Whitney U test) in more eTissues compared to variants with one GWAS category (Fig.
+\ref{fig:gwas_egene_etisue_per_variant}b).
+
+\subsubsection*{More GWAS per variant-eGene-eTissue?}
+
+My next hypothesis was that pleiotropic variants are associated to more GWAS categories even after taking into account differences in eGene and eTissue counts.
+
+To test this hypothesis, the GWAS category counts per variant-eGene-eTissue trios were counted.
+%
+Then these GWAS categorie counts per variant-eGene-eTissue trios were classified according to the GWAS category count per variant.
+%
+If the GWAS category count per variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS category count per variant.
+%
+I found means of GWAS category counts per variant-eGene-eTissue of 3, 2, 1.6, 1.4 and 1 for GWAS categories counts per variant of 5, 4, 3, 2 and 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+%
+This shows a significant larger number of GWAS categories even in unique trios of variant-eGene-eTissue (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+%
+This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
 
@@
 
-I hypothesized that pleiotropic variants are active in more etissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
+In Fig. \ref{fig:gwas_egene_etisue_per_variant}b, I observed that pairs of eQTL-eGene are active in more eTissues.
 %
-I looked at the number of unique transcription factors bound in a window of 100 bp around each variant.
+To explain this observation, I hypothesized that pleiotropic variants are active in more eTissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
 %
-I found a significant increase of unique transcription factors around pleagonzaleziotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
+To test this hypothesis, I counted the number of unique transcription factors bound in a radius of 50 bp around each variant (Window of 100 bp).
+%
+I found a significant larger number of transcription factors bound around pleiotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
 
-\subsubsection*{More GWAS per variant-egene-etissue?}
+Cis-regulatory modules (CRMs) are non-coding genomic regions with a higher density of bound cis-regulatory modules \citep{2021.Ballester.Hammal}.
+%
+I found that the odds ratio of variants annotated with CRMs vs non-annotated is significantly (Fisher's exact test) higher for variants with 2 and 3 GWAS category counts compared to category count 1 (Fig.
+\ref{fig:freq_tf_per_variant}b).
 
-Another molecular mechanism of pleiotropy is that egenes correlate with more GWAS categories even after taking into account variants and etissues.
+\subsection*{How do effect size and }
+
+Pleiotropic variants affect simultaneously more GWAS traits.
 %
-I computed the number of GWAS categories per variant-egene-etissue triplet.
+Therefore it is interesting to investigate how does variant causality and pleiotropy relate to effect size (beta) and significance (p-value) at the level of eQTL and GWAS traits.
 %
-Indeed I found a significant increase of GWAS categories for pleiotropic variant-egene-etissue triplets (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS category counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
 %
-This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
+Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS category counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 
-\subsection*{What is the effect size of pleiotropic eQTLs?}
+Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
+%
+I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS category counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
+%
+Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS category counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 
-Then I wondered, how does the effect size (beta) and significance (p-value) relate to pleiotropy.
-%
-I found a significant decrease of effect size of eQTLs and GWAS variants with the exception of the GWAS effect size of variants with 3 categories (Fig. \ref{fig:beta_pval}a,b).
+These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).
 
-Then I looked at the significance of pleiotropic eQTLs and GWAS variants.
-%
-eQTL significance significantly decreased for pleiotropic variants (Fig. \ref{fig:beta_pval}c).
-%
-By contrast, the significance of GWAS associations increased for GWAS variants (Fig. \ref{fig:beta_pval}d).
-
-In summary, pleiotropic eQTLs are weaker with a weaker beta and less significant p-values..
-%
-On the other hand, the effect size of GWAS seems smaller for pleiotropic variants but the significance is stronger.
+In summary, I observed opposite effects of pleiotropy on the eQTL and GWAS variant strengths.

### 2022-08-28 ms/ms05_floats.tex
ENH intro and discussion
ENH beta and pval improvements

@@
 %
-\begin{subfigure}[]{.66\textwidth}
+\begin{subfigure}[]{0.99\textwidth}
 \textbf{a}
@@
 
-\begin{subfigure}[]{.66\textwidth}
+\begin{subfigure}[]{.99\textwidth}
 \textbf{b}
@@
 %
-\begin{subfigure}[]{.33\textwidth}
+\begin{subfigure}[]{.32\textwidth}
 \textbf{a}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_beta.py/gwas_beta.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_logpval.py/eqtl.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_gwas_cat_y_logpval.py/gwas.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
 \end{subfigure}

### 2022-08-28 ms/ms05_floats.tex
ENH ms and refactoring

@@
 %
-\caption{Gene ontology analysis of egenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to egenes of variants associated to one GWAS category.} \label{fig:geneontology}
+\caption{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.} \label{fig:geneontology}
 \end{figure*}
@@
 %
-\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:geneontology}
+\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:tissues}
 \end{figure*}
@@
 %
-% Fig 1: Histograms variants vs GWAS, egene and etissues
+% Fig 1: Histograms variants vs GWAS, eGene and eTissues
 %
@@
 %
-\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS categories (\textbf{a}), egenes (\textbf{b}) and etissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
+\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS categories (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
 \end{figure*}
@@
 
-\caption{Count of GWAS categories (\textbf{a-c}), egenes (\textbf{d-f}) and etissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
+\caption{Count of GWAS categories (\textbf{a-c}), eGenes (\textbf{d-f}) and eTissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
 Genomic coordinates are given for the hg38 assembly.} \label{fig:region_gwas_egenes_tissues}
@@
 %
-% Fig 4: Violin plots. eGenes, etissues and gwas per variants
+% Fig 4: Violin plots. eGenes, eTissues and gwas per variants
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_etissue_y_egene.py/plt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_y_etissue.py/plt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_variant_egene_etissue_y_gwas.py/plt.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
 \end{subfigure}
 %
-\caption{Count of egenes per variant-tissue (\textbf{a}), etissues per variant-egene (\textbf{b}) and GWAS categories per variant-egene-etissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
+\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS categories per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
 %
@@
 \caption{\textbf{Model of regulatory variant pleiotropy.} I have investigated three possible mechanisms of pleiotropy.
-\textbf{Left}, Pleiotropic variants have more egenes that result in more functions and more phenotypes.
+\textbf{Left}, Pleiotropic variants have more eGenes that result in more functions and more phenotypes.
 This might arise from an enrichment of pleiotropic variants in splicing or 3' UTR regions.
-\textbf{Center}, I have also found that egenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
+\textbf{Center}, I have also found that eGenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
 This might be explained from variants being bound by more transcription factors.
-\textbf{Left} Triplets of variant-egene-etissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
+\textbf{Left} Triplets of variant-eGene-eTissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
 I have found that this might be explained by an enrichment of missense alleles.} \label{fig:beta}

### 2022-08-29 ms/ms04_discussion.tex
ENH beta and pval improvements
BUG fixed

@@
 
+% TODO Causal variants
+% If pleiotropic variants bind more transcription factors and are more frequently annotated with a CRM, then these variants are probably more likely causal variants.
+
+% TODO Relation between eQTL and GWAS and related publications.
+
 %% 6p21.3
@@
 
-I developed a colocalization pipeline based on a pipeline that is public at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
+My colocalization pipeline was initially based on another pipeline that is at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
 %
-First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants in a radius of 500 000 nt are retrieved.
+First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants with radius of 500,000 nt (Total window 1,000,000 nt) are retrieved.
 %

### 2022-08-29 ms/ms03_results.tex
ENH ms and refactoring
ENH ms

@@
 %
-This analysis explored 30 261 unique variants and 2,479,065 potential variant colocalizations (OSF URL).
+This analysis resulted in 30 261 unique variants and 2,479,065 potential variant colocalizations from 150 GWAS studies (OSF URL).
+%
+%TODO add PMID to supplementary table with loci_explained_perc
+For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
+%
+Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
 %
@@
 
-\subsection*{How do effect size and }
+\subsection*{How do effect size and significance relate to pleiotropy}
 
@@
 These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).
-
-In summary, I observed opposite effects of pleiotropy on the eQTL and GWAS variant strengths.

### 2022-08-29 ms/ms04_discussion.tex
ENH ms

@@
 
-My colocalization pipeline was initially based on another pipeline that is at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
+My colocalization pipeline was inspired by another pipeline that is at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
 %
-First lead eQTLs with FDR 0.05 in each sample are selected and surrounding eQTL and GWAS variants with radius of 500,000 nt (Total window 1,000,000 nt) are retrieved.
+GWAS were downloaded from OpenGWAS and converted to hg38 coordinates.
 %
-Colocalization between eQTL and GWAS variants is tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
+Permuted eQTLs with FDR 0.05 in each sample were selected as lead variants.
+%
+eQTL and GWAS variants surrounded the lead variants with radius of 500,000 nt (Total window 1,000,000 nt) were retrieved.
+%
+These variants were kept if MAF was strictly greater than 0 and lower than 1, the variants were not duplicated and there was no missing data.
+%
+Colocalization between eQTL and GWAS variants was tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
 

### 2022-08-30 poster_eccb22_barcelona/poster_eccb22_barcelona.tex
ENH poster

%\documentclass[landscape,a0b,final,a4resizeable]{a0poster}
%\documentclass[landscape,a0b,final]{a0poster}
%\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
%%% Option "a4resizeable" makes it possible ot resize the
%   poster by the command: psresize -pa4 poster.ps poster-a4.ps
%   For final printing, please remove option "a4resizeable" !!

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{natbib}
\usepackage{pstricks,pst-grad}
\usepackage{subfigure}
\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%           document formatting parameters
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% DEFINE POSTER WIDTH
\newenvironment{poster}{
  \begin{center}
  \begin{minipage}[c]{0.98\textwidth}
}{
  \end{minipage} 
  \end{center}
}
% DEFINITION OF SOME METRICS
\setlength{\columnsep}{3cm}
\setlength{\columnseprule}{2mm}
%\setlength{\parindent}{0.0cm}
% DEFINITION OF FUNCTION TO SET POSTER BACKGROUND GRADIENT COLORS
\newcommand{\backgroundcolor}[3]{
  \newrgbcolor{cgradbegin}{#1}
  \newrgbcolor{cgradend}{#2}
  \psframe[fillstyle=gradient,gradend=cgradend,
  gradbegin=cgradbegin,gradmidpoint=#3](0.,0.)(1.\textwidth,-1.\textheight)
}
% MAIN TITLE ENVIRONMENT
\newenvironment{pcolumn}[1]{
  \begin{minipage}{#1\textwidth}
  \begin{center}
}{
  \end{center}
  \end{minipage}
}
% SECTION TITLE DECORATION BOX (BLUE GRADIENT)
\newrgbcolor{lcolor}{0. 0. 0.80}
\newrgbcolor{gcolor1}{1. 1. 1.}
\newrgbcolor{gcolor2}{.80 .80 1.}
%
\newcommand{\pbox}[4]{
\psshadowbox[#3]{
\begin{minipage}[t][#2][t]{#1}
#4
\end{minipage}
}}

% MY FIGURES
% \myfig - replacement for \figure
% necessary, since in multicol-environment 171231_training_matrix.eps
% \figure won't work
\newcommand{\myfig}[3][0]{
\begin{center}
  \vspace{0.5cm}
  \includegraphics[width=#3\hsize,angle=#1]{#2}
  \nobreak\medskip
\end{center}}
%
% MY CAPTIONS
% \mycaption - replacement for \caption
% necessary, since in multicol-environment \figure and
% therefore \caption won't work
%\newcounter{figure}
\setcounter{figure}{1}
\newcommand{\mycaption}[1]{
  \vspace{0.25cm}
  \begin{quote}%\centering
    {{\bf Figure \arabic{figure}:} #1}
  \end{quote}
  \vspace{0.5cm}
  \stepcounter{figure}
}
% DEFINE FONT COMMANDS
\newcommand{\fontbody}{\Large}
%\newcommand{\fontbody}[1]{\textsf{#1}}
\newcommand{\fonttitle}[1]{\veryHuge \textsf{#1}} 
\newcommand{\fontheader}[1]{\Huge \textsf{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

% Set document size
\large

\backgroundcolor{1. 1. 1.}{1. 1. 1.}{0.5}

\vspace*{0.5cm}

\newrgbcolor{lightblue}{0. 0. 0.80}
\newrgbcolor{white}{1. 1. 1.}
\newrgbcolor{whiteblue}{.80 .80 1.}

\begin{poster}

%%%%%%%%%%%%%%%%%%%%%
%%% Header
%%%%%%%%%%%%%%%%%%%%%

\begin{center}
%\begin{pcolumn}{1}
\pbox{0.98\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightblue,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whiteblue,gradmidpoint=1.0,framesep=0.25em}{
\begin{minipage}[c][10cm][c]{0.1\textwidth}
  \begin{center}
    \includegraphics[width=12cm,angle=0]{img/logo_sciences.eps}
  \end{center}
\end{minipage}
\begin{minipage}[c][9cm][c]{0.7\textwidth}
  \begin{center}
%Microarray analysis of chicken and mouse presomitic mesoderm along the posterior-anterior axisS
%    {\sc \Huge Microarray analysis of chicken and mouse}\\[4.mm]
%    {\sc \Huge presomitic mesoderm along the posterior-anterior axis}\\[2.mm]
%\vspace{1cm}
    { \fonttitle{Analysis of gene regulatory properties} }\\[4.mm]
    { \fonttitle{underlying trait pleiotropy} }\\[2.mm]
    \huge Aitor Gonz\'alez$^*$\\[2.mm]
    \large TAGC Laboratory, Aix-Marseille Univ, INSERM UMR1090, 13288 Marseille, France\\[2.mm]
    \large $^*$Email: aitor.gonzalez@univ-amu.fr
  \end{center}
  \end{minipage}
\begin{minipage}[c][10cm][c]{0.1\textwidth}
  \begin{center}
    \includegraphics[width=12cm,angle=0]{img/logo_tagc_400x230_centered_1.eps}
    \includegraphics[width=12cm,angle=0]{img/logo_tagc_400x230_centered_1.eps}
  \end{center}
\end{minipage}
} % end pbox
%\end{pcolumn}
\end{center}

%%%%%%%%%%%%%%%%%%%%%
%%% Content
%%%%%%%%%%%%%%%%%%%%%

\vspace{0.5cm} % intersection space

%%% Begin of Multicols-Enviroment
\begin{multicols}{2}

%%%%%%%%%%%%%%%%%%%%%
%%% Background

\begin{center}\pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightblue,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whiteblue,gradmidpoint=1.0,framesep=0.5em}{\huge \begin{center}\fontheader{Background and Objectives}\end{center}}\end{center}

\section*{Background}

\begin{enumerate}
\item Genome-wide association studies (GWAS) map genotypes to traits.
\item The IEU OpenGWAS project is a database of GWAS:\\ \url{https://gwas.mrcieu.ac.uk}
\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (REF).
\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is well understood.
\item Expression quantitative trait loci (eQTL) map genotypes to expression phenotypes.
\item The EBI eQTL Catalogue is a database of uniformely processed eQTLs:\\ \url{https://www.ebi.ac.uk/eqtl}
\item The R package coloc performs colocalization tests between genetic traits:\\ \url{https://CRAN.R-project.org/package=coloc}
\end{enumerate}

\section*{Objectives}

\begin{enumerate}
\item To develop an eQTL/GWAS variant colocalization pipeline
\item To investigate gene regulatory properties underlying trait pleiotropy
\end{enumerate}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%
%%% Results

\begin{center}\pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightblue,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whiteblue,gradmidpoint=1.0,framesep=0.5em}{\begin{center}\fontheader{Results}\end{center}}\end{center}

- A pipeline of eQTL/GWAS variant colocalization (Screenshot pythonanywhere)
- Table with most pleiotropic variants (Table 1)
- Molecular mechanisms Fig 7 and fig 8a
- Model of mechanisms Fig 10

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURE 1
\begin{minipage}{1\columnwidth}
    \begin{center}
	    \begin{minipage}{0.6\columnwidth} 
		    \includegraphics[width=\columnwidth]{example-image-a}
	    \end{minipage}
	    \mycaption{\fontbody A set of pruned index SNPs in approximate linkage equilibrium is labeled as positive based on LD distance to associated SNPs from the GRASP database \citep{Leslie2014} and annotated with the LD distance to annotations. This training dataset is used for supervised classification based on the gradient boosting algorithm (XGBOOST software). Intergenic regions are annotated and scored with this model.}
    \end{center}
\end{minipage}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURE 2
\begin{minipage}{1\columnwidth}
    \begin{center}
	    \begin{minipage}{0.55\columnwidth} 
		    \includegraphics[width=\columnwidth]{example-image-a}
	    \end{minipage}
	    \begin{minipage}{0.35\columnwidth} 
		    \includegraphics[width=\columnwidth]{example-image-b}
	    \end{minipage}
	    \mycaption{\fontbody A GWAS SNP prediction model based on gene regulatory annotations. Left: AUC values of intergenic GWAS Catalog SNPs calculated with GWAVA, DeepSea and TAGOOS scores. Right: Odds ratio of GWAS Catalog SNP enrichment in TAGOOS-significant or cell-specific regulatory (TSS, promoter flanking region, weak or strong enhancer) chromatin state regions.}
    \end{center}
\end{minipage}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURE 3
\begin{minipage}{1\columnwidth}
    \begin{center}
	    \begin{center}
		    \begin{minipage}{0.45\columnwidth} 
			    \includegraphics[width=\columnwidth]{example-image-c}
		    \end{minipage}
		    \begin{minipage}{0.45\columnwidth} 
		    \mycaption{\fontbody Correlation of sample number of selected variables with the TAGOOS score. For each unseen GWAS Catalog SNPs, we calculated a sample proportion matrix with the proportion of samples out of the maximal sample number of a given molecular assay. Heatmap of the sample proportion matrix with the SNPs in the rows ordered by the TAGOOS score (Green column).}
		    \end{minipage}
	    \end{center}
    \end{center}
\end{minipage}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURE 3
\begin{minipage}{1\columnwidth}
    \begin{center}
	    \begin{center}
		    \begin{minipage}{0.5\columnwidth} 
			    \includegraphics[width=\columnwidth]{example-image-a}
		    \end{minipage}
		    \mycaption{\fontbody Percentage of unseen GWAS Catalog SNPs with significant and non-significant TAGOOS scores annotated with each transcriptional regulator from the ReMap database ordered by the decreasing percentage in significant TAGOOS SNPs ($P_{\text{Wilcox paired}}=2.2\times 10^{-16}$) \citep{Cheneby2017}.}
	    \end{center}
    \end{center}
\end{minipage}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURE 5
\begin{minipage}{1\columnwidth}
    \begin{center}
	    \begin{center}
		    \begin{minipage}{0.7\columnwidth} 
			    \includegraphics[width=\columnwidth]{example-image-b}
		    \end{minipage}
		    \mycaption{\fontbody Motif density computed with RSAT Matrix-scan in 10 nt bins in a 100 nt window around intergenic unseen GWAS Catalog SNPs split according to the TAGOOS significance and the reference versus alternative allele \citep{Medina-Rivera2015}}
	    \end{center}
    \end{center}
\end{minipage}

\vspace{0.5cm} % intersection space

%%%%%%%%%%%%%%%%%%%%%
%%% Conclusions and Future Directions

\begin{center}\pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightblue,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whiteblue,gradmidpoint=1.0,framesep=0.5em}{\begin{center}\fontheader{Conclusions}\end{center}}\end{center}

\begin{itemize}
\fontbody

\item We present a new method to learn a supervised model of associated SNPs based on regulatory features (Fig. 1).

\end{itemize}

\footnotesize
\bibliographystyle{plain}
\bibliography{biblio}

\end{multicols}

\end{poster}

\end{document}

### 2022-08-30 poster_eccb22_barcelona/poster_eccb22_barcelona.tex
ENH progressed poster

@@
 	\usepackage{hyperref}
+	\usepackage{csvsimple}
 
@@
 	% therefore \caption won't work
+
 	%\newcounter{figure}
 	\setcounter{figure}{1}
-\newcommand{\mycaption}[1]{
+	\newcommand{\mycaptionfig}[1]{
 	  \vspace{0.25cm}
@@
 	}
+
+	%\newcounter{table}
+	\setcounter{table}{1}
+	\newcommand{\mycaptiontab}[1]{
+	  \vspace{0.25cm}
+	  \begin{quote}%\centering
+	    {{\bf Table \arabic{table}:} #1}
+	  \end{quote}
+	  \vspace{0.5cm}
+	  \stepcounter{figure}
+	}
+
 	% DEFINE FONT COMMANDS
@@
 	  \begin{center}
-    \includegraphics[width=12cm,angle=0]{img/logo_sciences.eps}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_sciences.eps}
 	  \end{center}
@@
 	  \begin{center}
-    \includegraphics[width=12cm,angle=0]{img/logo_tagc_400x230_centered_1.eps}
-    \includegraphics[width=12cm,angle=0]{img/logo_tagc_400x230_centered_1.eps}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_tagc_400x230_centered_1.eps}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_tagc_400x230_centered_1.eps}
 	  \end{center}
@@
 	\item The IEU OpenGWAS project is a database of GWAS:\\ \url{https://gwas.mrcieu.ac.uk}
-\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (REF).
+	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (\citep{2019.Posthuma.Watanabe}).
 	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is well understood.
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	% FIGURE 1 web interface
+	\begin{minipage}{\columnwidth}
+	    \begin{center}
+		    \begin{minipage}{1\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img/web.eps}
+		    \end{minipage}
+		    \mycaptionfig{Web interface with the results of the colocalization pipeline for the most pleiotropic variant rs3184504.}
+	    \end{center}
+	\end{minipage}
+
+	\vspace{0.5cm} % intersection space
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	%
+	% Tab 1: Winner variants
+	%
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+	% full size table is table*
+	\begin{minipage}{\columnwidth}
+	   \begin{center}
+	\scriptsize
+	\csvreader[separator=tab,
+	tabular=ccrrp{0.4\textwidth},
+	head,
+	table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
+	]{../ms/fig/tab_rsid_most_pleiotropic.tsv}{}% use head of csv as column names
+	{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolvi}% specify your coloumns here
+	\large  % normal size
+		    \mycaptiontab{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}
+	    \end{center}
+	\end{minipage}
+
+	\vspace{0.5cm} % intersection space
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 	% FIGURE 1
@@
 	    \begin{center}
-	    \begin{minipage}{0.6\columnwidth} 
-		    \includegraphics[width=\columnwidth]{example-image-a}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_etissue_y_egene.eps}
 		    \end{minipage}
-	    \mycaption{\fontbody A set of pruned index SNPs in approximate linkage equilibrium is labeled as positive based on LD distance to associated SNPs from the GRASP database \citep{Leslie2014} and annotated with the LD distance to annotations.
-This training dataset is used for supervised classification based on the gradient boosting algorithm (XGBOOST software).
-Intergenic regions are annotated and scored with this model.}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_y_etissue.eps}
+		    \end{minipage}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_etissue_y_gwas.eps}
+		    \end{minipage}
+		    \mycaptionfig{Count of eGenes per variant-tissue (\textbf{Left}), eTissues per variant-eGene (\textbf{Center}) and GWAS categories per variant-eGene-eTissue (\textbf{Right}).}
+	    \end{center}
+	\end{minipage}
+
+	\vspace{0.5cm} % intersection space
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	% FIGURE 2 VEP
+	\begin{minipage}{1\columnwidth}
+	    \begin{center}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/missense_variant.eps}
+		    \end{minipage}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/splice_region_variant.eps}
+		    \end{minipage}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/3_prime_UTR_variant.eps}
+		    \end{minipage}
+		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (Left), splice region variant (Center) and 3’ UTR variant (Right)}
 	    \end{center}
@@
 	    \begin{center}
-	    \begin{minipage}{0.55\columnwidth} 
-		    \includegraphics[width=\columnwidth]{example-image-a}
+		    \begin{center}
+			    \begin{minipage}{0.35\columnwidth} 
+				    \includegraphics[width=\columnwidth]{img_tmp/bxplt_remaptf_per_rsid_flank_0.eps}
 			    \end{minipage}
-	    \begin{minipage}{0.35\columnwidth} 
-		    \includegraphics[width=\columnwidth]{example-image-b}
-	    \end{minipage}
-	    \mycaption{\fontbody A GWAS SNP prediction model based on gene regulatory annotations.
-Left: AUC values of intergenic GWAS Catalog SNPs calculated with GWAVA, DeepSea and TAGOOS scores.
-Right: Odds ratio of GWAS Catalog SNP enrichment in TAGOOS-significant or cell-specific regulatory (TSS, promoter flanking region, weak or strong enhancer) chromatin state regions.}
-    \end{center}
-\end{minipage}
-
-\vspace{0.5cm} % intersection space
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE 3
-\begin{minipage}{1\columnwidth}
-    \begin{center}
-	    \begin{center}
-		    \begin{minipage}{0.45\columnwidth} 
-			    \includegraphics[width=\columnwidth]{example-image-c}
-		    \end{minipage}
-		    \begin{minipage}{0.45\columnwidth} 
-		    \mycaption{\fontbody Correlation of sample number of selected variables with the TAGOOS score.
-For each unseen GWAS Catalog SNPs, we calculated a sample proportion matrix with the proportion of samples out of the maximal sample number of a given molecular assay.
-Heatmap of the sample proportion matrix with the SNPs in the rows ordered by the TAGOOS score (Green column).}
+			    \begin{minipage}{0.6\columnwidth} 
+			    \mycaptionfig{Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS category count. }
 			    \end{minipage}
@@
 	    \begin{center}
-	    \begin{center}
-		    \begin{minipage}{0.5\columnwidth} 
-			    \includegraphics[width=\columnwidth]{example-image-a}
+		    \begin{minipage}{1\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/graphical_summary.eps}
 		    \end{minipage}
-		    \mycaption{\fontbody Percentage of unseen GWAS Catalog SNPs with significant and non-significant TAGOOS scores annotated with each transcriptional regulator from the ReMap database ordered by the decreasing percentage in significant TAGOOS SNPs ($P_{\text{Wilcox paired}}=2.2\times 10^{-16}$) \citep{Cheneby2017}.}
-	    \end{center}
-    \end{center}
-\end{minipage}
-
-\vspace{0.5cm} % intersection space
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIGURE 5
-\begin{minipage}{1\columnwidth}
-    \begin{center}
-	    \begin{center}
-		    \begin{minipage}{0.7\columnwidth} 
-			    \includegraphics[width=\columnwidth]{example-image-b}
-		    \end{minipage}
-		    \mycaption{\fontbody Motif density computed with RSAT Matrix-scan in 10 nt bins in a 100 nt window around intergenic unseen GWAS Catalog SNPs split according to the TAGOOS significance and the reference versus alternative allele \citep{Medina-Rivera2015}}
-	    \end{center}
+		    \mycaptionfig{Model of regulatory variant pleiotropy. I have investigated
+	three possible mechanisms of pleiotropy. Left, Pleiotropic variants have more
+	eGenes that result in more functions and more phenotypes. This might arise
+	from an enrichment of pleiotropic variants in splicing or 3’ UTR regions.
+	Center, I have also found that eGenes of pleiotropic variants are active
+	in more etissues which result in more GWAS phenotypes. This might be
+	explained from variants being bound by more transcription factors. Left
+	Triplets of variant-eGene-eTissues are associated with more GWAS pheno-
+	types, which directly affect the number of GWAS phenotypes. I have found
+	that this might be explained by an enrichment of missense alleles.}
 	    \end{center}
@@
 
+	\item A pipeline to colocalize eQTLs and GWAS variantsWe present a new method to learn a supervised model of associated SNPs based on regulatory features (Fig. 1).
 	\item We present a new method to learn a supervised model of associated SNPs based on regulatory features (Fig. 1).
@@
 
-\footnotesize
+	\normalsize
 	\bibliographystyle{plain}
-\bibliography{biblio}
+	\bibliography{../ms/ms_pleiotropy.bib}
 

### 2022-08-31 poster_eccb22_barcelona/poster_eccb22_barcelona.tex
ENH improved

@@
 	  \vspace{0.5cm}
-	  \stepcounter{figure}
+	  \stepcounter{table}
 	}
@@
 	  \begin{center}
-	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_sciences.eps}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/tagc_logo.eps}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/inserm.eps}
 	  \end{center}
@@
 	\begin{minipage}[c][10cm][c]{0.1\textwidth}
+	    \includegraphics[width=12cm,angle=0]{img_tmp/DIRCOM-Logo-FSS.eps}
 	  \begin{center}
-	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_tagc_400x230_centered_1.eps}
-	    \includegraphics[width=12cm,angle=0]{img_tmp/logo_tagc_400x230_centered_1.eps}
 	  \end{center}
@@
 	\item Genome-wide association studies (GWAS) map genotypes to traits.
-	\item The IEU OpenGWAS project is a database of GWAS:\\ \url{https://gwas.mrcieu.ac.uk}
+	\item Here we use this GWAS database: The IEU OpenGWAS project\\ \url{https://gwas.mrcieu.ac.uk}
 	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (\citep{2019.Posthuma.Watanabe}).
-	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is well understood.
+	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is not well understood.
 	\item Expression quantitative trait loci (eQTL) map genotypes to expression phenotypes.
-	\item The EBI eQTL Catalogue is a database of uniformely processed eQTLs:\\ \url{https://www.ebi.ac.uk/eqtl}
+	\item Here we use this eQTL database: The EBI eQTL Catalogue\\ \url{https://www.ebi.ac.uk/eqtl}
 	\item The R package coloc performs colocalization tests between genetic traits:\\ \url{https://CRAN.R-project.org/package=coloc}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	% FIGURE 1
+	% FIGURE 2 downstream regulation
+	\begin{minipage}{1\columnwidth}
+	    \begin{center}
+		    \begin{minipage}{\columnwidth} 
+\includegraphics[width=\textwidth]{img_tmp/david_pleio_3.eps}
+\includegraphics[width=\textwidth]{img_tmp/david_pleio_2.eps}
+		    \end{minipage}
+		    \mycaptionfig{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.}
+	    \end{center}
+	\end{minipage}
+
+	\vspace{0.5cm} % intersection space
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	% FIGURE 1 VEP
+	\begin{minipage}{1\columnwidth}
+	    \begin{center}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/missense_variant.eps}
+		    \end{minipage}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/splice_region_variant.eps}
+		    \end{minipage}
+		    \begin{minipage}{0.32\columnwidth} 
+			    \includegraphics[width=\columnwidth]{img_tmp/3_prime_UTR_variant.eps}
+		    \end{minipage}
+		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (Left), splice region variant (Center) and 3’ UTR variant (Right)}
+	    \end{center}
+	\end{minipage}
+
+	\vspace{0.5cm} % intersection space
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	% FIGURE 2 downstream regulation
 	\begin{minipage}{1\columnwidth}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	% FIGURE 2 VEP
-	\begin{minipage}{1\columnwidth}
-	    \begin{center}
-		    \begin{minipage}{0.32\columnwidth} 
-			    \includegraphics[width=\columnwidth]{img_tmp/missense_variant.eps}
-		    \end{minipage}
-		    \begin{minipage}{0.32\columnwidth} 
-			    \includegraphics[width=\columnwidth]{img_tmp/splice_region_variant.eps}
-		    \end{minipage}
-		    \begin{minipage}{0.32\columnwidth} 
-			    \includegraphics[width=\columnwidth]{img_tmp/3_prime_UTR_variant.eps}
-		    \end{minipage}
-		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (Left), splice region variant (Center) and 3’ UTR variant (Right)}
-	    \end{center}
-	\end{minipage}
-
-	\vspace{0.5cm} % intersection space
-
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	% FIGURE 2
+	% FIGURE 3 TFs
 	\begin{minipage}{1\columnwidth}
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	% FIGURE 3
+	% FIGURE 4 Model
 	\begin{minipage}{1\columnwidth}
@@
 	\begin{itemize}	
-	\fontbody
-
-	\item A pipeline to colocalize eQTLs and GWAS variantsWe present a new method to learn a supervised model of associated SNPs based on regulatory features (Fig. 1).
-	\item We present a new method to learn a supervised model of associated SNPs based on regulatory features (Fig. 1).
-
+	\item We have developed a pipeline of eQTL/GWAS colocalization (Fig. 1).
+	\item These colocalization results are in an easy to use web site (Fig. 1).
+	\item We have computed a list of pleiotropic eQTLs (Table 1).
+	\item Pleiotropic variants are enriched in missense and splice region VEP effect consequences (Fig. 2).
+	\item Pleiotropic eQTLs have more eGenes per variant-eTissue, are active in more eTissues per variant-eGene and are associated to more GWAS categories per variant-eGene-eTissue (Fig. 3).
+	\item Pleiotropic variants bind transcription factors (Fig. 4).
+	\item Model explaining the reason of pleiotropy for eQTLs (Fig. 5).
 	\end{itemize}

### 2022-09-10 ms/ms04_discussion.tex
RFR to gwas2eqtl

@@
 %
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/eqtl2gwas} and \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
+The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/gwas2eqtl} and \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
 
@@
 
-The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/eqtl2gwas}.
+The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/gwas2eqtl}.
 %
-The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/eqtl2gwas_pleiotropy}.
+The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
 

### 2022-09-13 ms/ms05_floats.tex
ENH added remap plots
ENH updated based on gwas420

@@
 table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
-]{fig/tab_rsid_most_pleiotropic.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolvi}% specify your coloumns here
+]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
 \hline
@@
 %
-  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+  \caption{Colocalized eQTL/GWAS variants involved in 6 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
 \end{table*}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 1: GO of pleiotropic variants
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 1: GO of pleiotropic variants
+%\begin{figure*}[!tbp]
+%\centering
+%%
+%\begin{subfigure}[]{0.99\textwidth}
+%\textbf{a}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
+%\end{subfigure}
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{subfigure}[]{.99\textwidth}
+%\textbf{b}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
+%\end{subfigure}
+%%
+%\caption{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.} \label{fig:geneontology}
+%\end{figure*}
 
-\begin{figure*}[!tbp]
-\centering
-%
-\begin{subfigure}[]{0.99\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
-\end{subfigure}
-
-\begin{subfigure}[]{.99\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
-\end{subfigure}
-%
-\caption{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.} \label{fig:geneontology}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{figure*}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
-]{fig/tab_region_window_100000_pleio_highest.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolvi}% specify your coloumns here
+]{\floatRelativePath/cmpt_pleiotropic_regions.py/region_window_ms_100000.tsv}{}% use head of csv as column names
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
 \hline
@@
 \end{table*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 2: Length histogram of regions
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 2: Length histogram of regions
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}[!tbp]
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 1: Histograms variants vs GWAS, eGene and eTissues
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 1: Histograms variants vs GWAS, eGene and eTissues
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}[!tbp]
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 2: Scatter plots of winner regions
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 2: Scatter plots of winner regions
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 \begin{figure*}[!tbp]
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{e}
+\textbf{b}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{f}
+\textbf{c}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start132239645_end132497907_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{h}
+\textbf{b}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr6_start31034839_end32478149_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{i}
+\textbf{c}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start111395984_end111645358_categories11.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
 \end{subfigure}
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 3: VEP consequences
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 3: VEP consequences
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}[!tbp]
-\centering
+%\begin{figure*}[!tbp]
+%\centering
+%%
+%\begin{subfigure}[]{.32\textwidth}
+%\textbf{a}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/missense_variant.png}
+%%
+%\end{subfigure}
+%%
+%\begin{subfigure}[]{.32\textwidth}
+%\textbf{b}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/splice_region_variant.png}
+%%
+%\end{subfigure}
+%%
+%\begin{subfigure}[]{.32\textwidth}
+%\textbf{c}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
+%%
+%\end{subfigure}
+%%
+%\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
+%%
+%\end{figure*}
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/missense_variant.png}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 4: Violin plots. eGenes, eTissues and gwas per variants
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/splice_region_variant.png}
-%
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
-%
-\end{subfigure}
-%
-\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
-%
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 4: Violin plots. eGenes, eTissues and gwas per variants
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}[!tbp]
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 6: TF count per GWAS category count
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 6: TF count per GWAS category count
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_0.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_50.png}
 \end{subfigure}
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 7: beta and log pval
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 7: beta and log pval
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}[!tbp]
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Fig 7: graphical conclusions
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 7: graphical conclusions
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}[!tbp]

### 2022-09-13 scripts/pltsctr_x_per_rsid_y_etissue.py

@@ -47,7 +47,7 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 title = "Coloc. eQTL/GWAS variant"
 count_col_name = "etissue_label_count"
-ylim = [0, 40]
+ylim = [0, 20]
 c = 'blue'
 ylabel = "eTissue count"
 

### 2022-09-13 ms/ms05_floats.tex

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%
+%%% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
+%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\begin{figure*}[!tbp]
+%\centering
 %%
-%% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
+%\begin{subfigure}[]{.66\textwidth}
+%\textbf{a}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_tissue_enrich_in_pleio.py/plt.png}
+%\end{subfigure}
 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.66\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_tissue_enrich_in_pleio.py/plt.png}
-\end{subfigure}
-%
-\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:tissues}
-\end{figure*}
+%\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:tissues}
+%\end{figure*}
 %
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{b}
+\textbf{e}
 \\
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{c}
+\textbf{f}
 \\
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{b}
+\textbf{h}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
 \end{subfigure}
@@
 \begin{subfigure}[]{.32\textwidth}
-\textbf{c}
+\textbf{i}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
 \end{subfigure}
@@
 %
-%\begin{figure*}[!tbp]
-%\centering
-%%
-%\begin{subfigure}[]{.32\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/missense_variant.png}
-%%
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.32\textwidth}
-%\textbf{b}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/splice_region_variant.png}
-%%
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.32\textwidth}
-%\textbf{c}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
-%%
-%\end{subfigure}
-%%
-%\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Missense variant, \textbf{b}, splice region variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
-%%
-%\end{figure*}
+\begin{figure*}[!tbp]
+\centering
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/upstream_gene_variant.png}
+%
+\end{subfigure}
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/intron_variant.png}
+%
+\end{subfigure}
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{c}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
+%
+\end{subfigure}
+%
+\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Upstream gene variant, \textbf{b}, intron variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
+%
+\end{figure*}
 %

### 2022-10-01 scripts/plthtmp_disease_comorbidity_matrix.py
ENH Update filterh4

@@ -40,7 +40,7 @@ corr_df = corr_df.loc[corr_df.sum(axis=1) > 0, ]
 corr_df = corr_df[corr_df.columns[corr_df.sum(axis=0) > 0]]
 
 #%% filter row with at least a number of correlations >0.1 larger than 30
-mask = (corr_df >= 0.05).sum(axis=1) > 5
+mask = (corr_df >= 0.04).sum(axis=1) > 10
 corr_df = corr_df.loc[mask]
 corr_df = corr_df[corr_df.columns[mask]]
 
@@ -126,7 +126,7 @@ clustermap_args_dic['yticklabels'] = annotation_df['icd10_level1'].tolist()
 # clustermap_args_dic['dendrogram_ratio'] = (0.05, 0.05)
 
 g = seaborn.clustermap(dis_df, **clustermap_args_dic)
-seaborn.set(font_scale=0.5)
+seaborn.set(font_scale=1)
 # g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize = 10)
 # plt.xlabel('Years', fontsize = 10) # x-axis label with fontsize 15
 # plt.ylabel('Years', fontsize = 10) # x-axis label with fontsize 15
@@ -136,13 +136,13 @@ g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize = 6)
 
 for label in subset1_labels.unique():
     g.ax_col_dendrogram.bar(0, 0, color=subset1_lut[label], label=label, linewidth=0);
-l1 = g.ax_col_dendrogram.legend(title='subset1', loc="upper left", bbox_to_anchor=(0.0, 1), ncol=4, bbox_transform=gcf().transFigure)
+l1 = g.ax_col_dendrogram.legend(title='Category', loc="upper left", bbox_to_anchor=(0.05, 0.95), ncol=4, bbox_transform=gcf().transFigure)
 
 # for label in subset2_labels.unique():
 #     g.ax_row_dendrogram.bar(0, 0, color=subset2_lut[label], label=label, linewidth=0);
 # l2 = g.ax_row_dendrogram.legend(title='subset2', loc="upper left", bbox_to_anchor=(0.0, 0.9), ncol=3, bbox_transform=gcf().transFigure)
 
-plt.subplots_adjust(top=1.1, left=0.00, right=0.7, bottom=0.05)
[…]

### 2022-10-02 ms/ms05_floats.tex
ENH updated scatter regions

@@
 %
-  \caption{Colocalized eQTL/GWAS variants involved in 6 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+  \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
 \end{table*}
@@
 
-\begin{subfigure}[]{.32\textwidth}
+\begin{subfigure}[]{.19\textwidth}
 \textbf{a}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
+\begin{subfigure}[]{.19\textwidth}
 \textbf{b}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
+\begin{subfigure}[]{.19\textwidth}
 \textbf{c}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{d}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{e}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
 \end{subfigure}
 
-\begin{subfigure}[]{.32\textwidth}
-\textbf{d}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{f}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{e}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{g}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{f}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{h}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{i}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{j}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
 \end{subfigure}
 
-\begin{subfigure}[]{.32\textwidth}
-\textbf{g}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{k}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories9.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{h}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{l}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr17_start39548459_end40421164_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
 \end{subfigure}
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{i}
+\begin{subfigure}[]{.19\textwidth}
+\textbf{m}
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr17_start48116861_end49436349_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{n}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.19\textwidth}
+\textbf{o}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
 \end{subfigure}

### 2022-10-02 scripts/cmpt_count_per_rsid.py

@@ -51,7 +51,7 @@ gwas_df.to_csv(tsv_path, sep="\t", index=False)
 gwas_ms_df = gwas_df.copy()
 # gwas_ms_df['cytoband'] = gwas_ms_df['cytoband'].str.split('.', expand=True)[0]
 gwas_ms_df = gwas_ms_df.drop_duplicates('cytoband', keep='first')
-gwas_ms_df = gwas_ms_df.loc[gwas_ms_df['gwas_category_count'] >= 6]
+gwas_ms_df = gwas_ms_df.loc[gwas_ms_df['gwas_category_count'] >= 5]
 # format output
 gwas_ms_df.drop(['gwas_category_count'], inplace=True, axis=1)
 gwas_ms_df['gwas_category_lst'] = gwas_ms_df['gwas_category_lst'].str.replace(',', ', ')

### 2022-10-02 ms/ms05_floats.tex
ENH updated ms

@@
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/upstream_gene_variant.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
 %
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/intron_variant.png}
-%
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/3_prime_UTR_variant.png}
-%
-\end{subfigure}
-%
-\caption{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: \textbf{a}, Upstream gene variant, \textbf{b}, intron variant and \textbf{c}, 3' UTR variant.} \label{fig:vep_consequence}
+\caption{Odds ratio compared of the variants annotated with given variant effect predictor (VEP) consequence and given GWAS category count compared with non pleiotropic variants (GWAS category count one).
+The significance is given for the q-value (FDR 5\%).}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remaptf.py/bxplt_remaptf_per_rsid_flank_50.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_remap_crm_gwas_categories.py/remap_crm_fisher.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
 \end{subfigure}

### 2022-10-03 scripts/cmpt_public_web_files.py
ENH added coded to create web coloc data

@@ -1,0 +1,55 @@
+from sqlalchemy import create_engine, select, asc
+from gwas2eqtl_pleiotropy.Logger import Logger
+from gwas2eqtl_pleiotropy.db import meta, coloc, gwas_annot_tbl, ensg2symbol_tbl, rsid2cytoband_tbl
+
+import os
+import pandas
+import pathlib
+import sys
+
+
+#%%
+help_cmd_str = "todo"
+try:
+    db_sqlite = sys.argv[1]
+    coloc_tsv_gz_path = sys.argv[2]
+    if len(sys.argv) > 3:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+outdir_path = os.path.dirname(coloc_tsv_gz_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
+
+#%% Insert into coloc
+engine = create_engine("sqlite:///{}".format(db_sqlite), echo = False)
+meta.create_all(engine)
+
+#%%
+with engine.connect() as con:
+    for chrom in range(1, 23):  # loop over chroms to avoid memory problems
+        # print(chrom)
+        sel_cols = [coloc, gwas_annot_tbl.c.gwas_trait, gwas_annot_tbl.c.gwas_class, ensg2symbol_tbl.c.symbol, rsid2cytoband_tbl.c.cytoband]
+        sel = select(sel_cols).distinct()
+        sel = sel.where(coloc.c.gwas_id == gwas_annot_tbl.c.gwas_id)
+        sel = sel.where(coloc.c.egene == ensg2symbol_tbl.c.gene_id)
+        sel = sel.where(coloc.c.rsid == rsid2cytoband_tbl.c.rsid)
+        sel = sel.where(coloc.c.chrom == chrom)
+        sel = sel.order_by(asc(coloc.c.pos))
+
+        Logger.info("SQL query: chrom {}".format(chrom))
+        result_lst = con.execute(sel).fetchall()
+        if len(result_lst) == 0:
+            continue
+        coloc_df = pandas.DataFrame(result_lst)
+        coloc_df.drop(['id'], inplace=True, axis=1)
[…]

### 2022-10-03 scripts/cmpt_vep_consequence_fisher_bak.py

@@ -1,88 +1,0 @@
-import numpy
-import pandas
-import sys
-
-from gwas2eqtl_pleiotropy.Logger import Logger
-from scipy.stats import fisher_exact
-from statsmodels.stats import multitest as multitest
-
-
-#%%
-help_cmd_str = "todo"
-try:
-    vep_input_path = sys.argv[1]
-    vep_output_path = sys.argv[2]
-    upper_var_gwas_cat_count = int(sys.argv[3])
-    consequence_tsv_path = sys.argv[4]
-    if len(sys.argv) > 5:
-        print("""Two many arguments!
-        {}""".format(help_cmd_str))
-        sys.exit(1)
-except IndexError:
-    print("""Argument missing!
-    {}""".format(help_cmd_str))
-    sys.exit(1)
-
-
-#%%
-vep_input_column_lst = ['chrom', 'start', 'end', 'alleles', 'strand', 'rsid', 'gwas_category_count']
-vep_input_df = pandas.read_csv(vep_input_path, sep=" ", header=None, names=vep_input_column_lst)
-columns = ['#Uploaded_variation', 'Location', 'Allele', 'Gene', 'Feature', 'Feature_type', 'Consequence', 'cDNA_position', 'CDS_position', 'Protein_position', 'Amino_acids', 'Codons', 'Existing_variation', 'Extra']
-vep_output_df = pandas.read_csv(vep_output_path, sep="\t", comment='#', header=None, names=columns)
-
-
-#%%
-df0 = vep_input_df.merge(vep_output_df, left_on='rsid', right_on='#Uploaded_variation', how='left')
-df0 = df0[['rsid', 'gwas_category_count', 'Consequence']].drop_duplicates()
-
-#%%
-df0['Consequence'] = df0['Consequence'].str.split(',')
-df = df0.explode('Consequence').drop_duplicates()
-
-out_columns = ['consequence', 'gwas_category_count', 'a_pleio_x_with_consequence', 'b_pleio_1_with_consequence', 'c_pleio_x_wout_consequence', 'd_pleio_1_wout_consequence', 'oddsr', 'p']
-out_dic = dict(zip(out_columns, [[] for i in range(len(out_columns))]))
-out_df = pandas.DataFrame(out_dic)
-
-all_df = vep_input_df[['rsid', 'gwas_category_count']].drop_duplicates()
-Logger.info('Consequence list: {}'.format(sorted(df['Consequence'].unique())))
[…]

### 2022-10-14 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH presentation gold

\documentclass{beamer}
\usetheme{Boadilla}

\usepackage{graphicx}
\usepackage{subcaption}
% Change size of footnotes
\renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
\title{Beamer presentation}
%\subtitle{Using Beamer}
\author{Aitor Gonz\'alez}
\institute{TAGC/INSERM U1090}
\date{Internal Seminar: January 19, 2022}

% Add section slide
\AtBeginSection[]
{
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}

\titlepage

\end{frame}

\end{document}

### 2022-10-14 presentation_gold2022_paris/presentation_gold2022_paris_bak.tex
ENH presentation gold

\documentclass{beamer}
\usetheme{Boadilla}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{csvsimple}
 The R package coloc performs colocalization tests between genetic traits:
https://CRAN.R-project.org/package=coloc
% Change size of footnotes
\renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
\title{Analysis of gene regulatory properties underlying trait pleiotropy}
\subtitle{Meeting GOLD 2022}
\author{Aitor Gonz\'alez}
\institute{Aix Marseille Univ, INSERM, TAGC}
\date{Oct. 19, 2022}

% Add section slide
\AtBeginSection[]
{
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
}

\logo{\includegraphics[width=.3\textwidth]{fig_internal_seminal/tagc_inserm_amu.png}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}

\titlepage

\end{frame}

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Genome-wide association studies (GWAS)}

\begin{itemize}
\item Genetic loci with higher frequency in individuals with a given disease
\item A very active field with many studies
\item Many genetic variants are involved in several phenotypes
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Limitations of GWAS}

\begin{itemize}
\item Correlation between variant alleles
\item Which cell types are causal to the disease?
\item Over 90\% GWAS variants fall in non-coding regions
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Limitations of GWAS}

\begin{figure}[!]
\includegraphics[width=0.25\textwidth]{{presentation_goldwp2022_paris/fig/41586_2017_Article_BFnature24284_Fig1_HTML}
\caption{GWAS for breast cancer}
\end{figure}

\begin{itemize}
\item Correlation between variants, ie. LD
\item Which cell types are causal to the disease?
\item Over 90\% GWAS variants fall in non-coding regions
\end{itemize}
%
\vfill
%
These points can be partially addressed by looking at eQTLs in different tissues

\let\thefootnote\relax\footnotetext{doi:10.3389/fgene.2020.00424}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Integration of GWAS and eQTL variants:  Colocalisation analysis}

\begin{figure}[!]
\includegraphics[width=0.6\textwidth]{presentation_goldwp2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4.png}
\end{figure}

\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
\end{frame}

\section{Objectives} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Objectives}

\begin{itemize}
\item To identify eQTLs in immune cells that colocalise with frequent variants associated to cancer risk
\item To identify which immune cells are involved in cancer predisposition
\item To characterize these eQTLs involved in immune cells and cancer predisposition
\end{itemize}
\end{frame}

\section{Materials and Methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Data and Algorithm}

\begin{itemize}
\item IEA OpenGWAS database
\item EBI eQTLs database
\item CRAN - Package coloc
\end{itemize}

\let\thefootnote\relax\footnotetext{https://gwas.mrcieu.ac.uk}
\end{frame}

\section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Conclusions}

\begin{itemize}
\item A  colocalisation pipeline to link EBI Catalogue eQTLs and IEA OpenGWAS signals
\item Several colocalisation signals between immune cell eQTLs and cancer loci
\item Preliminary analyses of these colocalised signals
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Perspectives}

\begin{itemize}
\item Use H1 as background dataset: H1=Causal SNP in cancer w/out eQTL
\item Statistics of cell type enrichments
\item Causal SNP characterization: Motif analyses
\item Other applications:
\begin{itemize}
\item Pleiotropy of eQTLs: eQTLs affecting two genes or same gene in two cell-types
\item Causality between other cell type eQTLs and diseases: Immunity and Covid-19
\end{itemize}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{Acknowledgements}

\begin{itemize}
\item L\'eopoldine Lecerf (M1)
\item P Paul, P Rihet, M Michel
\item S Marquet, S Spicuglia
\item Funding:
\begin{itemize}
\item Institut Cancer et Immunologie (ICI)
\item Agence nationale de la recherche (ANR)
\end{itemize}
\end{itemize}

\end{frame}

\end{document}

### 2022-10-14 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH

@@
 \usepackage{subcaption}
+\usepackage{csvsimple}
+
 % Change size of footnotes
 \renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
-\title{Beamer presentation}
-%\subtitle{Using Beamer}
+\title{Analysis of gene regulatory properties underlying trait pleiotropy}
+\subtitle{Meeting GOLD 2022}
 \author{Aitor Gonz\'alez}
-\institute{TAGC/INSERM U1090}
-\date{Internal Seminar: January 19, 2022}
+\institute{Aix Marseille Univ, INSERM, TAGC}
+\date{Oct. 19, 2022}
 
@@
 
+\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Genome-wide association studies (GWAS)}
+
+\begin{itemize}
+\item Genetic loci with higher frequency in individuals with a given disease
+\item A very active field with many studies
+\item Many genetic variants are involved in several phenotypes
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Limitations of GWAS}
+
+\begin{figure}[!]
+\includegraphics[width=0.25\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/41586_2017_Article_BFnature24284_Fig1_HTML.jpg}
+\caption{GWAS for breast cancer}
+\end{figure}
+
+\begin{itemize}
+\item Correlation between variants, ie. LD
+\item Which cell types are causal to the disease?
+\item Over 90\% GWAS variants fall in non-coding regions
+\end{itemize}
+%
+\vfill
+%
+These points can be partially addressed by looking at eQTLs in different tissues
+
+\let\thefootnote\relax\footnotetext{doi:10.3389/fgene.2020.00424}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Integration of GWAS and eQTL variants:  Colocalisation analysis}
+
+\begin{figure}[!]
+\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4.png}
+\end{figure}
+
+\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+\end{frame}
+
+\section{Objectives} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Objectives}
+
+\begin{itemize}
+\item To carry out a large scale colocalization analysis between GWAS and eQTL variants
+\item To identify pleiotropic genetic loci
+\item To characterize the regulatory causas of this pleiotropy
+\end{itemize}
+\end{frame}
+
+\section{Materials and Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Data and Algorithm}
+
+\begin{itemize}
+\item IEA OpenGWAS database
+\item EBI eQTLs database
+\item CRAN - Package coloc
+\end{itemize}
+
+\let\thefootnote\relax\footnotetext{https://gwas.mrcieu.ac.uk}
+\end{frame}
+
+\section{Results}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Large scale GWAS/eQTL colocalization analysis}
+
+\begin{itemize}
+\item Colocalization of variants between 418 GWAS and 127 eQTL studies, ...
+\end{itemize}
+
+%\begin{figure}[!]
+%\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/poster_eccb22_barcelona/img/web.eps}
+%\end{figure}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do GWAS traits cluster coherently?}
+
+\begin{itemize}
+\item Distance between GWAS traits based on eQTL beta
+\item Clustering of immune diseases coherent
+\item eQTL beta of GWAS variants coherent
+\end{itemize}
+
+\begin{figure}[!]
+\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
+\end{figure}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{frame}
+%\frametitle{Which are the most pleiotropic variants?}
+
+%\begin{table}[!tbp]
+%\centering
+%\scriptsize
+%\hline
+%\csvreader[separator=tab,
+%tabular=ccrrp{0.4\textwidth},
+%head,
+%table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
+%]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+%{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
+%\hline
+%%
+%\vspace{15pt}
+%%
+%\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+%\end{table}
+
+%\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Which are the most pleiotropic variants?}
+
+\begin{itemize}
+\item Distance between GWAS traits based on eQTL beta
+\item Clustering of immune diseases coherent
+\item eQTL beta of GWAS variants coherent
+\end{itemize}
+
+\begin{figure}[!]
+\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
+\end{figure}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants have more egenes?}
+
+\begin{figure}[!tbp]
+\centering
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.32\textwidth}
+\textbf{c}
+\\
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
+\end{subfigure}
+%
+\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS categories per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
+%
+\end{figure}
+
+\end{frame}
+
+\section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Conclusions}
+
+\begin{itemize}
+\item A  colocalisation pipeline to link EBI Catalogue eQTLs and IEA OpenGWAS signals
+\item Several colocalisation signals between immune cell eQTLs and cancer loci
+\item Preliminary analyses of these colocalised signals
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Perspectives}
+
+\begin{itemize}
+\item Use H1 as background dataset: H1=Causal SNP in cancer w/out eQTL
+\item Statistics of cell type enrichments
+\item Causal SNP characterization: Motif analyses
+\item Other applications:
+\begin{itemize}
+\item Pleiotropy of eQTLs: eQTLs affecting two genes or same gene in two cell-types
+\item Causality between other cell type eQTLs and diseases: Immunity and Covid-19
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Acknowledgements}
+
+\begin{itemize}
+\item L\'eopoldine Lecerf (M1)
+\item P Paul, P Rihet, M Michel
+\item S Marquet, S Spicuglia
+\item Funding:
+\begin{itemize}
+\item Institut Cancer et Immunologie (ICI)
+\item Agence nationale de la recherche (ANR)
+\end{itemize}
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2022-10-14 presentation_gold2022_paris/presentation_gold2022_paris_bak.tex
ENH

@@
 \renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
-\title{Analysis of gene regulatory properties underlying trait pleiotropy}
+\title{Analysis of gene regulatory properties underlying trait pleiotropy2}
 \subtitle{Meeting GOLD 2022}
@@
 \begin{figure}[!]
-\includegraphics[width=0.25\textwidth]{{presentation_goldwp2022_paris/fig/41586_2017_Article_BFnature24284_Fig1_HTML}
+\includegraphics[width=0.25\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/41586_2017_Article_BFnature24284_Fig1_HTML.jpg}
 \caption{GWAS for breast cancer}

### 2022-10-15 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH presentation gold meeting

@@
 \end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Limitations of GWAS}
-
-\begin{figure}[!]
-\includegraphics[width=0.25\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/41586_2017_Article_BFnature24284_Fig1_HTML.jpg}
-\caption{GWAS for breast cancer}
-\end{figure}
-
+%
+\vfill
+%
+Limitations
+%
+\vfill
+%
 \begin{itemize}
@@
 %
-These points can be partially addressed by looking at eQTLs in different tissues
+These limitations can be partially addressed by looking at eQTLs in different tissues
 
-\let\thefootnote\relax\footnotetext{doi:10.3389/fgene.2020.00424}
 \end{frame}
@@
 
-\section{Objectives} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Materials and Methods}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}
-\frametitle{Data and Algorithm}
+\frametitle{Data and Algorithms}
 
@@
 \begin{frame}
-\frametitle{Which are the most pleiotropic variants?}
+\frametitle{How are pleiotropic variants distributed?}
 
-\begin{itemize}
-\item Distance between GWAS traits based on eQTL beta
-\item Clustering of immune diseases coherent
-\item eQTL beta of GWAS variants coherent
-\end{itemize}
-
-\begin{figure}[!]
-\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
-\end{figure}
+\begin{columns}
+\begin{column}{0.32\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+     \end{center}
+\end{column}
+\begin{column}{0.32\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+     \end{center}
+\end{column}
+\begin{column}{0.32\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+     \end{center}
+\end{column}
+\end{columns}
+%
+\vfill
+%
+Neighboring variants show similar pleiotropy.
 
@@
 \begin{frame}
-\frametitle{Do pleiotropic variants have more egenes?}
+\frametitle{Where are the pleiotropic variants?}
 
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{a}
-\\
+    \begin{center}
+\includegraphics[width=0.8\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
+     \end{center}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants associate to more egenes?}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_egenes.png}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
 \includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{b}
-\\
+     \end{center}
+\end{column}
+\end{columns}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants associate to more etissues?}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_etissues.png}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
 \includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
+     \end{center}
+\end{column}
+\end{columns}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_gwas.png}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
 \includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-\end{subfigure}
-%
-\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS categories per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
-%
-\end{figure}
+     \end{center}
+\end{column}
+\end{columns}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants bind more transcription factors?}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+     \end{center}
+\end{column}
+\end{columns}
 
@@
 
+Characteristics of pleiotropic variants
+%
 \begin{itemize}
-\item A  colocalisation pipeline to link EBI Catalogue eQTLs and IEA OpenGWAS signals
-\item Several colocalisation signals between immune cell eQTLs and cancer loci
-\item Preliminary analyses of these colocalised signals
+\item Variant pleiotropy is correlated in regions
+\item Enrichment of variants upstream (incl. 5'UTR) and downstream of genes (incl. 3'UTR)
+\item Enrichment of transcription factor and CRM binding
+\item Increased number of associated egenes and etissues
+\end{itemize}
+%
+\vfill
+%
+Other outcomes of this study
+%
+\begin{itemize}
+\item A large analysis of GWAS/eQTL colocalisations
+\item A list of pleiotropic variants and regions
+\item Analysis of regulatory properties of pleiotropic variants
 \end{itemize}
@@
 \begin{itemize}
-\item Use H1 as background dataset: H1=Causal SNP in cancer w/out eQTL
-\item Statistics of cell type enrichments
-\item Causal SNP characterization: Motif analyses
-\item Other applications:
-\begin{itemize}
-\item Pleiotropy of eQTLs: eQTLs affecting two genes or same gene in two cell-types
-\item Causality between other cell type eQTLs and diseases: Immunity and Covid-19
-\end{itemize}
+\item A web portail to share data
+\item Causality of egenes on gwas (Looking for experts ;-) )
 \end{itemize}
@@
 \item L\'eopoldine Lecerf (M1)
-\item P Paul, P Rihet, M Michel
-\item S Marquet, S Spicuglia
-\item Funding:
+\item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
+\end{itemize}
+%
+\vfill
+%
+Funding
+%
 \begin{itemize}
@@
 \item Agence nationale de la recherche (ANR)
-\end{itemize}
+\item GOLD
 \end{itemize}

### 2022-10-17 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH
ENH presentation o

@@
 %
-\vfill
-%
 \begin{itemize}
@@
 
+\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Integration of GWAS and eQTL variants:  Colocalisation analysis}
+\frametitle{Expression quantitative trait loci (eQTL)}
 
-\begin{figure}[!]
-\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4.png}
-\end{figure}
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}
+
+\end{column}
+\end{columns}
+
+\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Colocalisation analysis of GWAS and eQTL variants}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4bc.png}
+     \end{center}
+\end{column}
+\end{columns}
 
@@
 
-\begin{itemize}
-\item To carry out a large scale colocalization analysis between GWAS and eQTL variants
+\begin{enumerate}
+\item Large scale colocalization analysis between GWAS and eQTL variants
 \item To identify pleiotropic genetic loci
-\item To characterize the regulatory causas of this pleiotropy
-\end{itemize}
+\item To characterize the regulatory causes of trait pleiotropy
+\end{enumerate}
 \end{frame}
@@
 
+\begin{columns}
+\begin{column}{0.5\textwidth}
 \begin{itemize}
-\item IEA OpenGWAS database
-\item EBI eQTLs database
-\item CRAN - Package coloc
+\item IEA OpenGWAS database$^1$: 244,881e6 associations from 42e3 GWAS datasets
+\item EBI eQTLs database$^2$: Processed gene expression and splicing QTLs from all public studies on human.
+\item CRAN - Package coloc$^3$: Genetic colocalisation analysis of two phenotypes to look for shared common genetic causal variants
 \end{itemize}
+\end{column}
+\begin{column}{0.5\textwidth}
+\begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/pgen.1004383.g001.png}
+\end{center}
+\end{column}
+\end{columns}
 
-\let\thefootnote\relax\footnotetext{https://gwas.mrcieu.ac.uk}
+\let\thefootnote\relax\footnotetext{$^1$https://gwas.mrcieu.ac.uk}
+\let\thefootnote\relax\footnotetext{$^2$https://www.ebi.ac.uk/eqtl}
+\let\thefootnote\relax\footnotetext{$^3$https://cran.r-project.org/web/packages/coloc/index.html}
 \end{frame}
@@
 \begin{itemize}
-\item Colocalization of variants between 418 GWAS and 127 eQTL studies, ...
+\item Colocalization analysis of 418 GWAS and 127 eQTL studies (53e3 pairs)
+\item 143e3 shared causal variants with probability $\geq$ 0.8
+\item Percentage of explained GWAS loci: 0$\%$ - 77$\%$ (Average 33$\%$)
 \end{itemize}
@@
 
-\begin{itemize}
-\item Distance between GWAS traits based on eQTL beta
-\item Clustering of immune diseases coherent
-\item eQTL beta of GWAS variants coherent
-\end{itemize}
+Distance between GWAS traits based on eQTL beta
 
 \begin{figure}[!]
-\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
+\includegraphics[width=0.5\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
 \end{figure}
+
+Clustering coherent
 
@@
 \begin{frame}
-\frametitle{Where are the pleiotropic variants?}
+\frametitle{Where are the pleiotropic variants?$^1$}
 
@@
      \end{center}
+varianb
+\let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}
 
@@
 \begin{frame}
-\frametitle{Do pleiotropic variants bind more transcription factors?}
+\frametitle{Do pleiotropic variants bind more transcription factors?$^1$}
 
@@
 
+\let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}
+
 \section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Characteristics of pleiotropic variants
+Regulatory characteristics of pleiotropic variants
 %
@@
 \item A list of pleiotropic variants and regions
-\item Analysis of regulatory properties of pleiotropic variants
 \end{itemize}

### 2022-10-17 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH finished refactoring class

@@
     \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
      \end{center}
@@
     \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
      \end{center}

### 2022-10-17 scripts/pltbar_vep_consequence.py

@@ -53,7 +53,8 @@ for c in g.containers:
     gwas_class_count = c.get_label()
     labels = in_df.loc[in_df['gwas_class_count'] == gwas_class_count, 'signif'].tolist()
     # set the bar label
-    g.bar_label(c, labels=labels, label_type='edge', padding=5, verticalalignment='bottom')
+    # g.bar_label(c, labels=labels, label_type='edge', padding=5, verticalalignment='bottom')
+    g.bar_label(c, labels=labels, label_type='edge', padding=5)
 plt.legend(title='GWAS cat.', loc='best',title_fontsize='small')
 
 plt.tight_layout()

### 2022-10-17 scripts/pltsctr_x_per_rsid_y_egene.py

@@ -17,10 +17,11 @@ seaborn.set_theme(**seaborn_theme_dic)
 #%%
 help_cmd_str = "todo"
 try:
-    region_window_100000_tsv_path = sys.argv[1]
-    count_per_rsid_egene_tsv_path = sys.argv[2]
-    outdir_path = sys.argv[3]
-    if len(sys.argv) > 4:
+    max_gwas_class_count = int(sys.argv[1])
+    region_window_100000_tsv_path = sys.argv[2]
+    count_per_rsid_egene_tsv_path = sys.argv[3]
+    outdir_path = sys.argv[4]
+    if len(sys.argv) > 5:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -54,7 +55,7 @@ ylabel = "eGene count"
 count_per_rsid_df['pos'] = count_per_rsid_df['pos'].astype('int')
 
 #%% Loop over regions
-pleiotropic_regions_df = region_window_100000_df.loc[region_window_100000_df['gwas_class_count'] >= 6, ['chrom', 'start', 'end', 'gwas_class_count']]
+pleiotropic_regions_df = region_window_100000_df.loc[region_window_100000_df['gwas_class_count'] >= max_gwas_class_count, ['chrom', 'start', 'end', 'gwas_class_count']]
 for rowi, row in pleiotropic_regions_df.iterrows():
     chrom = row['chrom']
     start = row['start']

### 2022-10-18 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH

@@
 %\begin{frame}
-%\frametitle{Which are the most pleiotropic variants?}
+%\frametitle{Diseases?}
 
@@
 
-%\end{frame}
+\end{frame}
 
@@
     \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
      \end{center}
@@
 \end{columns}
-%
-\vfill
-%
-Neighboring variants show similar pleiotropy.
+
+\begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/chr5_131912097_132802472.png}
+\end{center}
+
+\small
+\begin{itemize}
+\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IL3, IL4, IL5, ...
+\item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
+\item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
+\end{itemize}
+\normalsize
 

### 2022-10-18 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH presentation

@@
 \begin{itemize}
-\item Genetic loci with higher frequency in individuals with a given disease
+\item Genetic alleles with higher frequency in individuals with a given disease
 \item A very active field with many studies
@@
 \begin{itemize}
-\item Correlation between variants, ie. LD
+\item Correlation between variants
 \item Which cell types are causal to the disease?
@@
 \item 143e3 shared causal variants with probability $\geq$ 0.8
-\item Percentage of explained GWAS loci: 0$\%$ - 77$\%$ (Average 33$\%$)
+\item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
 \end{itemize}
 
-%\begin{figure}[!]
-%\includegraphics[width=0.6\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/poster_eccb22_barcelona/img/web.eps}
-%\end{figure}
+\begin{figure}[!]
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/coloc_web.png}
+\end{figure}
 
@@
 
-Distance between GWAS traits based on eQTL beta
+Distance between GWAS traits based on eQTL/egene beta
 
@@
 
-Clustering coherent
+GWAS/eQTL colocalization analysis allows to cluster diseases coherently
 
@@
 \begin{itemize}
-\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IL3, IL4, IL5, ...
+\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
 \item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
@@
     \begin{center}
-\includegraphics[width=0.8\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
+\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
      \end{center}
-varianb
+
+Upstream and downstream of genes and in the intron
+
 \let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}
@@
 \begin{frame}
-\frametitle{Do pleiotropic variants associate to more egenes?}
+\frametitle{Why are these variants pleiotropic? Do pleiotropic variants associate to more egenes?}
 
@@
 \end{columns}
+
+More egenes per variant-etissue pair.
 
@@
 \frametitle{Do pleiotropic variants bind more transcription factors?$^1$}
+
+This could explain why pleiotropic variants active in more etissues and associated to more egenes
 

### 2022-10-19 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH

@@
 \renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
-\title{Analysis of gene regulatory properties underlying trait pleiotropy}
+\title{Analysis of gene regulatory properties underlying pleiotropy between complex traits}
 \subtitle{Meeting GOLD 2022}
@@
 %
-Limitations
+Questions:
 %
 \begin{itemize}
-\item Correlation between variants
-\item Which cell types are causal to the disease?
-\item Over 90\% GWAS variants fall in non-coding regions
+\item Correlation between variants. Which is the causal variant?
+\item Which cell type important for disease?
+\item Over 90\% GWAS variants are non-coding. What is their role?
 \end{itemize}
@@
 %
-These limitations can be partially addressed by looking at eQTLs in different tissues
+Questions partially addressed by looking at eQTLs in different tissues
 
@@
 \begin{frame}
-\frametitle{Colocalisation analysis of GWAS and eQTL variants}
+\frametitle{Annotation of GWAS variants with gene and tissue targets using colocalization analysis}
 
@@
 \begin{enumerate}
-\item Large scale colocalization analysis between GWAS and eQTL variants
-\item To identify pleiotropic genetic loci
-\item To characterize the regulatory causes of trait pleiotropy
+\item Large scale gene and tissue annotation of variants of complex traits using GWAS/eQTL colocalization analysis
+\item To identify genetic variants leading to pleiotropy of complex traits
+\item To characterize the gene regulatory mechanisms of trait pleiotropy
 \end{enumerate}
@@
 
-GWAS/eQTL colocalization analysis allows to cluster diseases coherently
+\begin{itemize}
+\item Disease clusters are coherent $\rightarrow$ Validates approche and data
+\item Definition of disease classes
+\end{itemize}
 
@@
 %\end{table}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{How are pleiotropic variants defined?}
+
+\begin{itemize}
+\item Manually annotate variants based on trait clustering and ICD in 129 classifications
+\item Aggregate colocalized variants based on trait classes
+\item Analysed variants in groupes based on the number of trait classes
+\end{itemize}
 
@@
 \begin{frame}
-\frametitle{Why are these variants pleiotropic? Do pleiotropic variants associate to more egenes?}
+\frametitle{Do pleiotropic variants associate to more eQTL genes?}
 
@@
 \begin{frame}
-\frametitle{Do pleiotropic variants associate to more etissues?}
+\frametitle{Do pleiotropic variants associate to more eQTL tissues?}
 
@@
 \begin{frame}
-\frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_gwas.png}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-     \end{center}
-\end{column}
-\end{columns}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
 \frametitle{Do pleiotropic variants bind more transcription factors?$^1$}
-
-This could explain why pleiotropic variants active in more etissues and associated to more egenes
 
@@
 
+\begin{itemize}
+\item This could explain why pleiotropic variants active in more etissues and associated to more egenes
+\end{itemize}
+
 \let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}
@@
 \begin{frame}
-\frametitle{Conclusions}
+\frametitle{Concluding remarks}
 
-Regulatory characteristics of pleiotropic variants
+Outcomes interesting for the GOLD population genotypes
+%
+\begin{itemize}
+\item A large set of causal GWAS variants annotated with target genes and tissues
+\item A list of pleiotropic variants and regions
+\end{itemize}
+%
+\vfill
+%
+Conclusion about regulatory characteristics of pleiotropic variants
 %
@@
 \item Increased number of associated egenes and etissues
-\end{itemize}
-%
-\vfill
-%
-Other outcomes of this study
-%
-\begin{itemize}
-\item A large analysis of GWAS/eQTL colocalisations
-\item A list of pleiotropic variants and regions
 \end{itemize}
@@
 \begin{itemize}
-\item Institut Cancer et Immunologie (ICI)
+\item Institut Cancer et Immunologie - Aix-Marseille Univ.
 \item Agence nationale de la recherche (ANR)
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}
+\frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
+
+\begin{columns}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_gwas.png}
+     \end{center}
+\end{column}
+\begin{column}{0.5\textwidth}  %%<--- here
+    \begin{center}
+\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
+     \end{center}
+\end{column}
+\end{columns}
+
+\end{frame}
+
 \end{document}

### 2022-11-12 ms/ms02_intro.tex

--- 2022-08-27 ms/ms02_intro.tex
@@
 
-Genome-wide association studies (GWAS) allow to define the impact of genetic variants on a trait in the population.
+% GWAS in general
+Genome-wide association studies (GWAS) aim at associating common genetic variants with common traits.
 %
-The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
+The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007...Nature,2018.Parkinson.Buniello}.
 %
@@
 
-However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
+% Short comings of GWAS
+However interpretation of the GWAS variants is complicate because of three reasons \citep{2020.Trynka.CanoGamez}.
 %
@@
 %
-Third, GWAS variants affect most often non-coding regions.
+Third, GWAS variants affect often non-coding regions.
 
+% Intererest of studying GWAS
 Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
@@
 %
-Also the distribution of eQTLs and GWAS variants in the genome have been suggested to be different because of negative selection.
+Moreover the distribution of eQTLs and GWAS variants in the genome have been suggested to be different because of negative selection \citep{2022.Pritchard.Mostafavi}.
 
-Despite these previous disadvantages eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (egenes) and tissues (etissues) (REF).
+Nevertheless eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (eGenes) and tissues (eTissues) \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
 %
-In addition colocalization anlaysis can pinpoint more precisely the causal variants that either data set independently (REF).
+In addition colocalization analysis can help prioritize causal variants.
 
-The large number of GWAS has allowed to discover that many genetic variants are involved in different traits.
+The large number of GWAS has allowed to observe that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
 %
-TODO explain Watanabe.
+In that work, the analysed loci covered 60\% of the genome and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
 %
-Because GWAS variants affect mostly non-coding regions, it makes sense to investigate trait pleiotropy in terms of non-coding gene regulatory functions.
+They also found that pleiotropic genes and variants where less tissue specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
+%
+However the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature}.
+%
+In addition, no hypothesis have been directly tested and proposed for the cause of pleiotropy.
 
-For these reasons, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
+In this work, I have taken advantage of two well formatted eQTL and GWAS data sets from the EBI eQTL Catalogue and the IEU OpenGWAS to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
 %
-I have used these colocalization results to investigate the gene regulatory properties of pleiotropic variants.
+I have developed a pipeline that takes these data and carries out colocalization analysis between all pairs of eQTLs studies and GWAS
 %
-This global colocalization analysis is also useful to help interpret GWAS variants.
+I have used the colocalization results to investigate how gene regulatory properties of pleiotropic variant contribute to trait pleiotropy.
+%
+These results are also helpful to prioritize the causal variants and to interpret the molecular mechanisms of functional variants.
+%
+Therefore I provide the whole results as an interactive table that can be easily browsed by biologists.

### 2022-11-12 ms/ms05_floats.tex
ENH added coded to create web coloc data

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 1: Comorbidity matrix
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\begin{figure}[!tbp]
+\centering
+%
+\includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr.png}
+%
+\caption{Commorbidity matrix.}
+%
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% full size table is table*
-\begin{table*}[!tbp]
+% full size table is table
+\begin{table}[!tbp]
 \centering
 \scriptsize
-\hline
+%\hline
 \csvreader[separator=tab,
@@
 {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
-\hline
+%\hline
 %
@@
 \caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
-\end{table*}
+\end{table}
 
@@
 %
-%\begin{figure*}[!tbp]
+%\begin{figure}[!tbp]
 %\centering
@@
 %\caption{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.} \label{fig:geneontology}
-%\end{figure*}
+%\end{figure}
 
@@
 %\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:tissues}
-%\end{figure*}
+%\end{figure}
 %
@@
 
-% full size table is table*
-\begin{table*}[!tbp]
+% full size table is table
+\begin{table}[!tbp]
 \centering
 \scriptsize
-\hline
+%\hline
 \csvreader[
@@
 \caption{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
-\end{table*}
+\end{table}
 %
@@
 %
-\begin{figure*}[!tbp]
+\begin{figure}[!tbp]
 \centering
@@
 \textbf{b}, Cumulative sum ordered inversely with the number of GWAS categories.} \label{fig:pleiotropy_region_distribution}
-\end{figure*}
+\end{figure}
 %
@@
 \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS categories (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
-\end{figure*}
+\end{figure}
 %
@@
 %
-\begin{figure*}[!tbp]
+\begin{figure}[!tbp]
 
-\begin{subfigure}[]{.19\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-\end{subfigure}
+%\begin{subfigure}[]{.19\textwidth}
+%\textbf{a}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
+%\end{subfigure}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
 \end{subfigure}
 
-\begin{subfigure}[]{.19\textwidth}
-\textbf{f}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-\end{subfigure}
+%\begin{subfigure}[]{.19\textwidth}
+%\textbf{f}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
+%\end{subfigure}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
 \end{subfigure}
 
-\begin{subfigure}[]{.19\textwidth}
-\textbf{k}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-\end{subfigure}
+%\begin{subfigure}[]{.19\textwidth}
+%\textbf{k}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
+%\end{subfigure}
 %
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start55721801_end56171027_categories7.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
 \end{subfigure}
@@
 \\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr19_start48668049_end48756272_categories8.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
 \end{subfigure}
@@
 %
-\end{figure*}
+\end{figure}
 %
@@
 
-\begin{figure*}[!tbp]
+\begin{figure}[!tbp]
 \centering
@@
 %
-\end{figure*}
+\end{figure}

### 2022-11-12 presentation_gold2022_paris/presentation_gold2022_paris.tex
ENH

@@
 \begin{itemize}
-\item Disease clusters are coherent $\rightarrow$ Validates approche and data
+\item Disease clusters are coherent
 \item Definition of disease classes
@@
 
-More egenes per variant-etissue pair.
-
 \end{frame}
@@
 
+\let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}
+
 \end{frame}
-
-\begin{itemize}
-\item This could explain why pleiotropic variants active in more etissues and associated to more egenes
-\end{itemize}
-
-\let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}
 
@@
 %
-Conclusion about regulatory characteristics of pleiotropic variants
+What are the gene regulatory properties underlying pleiotropy between complex traits?
 %
@@
 \item A web portail to share data
-\item Causality of egenes on gwas (Looking for experts ;-) )
+\item Causality of eQTL genes on GWAS (Looking for experts ;-) )
 \end{itemize}

### 2022-11-12 ms/ms01_titlepage.tex
ENH ms and refactoring
ENH updated with 418
MS

@@
 
-\title{Trait pleiotropy of regulatory variants increases with the variety of bound transcription factors, regulated genes and active tissues}
+\title{Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors}
 %\shorttitle{Pleiotropy and gene regulatory variants}
@@
 %
-In this work, I investigate the link between gene regulation and trait pleiotropy for frequent variants.
+In this work, I investigate the link between frequent gene regulatory variants and pleiotropy.
 
 % Verify numbers
-First I have computed the colocalization of 413 GWAS studies and 127 eQTL studies and have observed 143,119 colocalized pairs with a probability above 0.8.
+First I have computed the colocalization probability of 418 GWAS and 127 eQTL studies and I have observed 10,334 GWAS/eQTL pairs with at least one colocalized variants with a probability above 0.8.
+%SELECT count(DISTINCT concat(`gwas_id`, `eqtl_id`)) FROM `gwas2eqtl_modelgwas2eqtl` where `PP_H4_abf`>=0.8; 10334
+%Pourcentage of GWAS
 %
@@
 
-\begin{keywords2}
-eGenes; eQTLs
-\end{keywords2}
+% Keywords
+% eGenes; eQTLs

### 2022-11-12 ms/ms03_results.tex
MS
ENH

@@
 %
-These filters resulted in 413 GWAS (Supplementary table 2).
-
+These filters resulted in 418 GWAS (Supplementary table 2).
+%
 I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
 %
-This analysis resulted in 30 261 unique variants and 2,479,065 potential variant colocalizations from 150 GWAS studies (OSF URL).
+This analysis tested 139,069 unique variants and ? potential variant/GWAS/eQTL sample and e-gene colocalizations from 418 GWAS studies (OSF URL).
+%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc); 139,069
+%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc where `PP.H4.abf`>=0.8); 103,551
+%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc);
+%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc where `PP.H4.abf`>=0.8); 2,477,640
+%
+I found 103,551 variants or from ? GWAS that colocalized with at least one eQTL.
+%select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
 %
@@
 
-I manually assigned the 413 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).
+I manually assigned the 418 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).
+
 %

### 2022-11-12 gwas2eqtl_pleiotropy/db.py

@@ -47,3 +47,23 @@ coloc = Table(
    Column('coloc_lead_rsid', String(50), nullable=False),
    Column('coloc_region', String(50), nullable=False),
 )
+
+tophits = Table(
+   'tophits', meta,
+   Column('id', Integer, primary_key = True, autoincrement=True),
+   Column('chrom', Integer, nullable=False),
+   Column('pos', Integer, nullable=False),
+   Column('rsid', String(50), ForeignKey("rsid2cytoband.rsid"), nullable=False),
+   Column('nea', String(50), nullable=False),
+   Column('ea', String(50), nullable=False),
+   Column('pval', Float, nullable=False),
+   Column('beta', Float, nullable=False),
+   Column('se', Float, nullable=False),
+   Column('eaf', Float, nullable=True),
+   Column('n', Float, nullable=True),
+   Column('se', Float, nullable=False),
+   Column('gwas_id', String(50), ForeignKey("gwas_annot.gwas_id"), nullable=False),
+   __table_args__ = (
+           UniqueConstraint('chrom', 'pos', 'rsid', 'gwas_id', name='unique_cnstnt'),
+       )
+)

### 2022-11-13 ms/ms03_results.tex
MS

@@
 To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
-
+%
 127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
@@
 %
-The fourth criterion was to keep only GWAS studies with at least 10000 subjects, 2000 controls and 2000 cases (Supplementary table 2).
+The fourth criterion was to keep only GWAS studies with at least 10,000 subjects, 2,000 controls and 2,000 cases (Supplementary table 2).
 %
-These filters resulted in 418 GWAS (Supplementary table 2).
+These filters resulted in 335 GWAS with 10,627 clumped leading variants and a p-value below 5e-8 (Supplementary table 2).
+% sqlite
+% select count(distinct gwas_id) from tophits; 335
+% select count(*) from (select DISTINCT chrom, pos, rsid, nea, ea from tophits) limit 50; 10627
+
+I found eQTL colocalizations for 3,849 or 36$\%$ of the leading GWAS variants with 2,348 eQTL genes and 127 eQTL samples.
+% sqlite
+% select count(distinct tophits.rsid) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8; 3849
+% alternative: select count(distinct tophits.rsid) from tophits, coloc where coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8; 3849
+% select count(distinct coloc.egene) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8
+% select count(distinct coloc.eqtl_id) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8
+
+The largests classes of diseases by number of leading GWAS variants are autoimmune diseases (2107), cancer of breast (1976) and allergy (1321)
 %
-I developed a colocalization pipeline that explored all the 51 625 combinations of eQTL studies and GWAS (Section Methods).
+and the percentage of variants with a colocalized eQTL is 37\%, 21\% and 42\%, respectively (Supplementary table ST6).
 %
-This analysis tested 139,069 unique variants and ? potential variant/GWAS/eQTL sample and e-gene colocalizations from 418 GWAS studies (OSF URL).
-%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc); 139,069
-%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc where `PP.H4.abf`>=0.8); 103,551
-%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc);
-%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc where `PP.H4.abf`>=0.8); 2,477,640
+%/* count tophits.rsid group by gwas_annot.gwas_class where coloc.`PP.H4.abf`>=0.8*/
+%select gwas_annot.gwas_class, count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8 GROUP by gwas_annot.gwas_class;
+%/* count tophits.rsid group by gwas_annot.gwas_class*/
+%select gwas_annot.gwas_class, count(distinct tophits.rsid) from tophits, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id GROUP by gwas_annot.gwas_class order by gwas_annot.gwas_class;
 %
-I found 103,551 variants or from ? GWAS that colocalized with at least one eQTL.
+For the autoimmune diseases, this is in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
+
+For the autoimmune diseases, ? of the ? leading variants had an annotation with probability PP.H4.abf$\ge$0.8.
+% sqlite
+% select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_class='Autoimmune dis.' and gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8;
+% 778
+% grouped H4 0.8: select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8 GROUP by  gwas_annot.gwas_class;
+%
+This is perc of the
+% grouped all: select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid GROUP by  gwas_annot.gwas_class;%
+%
+%%TODO add PMID to supplementary table with loci_explained_perc
+%For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
+%%
+%Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
+%%
+%Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
+
+This pipeline has resulted in 103,551 variants with or from 246 GWAS that colocalized (PP.H4.abf 0.8) with at least one eQTL.
 %select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
 %
-%TODO add PMID to supplementary table with loci_explained_perc
-For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
+This is a useful annotation ressource that I have made available through a web application (TODO).
 %
-Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
+%This analysis tested 139,069 unique variants in windows of one million nucleotides around the GWAS leading variants and ?
+potential variant/GWAS/eQTL sample and e-gene colocalizations from 418 GWAS studies (OSF URL).
+%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc); 139,069
+%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc); 17,595,868
 %
-Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
+%There are 103,551 variants from 246 GWAS with a least one colocalized eQTL and probability PP.H4.abf$\ge$0.8.
+%select count(DISTINCT gwas_id) from coloc where `PP.H4.abf`>=0.8; 246
+%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc where `PP.H4.abf`>=0.8); 103,551
+%
+%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc where `PP.H4.abf`>=0.8); 2,477,640
 
@@
 
+In the EBI eQTL Catalogue, immune cell types or neural tissues are annotated to great detail while other tissues are annotated at lower resolution.
 %
-In the EBI eQTL Catalogue, immune cell types are annotated to great detail while other tissues are annotated at lower resolution.
-%
-Therefore I manually defined 36 categories to annotate eQTL biological samples at equivalent resolution (Supplementary table 1).
+Therefore I manually defined 36 categories to annotate eQTL samples at equivalent resolution (Supplementary table 1).
 
@@
 %
-Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+These aggregated variants were highly pleiotropic.
+%
+For instance the variant rs2522051, which is located downstream of the IRF1 gene,
+
+re located upstream and downstream of the IRF1 gene, respectively.
+
+%Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+Pleiotropic variants aggregate in some cytobands such as 5q31.1, 19q13.33, 3q23, 9p21.3 and 15q24.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+
+% 5q31.1
 

### 2022-11-14 ms/ms02_intro.tex
MS intro

--- 2022-08-28 ms/ms02_intro.tex
@@
 
-Genome-wide association studies (GWAS) allow to define the impact of genetic variants on a trait in the population.
+% GWAS in general
+Genome-wide association studies (GWAS) have allowed the association of frequent genetic variants with common traits.
 %
-The number of GWAS has rapidly increased since the first GWAS in 2010 and has result in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2018.Parkinson.Buniello}.
+The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007...Nature,2018.Parkinson.Buniello}.
 %
-For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71673 variant-trait associations from 3567 publications \citep{2018.Parkinson.Buniello}.
+For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71,673 variant-trait associations from 3,567 publications \citep{2018.Parkinson.Buniello}.
 
-However the molecular mechanism of GWAS variants is most often unclear because of three reasons \citep{2020.Trynka.CanoGamez}.
+% Shortcomings of GWAS
+However, the interpretation of the GWAS variants is complex because of three reasons \citep{2020.Trynka.CanoGamez}.
 %
-First, identification of the causaul variant is difficult because of linkage disequilibrium.
+First, the causal variant is hidden in regions with high linkage disequilibrium (LD).
 %
-Second, GWAS do not provide hints regarding the tissues involved in the phenotypes.
+Second, GWAS do not inform about the genes, cell types, and tissues involved in the phenotypes.
 %
-Third, GWAS variants affect most often non-coding regions.
+Third, GWAS variants often affect non-coding regions, where the molecular cause is more difficult to define.
 
-Expression QTLs (eQTLs) are variants that change the expression of a given gene in a tissue or cell type.
+% Interest in studying GWAS
+Expression QTLs (eQTLs) are variants that change the expression of a given gene in a cell type or tissue.
 %
-eQTLs share some of the same shortcommings as GWAS variants such as inability to identify the causal variant.
+eQTLs share some of the same shortcomings as GWAS variants such as the inability to identify the causal variant in regions of high LD.
 %
-Also the distribution of eQTLs and GWAS variants in the genome have been suggested to be different because of negative selection.
+Moreover, the distribution of eQTLs and GWAS variants in the genome are different because of negative selection \citep{2022.Pritchard.Mostafavi}.
+%
+Nevertheless, eQTLs are still important data sets to interpret GWAS variants, because they provide gene regulatory information about the eQTL genes and cell types and tissues \citep{2021.Li.Mu}.
+%
+%In addition, colocalization analysis can help prioritize causal variants.
 
-Nevertheless eQTLs are still important data sets to interprete GWAS variants, because they provide gene regulatory information about the eQTL genes (eGenes) and tissues (etissues) (REF).
+Many GWAS have allowed observing that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
 %
-In addition colocalization anlaysis can pinpoint more precisely the causal variants that either data set independently (REF).
+In that work, the analyzed loci covered 60\% of the genome, and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
+%
+They also found that pleiotropic genes and variants were less tissue-specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
+%
+However, the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu,2018.Vijayanand.Schmiedel,2017gtex.nature}.
+%
+In addition, the mechanism of pleiotropy is not well understood yet.
 
-The large number of GWAS has allowed to discover that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
+In this work, I have taken advantage of two large eQTL and GWAS data sets from the EBI eQTL Catalogue and the IEU OpenGWAS, respectively,
+to develop a global colocalization analysis based on 127 eQTL studies and 418 GWAS across many cell types, tissues, and phenotypes.
 %
-In this work, the analysed loci covered 60\% of the genome and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
+I have used these eQTL/GWAS variants to investigate mechanisms of pleiotropic gene regulatory variants.
 %
-They also found that pleiotropic genes and variants where less tissue specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
+These eQTL/GWAS variants might be a useful resource for the annotation and prioritization of frequent variants.
 %
-The question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu.GenomeBiology.impactcelltype,2018.Vijayanand.Schmiedel.Cell.ImpactGeneticPolymorphisms,2017...Nature}.
-
-In this work, I have taken advantage of two well formatted eQTL and GWAS data sets to develop a global colocalization analysis that includes 127 eQTL studies and 413 GWAS.
-%
-I have used these colocalization results to investigate molecular differences between pleiotropic variants in terms of gene regulatory properties.
-%
-We provide a model to mechanistically understand our results.
-%
-This global colocalization analysis is also useful to help interpret GWAS variants.
+Therefore, I have developed a web application to query the variants.

### 2022-11-16 ms/ms03_results.tex
ENH

@@
 %
-These filters resulted in 335 GWAS with 10,627 clumped leading variants and a p-value below 5e-8 (Supplementary table 2).
+These filters resulted in 418 GWAS.
+%
+Among these studies, there 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary table 2).
 % sqlite
@@
 
-The largests classes of diseases by number of leading GWAS variants are autoimmune diseases (2107), cancer of breast (1976) and allergy (1321)
+The largest classes of diseases by number of leading GWAS variants are autoimmune diseases (2107), cancer of breast (1976) and allergy (1321)
 %
-and the percentage of variants with a colocalized eQTL is 37\%, 21\% and 42\%, respectively (Supplementary table ST6).
+and the percentage of variants with a colocalized eQTL is 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST6).
 %

### 2022-11-16 scripts/cmpt_supplementary_files.py

@@ -1,17 +1,17 @@
-import pathlib
-
 from gwas2eqtl_pleiotropy.PathManager import PathManager
 
 import os
 import pandas
+import pathlib
 import sys
 
 #%%
 help_cmd_str = "todo"
 try:
-    wdir_path = sys.argv[1]
-    supp_tabl_xlsx_path = sys.argv[2]
-    if len(sys.argv) > 3:
+    gwas_class_ods_path = sys.argv[1]
+    wdir_path = sys.argv[2]
+    supp_tabl_xlsx_path = sys.argv[3]
+    if len(sys.argv) > 4:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -37,13 +37,13 @@ description_lst = ['Classification of eQTL tissues and cell types. The first six
 sheet_name_lst.append("ST2")
 description_lst.append("Metadata and classification of GWAS")
 #
-sheet_name_lst.append("ST4")
+sheet_name_lst.append("ST3")
 description_lst.append("Count and list of GWAS phenotypes, egenes and etissues for each eQTL/GWAS variant")
 #
-sheet_name_lst.append("ST5")
+sheet_name_lst.append("ST4")
 description_lst.append("Count and list of GWAS phenotypes for each pleiotropic region")
 #
-sheet_name_lst.append("ST6")
+sheet_name_lst.append("ST5")
 description_lst.append("Percentage of explained GWAS loci")
 
 capt_df = pandas.DataFrame({'Supp. Tab.': sheet_name_lst, 'Description': description_lst})
@@ -55,18 +55,20 @@ etissue_class_ods_path = os.path.join(PathManager.get_project_path(), "config", 
 st_df = pandas.read_excel(etissue_class_ods_path, index_col=None, header=0)
 st_df.drop(['Unnamed: 8', 'etissue_class.1', 'count'], axis=1, inplace=True)
 st_df.to_excel(st_writer, sheet_name=sheet_name, index=False, header=True)
-sheet_counter += 1
 
 #%% ST2
 sheet_name = 'ST2'
-tsv_path = os.path.join(wdir_path, "annot_gwas_metadata.py/gwas413_metadata.tsv")
-st_df = pandas.read_csv(tsv_path, sep="\t", header=0)
[…]

### 2022-11-16 ms/ms03_results.tex
MS

@@
 %
-Among these studies, there 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary table 2).
+Among these studies, there were 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary table 2).
 % sqlite
@@
 %
-and the percentage of variants with a colocalized eQTL is 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST6).
+and the percentages of variants with a colocalized eQTL are 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST6).
 %
@@
 %
-For the autoimmune diseases, this is in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
+These results agree with previous work for autoimmune diseases \citep{2021.Li.Mu}.
 
-For the autoimmune diseases, ? of the ? leading variants had an annotation with probability PP.H4.abf$\ge$0.8.
-% sqlite
-% select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_class='Autoimmune dis.' and gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8;
-% 778
-% grouped H4 0.8: select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8 GROUP by  gwas_annot.gwas_class;
+This pipeline has resulted in 103,551 variants from 246 GWAS that colocalized (PP.H4.abf$\geq$0.8) with at least one eQTL.
+%select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
+%select count(*) from (select DISTINCT rsid FROM coloc where `PP.H4.abf`>=0.8);
 %
-This is perc of the
-% grouped all: select count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid GROUP by  gwas_annot.gwas_class;%
-%
-%%TODO add PMID to supplementary table with loci_explained_perc
-%For GWAS studies with more than 10 loci, the explained of loci variate from 0\% to 77\% with an average of 33\% (Supplementary table ST6).
-%%
-%Autoimmune diseases such as inflammatory bowel disease, Crohn's disease and Ulcerative colitis showed a colocalization average of 45.4\% in agreement with previous work \citep{2021.Li.Mu.GenomeBiology.impactcelltype}.
-%%
-%Selection of colocalizations with PP.H4.abf greater than 0.8 resulted in 9 758 variants and 143 119 colocalizations (Supplementary table 3).
-
-This pipeline has resulted in 103,551 variants with or from 246 GWAS that colocalized (PP.H4.abf 0.8) with at least one eQTL.
-%select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
-%
-This is a useful annotation ressource that I have made available through a web application (TODO).
-%
-%This analysis tested 139,069 unique variants in windows of one million nucleotides around the GWAS leading variants and ?
-potential variant/GWAS/eQTL sample and e-gene colocalizations from 418 GWAS studies (OSF URL).
-%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc); 139,069
-%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc); 17,595,868
-%
-%There are 103,551 variants from 246 GWAS with a least one colocalized eQTL and probability PP.H4.abf$\ge$0.8.
-%select count(DISTINCT gwas_id) from coloc where `PP.H4.abf`>=0.8; 246
-%select count(*) from (select DISTINCT chrom,pos,rsid FROM coloc where `PP.H4.abf`>=0.8); 103,551
-%
-%select count(*) from (select DISTINCT chrom,pos,rsid,gwas_id,eqtl_id,egene FROM coloc where `PP.H4.abf`>=0.8); 2,477,640
+I have developed a web application where I expose 116,736 variants from 246 GWAS that colocalize with at least one eQTL at a lower threshold value PP.H4.abf$\geq$0.5.
+% select count(*) from (select DISTINCT rsid FROM coloc where `PP.H4.abf`>=0.5);
+% select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.5);
 
@@
 
-I manually assigned the 418 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 2).
+I manually assigned the 418 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 1).
+%
+Then I verified whether GWAS traits cluster coherently within these classes.
+%
+I computed distances between GWAS traits based on the Spearman correlation of the colocalized eQTL beta of the different eQTL genes and samples.
+%
+I plot only classes with more than a minimum number of GWAS traits (Fig. \ref{fig:gwas_distance}).
+%
+GWAS traits cluster together in autoimmune diseases and circulatory system diseases.
+%
+Due to the large number of breast cancer GWAS, we have created a separeted class from cancer.
+Nevertheless a weaker specific signal is observed between the breast cancer class and other types of cancers.
 
@@
 %
-Therefore I manually defined 36 categories to annotate eQTL samples at equivalent resolution (Supplementary table 1).
+Therefore I manually defined 36 categories to annotate eQTL samples at equivalent resolution (Supplementary table 2).
 
@@
 %
-For instance the variant rs2522051, which is located downstream of the IRF1 gene,
-
-re located upstream and downstream of the IRF1 gene, respectively.
-
-%Pleiotropic variants aggregate in some cytobands such as 3q23, 5q31.1, 11q13.5, 12q24.12 and 15q26.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-Pleiotropic variants aggregate in some cytobands such as 5q31.1, 19q13.33, 3q23, 9p21.3 and 15q24.1 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
-
-% 5q31.1
-
-% 12q24.12
-For instance, the most pleiotropic variants in cytoband 12q24.12 are involved in allergies, cancer, cardiovascular and autoimmune diseases (Table \ref{tab:pleitropic_variants}).
+For instance the variant rs2522051 is located upstream of the IRF1 gene in cytoband 5q31.1 and it is involved in six classes of traits, namely,
+allergy, autoimmune diseases, cancer of breast, circulatory system diseases, hypertension and respiratory system diseases
+(Table \ref{tab:pleitropic_variants} and Supplementary table 3).
 %
-These variants control eGenes ALDH2, LINC01405, MAPKAPK5, SH2B3 that are involved in alcohol-related disorders (ALDH2), cancer (LINC01405), multiple congenital anomalies syndromes (MAPKAPK5),  inflammation and hematological disorders (SH2B3).
+This variant rs2522051 is active in 23 classes of cell type and tissues including adipose, brain, digestive cells, immune cells and sexual organs.
 %
-These eQTLs are active in adipose tissue, arteries, blood, colon, immune cells, skin and induced pluripotent stem cells (iPSC).
+This variant rs2522051 regulates several genes such as IL13, PDLIM4 and RAD50.
 
@@
 
-Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
-%
-I analyzed the ontology of eGenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
-%
-I found that eGenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-For instance, eGenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
-%
-IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
-%
-Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
-%
-These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
+%Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
+%%
+%I analyzed the ontology of eGenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
+%%
+%I found that eGenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
+%For instance, eGenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
+%%
+%IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
+%%
+%Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
+%%
+%These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
 
@@
 
-Pleiotropic variants are aggregated in few cytobands.
+Pleiotropic variants are aggregated in genomic regions.
 %
@@
 %
-I found 453 regions with 441 regions with 2 or more categories, 114 regions with 3 or more categories, 37 regions with 4 or more categories and 18 regions with 5 or more categories (Supplementary table 5).
+I found 13 regions with 5 or more GWAS classes, 30 regions with 4 or more GWAS classes and 80 regions with 3 or more GWAS.
 %

### 2022-11-17 ms/ms03_results.tex
MS

@@
 
-I found eQTL colocalizations for 3,849 or 36$\%$ of the leading GWAS variants with 2,348 eQTL genes and 127 eQTL samples.
+My pipeline uncovered eQTL colocalizations for 3,849 or 36$\%$ of the leading GWAS variants with 2,348 eQTL genes and 127 eQTL samples.
 % sqlite
@@
 
-This pipeline has resulted in 103,551 variants from 246 GWAS that colocalized (PP.H4.abf$\geq$0.8) with at least one eQTL.
+This pipeline resulted in 103,551 variants from 246 GWAS that colocalized (PP.H4.abf$\geq$0.8) with at least one eQTL.
 %select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
@@
 %
-I have developed a web application where I expose 116,736 variants from 246 GWAS that colocalize with at least one eQTL at a lower threshold value PP.H4.abf$\geq$0.5.
+I have also developed a web application where I expose 116,736 variants from 246 GWAS that colocalize with at least one eQTL at a lower threshold value PP.H4.abf$\geq$0.5.
 % select count(*) from (select DISTINCT rsid FROM coloc where `PP.H4.abf`>=0.5);
@@
 
-These category annotations were used to investigate the variant pleiotropy for GWAS traits.
+These class annotations were used to investigate the variant pleiotropy.
 %
@@
 %
-These aggregated variants were highly pleiotropic.
+The aggregated variants are highly pleiotropic.
 %
-For instance the variant rs2522051 is located upstream of the IRF1 gene in cytoband 5q31.1 and it is involved in six classes of traits, namely,
-allergy, autoimmune diseases, cancer of breast, circulatory system diseases, hypertension and respiratory system diseases
+For instance the variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1 and it is involved in six classes of traits, namely,
+allergy, autoimmune diseases, cancer of breast, circulatory system diseases, hypertension, and respiratory system diseases
 (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
 %
-This variant rs2522051 is active in 23 classes of cell type and tissues including adipose, brain, digestive cells, immune cells and sexual organs.
+This variant rs2522051 is active in 23 classes of cell types and tissues including adipose, brain, digestive cells, immune cells, and sexual organs.
 %
@@
 
-\subsubsection*{What is the function of pleiotropic variants?}
-
-%Then I evaluated the function of pleiotropic variants with 5, 4, 3 and 2 GWAS categories using gene ontology analysis.
-%%
-%I analyzed the ontology of eGenes with 5, 4, 3 and 2 GWAS categories using the DAVID web services \citep{2008.Lempicki.Huang,2008.Lempicki.Huang.NucleicAcidsResearch}.
-%%
-%I found that eGenes of pleiotropic variants with 3 and 2 GWAS categories were significantly enriched in functions of the immune system (Fig. \ref{s:results}{fig:geneontology}).
-%For instance, eGenes in the cytoband 5q31.1 include IRF1 and IL4, which are important factors of the immune system.
-%%
-%IRF1 is a response protein to the presence of virus and oncogenic proteins and IL4 is required to stimulate proliferation of activated B and T-cells.
-%%
-%Variant rs17622656 is located in the intron of the IRF1 gene whereas variants rs736801 and rs2522051 are located upstream and downstream of the IRF1 gene, respectively.
-%%
-%These variants are associated with allergy, asthma, cardiovascular diseases, hypertension and ulcerative colitis and are active in a large number of tissues such as arteries, blood, brain, breast, digestive and immune system, muscle, skin, ovary, testis and thyroid.
-
-%Variants of the 3q23 cytoband are involved in allergy, cancer and height traits (Supplementary table 3).
-%%
-%Interestingly all these variants target the eGene ZBTB3.
-%%
-%ZBTB3 is a transcription factor that controls pro-inflammatory factors such as IRF5 (10.1016/j.jaut.2016.08.001).
-%%
-%These variants are active in tissues such as blood, esophagus, immune cell, LCLs and muscle (Supplementary table 3).
-
-%%11q13.5
-%The variants in cytoband 11q13.5 are involved in allergy and autoimmune disease (Supplementary table 3).
-%%
-%All these variants target the eGene EMSY-DT, which is a long noncoding RNA upstream of EMSY.
-%%
-%These variants are active in LCL, Skin and iPSC (Supplementary table 3).
-%
-%In summary, I have identified pleiotropic variants with gene regulatory function.
-%%
-%Some of these loci are strongly involved in the immune system.
-
 \subsection*{Pleiotropic regions}
 
-Pleiotropic variants are aggregated in genomic regions.
+Pleiotropic variants are aggregated in genomic regions, for instance, in cytobands 3q23, 5q31.1, 9p21.3, 15q24.1, and 19q13.33 for the variants involved on five or more classes (Supplementary table 3).
 %
-Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions}).
+Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
 %
-I found 13 regions with 5 or more GWAS classes, 30 regions with 4 or more GWAS classes and 80 regions with 3 or more GWAS.
+I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait classes and 80 regions with 3 or more trait classes (Supplementary table 4).
 %
-75\% of regions are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
+% TODO verify this plot, is plot 2b used somewhere?
+75\% of regions involved in two or more trait classes are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig.
+\ref{fig:pleiotropy_region_distribution}a).
 
-The most pleiotropic region is 12:111,395,984-111,645,358 in cytoband 12q24.12 that we discussed above (Table \ref{tab:pleiotropic_regions}).
+% TODO continue here
+The most pleiotropic region is 5:131,912,097-132,802,472 in cytoband 5q31.1 with a high concentration of cytokines genes (Table \ref{tab:pleiotropic_regions}).
 %

### 2022-11-17 ms/ms01_titlepage.tex
MS

@@
 
-\title{Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors}
+\title{Pleiotropy of expression quantitative trait loci increases with the number of regulated genes, active cell types and tissues and bound transcription factors}
 %\shorttitle{Pleiotropy and gene regulatory variants}

### 2022-11-17 ms/ms03_results.tex
MS

@@
 
-\subsection*{Pleiotropic variants colocalized with eQTLs}
+%\subsection*{Pleiotropic variants colocalized with eQTLs}
 
-\subsubsection*{Which variants are the most pleiotropic?}
+\subsection*{Which eQTLs are the most pleiotropic?}
 
-I manually assigned the 418 GWAS to 96 categories to aggregate identical or similar traits (Supplementary table 1).
+I manually assigned the 418 GWAS to 96 classes to aggregate identical or similar traits (Supplementary table 1).
 %
@@
 
-In the EBI eQTL Catalogue, immune cell types or neural tissues are annotated to great detail while other tissues are annotated at lower resolution.
+In the EBI eQTL Catalogue, some samples have been analysed at the cell type level while other samples are analyzed at the tissue level.
 %
-Therefore I manually defined 36 categories to annotate eQTL samples at equivalent resolution (Supplementary table 2).
+For instance, there is much detail resolution regarding the immune cell types.
+%
+Therefore I manually defined 36 classes to aggregate eQTL samples to equivalent resolution (Supplementary table 2).
+%
+For instance the different immune cell types such as monocyte, T-cell, etc are aggregated to an ImmuneCell class (Supplementary table 2).
 
@@
 %
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS categories they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+Colocalized eQTL/GWAS variants were classified according to the number of GWAS classes they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
 %
@@
 
-\subsection*{Pleiotropic regions}
+\subsection*{Which genomic regions are pleiotropic regarding eQTLs?}
 
-Pleiotropic variants are aggregated in genomic regions, for instance, in cytobands 3q23, 5q31.1, 9p21.3, 15q24.1, and 19q13.33 for the variants involved on five or more classes (Supplementary table 3).
+Pleiotropic eQTLs are aggregated in genomic regions, for instance, in cytobands 3q23, 5q31.1, 9p21.3, 15q24.1, and 19q13.33 for the eQTLs involved on five or more classes (Supplementary table 3).
 %
-Therefore I computed regions that include a large number of pleiotropic variants (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
+Therefore I computed regions that include a large number of pleiotropic eQTLs (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
 %
@@
 %
-% TODO verify this plot, is plot 2b used somewhere?
 75\% of regions involved in two or more trait classes are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig.
@@
 
-% TODO continue here
-The most pleiotropic region is 5:131,912,097-132,802,472 in cytoband 5q31.1 with a high concentration of cytokines genes (Table \ref{tab:pleiotropic_regions}).
+The most pleiotropic region is 5:131,912,097-132,802,472 in cytoband 5q31.1 which contains the interferon response factor 1 (IRF1) and interleukins IL3, IL4, IL5 and IL13 (Table \ref{tab:pleiotropic_regions}).
 %
-The largest region is 11:13,260,511-17,396,930 in cytoband 11p15.2 with a length of 4,136,419 bp and variants involved in cardiovascular and hypertension traits (Supplementary table 5).
+Interferons and interleukins are very important factors for the immune system and anti-viral, inflammation and cancer responses.
 %
-The second largest region is 2:187,235,912-191,066,738 in cytoband 2q32.2 with a length of 3,830,826 bp and variants involved in cardiovascular, hypertension and autoimmune traits  (Supplementary table 5).
+The largest region is 7:2,712,518-7,254,268 in cytoband 7p22.3 with a length of 4,541,751 bp and variants associated to autoimmune and respiratory diseases and height (Supplementary table 5).
 
-Very pleiotropic regions with 6 GWAS categories or more make less than 1Mb of the genome, while moderate pleiotropic regions with 3 GWAS categories or more make 2 Mb of the genome (Fig.
+Very pleiotropic regions with 5 trait classes or more make less than 1 Mb of the genome, while moderate pleiotropic regions with 3 trait classes or more make round 3 Mb of the genome (Fig.
 \ref{fig:pleiotropy_region_distribution}b.
 
-\subsection*{How specific are variants to GWAS traits, eGenes and eTissues?}
+\subsection*{How specific to traits, genes and cell type/tissues? are colocalized eQTLs/GWAS variants}
 
-Next I explored the trait, eGene and eTissue specificity of eQTL/GWAS variants.
+Next I explored the trait, gene and cell type/tissue specificity of colocalized eQTL/GWAS variants.
 %
-83\% of variants are involved in one GWAS category, 14\% are involved in 2 categories and the remaining variants with 3 or more trait categories make 3\% or less (Fig.
+74\% of eQTLs are involved in one GWAS class, 21\% are involved in 2 classes and the remaining variants with 3 or more trait classes make less that 4\% of eQTLs (Fig.
 \ref{fig:hist_gwas_egene_etissue}(a).
 %
-Regarding eGenes, half of variants modulate one eGene, and the other half modulate two or more eGenes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
+Regarding target genes, 36\% of the colocalized eQTLs modulate a single gene, 24\% modulate two genes, and the remaining 40\% modulate three or more genes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
 %
-In the case of eTissues, only 40\% of variants are specific to a single tissue while the other 60\% are active in two or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}(c).
+Concerning cell types and tissues classes, 27\% are active in a single tissue, 16\% in two tissues, 11\% in three tissues, and the remaining 46\% in four or more tissues (Fig.
+\ref{fig:hist_gwas_egene_etissue}(c).
 %
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait. By contrast, half of them only module one specific eGene, and less than half are active in a single tissue.
+In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait.
+%
+By contrast, a minority of colocalized eQTL/GWAS variants are specific to one single tissue.
 
@@
 
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS categories, eGenes and eTissues for regulatory variants in three pleiotropic regions.
+In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS classes, eGenes and eTissues for regulatory variants in three pleiotropic regions.
 %
-All three regions are very pleiotropic with very different associated trait categories such as cancer, cardiovascular and autoimmune diseases.
+All three regions are very pleiotropic with very different associated trait classes such as cancer, cardiovascular and autoimmune diseases.
 %
@@
 
-\subsubsection*{More severe variant effect consequences?}
+\subsection*{Which are the effect consequence of pleiotropic eQTLs?}
 
@@
 %
-I separated variants according to the count of GWAS categories and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
+I separated variants according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
 %
-I found a significant larger number of missense variants among variants with 2, 3 and 4 categories compared to variants with 1 category (Fig. \ref{fig:vep_consequence}a).
+I found a significant larger number of missense variants among variants with 2, 3 and 4 classes compared to variants with 1 class (Fig. \ref{fig:vep_consequence}a).
 %
-I also found a significant larger number of splicing variants among variants with 3 categories (Fig. \ref{fig:vep_consequence}b).
+I also found a significant larger number of splicing variants among variants with 3 classes (Fig. \ref{fig:vep_consequence}b).
 %
-Finally, there is also a larger of the 3'-UTR variants among variants with 4 categories (Fig. \ref{fig:vep_consequence}c).
+Finally, there is also a larger of the 3'-UTR variants among variants with 4 classes (Fig. \ref{fig:vep_consequence}c).
 %
@@
 
-\subsubsection*{More eGenes per variant-eTissue?}
+\subsection*{Do pleiotropic eQTLs modulate more genes?}
 
@@
 %
-Then these eGene counts were classified according to the GWAS category count.
+Then these eGene counts were classified according to the GWAS class count.
 %
-If the GWAS category count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal one.
+If the GWAS class count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal one.
 %
-I found means of eGene counts of 1.4, 1.7, 1.7 and 1.6 for GWAS categories counts 5, 4, 3 and 2 compared to eGene count mean of 1.5 for GWAS category count 1 (Fig.
-\ref{fig:gwas_egene_etisue_per_variant}a).
+I found means of eGene counts of 1.4, 1.7, 1.7 and 1.6 for GWAS classes counts 5, 4, 3 and 2 compared to eGene count mean of 1.5 for GWAS class count 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
 %
-This suggests that pleiotropic variant-eTissue pairs have slightly but significantly (Mann–Whitney U test) more eGenes compared to variants with one GWAS category (Fig.
+This suggests that pleiotropic variant-eTissue pairs have slightly but significantly (Mann–Whitney U test) more eGenes compared to variants with one GWAS class (Fig.
 \ref{fig:gwas_egene_etisue_per_variant}a).
@@
 
-\subsubsection*{More eTissues per variant-eGene?}
+\subsection*{Are pleiotropic eQTLs active in more tissues?}
 
-My next hypothesis was that pleiotropic variant-eGene pairs are active in more eTissues compared with variant-eGene pairs with one GWAS category and this increases their probability to affect more GWAS categories.
+My next hypothesis was that pleiotropic variant-eGene pairs are active in more eTissues compared with variant-eGene pairs with one GWAS class and this increases their probability to affect more GWAS classes.
 %
@@
 %
-Then these eTissue counts were classified according to the GWAS category count.
+Then these eTissue counts were classified according to the GWAS class count.
 %
-If the GWAS category count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS category count.
+If the GWAS class count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS class count.
 %
-I found means of eTissue counts of 2.2, 3, 3 and 2.6 for GWAS categories counts 5, 4, 3 and 2 compared to eTissue count mean of 2.5 for GWAS category count 1 (Fig.
-\ref{fig:gwas_egene_etisue_per_variant}b).
+I found means of eTissue counts of 2.2, 3, 3 and 2.6 for GWAS classes counts 5, 4, 3 and 2 compared to eTissue count mean of 2.5 for GWAS class count 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
 %
-This suggests that pleiotropic variant-eGene pairs are active slightly but significantly (Mann–Whitney U test) in more eTissues compared to variants with one GWAS category (Fig.
+This suggests that pleiotropic variant-eGene pairs are active slightly but significantly (Mann–Whitney U test) in more eTissues compared to variants with one GWAS class (Fig.
 \ref{fig:gwas_egene_etisue_per_variant}b).
 
-\subsubsection*{More GWAS per variant-eGene-eTissue?}
+\subsection*{Are pleiotropic eQTLs associated to more traits in unique gene-tissue pairs?}
 
-My next hypothesis was that pleiotropic variants are associated to more GWAS categories even after taking into account differences in eGene and eTissue counts.
+My next hypothesis was that pleiotropic variants are associated to more GWAS classes even after taking into account differences in eGene and eTissue counts.
 
-To test this hypothesis, the GWAS category counts per variant-eGene-eTissue trios were counted.
+To test this hypothesis, the GWAS class counts per variant-eGene-eTissue trios were counted.
 %
-Then these GWAS categorie counts per variant-eGene-eTissue trios were classified according to the GWAS category count per variant.
+Then these GWAS categorie counts per variant-eGene-eTissue trios were classified according to the GWAS class count per variant.
 %
-If the GWAS category count per variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS category count per variant.
+If the GWAS class count per variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS class count per variant.
 %
-I found means of GWAS category counts per variant-eGene-eTissue of 3, 2, 1.6, 1.4 and 1 for GWAS categories counts per variant of 5, 4, 3, 2 and 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+I found means of GWAS class counts per variant-eGene-eTissue of 3, 2, 1.6, 1.4 and 1 for GWAS classes counts per variant of 5, 4, 3, 2 and 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
 %
-This shows a significant larger number of GWAS categories even in unique trios of variant-eGene-eTissue (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+This shows a significant larger number of GWAS classes even in unique trios of variant-eGene-eTissue (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
 %
@@
 
-\subsubsection*{How to explain that pleiotropic variants are active in more tissues?}
+\subsection*{Are pleiotropic eQTLs bound by more transcription factors?}
 
@@
 %
-I found that the odds ratio of variants annotated with CRMs vs non-annotated is significantly (Fisher's exact test) higher for variants with 2 and 3 GWAS category counts compared to category count 1 (Fig.
+I found that the odds ratio of variants annotated with CRMs vs non-annotated is significantly (Fisher's exact test) higher for variants with 2 and 3 GWAS class counts compared to class count 1 (Fig.
 \ref{fig:freq_tf_per_variant}b).
 
-\subsection*{How do effect size and significance relate to pleiotropy}
+\subsection*{Do pleiotropic eQTLs have stronger effect size on the gene and trait?}
 
@@
 %
-I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS category counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
+I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
 %
-Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS category counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 
@@
 %
-I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS category counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
+I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
 %
-Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS category counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 

### 2022-11-17 ms/ms05_floats.tex
MS

@@
 %
-\caption{Commorbidity matrix.}
+\caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different gene and samples.}
+\label{fig:gwas_distance}
 %
@@
 %
-% Tab 1: Winner variants
+% Tab 1: Winner eQTLs
 %
@@
 %
-\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
 \end{table}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 1: GO of pleiotropic variants
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{figure}[!tbp]
-%\centering
-%%
-%\begin{subfigure}[]{0.99\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_3.png}
-%\end{subfigure}
-%
-%\begin{subfigure}[]{.99\textwidth}
-%\textbf{b}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_davidgo.py/david_pleio_2.png}
-%\end{subfigure}
-%%
-%\caption{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.} \label{fig:geneontology}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%
-%%% Fig 2: Fisher stat of tissues in variants as a function of gwas category count
-%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{figure}[!tbp]
-%\centering
-%%
-%\begin{subfigure}[]{.66\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_tissue_enrich_in_pleio.py/plt.png}
-%\end{subfigure}
-%%
-%\caption{Tissue category odds ratio in pleitropic variants compared with variant with GWAS category count 1.} \label{fig:tissues}
-%\end{figure}
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \vspace{15pt}
-\caption{Pleiotropic regions involving 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
+\caption{Pleiotropic regions involving 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
 \end{table}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \textbf{a}, Length distribution of pleiotropic regions.
-\textbf{b}, Cumulative sum ordered inversely with the number of GWAS categories.} \label{fig:pleiotropy_region_distribution}
+\textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
 \end{figure}
@@
 %%
-%% Fig 1: Histograms variants vs GWAS, eGene and eTissues
+%% Fig 3: Histograms variants vs GWAS, eGene and eTissues
 %%
@@
 %
-\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS categories (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
+\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
 \end{figure}
@@
 %%
-%% Fig 2: Scatter plots of winner regions
+%% Fig 4: Scatter plots of winner regions
 %%
@@
 
-%\begin{subfigure}[]{.19\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-%\end{subfigure}
-%
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{b}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{c}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{d}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{e}
@@
 
-%\begin{subfigure}[]{.19\textwidth}
-%\textbf{f}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-%\end{subfigure}
-%
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{g}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{h}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{i}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{j}
@@
 
-%\begin{subfigure}[]{.19\textwidth}
-%\textbf{k}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr2_start43222246_end43961707_categories7.png}
-%\end{subfigure}
-%
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{l}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{m}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{n}
@@
 %
-\begin{subfigure}[]{.19\textwidth}
+\begin{subfigure}[]{.24\textwidth}
 \textbf{o}
@@
 
-\caption{Count of GWAS categories (\textbf{a-c}), eGenes (\textbf{d-f}) and eTissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
-Genomic coordinates are given for the hg38 assembly.} \label{fig:region_gwas_egenes_tissues}
+\caption{Count of GWAS classes (\textbf{a-c}), eGenes (\textbf{d-f}) and eTissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
+Genomic coordinates are given for the hg38 assembly.}
+\label{fig:region_gwas_egenes_tissues}
 %
@@
 %%
-%% Fig 3: VEP consequences
+%% Fig 5: VEP consequences
 %%
@@
 %
-\caption{Odds ratio compared of the variants annotated with given variant effect predictor (VEP) consequence and given GWAS category count compared with non pleiotropic variants (GWAS category count one).
+\caption{Odds ratio compared of the variants annotated with given variant effect predictor (VEP) consequence and given GWAS class count compared with non pleiotropic variants (GWAS class count one).
 The significance is given for the q-value (FDR 5\%).}
+\label{fig:vep_consequence}
 %
@@
 %%
-%% Fig 4: Violin plots. eGenes, eTissues and gwas per variants
+%% Fig 6: Violin plots. eGenes, eTissues and gwas per variants
 %%
@@
 %
-\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS categories per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
+\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS classes per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
 %
 \end{figure}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%
-%% Fig 6: TF count per GWAS category count
+%% Fig 7: TF count per GWAS class count
 %%
@@
 \caption{Analysis of transcription factor binding.
-\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS category count.
+\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS class count.
 \textbf{b}, Odds ratio of cis regulatory module-annotated vs non-annotated variants.} \label{fig:freq_tf_per_variant}
@@
 \end{figure}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%
-%% Fig 7: beta and log pval
+%% Fig 8: beta and log pval
 %%
@@
 
-\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS category count.} \label{fig:beta_pval}
+\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
 %
@@
 %%
-%% Fig 7: graphical conclusions
+%% Fig 9: graphical conclusions
 %%

### 2022-11-19 ms/ms02_intro.tex
NEH
MS

@@
 %
-The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007...Nature,2018.Parkinson.Buniello}.
+The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007burton.barret,2018.Parkinson.Buniello}.
 %
@@
 %
-Therefore, I have developed a web application to query the variants.
+Therefore, I have developed an easy-to-use web application to expose these colocalization data.

### 2022-11-19 ms/ms03_results.tex
MS

@@
 
-\subsection*{Which genomic regions are pleiotropic regarding eQTLs?}
+\subsection*{Which genomic regions contain the most pleiotropic eQTLs?}
 
-Pleiotropic eQTLs are aggregated in genomic regions, for instance, in cytobands 3q23, 5q31.1, 9p21.3, 15q24.1, and 19q13.33 for the eQTLs involved on five or more classes (Supplementary table 3).
+Pleiotropic eQTLs are concentrated in genomic regions such as in cytobands 3q23,
+5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Supplementary table 3).
 %
-Therefore I computed regions that include a large number of pleiotropic eQTLs (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
+Therefore I computed regions that include at least one pleiotropic eQTL in
+sliding windows of 100,000 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
 %
-I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait classes and 80 regions with 3 or more trait classes (Supplementary table 4).
+I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait
+classes and 80 regions with 3 or more trait classes (Supplementary table 4).
 %
-75\% of regions involved in two or more trait classes are shorter than 100 kb, 15\% of regions are between 100 kb and 200 kb, and less than 5\% are larger than 200 kb (Fig.
-\ref{fig:pleiotropy_region_distribution}a).
+50\% of regions with at least two trait classes are shorter than 100 kb, 20\% of regions are
+between 100 kb and 200 kb, 10\% between 200 kb and 300 kb, and 20\% are larger than 300 kb (Fig. \ref{fig:pleiotropy_region_distribution}a).
 
@@
 
-Very pleiotropic regions with 5 trait classes or more make less than 1 Mb of the genome, while moderate pleiotropic regions with 3 trait classes or more make round 3 Mb of the genome (Fig.
-\ref{fig:pleiotropy_region_distribution}b.
+Pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
+Very pleiotropic regions with 5 trait classes or more make around 0.6 Mb of the genome,
+regions with 4 GWAS classes make an additional 1 Mb and regions with 3 GWAS classes make an additional 1.5 Mb (Fig. \ref{fig:pleiotropy_region_distribution}b).
 
-\subsection*{How specific to traits, genes and cell type/tissues? are colocalized eQTLs/GWAS variants}
+\subsection*{How specific are pleiotropic eQTLs to traits, genes and tissues?}
 

### 2022-11-19 scripts/cmpt_pleiotropic_regions.py

@@ -1,3 +1,5 @@
+import numpy
+
 from gwas2eqtl_pleiotropy.constants import region_bin, label_fontsize, tick_fontsize
 from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 from matplotlib import pyplot as plt
@@ -146,24 +148,38 @@ for count_pleio in range(1, 6):
     region_pleio_i_df = regions_pleio_df.loc[regions_pleio_df['gwas_class_count'] == count_pleio,]
     region_pleio_i_df.to_csv(region_pleio_i_bed_path, sep="\t", index=False, header=False)
 
+#############################################
+# Distribution of region by length
+# % region <= 100kb, % >100kb and <=200kb, ...
+region_length_ser = (regions_pleio_df['end'] - regions_pleio_df['start'])
+data_ser = region_length_ser / 100000
 
+ax = plt.hist(data_ser, bins=range(50), density=True)
+plt.xlim(0, 10)
+plt.tight_layout()
+plt.savefig(png_path)
+plt.close()
 
 #%##############
 # histogram
 plt.rcParams["figure.figsize"] = (8, 6)
-region_lenght_ser = (regions_pleio_df['end'] - regions_pleio_df['start'])
+region_length_ser = (regions_pleio_df['end'] - regions_pleio_df['start'])
 
-ylabel = "Proportion"
-title = "Lengths of pleiotropic regions"
-ylim=[1e-10, 1]
+ylabel = "Cumulative proportion"
+title = "Length of pleiotropic regions"
+ylim=[0, 1]
 edgecolor='k'
 linewidth = 2
 
 hist_kwargs = {'density': False, 'edgecolor': edgecolor, 'linewidth': linewidth}
 
 #%%
-data_ser = region_lenght_ser/100000
-seaborn.histplot(data_ser, stat='percent', discrete=True)
+data_ser = region_length_ser / 100000
+
+# import pdb; pdb.set_trace()
+# seaborn.histplot(data_ser, stat='percent', discrete=True)
+ax = plt.hist(data_ser, bins=range(50), density=True, cumulative=True)
+plt.xlim(0, 10)
 
 plt.grid(visible=True, axis='y')
 plt.title(title, fontsize=label_fontsize)

### 2022-11-21 ms/ms03_results.tex
MS

@@
 
-\subsection*{How specific are pleiotropic eQTLs to traits, genes and tissues?}
+\subsection*{How specific are eQTLs regarding traits, genes and tissues?}
 
-Next I explored the trait, gene and cell type/tissue specificity of colocalized eQTL/GWAS variants.
+Next I verified whether eQTLs tend to be specific to traits, eQTL genes and eQTL tissues.
 %
-74\% of eQTLs are involved in one GWAS class, 21\% are involved in 2 classes and the remaining variants with 3 or more trait classes make less that 4\% of eQTLs (Fig.
-\ref{fig:hist_gwas_egene_etissue}(a).
+74\% of eQTLs are associated to one GWAS class, 21\% are associated to 2 classes and the remaining variants associated to 3 or more trait classes make less that 4\% of eQTLs (Fig.
+\ref{fig:hist_gwas_egene_etissue}a).
 %
-Regarding target genes, 36\% of the colocalized eQTLs modulate a single gene, 24\% modulate two genes, and the remaining 40\% modulate three or more genes (Fig. \ref{fig:hist_gwas_egene_etissue}(b).
+Regarding eQTL genes, 36\% of the colocalized eQTL/GWAS variant modulate a single gene, 24\% modulate two genes, and the remaining 40\% modulate three or more genes (Fig.
+\ref{fig:hist_gwas_egene_etissue}b).
 %
-Concerning cell types and tissues classes, 27\% are active in a single tissue, 16\% in two tissues, 11\% in three tissues, and the remaining 46\% in four or more tissues (Fig.
-\ref{fig:hist_gwas_egene_etissue}(c).
+Concerning tissues, 27\% colocalized eQTL/GWAS variants are active in a single tissue, 16\% in two tissues, 11\% in three tissues, and the remaining 46\% in four or more tissues (Fig.
+\ref{fig:hist_gwas_egene_etissue}c).
 %
-In conclusion, most colocalized eQTL/GWAS variants are specific to one GWAS trait.
-%
-By contrast, a minority of colocalized eQTL/GWAS variants are specific to one single tissue.
+In conclusion, trait pleiotropy is rather rare, while, regulation of least two gene in at least two tissues is rather common (Fig. \ref{fig:hist_gwas_egene_etissue}).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{How do trait frequencies relates to eQTL and eTissue frequency at the loci level?}
+\subsection*{Trait, eQTL gene and tissue frequency of individual eQTLs in pleiotropic loci}
 
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of GWAS classes, eGenes and eTissues for regulatory variants in three pleiotropic regions.
+In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues for regulatory variants in three pleiotropic regions.
 %
@@
 %
-However these plots show that the relationship between the number of traits, eGenes and eTissues can be very different.
+However these plots show that the relationship between pleiotropy and the number of eQTL genes and tissues is not direct.
+
+Some regions have high number of eQTL genes and intermediate number of tissues like the interleukin locus in region 5:131912097-132802472 (5q31.1),
+intermediate number of eQTL gene and variable number of tissues like the 11:65488118-65916060 locus (11q13.1),
+high number of both eQTL genes and tissues like the 12:55,721,801-56,171,027 (12q13.2) region
+and relatively low number of eQTL genes and tissue like the 19:48,668,049-48,756,272 locus (19q13.33) (Fig. \ref{fig:region_gwas_egenes_tissues}).
+
+In summary, trait pleiotropy can arise in locus with variable absolute numbers of genes and tissues.
+
+\subsection*{Where are the pleiotropic variants?}
+
+%\subsection*{What are the mechanisms of pleiotropy}
+
+In the following sections, I wonder about the the molecular mechanism of pleiotropic eQTLs.
 %
-The SLC22A5 locus in region 5:132,239,645-132,497,907 (5q31.1) has a low number of eGenes but a high number of eTissues.
+%I hypothesized that pleiotropy arises from bias in regulatory effects of variants.
+%%
+%For instance, pleiotropic variants might significantly affect some molecular functions more often.
+%%
+%Another possibility is that pleiotropic variants affect more eGenes in more eTissues, which affect more GWAS traits.
+
+First, I analysed whether there are significant differences of some effect consequences between more and less pleiotropic variants.
 %
-By contrast, the MHC locus in region 6:31,034,839-32,478,149 (6p21.33) has a high number of both eGenes and eTissues.
+I separated variants according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence (\citep{2016.Cunningham.McLaren}).
 %
-Finally, the ATXN2 locus in region 12:111,395,984-111,645,358 (12q24.12) shows the highest trait pleiotropy but the number of eGenes and eTissues is very low.
+I found a significant larger number of upstream, 5'-UTR, downstream and 3'-UTR variants
+with 2, 3 and 4 and more classes compared to variants with one class (Fig. \ref{fig:vep_consequence}).
 %
-In summary, trait pleiotropy can arise from different situations: few eGenes and eTissues, many eGenes and eTissues or a mix of both.
-
-\subsection*{What are the mechanisms of pleiotropy}
-
-Then, I wanted to understand the molecular mechanism of pleiotropic variants and regions.
+I also found a significant larger number of intron, and non-coding transcript variants
+with 3 and 4 and more classes (Fig. \ref{fig:vep_consequence}).
 %
-I hypothesized that pleiotropy arises from bias in regulatory effects of variants.
-%
-For instance, pleiotropic variants might significantly affect some molecular functions more often.
-%
-Another possibility is that pleiotropic variants affect more eGenes in more eTissues, which affect more GWAS traits.
-
-\subsection*{Which are the effect consequence of pleiotropic eQTLs?}
-
-I first analysed whether there are significant differences of some effect consequences between more and less pleiotropic variants.
-%
-I separated variants according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence \citep{2016.Cunningham.McLaren}.
-%
-I found a significant larger number of missense variants among variants with 2, 3 and 4 classes compared to variants with 1 class (Fig. \ref{fig:vep_consequence}a).
-%
-I also found a significant larger number of splicing variants among variants with 3 classes (Fig. \ref{fig:vep_consequence}b).
-%
-Finally, there is also a larger of the 3'-UTR variants among variants with 4 classes (Fig. \ref{fig:vep_consequence}c).
-%
-% TODO Discuss sQTLs in 2021.Li.Mu.GenomeBiology.impactcelltype paper
-These analyses suggest that more pleiotropic variants have a stronger effect on the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
+These analyses suggest that more pleiotropic variants have a stronger effect on
+the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
 
@@
 
-Then I hypothesized that pleiotropic variants regulate more eGene even when we keep fixed the eTissue.
+Then I hypothesized that pleiotropic variants modulate more genes even within the same tissues.
 %
-To test this hypothesis, the eGenes per variant-eTissues pairs were counted.
+To test this hypothesis, I computed the number of eQTL genes per variant-tissue pair for different trait class counts.
 %
-Then these eGene counts were classified according to the GWAS class count.
+%If the GWAS class count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal one.
 %
-If the GWAS class count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal one.
-%
-I found means of eGene counts of 1.4, 1.7, 1.7 and 1.6 for GWAS classes counts 5, 4, 3 and 2 compared to eGene count mean of 1.5 for GWAS class count 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
-%
-This suggests that pleiotropic variant-eTissue pairs have slightly but significantly (Mann–Whitney U test) more eGenes compared to variants with one GWAS class (Fig.
+eQTLs involved in 1, 2, 3, 4, and 5 and more GWAS classes modulate an average of 1.3, 1.5, 1.6, 1.9 and 1.8 genes, respectively (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+%I found means of eGene counts of 1.4, 1.7, 1.7 and 1.6 for GWAS classes counts 5, 4, 3 and 2 compared to eGene count mean of 1.5 for GWAS class count 1 (Fig.
 \ref{fig:gwas_egene_etisue_per_variant}a).
 %
-This observation could be explained by my previous observation that pleiotropic variants have more often an effect on the splicing and 3'UTR regions (Figure \ref{fig:vep_consequence}b,c).
+This suggests that pleiotropic eQTLs modulate more genes compared to eQTLs involved in only one GWAS class (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+%
+More target genes of eQTLs provides an explanation to the observed pleiotropy.
 
@@
 
-My next hypothesis was that pleiotropic variant-eGene pairs are active in more eTissues compared with variant-eGene pairs with one GWAS class and this increases their probability to affect more GWAS classes.
+My next hypothesis was that pleiotropic eQTL-gene pairs are active in more tissues than non-pleiotropic eQTLs.
 %
-To test this hypothesis, the eTissues per variant-eGene pairs were counted.
+To test this hypothesis, the tissues per eQTL-gene pair was counted for eQTL involved in different GWAS class counts.
 %
-Then these eTissue counts were classified according to the GWAS class count.
+eQTLs involved in 1, 2, 3, 4, and 5 and more GWAS classes are active in an average of 2.4, 2.5, 2.8, 3.4 and 4.3 tissues, respectively (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
 %
-If the GWAS class count of a variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS class count.
+This suggests that pleiotropic eQTL are active in more tissues (Mann–Whitney U test) compared to non-pleiotropic eQTLs (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
+
+%\subsection*{Are pleiotropic eQTLs associated to more traits in unique gene-tissue pairs?}
 %
-I found means of eTissue counts of 2.2, 3, 3 and 2.6 for GWAS classes counts 5, 4, 3 and 2 compared to eTissue count mean of 2.5 for GWAS class count 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}b).
+%My next hypothesis was that pleiotropic variants are associated to more GWAS classes even after taking into account differences in eGene and eTissue counts.
 %
-This suggests that pleiotropic variant-eGene pairs are active slightly but significantly (Mann–Whitney U test) in more eTissues compared to variants with one GWAS class (Fig.
-\ref{fig:gwas_egene_etisue_per_variant}b).
-
-\subsection*{Are pleiotropic eQTLs associated to more traits in unique gene-tissue pairs?}
-
-My next hypothesis was that pleiotropic variants are associated to more GWAS classes even after taking into account differences in eGene and eTissue counts.
-
-To test this hypothesis, the GWAS class counts per variant-eGene-eTissue trios were counted.
-%
-Then these GWAS categorie counts per variant-eGene-eTissue trios were classified according to the GWAS class count per variant.
-%
-If the GWAS class count per variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS class count per variant.
-%
-I found means of GWAS class counts per variant-eGene-eTissue of 3, 2, 1.6, 1.4 and 1 for GWAS classes counts per variant of 5, 4, 3, 2 and 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
-%
-This shows a significant larger number of GWAS classes even in unique trios of variant-eGene-eTissue (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
-%
-This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
+%To test this hypothesis, the GWAS class counts per variant-eGene-eTissue trios were counted.
+%%
+%Then these GWAS categorie counts per variant-eGene-eTissue trios were classified according to the GWAS class count per variant.
+%%
+%If the GWAS class count per variant changed in different variant-eGene-eTissue trios, then we kept the maximal GWAS class count per variant.
+%%
+%I found means of GWAS class counts per variant-eGene-eTissue of 3, 2, 1.6, 1.4 and 1 for GWAS classes counts per variant of 5, 4, 3, 2 and 1 (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+%%
+%This shows a significant larger number of GWAS classes even in unique trios of variant-eGene-eTissue (Fig. \ref{fig:gwas_egene_etisue_per_variant}c).
+%%
+%This observation could be explained by my previous observation that pleiotropic variants have more often missense effects (Fig. \ref{fig:vep_consequence}a).
 
@@
 
-In Fig. \ref{fig:gwas_egene_etisue_per_variant}b, I observed that pairs of eQTL-eGene are active in more eTissues.
+In Fig. \ref{fig:gwas_egene_etisue_per_variant}b, I observed that pairs of eQTL-gene are active in more tissues.
 %
-To explain this observation, I hypothesized that pleiotropic variants are active in more eTissues, because genomic regions around pleiotropic variants bind more transcription factors, which upregulate the pleiotropic regions in more tissues.
+To explain this observation, I hypothesized that pleiotropic eQTLs are active in more tissues,
+because genomic regions around pleiotropic variants bind more transcription factors,
+which upregulate the pleiotropic regions in more tissues.
 %
-To test this hypothesis, I counted the number of unique transcription factors bound in a radius of 50 bp around each variant (Window of 100 bp).
+To test this hypothesis, I counted the number of unique transcription factors bound in a radius of 50 bp around each eQTL (Window of 100 bp).
 %
-I found a significant larger number of transcription factors bound around pleiotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
+eQTLs involved in 1, 2, 3 and 4 and more GWAS classes bind a mean number of 17,
+19, 23 and 27 transcription factors, respectively (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+%I found a significant larger number of transcription factors bound around pleiotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).
 
@@
 
-\subsection*{Do pleiotropic eQTLs have stronger effect size on the gene and trait?}
+In summary, pleitropic eQTLs bind more unique transcription factors and are more
+likely to belong to cis-regulatory modules.
 
-Pleiotropic variants affect simultaneously more GWAS traits.
+\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
+
+Pleiotropic eQTLs affect simultaneously more trait classes.
 %
-Therefore it is interesting to investigate how does variant causality and pleiotropy relate to effect size (beta) and significance (p-value) at the level of eQTL and GWAS traits.
+Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
 %

### 2022-11-21 ms/ms05_floats.tex
MS

@@
 %
-\caption{Odds ratio compared of the variants annotated with given variant effect predictor (VEP) consequence and given GWAS class count compared with non pleiotropic variants (GWAS class count one).
-The significance is given for the q-value (FDR 5\%).}
+\caption{Odds ratio compared of the variants annotated with given variant effect
+predictor (VEP) consequence and given GWAS class count compared with non
+pleiotropic variants (GWAS class count one). The significance is given for the
+q-value (False discovery rate 5\%).}
 \label{fig:vep_consequence}
@@
 %
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-\end{subfigure}
+%\begin{subfigure}[]{.32\textwidth}
+%\textbf{c}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
+%\end{subfigure}
 %
-\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS classes per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
+\caption{Mean number of genes in eQTL-tissue pairs (\textbf{a}) and mean number of tissues in eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
 %
@@
 \caption{Analysis of transcription factor binding.
-\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS class count.
-\textbf{b}, Odds ratio of cis regulatory module-annotated vs non-annotated variants.} \label{fig:freq_tf_per_variant}
+\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs (Radius 50 bp) as a function of the GWAS class count.
+\textbf{b}, Odds ratio of eQTLs annotated with cis regulatory modules vs non-annotated.} \label{fig:freq_tf_per_variant}
 %
@@
 \end{figure}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-11-27 ms/ms03_results.tex
ENH

@@
 
-\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}
+\subsection*{Systematic colocalization of variants from genome-wide associations studies and expression quantitative trait loci}
 
-To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
+To better understand the role of gene regulatory variants for variant pleiotropy,
+I developed a pipeline to systematically colocalize variants from genome-wide associations
+studies (GWAS) and expression quantitative trait loci (eQTL) (Section Methods).
 %
@@
 %	
-The second criterion was to include only the European population, because most samples from the EBI eQTL catalogue belong to the European population.
+The second criterion was to include only the European population, because most
+samples from the EBI eQTL catalogue belong to the European population.
 %
-The third criterion was to keep only well-defined medical or physiological conditions and exclude environmental traits such "employment status" or "self-reported" medical conditions.
+The third criterion was to keep only well-defined medical or physiological
+conditions and exclude environmental traits such "employment status" or "self-reported" medical conditions.
 %
@@
 
-\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
-
-Pleiotropic eQTLs affect simultaneously more trait classes.
+%\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
 %
-Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
+%Pleiotropic eQTLs affect simultaneously more trait classes.
+%%
+%Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
+%%
+%I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
+%%
+%Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 %
-I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
+%Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
+%%
+%I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
+%%
+%Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
 %
-Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
-
-Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
-%
-I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
-%
-Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
-
-These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).
+%These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).

### 2022-11-27 ms/ms06_floats.tex
ENH

--- 2022-11-21 ms/ms05_floats.tex
@@
 %
-\begin{figure}[!tbp]
-\centering
+%\begin{figure}[!tbp]
+%\centering
+%%
+%\begin{subfigure}[]{.33\textwidth}
+%\textbf{a}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
+%\end{subfigure}
+%%
+%\begin{subfigure}[]{.33\textwidth}
+%\textbf{b}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
+%\end{subfigure}
 %
-\begin{subfigure}[]{.33\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
-\end{subfigure}
+%\begin{subfigure}[]{.33\textwidth}
+%\textbf{c}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
+%\end{subfigure}
+%%
+%\begin{subfigure}[]{.33\textwidth}
+%\textbf{d}
+%\\
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
+%\end{subfigure}
 %
-\begin{subfigure}[]{.33\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
-\end{subfigure}
-
-\begin{subfigure}[]{.33\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{d}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
-\end{subfigure}
-
-\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
-%
-\end{figure}
+%\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
+%%
+%\end{figure}
 

### 2022-11-28 ms/ms03_results.tex
MS

@@
 
-Public GWAS summary statistics were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
+GWAS traits were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
 %
@@
 %
-The fourth criterion was to keep only GWAS studies with at least 10,000 subjects, 2,000 controls and 2,000 cases (Supplementary table 2).
+The fourth criterion was to keep only GWAS studies with at least 10,000 subjects, 2,000 controls and 2,000 cases.
 %
-These filters resulted in 418 GWAS.
+These filters resulted in 418 GWAS traits (Supplementary table 1).
 %

### 2022-11-28 scripts/cmpt_supplementary_files.py

@@ -57,6 +57,8 @@ class_pleio_df = class_pleio_df.loc[~class_pleio_df.isna().any(axis=1)]
 class_pleio_df.rename({'icd10_code_level1.1': 'icd10_code_level1'}, axis=1, inplace=True)
 #
 st_df = gwas_class_df[['id', 'trait', 'icd10_code_level1']].merge(class_pleio_df, on='icd10_code_level1')
+st_df = st_df.drop_duplicates()
+st_df = st_df.sort_values(st_df.columns.tolist())
 st_df.drop(['icd10_code_level1'], axis=1, inplace=True)
 st_df.to_excel(st_writer, sheet_name=sheet_name, index=False, header=True)
 

### 2022-11-28 ms/ms03_results.tex
MS

@@
 %
-127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.
+127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many
+tissues and cell types \citep{2021.Alasoo.Kerimov} (Supplementary table 1).
 
@@
 %
-These filters resulted in 418 GWAS traits (Supplementary table 1).
+These filters resulted in 418 GWAS traits (Supplementary table 2).
 %
-Among these studies, there were 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary table 2).
+Among these studies, there were 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
 % sqlite
@@
 % select count(*) from (select DISTINCT chrom, pos, rsid, nea, ea from tophits) limit 50; 10627
+% Supplementary file is db.sqlite that can go to OSF
 
@@
 %
-and the percentages of variants with a colocalized eQTL are 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST6).
+In these diseases, the percentages of variants with a colocalized eQTL are 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST3).
 %
@@
 
-\subsection*{Which eQTLs are the most pleiotropic?}
+\subsection*{Traits classification based on the colocalized eQTL effect size}
 
-I manually assigned the 418 GWAS to 96 classes to aggregate identical or similar traits (Supplementary table 1).
+I manually assigned the 418 GWAS to 96 classes to aggregate identical or similar traits (Supplementary table 2).
 %
@@
 %
-I computed distances between GWAS traits based on the Spearman correlation of the colocalized eQTL beta of the different eQTL genes and samples.
+I computed distances between GWAS traits based on the Spearman correlation of the colocalized eQTL beta for every target gene, cell type and tissue.
 %
@@
 %
-Due to the large number of breast cancer GWAS, we have created a separeted class from cancer.
-Nevertheless a weaker specific signal is observed between the breast cancer class and other types of cancers.
+Due to the large number of breast cancer GWAS, we have created a separated class from cancer.
+Nevertheless, we can observe weaker correlations between the breast cancer class and other types of cancers.
 
-In the EBI eQTL Catalogue, some samples have been analysed at the cell type level while other samples are analyzed at the tissue level.
+\subsection*{Which eQTLs are the most pleiotropic?}
+
+In the EBI eQTL Catalogue, some samples have been analyzed at the cell type level and others at the tissue level.
 %
-For instance, there is much detail resolution regarding the immune cell types.
+For instance, immune cell types are analyzed at very high resolution that includes different stimulation of the same cell type.
 %
-Therefore I manually defined 36 classes to aggregate eQTL samples to equivalent resolution (Supplementary table 2).
+Therefore I manually defined 36 classes to aggregate eQTL cell types into tissues (Supplementary table 1).
 %
-For instance the different immune cell types such as monocyte, T-cell, etc are aggregated to an ImmuneCell class (Supplementary table 2).
+For instance, the different immune cell types such as monocyte, T-cell, etc are aggregated into an ImmuneCell class (Supplementary table 1).
 
@@
 %
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS classes they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+Colocalized eQTL/GWAS variants were classified according to the number of GWAS classes they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
 %
-The aggregated variants are highly pleiotropic.
+The most pleiotropic variants belong to 6 classes that include different traits such as autoimmune diseases, circulatory disases and cancer.
 %
-For instance the variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1 and it is involved in six classes of traits, namely,
-allergy, autoimmune diseases, cancer of breast, circulatory system diseases, hypertension, and respiratory system diseases
-(Table \ref{tab:pleitropic_variants} and Supplementary table 3).
+For instance, the variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1
 %
-This variant rs2522051 is active in 23 classes of cell types and tissues including adipose, brain, digestive cells, immune cells, and sexual organs.
+This variant rs2522051 is involved in six classes of traits, namely,
+allergy, autoimmune diseases, cancer of the breast, circulatory system diseases, hypertension, and respiratory system diseases
+(Table \ref{tab:pleitropic_variants} and Supplementary table 4).
 %
-This variant rs2522051 regulates several genes such as IL13, PDLIM4 and RAD50.
+This variant rs2522051 is active in 23 classes of tissues including adipose, brain, digestive cells, immune cells, and sexual organs.
+%
+This variant rs2522051 regulates several genes such as IL13, PDLIM4, and RAD50.
 
@@
 Pleiotropic eQTLs are concentrated in genomic regions such as in cytobands 3q23,
-5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Supplementary table 3).
+5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
 %
-Therefore I computed regions that include at least one pleiotropic eQTL in
-sliding windows of 100,000 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 4).
+Therefore, I wanted to compute pleiotropic regions that include concentrations of pleiotropic eQTLs.
+%
+To compute these pleiotropic regions, I aggregated genomic regions with at least one pleiotropic eQTL in a sliding
+window of 100,000 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 5).
 %
 I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait
-classes and 80 regions with 3 or more trait classes (Supplementary table 4).
+classes and 80 regions with 3 or more trait classes (Supplementary table 5).
 %
@@
 
-The most pleiotropic region is 5:131,912,097-132,802,472 in cytoband 5q31.1 which contains the interferon response factor 1 (IRF1) and interleukins IL3, IL4, IL5 and IL13 (Table \ref{tab:pleiotropic_regions}).
+The most pleiotropic region is 5:131,912,097-132,802,472 in cytoband 5q31.1 which contains the interferon response factor 1
+(IRF1) and interleukins IL3, IL4, IL5 and IL13 (Table \ref{tab:pleiotropic_regions}).
 %
@@
 %
-The largest region is 7:2,712,518-7,254,268 in cytoband 7p22.3 with a length of 4,541,751 bp and variants associated to autoimmune and respiratory diseases and height (Supplementary table 5).
+The largest region is 7:2,712,518-7,254,268 in cytoband 7p22.3 with a length of 4,541,751 bp and variants associated to
+autoimmune and respiratory diseases and height (Supplementary table 5).
 
-Pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
 Very pleiotropic regions with 5 trait classes or more make around 0.6 Mb of the genome,
-regions with 4 GWAS classes make an additional 1 Mb and regions with 3 GWAS classes make an additional 1.5 Mb (Fig. \ref{fig:pleiotropy_region_distribution}b).
+regions with 4 GWAS classes make an additional 1 Mb and regions with 3 trait classes make an additional 1.5 Mb (Fig. \ref{fig:pleiotropy_region_distribution}b).
+%
+Altogether, pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
 
@@
 
-Next I verified whether eQTLs tend to be specific to traits, eQTL genes and eQTL tissues.
+Next I verified whether the colocalized eQTL/GWAS variants tend to be specific to traits, eQTL genes and eQTL tissues.
 %
-74\% of eQTLs are associated to one GWAS class, 21\% are associated to 2 classes and the remaining variants associated to 3 or more trait classes make less that 4\% of eQTLs (Fig.
-\ref{fig:hist_gwas_egene_etissue}a).
+74\% of colocalized eQTL/GWAS variants are associated to one GWAS class, 21\% are associated to 2 classes and the remaining variants associated to
+3 or more trait classes make less that 4\% of colocalized eQTL/GWAS variants (Fig. \ref{fig:hist_gwas_egene_etissue}a).
 %
-Regarding eQTL genes, 36\% of the colocalized eQTL/GWAS variant modulate a single gene, 24\% modulate two genes, and the remaining 40\% modulate three or more genes (Fig.
-\ref{fig:hist_gwas_egene_etissue}b).
+Regarding eQTL genes, 36\% of the colocalized eQTL/GWAS variants modulate a single gene, 24\% modulate two genes, and the
+remaining 40\% modulate three or more genes (Fig. \ref{fig:hist_gwas_egene_etissue}b).
 %
-Concerning tissues, 27\% colocalized eQTL/GWAS variants are active in a single tissue, 16\% in two tissues, 11\% in three tissues, and the remaining 46\% in four or more tissues (Fig.
-\ref{fig:hist_gwas_egene_etissue}c).
+Concerning tissues, 27\% colocalized eQTL/GWAS variants are active in a single tissue, 16\% in two tissues, 11\% in three
+tissues, and the remaining 46\% in four or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}c).
 %
-In conclusion, trait pleiotropy is rather rare, while, regulation of least two gene in at least two tissues is rather common (Fig. \ref{fig:hist_gwas_egene_etissue}).
+In conclusion, trait pleiotropy is rather rare among colocalized eQTL/GWAS variants, while, regulation of at least two
+genes in at least two tissues is rather common (Fig. \ref{fig:hist_gwas_egene_etissue}).
 
-\subsection*{Trait, eQTL gene and tissue frequency of individual eQTLs in pleiotropic loci}
+\subsection*{Trait, eQTL gene and tissue frequency of individual eQTLs in pleiotropic regions}
 
+% TODO continue here
 In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues for regulatory variants in three pleiotropic regions.

### 2022-11-28 ms/ms06_floats.tex
MS

@@
 %
-\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
+\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
+    (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
 \end{figure}
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{b}
+\textbf{a}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{c}
+\textbf{b}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{d}
+\textbf{c}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{e}
+\textbf{d}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{g}
+\textbf{e}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{h}
+\textbf{f}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{i}
+\textbf{g}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{j}
+\textbf{h}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{l}
+\textbf{i}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{m}
+\textbf{j}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{n}
+\textbf{k}
 \\
@@
 \begin{subfigure}[]{.24\textwidth}
-\textbf{o}
+\textbf{l}
 \\
@@
 
-\caption{Count of GWAS classes (\textbf{a-c}), eGenes (\textbf{d-f}) and eTissues (\textbf{g-i}) in three pleiotropic genomic regions: \textbf{a,d,g}: 5:132,239,645-132,497,907 (5q31.1), \textbf{b,e,h}: 6:31,034,839-32,478,149 (6p21.33) and \textbf{c,g,i}: 12:111,395,984-111,645,358 (12q24.1).
+\caption{Count of GWAS classes (\textbf{a-c}), eQTL genes (\textbf{d-f}) and eQTL tissues (\textbf{g-i}) in three
+pleiotropic genomic regions: \textbf{a,e,i}: 5:131,912,097-132,802,472 (5q31.1), \textbf{b,f,j}: 11:65,488,118-65,916,060 (11q13.1),
+    \textbf{c,g,k}: 12:55,721,801-56,171,027 (12q13.2), and \textbf{d,h,l}: 19:48,668,049-48,756,272 (19q13.33).
     Genomic coordinates are given for the hg38 assembly.}

### 2022-11-28 ms/ms03_results.tex
MS

@@
 
-% TODO continue here
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues for regulatory variants in three pleiotropic regions.
+In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues of
+eQTLs in three pleiotropic regions.
 %
-All three regions are very pleiotropic with very different associated trait classes such as cancer, cardiovascular and autoimmune diseases.
+All three regions are very pleiotropic with very different associated trait classes such as cancer, cardiovascular and
+autoimmune diseases (Table \ref{tab:pleiotropic_regions} and Supplementary table 5).
 %

### 2022-11-28 ms/ms06_floats.tex
MS

@@
 %
-\caption{Odds ratio compared of the variants annotated with given variant effect
+\caption{Odds ratio of the variants annotated with given variant effect
 predictor (VEP) consequence and given GWAS class count compared with non
-pleiotropic variants (GWAS class count one). The significance is given for the
-q-value (False discovery rate 5\%).}
+pleiotropic variants with GWAS class count one. The significance is given for the
+q-value with False discovery rate 5\%.}
 \label{fig:vep_consequence}

### 2022-11-28 ms/ms03_results.tex
MS

@@
 
-\subsection*{Trait, eQTL gene and tissue frequency of individual eQTLs in pleiotropic regions}
+\subsection*{Relation between trait, eQTL gene and tissue counts}
 
-In Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues of
-eQTLs in three pleiotropic regions.
+In Supplementary Figure \ref{fig:region_gwas_egenes_tissues}, I have plotted the number of trait classes, eQTL genes and tissues of
+eQTLs in four pleiotropic regions.
 %
@@
 However these plots show that the relationship between pleiotropy and the number of eQTL genes and tissues is not direct.
+%
+These plots suggest that counts of eQTL genes and tissues per variant correlate better between them than with the counts of the trait classes.
+%
+Indeed, Spearman correlation between eQTL gene and tissue counts is 0.75 while their correlation with the count of GWAS classes is 0.29 and 0.25, respectively (Fig.
+\ref{fig:region_gwas_egenes_tissues}).
 
-Some regions have high number of eQTL genes and intermediate number of tissues like the interleukin locus in region 5:131912097-132802472 (5q31.1),
-intermediate number of eQTL gene and variable number of tissues like the 11:65488118-65916060 locus (11q13.1),
-high number of both eQTL genes and tissues like the 12:55,721,801-56,171,027 (12q13.2) region
-and relatively low number of eQTL genes and tissue like the 19:48,668,049-48,756,272 locus (19q13.33) (Fig. \ref{fig:region_gwas_egenes_tissues}).
-
-In summary, trait pleiotropy can arise in locus with variable absolute numbers of genes and tissues.
+In summary, counts of eQTL genes and tissues is highly correlated, while correlation with GWAS classes is much lower.
 
@@
 
-In the following sections, I wonder about the the molecular mechanism of pleiotropic eQTLs.
+In the following sections, I studied different molecular mechanisms of pleiotropic eQTLs.
 %
@@
 
-First, I analysed whether there are significant differences of some effect consequences between more and less pleiotropic variants.
+First, I analysed whether there are significant differences of some effect consequences between more and less pleiotropic eQTLs.
 %
-I separated variants according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence (\citep{2016.Cunningham.McLaren}).
+I separated eQTLs according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence (\citep{2016.Cunningham.McLaren}).
 %
-I found a significant larger number of upstream, 5'-UTR, downstream and 3'-UTR variants
-with 2, 3 and 4 and more classes compared to variants with one class (Fig. \ref{fig:vep_consequence}).
+I found a significant larger number of upstream, 5'-UTR, downstream and 3'-UTR eQTLs
+with 2, 3 and 4 and more classes compared to eQTLs with one class (Fig. \ref{fig:vep_consequence}).
 %
-I also found a significant larger number of intron, and non-coding transcript variants
+I also found a significant larger number of intron, and non-coding transcript eQTLs
 with 3 and 4 and more classes (Fig. \ref{fig:vep_consequence}).
 %
-These analyses suggest that more pleiotropic variants have a stronger effect on
+These analyses suggest that more pleiotropic eQTLs have a stronger effect on
 the coding sequence and splicing regions, which might explain partly their more pleiotropic function.
@@
 
-Then I hypothesized that pleiotropic variants modulate more genes even within the same tissues.
+Then I hypothesized that pleiotropic eQTLs modulate more genes even after taking into account the same tissues.
 %
-To test this hypothesis, I computed the number of eQTL genes per variant-tissue pair for different trait class counts.
+To test this hypothesis, I computed the number of eQTL genes per eQTL-tissue pair for different trait class counts.
 %
@@
 %
-To test this hypothesis, the tissues per eQTL-gene pair was counted for eQTL involved in different GWAS class counts.
+To test this hypothesis, the number of tissues per eQTL-gene pair was counted for eQTL involved in different GWAS class counts.
 %
@@
 %
-To explain this observation, I hypothesized that pleiotropic eQTLs are active in more tissues,
-because genomic regions around pleiotropic variants bind more transcription factors,
-which upregulate the pleiotropic regions in more tissues.
+One explanation might be that pleiotropic eQTLs bound by more transcription factors, which in turn, upregulate the pleiotropic eQTLs in more tissues.
 %
@@
 eQTLs involved in 1, 2, 3 and 4 and more GWAS classes bind a mean number of 17,
-19, 23 and 27 transcription factors, respectively (Fig. \ref{fig:gwas_egene_etisue_per_variant}a).
+19, 23 and 27 transcription factors, respectively (Fig. \ref{fig:freq_tf_per_variant}a).
 %I found a significant larger number of transcription factors bound around pleiotropic variants (Fig. \ref{fig:freq_tf_per_variant}a).

### 2022-11-28 ms/ms04_discussion.tex
ENH updated with 418
MS

@@
 
-In this article, I present the results of a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
+In this article, I present the results of a colocalization pipeline for eQTLs and GWAS variants that I have applied to 418 GWAS and 127 eQTL studies.
 %
@@
 %
-These results are provided as a downloadable and an interactive table that can be used to explore variants with the GWAS traits, etissues and egenes.
+We start with 10,627 tag variants with a significant p-value below 5e-8 and we observe a colocalization for 3,849 or 36\% of the leading variants from 335 GWAS.
 %
-For a posterior probability PP.H4.abf$\ge$0.8 of common causal eQTL and GWAS variants, we found 100,526 loci.
+More preciseley, in autoimmune diseases, we find colocalization for 37\% of leading GWAS variants, which is comparable
+to a previous study that found colocalization for 40.4\% of GWAS loci \citep{2021.Li.Mu}.
 %
-If I focus on causal variants with PP.H4.abf$\ge$0.8 and SNP.PP.H4$\ge$0.5, then I find 4684 variants.
+The difference is that many colocalization by the previous study were contributed by RNA splicing, whereas our study entirely focuses on eQTLs.
 
-This interactive table can be used for different applications.
+These results are provided as a downloadable table that can be used to explore variants with the GWAS traits, eQTL tissues and genes.
 %
-For instance, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of GWAs traits.
-%
-Colocalized loci provide predictions of potential causal variants in a given loci of GWAS.
-%
-The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
+We found 103,551 variants from 246 GWAS that colocalized with a posterior probability PP.H4.abf$\ge$0.8 with at least one eQTL.
+
+%TODO continue
 
@@
 On the other hand, the effect on GWAS traits is more significant suggesting that the additive effect of several egenes and etissues results in stronger effects at the level of the GWAS trait (Fig TODO).
+
+% TODO comparison with watanabe pleiotropy study
+
+% TODO comparison with genome biology colocalization paper
+
+% This interactive table can be used for different applications.
+% %
+% For instance, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of GWAs traits.
+% %
+% Colocalized loci provide predictions of potential causal variants in a given loci of GWAS.
+% %
+% The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
+Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
 
@@
 %
-A website to access colocalization variants with PP.H4.abf $\geq$ 0.8 is available at this URL \url{https://agonzalez.pythonanywhere.com}.
+A website to access colocalization variants with PP.H4.abf $\geq$ 0.5 is available at this URL: TODO.
 

### 2022-11-28 ms/ms06_floats.tex
MS

@@
 \end{figure}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%
-%% Fig 4: Scatter plots of winner regions
+%% Fig 4: Correlation between GWAS, genes and tissue counts
 %%
@@
 %
+\begin{figure}[!tbp]
+\centering
 %
-\begin{figure}[!tbp]
-
-\begin{subfigure}[]{.24\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-\end{subfigure}
+\includegraphics[width=0.7\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
 %
-\begin{subfigure}[]{.24\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{d}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
-\end{subfigure}
-
-\begin{subfigure}[]{.24\textwidth}
-\textbf{e}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{f}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{g}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{h}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
-\end{subfigure}
-
-\begin{subfigure}[]{.24\textwidth}
-\textbf{i}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{j}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr11_start65488118_end65916060_categories7.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{k}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr12_start55721801_end56171027_categories6.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.24\textwidth}
-\textbf{l}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr19_start48668049_end48756272_categories6.png}
-\end{subfigure}
-
-\caption{Count of GWAS classes (\textbf{a-c}), eQTL genes (\textbf{d-f}) and eQTL tissues (\textbf{g-i}) in three
-pleiotropic genomic regions: \textbf{a,e,i}: 5:131,912,097-132,802,472 (5q31.1), \textbf{b,f,j}: 11:65,488,118-65,916,060 (11q13.1),
-    \textbf{c,g,k}: 12:55,721,801-56,171,027 (12q13.2), and \textbf{d,h,l}: 19:48,668,049-48,756,272 (19q13.33).
-    Genomic coordinates are given for the hg38 assembly.}
-\label{fig:region_gwas_egenes_tissues}
+\caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL tissues.}
+\label{fig:correlation_counts}
 %
 \end{figure}
-%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \caption{Analysis of transcription factor binding.
-\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs (Radius 50 bp) as a function of the GWAS class count.
-\textbf{b}, Odds ratio of eQTLs annotated with cis regulatory modules vs non-annotated.} \label{fig:freq_tf_per_variant}
+\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
+\textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs non-annotated.} \label{fig:freq_tf_per_variant}
 %
@@
 %\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
-%%
+%%Correlation
 %\end{figure}

### 2022-11-28 ms/ms06_suppl.tex
MS

--- 2022-11-17 ms/ms05_floats.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 1: Comorbidity matrix
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr.png}
-%
-\caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different gene and samples.}
-\label{fig:gwas_distance}
-%
-\end{figure}
+%  CONFIGURE NEW SINGLE-PAGE FORMAT
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Tab 1: Winner eQTLs
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\onecolumn % go back to one column
+\fancyhead{} % make sure we get no headers
+\renewcommand{\floatpagefraction}{0.1}
+\lfoot[\bSupInf]{\dAuthor}
+\rfoot[\dAuthor]{\cSupInf}
+\newpage
 
-% full size table is table
-\begin{table}[!tbp]
-\centering
-\scriptsize
-%\hline
-\csvreader[separator=tab,
-tabular=ccrrp{0.4\textwidth},
-head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
-]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
-%\hline
-%
-\vspace{15pt}
-%
-\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
-\end{table}
+\captionsetup*{format=largeformat} % make figure legend slightly larger than in the paper
+\setcounter{figure}{0} % reset figure counter for Supp. Figures
+\setcounter{equation}{0} % reset equation counter for Supp. Equations
+%\setcounter{page}{1} % reset page count
+\makeatletter
+\renewcommand{\thefigure}{S\@arabic\c@figure} % make Figure legend start with Figure S
+\makeatother
+\def\theequation{S\arabic{equation}}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Tab 2: Winner regions
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  MAIN TEXT
 
-% full size table is table
-\begin{table}[!tbp]
-\centering
-\scriptsize
-%\hline
-\csvreader[
-separator=tab,
-tabular=ccrrp{0.45\textwidth},
-head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
-]{\floatRelativePath/cmpt_pleiotropic_regions.py/region_window_ms_100000.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
-%\hline
-%
-\vspace{15pt}
-\caption{Pleiotropic regions involving 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
-\end{table}
+\newpage
+\section*{Supplementary Information}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 2: Length histogram of regions
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
-\end{subfigure}
-%
-\caption{Analysis of pleiotropic regions.
-\textbf{a}, Length distribution of pleiotropic regions.
-\textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 3: Histograms variants vs GWAS, eGene and eTissues
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
-\end{subfigure}
-%
-\caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eGenes (\textbf{b}) and eTissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
-\end{figure}
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
+
 \begin{figure}[!tbp]
@@
 \end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 5: VEP consequences
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-%
-\caption{Odds ratio compared of the variants annotated with given variant effect predictor (VEP) consequence and given GWAS class count compared with non pleiotropic variants (GWAS class count one).
-The significance is given for the q-value (FDR 5\%).}
-\label{fig:vep_consequence}
-%
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 6: Violin plots. eGenes, eTissues and gwas per variants
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.32\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-\end{subfigure}
-%
-\caption{Count of eGenes per variant-tissue (\textbf{a}), eTissues per variant-eGene (\textbf{b}) and GWAS classes per variant-eGene-eTissue (\textbf{c}).} \label{fig:gwas_egene_etisue_per_variant}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 7: TF count per GWAS class count
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
-\end{subfigure}
-%
-\caption{Analysis of transcription factor binding.
-\textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic variants (Radius 50 bp) as a function of the GWAS class count.
-\textbf{b}, Odds ratio of cis regulatory module-annotated vs non-annotated variants.} \label{fig:freq_tf_per_variant}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 8: beta and log pval
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
-\end{subfigure}
-
-\begin{subfigure}[]{.33\textwidth}
-\textbf{c}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{d}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
-\end{subfigure}
-
-\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
-%
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 9: graphical conclusions
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{\textwidth}
-
-\includegraphics[width=\textwidth]{fig/graphical_summary.png}
-\end{subfigure}
-
-\caption{\textbf{Model of regulatory variant pleiotropy.} I have investigated three possible mechanisms of pleiotropy.
-\textbf{Left}, Pleiotropic variants have more eGenes that result in more functions and more phenotypes.
-This might arise from an enrichment of pleiotropic variants in splicing or 3' UTR regions.
-\textbf{Center}, I have also found that eGenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
-This might be explained from variants being bound by more transcription factors.
-\textbf{Left} Triplets of variant-eGene-eTissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
-I have found that this might be explained by an enrichment of missense alleles.} \label{fig:beta}
-%
-\end{figure}

### 2022-12-11 ms/ms06_floats.tex
ENH working based on postgres db

@@
 %
-\begin{figure}[!tbp]
-\centering
-%
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-%
-\caption{Odds ratio of the variants annotated with given variant effect
-predictor (VEP) consequence and given GWAS class count compared with non
-pleiotropic variants with GWAS class count one. The significance is given for the
-q-value with False discovery rate 5\%.}
-\label{fig:vep_consequence}
-%
-\end{figure}
+% \begin{figure}[!tbp]
+% \centering
+% %
+% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
+% %
+% \caption{Odds ratio of the variants annotated with given variant effect
+% predictor (VEP) consequence and given GWAS class count compared with non
+% pleiotropic variants with GWAS class count one. The significance is given for the
+% q-value with False discovery rate 5\%.}
+% \label{fig:vep_consequence}
+% %
+% \end{figure}
 

### 2022-12-11 gwas2eqtl_pleiotropy/db.py

@@ -53,7 +53,7 @@ tophits = Table(
    Column('id', Integer, primary_key = True, autoincrement=True),
    Column('chrom', Integer, nullable=False),
    Column('pos', Integer, nullable=False),
-   Column('rsid', String(50), ForeignKey("rsid2cytoband.rsid"), nullable=False),
+   Column('rsid', String(50), nullable=False),
    Column('nea', String(50), nullable=False),
    Column('ea', String(50), nullable=False),
    Column('pval', Float, nullable=False),
@@ -62,7 +62,7 @@ tophits = Table(
    Column('eaf', Float, nullable=True),
    Column('n', Float, nullable=True),
    Column('se', Float, nullable=False),
-   Column('gwas_id', String(50), ForeignKey("gwas_annot.gwas_id"), nullable=False),
+   Column('gwas_id', String(50), nullable=False),
    __table_args__ = (
            UniqueConstraint('chrom', 'pos', 'rsid', 'gwas_id', name='unique_cnstnt'),
        )

### 2022-12-11 scripts/annot_gwas_metadata.py

@@ -8,9 +8,6 @@ import pandas
 import pathlib
 import sys
 
-# gwas_metadata_ods = "config/gwas418.ods"
-# gwas_annot_tsv_path = "t.tsv"
-
 #%%
 help_cmd_str = "todo"
 try:
@@ -26,8 +23,6 @@ except IndexError:
     sys.exit(1)
 
 # #%% Outdir
-# if not '__file__' in locals():
-#     __file__ = "dwnld_gwas_info.py"
 outdir_path = os.path.dirname(gwas_annot_ods_path)
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
@@ -35,7 +30,6 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 metadata_df = pandas.read_excel(gwas_metadata_ods, engine='odf')
 
 #%%
-# import pdb; pdb.set_trace()
 metadata_df.drop(['Unnamed: 4'], axis=1, inplace=True)
 
 #%%
@@ -68,6 +62,5 @@ metadata_mrcieu_df.rename({'id': 'gwas_id'}, axis=1, inplace=True)
 
 #%%
 gwas_annot_df = metadata2_df.merge(metadata_mrcieu_df, on='gwas_id')
-# gwas_annot_df.to_csv(gwas_annot_tsv_path, sep="\t", header=True, index=False)
 with pandas.ExcelWriter(gwas_annot_ods_path) as fout:
     gwas_annot_df.to_excel(fout, index=False)

### 2022-12-11 scripts/cmpt_count_per_rsid.py

@@ -10,8 +10,8 @@ from matplotlib import pyplot as plt
 #%%
 help_cmd_str = "todo"
 try:
-    annotated_tsv_path = sys.argv[1]
-    max_gwas_class_count = int(sys.argv[2])
+    max_gwas_class_count = int(sys.argv[1])
+    url = sys.argv[2]
     count_per_rsid_gwas_tsv_path = sys.argv[3]
     count_per_rsid_gwas_egene_etissue_corr_png = sys.argv[4]
     if len(sys.argv) > 5:
@@ -26,28 +26,34 @@ except IndexError:
 outdir_path = os.path.dirname(count_per_rsid_gwas_tsv_path)
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
+# import pdb; pdb.set_trace()
+sql = 'select * from colocpleio'
+columns = ['chrom', 'pos38', 'cytoband', 'rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id', 'etissue_class']
+coloc_df = pandas.read_sql(sql, con=url, columns=columns).drop_duplicates()
+
+
 #%%
-coloc_df = pandas.read_csv(annotated_tsv_path, sep="\t")
-coloc_df['chrom'] = coloc_df['chrom'].astype(int)
+# coloc_df = pandas.read_csv(annotated_tsv_path, sep="\t")
+# coloc_df['chrom'] = coloc_df['chrom'].astype(int)
 
 #%% etissue pleiotropy
-etissue_df = coloc_df[['chrom', 'cytoband', 'pos', 'rsid', 'etissue_class']].drop_duplicates().groupby(['chrom', 'cytoband', 'pos', 'rsid']).agg({'etissue_class': ['size', (lambda x: (",".join(sorted(x))))]}).reset_index()
-etissue_df.columns = ['chrom', 'cytoband', 'pos', 'rsid', 'etissue_label_count', 'etissue_class_lst']
-etissue_df = etissue_df.sort_values(by=['etissue_label_count', 'chrom', 'pos', 'etissue_class_lst', 'rsid'], ascending=[False, True, True, True, True])
[…]

### 2022-12-11 scripts/cmpt_pleiotropic_regions.py

@@ -37,7 +37,7 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
 
 #%% Do it forward
-df.sort_values(by=['chrom', 'pos'], inplace=True, ascending=[True, True])
+df.sort_values(by=['chrom', 'pos38'], inplace=True, ascending=[True, True])
 more_than_1 = 0
 chrom_pleio_latest = 0
 pos_pleio_latest = -99999
@@ -46,17 +46,17 @@ df['region_pleio_fwd'] = False
 #%%
 for i, row in df.iterrows():
     chrom = row['chrom']
-    pos = row['pos']
+    pos38 = row['pos38']
     # import pdb; pdb.set_trace()
     if row['gwas_class_count'] > 1:
         chrom_pleio_latest = chrom
-        pos_pleio_latest = pos
-    if chrom == chrom_pleio_latest and (pos - pos_pleio_latest) <= region_bin:
+        pos_pleio_latest = pos38
+    if chrom == chrom_pleio_latest and (pos38 - pos_pleio_latest) <= region_bin:
         df.loc[i, 'region_pleio_fwd'] = True
 # df.to_csv("df.tsv", sep="\t", index=False)
 
 #%% Do it reversed
-df.sort_values(by=['chrom', 'pos'], inplace=True, ascending=[True, False])
+df.sort_values(by=['chrom', 'pos38'], inplace=True, ascending=[True, False])
 more_than_1 = 0
 chrom_pleio_latest = 0
 pos_pleio_latest = 999999999
@@ -65,18 +65,18 @@ df['region_pleio_rev'] = False
 #%%
 for i, row in df.iterrows():
     chrom = row['chrom']
-    pos = row['pos']
+    pos38 = row['pos38']
     df.loc[i, 'region_pleio'] = False
     if row['gwas_class_count'] > 1:  # update pos_pleio_latest
         chrom_pleio_latest = chrom
-        pos_pleio_latest = pos
-    if chrom == chrom_pleio_latest and (pos_pleio_latest - pos) <= region_bin:
+        pos_pleio_latest = pos38
+    if chrom == chrom_pleio_latest and (pos_pleio_latest - pos38) <= region_bin:
         if df.loc[i, 'region_pleio_fwd']:
             df.loc[i, 'region_pleio'] = True
 
 #%%
[…]

### 2022-12-11 scripts/cmpt_public_web_files.py

@@ -58,7 +58,7 @@ with engine.connect() as con:
                 continue
             coloc_df = pandas.DataFrame(result_lst)
             coloc_df.drop(['id'], inplace=True, axis=1)
-            coloc_df.rename({'symbol': 'egene_symbol'}, axis=1, inplace=True)
+            coloc_df.rename({'symbol': 'eqtl_gene_symbol'}, axis=1, inplace=True)
             col_name1_lst = []
             col_name1_lst.append('chrom')
             col_name1_lst.append('pos')
@@ -69,7 +69,7 @@ with engine.connect() as con:
             col_name1_lst.append('gwas_trait')
             col_name1_lst.append('gwas_class')
             col_name1_lst.append('gwas_beta')
-            col_name1_lst.append('egene_symbol')
+            col_name1_lst.append('eqtl_gene_symbol')
             col_name1_lst.append('eqtl_beta')
             col_name1_lst.append('eqtl_id')
             col_name1_lst.append('egene')

### 2022-12-11 scripts/insrt_etissue_class.py

@@ -1,0 +1,44 @@
+from gwas2eqtl_pleiotropy.Logger import Logger
+from psycopg2.extras import NumericRange
+from gwas2eqtl_pleiotropy.db2 import Base
+from sqlalchemy import create_engine
+
+import pandas
+import sys
+
+#%%
+
+help_cmd_str = "todo"
+try:
+    url = sys.argv[1]
+    etissue_cat_ods_path = sys.argv[2]
+    if len(sys.argv) > 3:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+#%% Create all tables
+engine = create_engine(url)
+Base.metadata.create_all(engine)
+
+eqtl_annot_df = pandas.read_excel(etissue_cat_ods_path, index_col=None, header=0)
+eqtl_annot_df.drop(['Unnamed: 8', 'etissue_class.1', 'count'], axis=1, inplace=True)
+
+eqtl_url = "https://raw.githubusercontent.com/eQTL-Catalogue/eQTL-Catalogue-resources/master/tabix/tabix_ftp_paths.tsv"
+eqtl_df = pandas.read_csv(eqtl_url, sep="\t", header=0)
+eqtl_df = eqtl_df.loc[eqtl_df['ftp_path'].str.contains('/ge/|/microarray/',regex=True,na=False), ]
+eqtl_df['index'] = (eqtl_df['ftp_path'].str.replace('.all.tsv.gz', '', regex=True)).str.split('/', expand=True)[10].tolist()
+eqtl_df.set_index('index', drop=True, verify_integrity=True, inplace=True)
+
+eqtl_annot_df = eqtl_df.merge(eqtl_annot_df, on=['study', 'qtl_group', 'tissue_ontology_id', 'tissue_ontology_term', 'tissue_label', 'condition_label'])
+eqtl_annot_df.set_index(eqtl_df.index, verify_integrity=True, inplace=True)
+eqtl_annot_df.index.rename('eqtl_id', inplace=True)
+
+Logger.info("Insert eqtl_annot")
+# import pdb; pdb.set_trace()
+engine.execute((Base.metadata.tables['eqtl_annot']).delete())
+eqtl_annot_df.to_sql('eqtl_annot', con=engine, if_exists='append', index=True)

### 2022-12-11 scripts/pltbar_x_per_variant_etissue_y_egene.py

@@ -19,9 +19,10 @@ seaborn.set_theme(**seaborn_theme_dic)
 #%%
 help_cmd_str = "todo"
 try:
-    h4_annot_tsv_path = sys.argv[1]
-    count_per_rsid_gwas_tsv_path = sys.argv[2]
-    max_gwas_class_count = int(sys.argv[3])
+    # h4_annot_tsv_path = sys.argv[1]
+    max_gwas_class_count = int(sys.argv[1])
+    url = sys.argv[2]
+    count_per_rsid_gwas_tsv_path = sys.argv[3]
     vlnplt_png_path = sys.argv[4]
     if len(sys.argv) > 5:
         print("""Two many arguments!
@@ -32,10 +33,10 @@ except IndexError:
     {}""".format(help_cmd_str))
     sys.exit(1)
 
-#%% Input1
-if not os.path.isfile(h4_annot_tsv_path):
-    print("input file does not exit")
-    sys.exit(1)
+# #%% Input1
+# if not os.path.isfile(h4_annot_tsv_path):
+#     print("input file does not exit")
+#     sys.exit(1)
 
 #%% Input2
 if not os.path.isfile(count_per_rsid_gwas_tsv_path):
@@ -46,21 +47,24 @@ outdir_path = os.path.dirname(vlnplt_png_path)
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 #%%
-h4_df = pandas.read_csv(h4_annot_tsv_path, sep="\t")
+# h4_df = pandas.read_csv(h4_annot_tsv_path, sep="\t")
+sql = 'select * from colocpleio'
+# columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
+h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
 
 #%%
 count_per_rsid_gwas_df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
 gwas_class_count_max_int = count_per_rsid_gwas_df['gwas_class_count'].max()
 
 #%%
-m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos', 'rsid'])
+m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos38', 'rsid'])
 
 # %%
-sel_cols = ['rsid', 'egene', 'etissue_class']  # egene per variant-etissuecategory
[…]

### 2022-12-11 scripts/plthtmp_disease_comorbidity_matrix.py

@@ -79,18 +79,20 @@ clustermap_args_dic['row_cluster'] = False
 clustermap_args_dic['col_cluster'] = False
 clustermap_args_dic['xticklabels'] = False
 
-clustermap_args_dic['yticklabels'] = annotation_df['trait'].tolist()
+clustermap_args_dic['yticklabels'] = [s[0:40] for s in annotation_df['trait'].tolist()]
 
 g = seaborn.clustermap(dis_df, **clustermap_args_dic)
+
 seaborn.set(font_scale=1)
 g.cax.set_visible(False)
-g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize = 6)
+g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize=10)
+g.ax_heatmap.set(ylabel='Trait')
 
 for label in class_labels.unique():
     g.ax_col_dendrogram.bar(0, 0, color=subset1_lut[label], label=label, linewidth=0);
 l1 = g.ax_col_dendrogram.legend(title='Class', loc="upper left", bbox_to_anchor=(0.05, 0.95), ncol=4, bbox_transform=gcf().transFigure)
 
-plt.subplots_adjust(top=1, right=0.8, bottom=0.05, left=-0.1)
+plt.subplots_adjust(top=1., right=0.6, bottom=0.05, left=-0.15)
 plt.savefig(htmp_disease_corr_png_path, dpi=600)
 plt.clf()
 plt.close()

### 2022-12-11 scripts/pltsctr_x_per_rsid_y_egene.py

@@ -52,7 +52,7 @@ ylim = [0, 20]
 c = 'blue'
 ylabel = "eGene count"
 
-count_per_rsid_df['pos'] = count_per_rsid_df['pos'].astype('int')
+count_per_rsid_df['pos38'] = count_per_rsid_df['pos38'].astype('int')
 
 #%% Loop over regions
 pleiotropic_regions_df = region_window_100000_df.loc[region_window_100000_df['gwas_class_count'] >= max_gwas_class_count, ['chrom', 'start', 'end', 'gwas_class_count']]
@@ -63,12 +63,12 @@ for rowi, row in pleiotropic_regions_df.iterrows():
     gwas_class_count = row['gwas_class_count']
     count_per_rsid_gwas_region_df = count_per_rsid_df.copy()
     count_per_rsid_gwas_region_df = count_per_rsid_gwas_region_df.loc[count_per_rsid_gwas_region_df['chrom'] == chrom, ]
-    count_per_rsid_gwas_region_df = count_per_rsid_gwas_region_df.loc[count_per_rsid_gwas_region_df['pos'] >= start, ]
-    count_per_rsid_gwas_region_df = count_per_rsid_gwas_region_df.loc[count_per_rsid_gwas_region_df['pos'] <= end, ]
+    count_per_rsid_gwas_region_df = count_per_rsid_gwas_region_df.loc[count_per_rsid_gwas_region_df['pos38'] >= start, ]
+    count_per_rsid_gwas_region_df = count_per_rsid_gwas_region_df.loc[count_per_rsid_gwas_region_df['pos38'] <= end, ]
 
     #%%
-    # plt.scatter(count_per_rsid_gwas_region_df['pos'] / 1000000, count_per_rsid_gwas_region_df[count_col_name], c='blue', s=scatter_dot_size)
-    seaborn.scatterplot(x=count_per_rsid_gwas_region_df['pos'] / 1000000, y=count_per_rsid_gwas_region_df[count_col_name], s=scatter_dot_size)
+    # plt.scatter(count_per_rsid_gwas_region_df['pos38'] / 1000000, count_per_rsid_gwas_region_df[count_col_name], c='blue', s=scatter_dot_size)
+    seaborn.scatterplot(x=count_per_rsid_gwas_region_df['pos38'] / 1000000, y=count_per_rsid_gwas_region_df[count_col_name], s=scatter_dot_size)
 
     plt.grid(True)
     plt.title(title, fontsize=label_fontsize)

### 2022-12-11 scripts/cmpt_comorbidites.py
ENH

@@ -27,9 +27,9 @@ except IndexError:
 outdir_path = os.path.dirname(disease_corr_tsv_path)
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
-sql = 'select * from coloc'
+sql = 'select * from colocpleio'
 columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
-d_df = pandas.read_sql(sql, con=url, index_col='id', columns=columns).drop_duplicates()
+d_df = pandas.read_sql(sql, con=url, columns=columns).drop_duplicates()
 
 #%%
 # Logger.info("Reading {}".format(annotated_tsv_path))

### 2022-12-12 ms/ms06_floats.tex
ENH

@@
 ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv}% specify your coloumns here
 %\hline
@@
 separator=tab,
-tabular=ccrrp{0.45\textwidth},
+tabular=p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.45\textwidth},
 head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
-]{\floatRelativePath/cmpt_pleiotropic_regions.py/region_window_ms_100000.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
+table head=\bfseries Chrom & \bfseries Pleio RSID & \bfseries Pleio pos. & \bfseries Cytob. & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
+]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi\ & \csvcolvii}% specify your coloumns here
 %\hline
@@
 \vspace{15pt}
-\caption{Pleiotropic regions involving 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
+\caption{Pleiotropic regions involving 7 or more GWAS classes.
+These regions were built with a sliding window of 1e5 nt.
+Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
 \end{table}
@@
 %
-\begin{figure}[!tbp]
-\centering
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{a}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
-\end{subfigure}
-%
-\begin{subfigure}[]{.33\textwidth}
-\textbf{b}
-\\
-\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
-\end{subfigure}
-%
-\caption{Analysis of pleiotropic regions.
+% \begin{figure}[!tbp]
+% \centering
+% %
+% \begin{subfigure}[]{.33\textwidth}
+% \textbf{a}
+% \\
+% \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
+% \end{subfigure}
+% %
+% \begin{subfigure}[]{.33\textwidth}
+% \textbf{b}
+% \\
+% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
+% \end{subfigure}
+% %
+% \caption{Analysis of pleiotropic regions.
 \textbf{a}, Length distribution of pleiotropic regions.
 \textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
-\end{figure}
+% \end{figure}
 %

### 2022-12-12 scripts/cmpt_pleiotropic_regions.py

@@ -1,3 +1,5 @@
+import pdb
+
 from gwas2eqtl_pleiotropy.constants import region_bin, label_fontsize, tick_fontsize
 from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 from matplotlib import pyplot as plt
@@ -31,6 +33,8 @@ except IndexError:
 outdir_path = os.path.dirname(pleio_tsv_path)
 pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
+manuscript_pleio_cutoff = 7
+
 #%%
 df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
 
@@ -46,7 +50,7 @@ df['gwas_class_count'] = 1
 # window = 3e4
 # window = 1e5
 for chrom in sorted(df['chrom'].unique()):
-    print(chrom)
+    # print(chrom)
     df_pleio_chrom = df.loc[df['chrom'] == chrom].copy()
     df_pleio_chrom['dist'] = [0] + [df_pleio_chrom.iloc[i]['pos38'] - df_pleio_chrom.iloc[i - 1]['pos38'] for i in range(1, df_pleio_chrom.shape[0])]
     region_pleio = 0
@@ -77,16 +81,19 @@ pleio_df['start'] = pleio_df['start'].astype(int)
 pleio_df['end'] = pleio_df['end'].astype(int)
 
 
-pleio_df['name'] = None
+pleio_df['pleio_rsid'] = None
+pleio_df['pleio_pos38'] = None
+# pleio_df['name'] = None
 for i, row in pleio_df.iterrows():
     start = row['start']
     end = row['end']
-    variant_lead = (df.loc[(df['pos38'] >= start) & (df['pos38'] <= end)]).head(1)
-    region_name = \
-    (variant_lead['cytoband'] + "_" + variant_lead['pos38'].astype(str) + "_" + variant_lead['rsid']).values[0]
-    pleio_df.loc[i, 'name'] = region_name
+    variant_pleio = (df.loc[(df['pos38'] >= start) & (df['pos38'] <= end)]).head(1)
+    # region_name = \
+    # (variant_lead['cytoband'] + "_" + variant_lead['pos38'].astype(str) + "_" + variant_lead['rsid']).values[0]
+    pleio_df.loc[i, 'pleio_rsid'] = variant_pleio['rsid'].values[0]
+    pleio_df.loc[i, 'pleio_pos38'] = variant_pleio['pos38'].values[0]
 
[…]

### 2022-12-13 ms/ms06_floats.tex
ENH

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 5: VEP consequences
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% \begin{figure}[!tbp]
-% \centering
-% %
-% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-% %
-% \caption{Odds ratio of the variants annotated with given variant effect
-% predictor (VEP) consequence and given GWAS class count compared with non
-% pleiotropic variants with GWAS class count one. The significance is given for the
-% q-value with False discovery rate 5\%.}
-% \label{fig:vep_consequence}
-% %
-% \end{figure}
+% Fig 5: VEP consequences
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!tbp]
+\centering
+%
+\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
+%
+\caption{Odds ratio of the variants annotated with given variant effect
+predictor (VEP) consequence and given GWAS class count compared with non
+pleiotropic variants with GWAS class count one. The significance is given for the
+q-value with False discovery rate 5\%.}
+\label{fig:vep_consequence}
+%
+\end{figure}
 

### 2022-12-13 scripts/pltbar_vep_consequence.py

@@ -42,6 +42,7 @@ in_df.loc[in_df['pfdr5perc'] <= 5.00e-02, 'signif'] = '*'
 in_df.loc[in_df['pfdr5perc'] <= 1.00e-02, 'signif'] = '**'
 in_df.loc[in_df['pfdr5perc'] <= 1.00e-03, 'signif'] = '***'
 in_df.loc[in_df['pfdr5perc'] <= 1.00e-04, 'signif'] = '****'
+in_df.loc[in_df['gwas_class_count'] == max_gwas_class_count, 'gwas_class_count'] = '≥{}'.format(max_gwas_class_count)
 
 ################################################################################
 # Draw a nested barplot by species and sex

### 2022-12-17 ms/ms05_methods.tex

--- 2022-09-10 ms/ms04_discussion.tex
@@
-\section*{Discussion}
-
-In this article, I present the results of a colocalization pipeline for eQTLs and GWAS variants that I have applied to a large number of 413 GWAS and 127 eQTL studies.
-%
-I have taken advantage of two public resources of GWAS and eQTL studies with summary statistics to compute a large number of colocalizations between eQTLs and GWAS variants.
-%
-These results are provided as a downloadable and an interactive table that can be used to explore variants with the GWAS traits, etissues and egenes.
-%
-For a posterior probability PP.H4.abf$\ge$0.8 of common causal eQTL and GWAS variants, we found 100,526 loci.
-%
-If I focus on causal variants with PP.H4.abf$\ge$0.8 and SNP.PP.H4$\ge$0.5, then I find 4684 variants.
-
-This interactive table can be used for different applications.
-%
-For instance, egenes and etissues of colocalized eQTL variants provide potential predictions of the molecular mechanism of GWAs traits.
-%
-Colocalized loci provide predictions of potential causal variants in a given loci of GWAS.
-%
-The variant rs2107595 is in locus 7p21.1 downstream of the histone deacetylase 9 (HDAC9) is cardiovascular.
-Interestingly, it controls expression of lncRNA XXX that is not known in the context of cardiovascular diseases.
-
-I have extensively compared molecular properties of pleiotropic variants involved in several GWAS traits.
-%
-Our analysis provides a number of hints about pleiotropic variants and regions.
-%
-I have found that these pleiotropic variants are under the control of more transcription factors and regulate more egenes in more etissues.
-%
-By contrast, the eQTLs of pleiotropic variants show lower beta and less significant p-values (Fig. TODO).
-%
-This can make sense, because these variants target a larger number of egenes and etissues, and larger effects are maybe negatively selected.
-%
-It has been recently observed that eQTLs are negatively selected from essential genes (Ref bioarxiv TODO).
-%
-On the other hand, the effect on GWAS traits is more significant suggesting that the additive effect of several egenes and etissues results in stronger effects at the level of the GWAS trait (Fig TODO).
-
-At the molecular level, I have found that the immune system might be partly involved in pleiotropy.
-%
-Immune related gene ontology is significantly found in pleiotropic egenes (Fig 1) and etissues related to immune cells are overrepresented in pleiotropic loci (Fig 2).
-%
-The most pleiotropic locus 12q24.12 includes the SH2B3 gene that regulates cytokine signaling.
-%
-This locus has been used to explain links between inflammation and hypertension (10.1097/MNH.0000000000000196), diabetes and autoimmunity (10.4239/wjd.v5.i3.316).
-%
-There is also the 5q31.1 locus that includes important genes of the immune system such as IRF1 and IL4 and the MHC locus that was previously discussed (REF watanaba discussion TODO).
-
-% TODO Causal variants
-% If pleiotropic variants bind more transcription factors and are more frequently annotated with a CRM, then these variants are probably more likely causal variants.
-
-% TODO Relation between eQTL and GWAS and related publications.
-
-%% 6p21.3
-%% TODO
-%
-%%15q26.1
-%The next pleiotropic locus is in cytoband 15q26.1 involved in cardiovascular, hypertension and psyciatric diseases with variants rs2071382, rs8039305 and rs6224.
-%%
-%eQTLs in cytoband 15q26.1 control genes like Furin, oncogene Fes and UNC45A involved in Osteootohepatoenteric syndrome (OMIM DB).
-%%
-%eQTLs are active in a very large number of cells including blood, artery, brain and immune cells (See supplementary table).
-%
-%% 1p13.3
-%The locus in cytoband 1p13.3 is mainly related to cardiovascular diseases.
-%Interestingly, these variants are active in many tissues (Supp table h4_annotated.ods).
-%
-%The variant rs11236797 and the two others in locus 11q13.5 is mainly autoimmune.
-%All three regulate the gene EMSY-DT and are active induced pluripotent cells, LCDs and skin.
-%
-%Another pleiotropic variants rs301806
-%rs301807
-%rs301802
-%rs301817
-%rs159963
-%rs301816
-%in the cytoband 1p36.23 that are involved in allergy, cardiovascular and hypertension.
-%This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\
-
 \section*{Methods}\label{sec:methods}
@@
 Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
-%
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/gwas2eqtl} and \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
 
@@
 
-My colocalization pipeline was inspired by another pipeline that is at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
+Full association data were downloaded from OpenGWAS and converted to hg38 coordinates using Picard and Crossmap \citep{2021.Marcora.Lyon,Picard2019toolkit,2013.Wang.Zhao}.
 %
-GWAS were downloaded from OpenGWAS and converted to hg38 coordinates.
+Top hit variants from the corresponding GWAS were download based on p-value threshold 0.8, clumping parameter r$^2$=0.1 and a 1 Mb clumping window.
 %
-Permuted eQTLs with FDR 0.05 in each sample were selected as lead variants.
+eQTL and GWAS variants in a total window with diameter 1 Mb around the top hits variants.
 %
-eQTL and GWAS variants surrounded the lead variants with radius of 500,000 nt (Total window 1,000,000 nt) were retrieved.
+Missing minor allele frequency variants (MAF) were retrieved from the European population in the 1000 genome database \citep{2015.Abecasis.Auton}.
 %
-These variants were kept if MAF was strictly greater than 0 and lower than 1, the variants were not duplicated and there was no missing data.
+Variants were kept if MAF was strictly between 0 and 1, the variants were not duplicated and there was no missing data.
 %
 Colocalization between eQTL and GWAS variants was tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
+%
+The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/gwas2eqtl} and \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
 

### 2022-12-18 ms/ms03_results.tex
ENH PMIDs in disease matrix

@@
 
-To better understand the role of gene regulatory variants for variant pleiotropy,
-I developed a pipeline to systematically colocalize variants from genome-wide associations
-studies (GWAS) and expression quantitative trait loci (eQTL) (Section Methods).
+To annotate traits from genome-wide associations studies (GWAS) with genes and tissues,
+I developed a pipeline that systematically colocalizes variants from GWAS and expression quantitative trait loci (eQTLs) (Section Methods).
 %
+% TODO Supplementary table 1 for newest pipeline
 127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many
@@
 %
-The fourth criterion was to keep only GWAS studies with at least 10,000 subjects, 2,000 controls and 2,000 cases.
+The fourth criterion was to keep only GWAS with at least 10,000 subjects, 2,000 controls and 2,000 cases.
 %
+% TODO Supplementary table 2 for newest pipeline
 These filters resulted in 418 GWAS traits (Supplementary table 2).
 %
-Among these studies, there were 10,627 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
-% sqlite
-% select count(distinct gwas_id) from tophits; 335
-% select count(*) from (select DISTINCT chrom, pos, rsid, nea, ea from tophits) limit 50; 10627
-% Supplementary file is db.sqlite that can go to OSF
+Among these studies, there were 8,580 clumped leading variants with a p-value below 5e-8 from 284 GWAS (Supplementary file).
+% select count(distinct gwas_id) from tophits; 284
+% select count(*) from (select DISTINCT chrom, pos, rsid, ea from tophits) as sel; 8,580
+% tophits goes to OSF
+% colocpleio goes to OSF
 
-My pipeline uncovered eQTL colocalizations for 3,849 or 36$\%$ of the leading GWAS variants with 2,348 eQTL genes and 127 eQTL samples.
-% sqlite
-% select count(distinct tophits.rsid) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8; 3849
-% alternative: select count(distinct tophits.rsid) from tophits, coloc where coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8; 3849
-% select count(distinct coloc.egene) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8
-% select count(distinct coloc.eqtl_id) from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.`PP.H4.abf`>=0.8
+My pipeline uncovered eQTL colocalizations for 5,484 or 64$\%$ of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
+% select count(*) from (select distinct tophits.rsid from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
+% select count(*) from (select distinct eqtl_gene_id from coloc where pp_h4_abf>=0.75) as sel;  7040
+% select count(*) from (select distinct eqtl_id from coloc where pp_h4_abf>=0.75) as sel;  127
 
@@
 In these diseases, the percentages of variants with a colocalized eQTL are 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST3).
-%
+% TODO
+% Number of tophits rsid per GWAS class
+% select count(distinct tophits.rsid),gwas_class from gwas_annot, tophits where tophits.gwas_id=gwas_annot.gwas_id group by gwas_class order by count desc;
+% Number of tophits rsid per GWAS class that is also preset in coloc table
+% select count(distinct tophits.rsid),gwas_class from gwas_annot, tophits, coloc where tophits.gwas_id=gwas_annot.gwas_id and tophits.rsid=coloc.rsid group by gwas_class order by count desc;
+% OLD SQLITE
+% select count(distinct tophits.rsid),gwas_class from gwas_annot join tophits on tophits.gwas_id=gwas_annot.gwas_id group by gwas_class order by count desc limit 5;
 %/* count tophits.rsid group by gwas_annot.gwas_class where coloc.`PP.H4.abf`>=0.8*/
@@
 %select gwas_annot.gwas_class, count(distinct tophits.rsid) from tophits, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id GROUP by gwas_annot.gwas_class order by gwas_annot.gwas_class;
-%
-These results agree with previous work for autoimmune diseases \citep{2021.Li.Mu}.
+
+These results are consistent with previous colocalization analysis that found explained percentages of work for autoimmune diseases \citep{2021.Li.Mu}.....
+% TODO
+% Discuss 25% in Chun et al 2017, Nature Genetics
+% Discuss 41% (eQTL + sQTLs) in Mu et al 2021, Genome Biology
+% Discuss 23% in Chen et al 2016, Cell. Genetic Drivers of Epigenetic and Transcriptional Variation in Human Immune Cells
 
 This pipeline resulted in 103,551 variants from 246 GWAS that colocalized (PP.H4.abf$\geq$0.8) with at least one eQTL.
-%select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.8);
-%select count(*) from (select DISTINCT rsid FROM coloc where `PP.H4.abf`>=0.8);
+% TODO
+%select count(coloc.rsid) from (select DISTINCT gwas_id FROM coloc where pp_h4_abf>=0.8);
+%select count(*) from (select distinct rsid from coloc where pp_h4_abf>=0.8) as sel;
 %
-I have also developed a web application where I expose 116,736 variants from 246 GWAS that colocalize with at least one eQTL at a lower threshold value PP.H4.abf$\geq$0.5.
-% select count(*) from (select DISTINCT rsid FROM coloc where `PP.H4.abf`>=0.5);
-% select count(*) from (select DISTINCT gwas_id FROM coloc where `PP.H4.abf`>=0.5);
+This might be a useful resource to annotate GWAS variants with genes and biosamples.
+Therefore I have developed a web application where I expose 116,736??? variants from 246??? GWAS that colocalize with at least one eQTL at a threshold value PP.H4.abf$\geq$0.75.
+% TODO
+% select count(*) from (select DISTINCT chrom,pos38,alt FROM colocweb) as sel;
+% select count(*) from (select DISTINCT gwas_id FROM coloc) as sel;
 
-%\subsection*{Pleiotropic variants colocalized with eQTLs}
-
+% Work with PP.H4.abf`>=0.75 and SNP.PP.H4>=0.5
 \subsection*{Traits classification based on the colocalized eQTL effect size}
 
+In the previous colocalization analysis, we found many eQTLs that are associated to very different traits.
+%
+To systematically identify pleiotropic eQTLs, I manually annotated the GWAS traits in different classes.
+
+% TODO Check if I aggregate different coloc windows or not.
+I also computed trait distances based on the spearman correlation of eQTL effect size for distinct eQTLs given by
+distinct chromosome, position, alternative allele, eQTL gene and eQTL biosample.
+%
+From this point on I defined colocalization with parameters PP.H4.abf>=0.75 and SNP.PP.H4>=0.5.
+%
+The argument for selecting a cutoff SNP.PP.H4>=0.5 is given below.
+%
+I plotted these trait distances as a hierarchically-clustered heatmap.
+%
+I plotted only trait classes with more than a minimum number of traits (Fig. \ref{fig:gwas_distance}).
+%
+Due to the large number of datasets, I have create a separate class for some cancers such as breast and skin.
+%
+This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases and cancers (Fig. \ref{fig:gwas_distance}).
+%
+
+Nevertheless, we can observe weaker correlations between the breast cancer class and other types of cancers.
+
+% TODO create supplementary table 2
 I manually assigned the 418 GWAS to 96 classes to aggregate identical or similar traits (Supplementary table 2).
@@
 %
-Altogether, pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
+Altogether, pleiotropic regions with 3 or Routinesmore trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
 

### 2022-12-18 scripts/insrt_gwas_annot.py

@@ -1,5 +1,3 @@
-import sqlalchemy
-
 from gwas2eqtl_pleiotropy.Logger import Logger
 from gwas2eqtl_pleiotropy.db2 import Base
 from sqlalchemy import create_engine
@@ -9,7 +7,6 @@ import sys
 
 
 #%%
-
 help_cmd_str = "todo"
 try:
     url = sys.argv[1]

### 2022-12-19 ms/ms03_results.tex
M

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Systematic colocalization of variants from genome-wide associations studies and expression quantitative trait loci}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % Work with PP.H4.abf`>=0.75 and SNP.PP.H4>=0.5
-\subsection*{Traits classification based on the colocalized eQTL effect size}
+\subsection*{Traits classification based on the effect size of eQTLs}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases and cancers (Fig. \ref{fig:gwas_distance}).
+This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases, cancers and allergies (Fig. \ref{fig:gwas_distance}).
 %
-
-Nevertheless, we can observe weaker correlations between the breast cancer class and other types of cancers.
+Breast and skin cancer are found in two different cluster with some shared signal.
+%
+Multiple sclerosis cluster with autoimmune diseases of the digestive system.
 
 % TODO create supplementary table 2
-I manually assigned the 418 GWAS to 96 classes to aggregate identical or similar traits (Supplementary table 2).
-%
-Then I verified whether GWAS traits cluster coherently within these classes.
-%
-I computed distances between GWAS traits based on the Spearman correlation of the colocalized eQTL beta for every target gene, cell type and tissue.
-%
-I plot only classes with more than a minimum number of GWAS traits (Fig. \ref{fig:gwas_distance}).
-%
-GWAS traits cluster together in autoimmune diseases and circulatory system diseases.
-%
-Due to the large number of breast cancer GWAS, we have created a separated class from cancer.
-Nevertheless, we can observe weaker correlations between the breast cancer class and other types of cancers.
-
-\subsection*{Which eQTLs are the most pleiotropic?}
+I manually assigned the 418 GWAS to 112 classes to aggregate identical or similar traits (Supplementary table 2).
 
@@
 %
-Therefore I manually defined 36 classes to aggregate eQTL cell types into tissues (Supplementary table 1).
+Therefore I manually defined 35 classes to aggregate eQTL cell types into tissues (Supplementary table 1).
 %
@@
 
-These class annotations were used to investigate the variant pleiotropy.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-Colocalized eQTL/GWAS variants were classified according to the number of GWAS classes they belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
+\subsection*{Which eQTLs are the most pleiotropic?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These trait class annotations were used to investigate pleiotropy of eQTLs.
+%
+Different trait classes of eQTLs were aggregated and sorted according to the number of GWAS classes the eQTLs belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
 %
@@
 %
-For instance, the variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1
+The most pleiotropic eQTLs are found in cytobands 1p13.2, 12q24.12 and 19.q13.33.
 %
+The eQTL rs2476601 is well known for rheumatoid arthritis, juvenile athritis, diabetes and cancer of skin REFs.
+%
+Our analysis annotates this variant with eQTL genes AP4B1,AP4B1-AS1,BCL2L15,DCLRE1B,MAGI3,PHTF1 and PTPN22 tissues Adipose,Brain,Colon,Fibroblast,ImmuneCell LCL and iPSC
+
+% TODO add rs2522051
 This variant rs2522051 is involved in six classes of traits, namely,
@@
 (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
+%
+The variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1
 %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Which genomic regions contain the most pleiotropic eQTLs?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %
-Therefore, I wanted to compute pleiotropic regions that include concentrations of pleiotropic eQTLs.
+Therefore, I wanted to compute pleiotropic regions with high concentration of pleiotropic eQTLs.
 %
-To compute these pleiotropic regions, I aggregated genomic regions with at least one pleiotropic eQTL in a sliding
-window of 100,000 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 5).
+To compute these pleiotropic regions, I use a sliding window to aggregate regions with pleiotropic eQTLs at distances
+of less than 1e5 nt (See Methods) (Table \ref{tab:pleiotropic_regions} and Supplementary table 5).
 %
-I found 13 regions with 5 or more trait classes, 30 regions with 4 or more trait
-classes and 80 regions with 3 or more trait classes (Supplementary table 5).
+% TODO add data
+I tested three SNP.H4.PP cutoffs, namely, 0.25, 0.5 and 0.75, to compute these regions.
 %
+% TODO
+For the SNP.PP.H4 cutoff 0.75, we obtain very few regions.
+%
+A less stringent cutoff 0.5, gives more pleiotropic regions, but still contains the pleiotropic regions found for a
+cutoff SNP.PP.H4 0.75.
+%
+A lower cutoff 0.25 gives more pleiotropic regions, but these regions are different than for the higher cutoffs.
+%
+Therefore, we fixed our analysis at SNP.PP.H4 0.5.
+
+%
+% TODO Verify values
+I found seven regions with seven or more trait classes, 30?? regions with 4?? or more trait
+classes and 80?? regions with 3?? or more trait classes (Supplementary table 5).
+%
+% TODO Verify values
 50\% of regions with at least two trait classes are shorter than 100 kb, 20\% of regions are
@@
 %
-Altogether, pleiotropic regions with 3 or Routinesmore trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
+Altogether, pleiotropic regions with 3 or more trait classes make around 3 Mb of the genome (Fig. \ref{fig:pleiotropy_region_distribution}b).
 
-\subsection*{How specific are eQTLs regarding traits, genes and tissues?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\subsection*{How specific are eQTLs for traits, genes and tissues?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Relation between trait, eQTL gene and tissue counts}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection*{Where are the pleiotropic variants?}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\subsection*{What is the impact of pleiotropic eQTLs?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Do pleiotropic eQTLs modulate more genes?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Are pleiotropic eQTLs active in more tissues?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \subsection*{Are pleiotropic eQTLs bound by more transcription factors?}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2022-12-19 ms/ms06_floats.tex
ENH improve MS
ENH PMIDs in disease matrix
M

@@
 %
-\includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr.png}
+\includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
 %
@@
 \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
-\textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs non-annotated.} \label{fig:freq_tf_per_variant}
+\textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.} \label{fig:freq_tf_per_variant}
 %

### 2022-12-19 gwas2eqtl_pleiotropy/EBIeQTLinfo.py

@@ -1,34 +1,0 @@
-import pdb
-
-from gwas2eqtl_pleiotropy.Logger import Logger
-from gwas2eqtl_pleiotropy.PathManager import PathManager
-from pandas import ExcelWriter
-import os
-import pandas
-import pathlib
-import requests
-
-from gwas2eqtl_pleiotropy.URL import URL
-from gwas2eqtl_pleiotropy.constants import eqtl_metadata_url
-
-
-class EBIeQTLinfo:
-
-    def __init__(self):
-
-        self._df = None
-
-    @property
-    def df(self):
-
-        """Loads file as DF with metadata compatible header"""
-
-        if self._df is None:
-
-            self.eqtl_tsv_path = URL(eqtl_metadata_url).download()
-            self._df = pandas.read_csv(self.eqtl_tsv_path, sep="\t")
-            eqtl_identifier_lst = \
-            (self._df['ftp_path'].str.replace('.all.tsv.gz', '', regex=True)).str.split('/', expand=True)[10].tolist()
-            self._df['identifier'] = eqtl_identifier_lst
-
-        return self._df

### 2022-12-19 gwas2eqtl_pleiotropy/PyTabix.py

@@ -1,31 +1,0 @@
-import subprocess
-from subprocess import Popen, PIPE
-
-def bgzip(filename):
-    """Call bgzip to compress a file."""
-    subprocess.run(['bgzip', '-f', filename])
-
-def tabix_index(filename, preset=None, chrom=None, start=None, end=None, skip=0, comment="#"):
-    command_lst = ['tabix', '-S', str(skip), '-c', comment]
-    if not preset is None:
-        command_lst.append('-p')
-        command_lst.append(preset)
-    if not chrom is None:
-        command_lst.append('-s')
-        command_lst.append(str(chrom))
-    if not start is None:
-        command_lst.append('-b')
-        command_lst.append(str(start))
-    if not end is None:
-        command_lst.append('-e')
-        command_lst.append(str(end))
-    command_lst.append(filename)
-    """Call tabix to create an index for a bgzip-compressed file."""
-    subprocess.run(command_lst)
-
-def tabix_query(filename, chrom, start, end):
-    """Call tabix and generate an array of strings for each line it returns."""
-    query = '{}:{}-{}'.format(chrom, start, end)
-    process = subprocess.run(['tabix', '-f', filename, query], stdout=PIPE)
-    for line in process.stdout:
-        yield line.strip().split()

### 2022-12-19 gwas2eqtl_pleiotropy/db.py

@@ -1,69 +1,90 @@
-from sqlalchemy import Table, Column, Integer, String, MetaData, Float, UniqueConstraint, ForeignKey
+from sqlalchemy import Column, Integer, String, SmallInteger, Float, UniqueConstraint
+from sqlalchemy.orm import declarative_base
+from sqlalchemy.dialects.postgresql import INT4RANGE
 
-meta = MetaData()
+# declarative base class
+Base = declarative_base()
 
-rsid2cytoband_tbl = Table(
-   'rsid2cytoband', meta,
-   Column('rsid', String(50), primary_key=True),
-   Column('cytoband', String(50), nullable=False),
-)
 
-ensg2symbol_tbl = Table(
-   'ensg2symbol', meta,
-   Column('gene_id', String(50), primary_key=True),
-   Column('symbol', String(50), nullable=False),
-)
+class pos19(Base):
+   """scripts/insrt_pos19.py"""
+   __tablename__ = "pos19"
+   __table_args__ = (UniqueConstraint('chrom', 'pos', name='_pos19_uc'),)
 
-gwas_annot_tbl = Table(
-   'gwas_annot', meta,
-   Column('gwas_id', String(50), primary_key=True),
-   Column('gwas_class', String(50), nullable=False),
-   Column('gwas_trait', String(50), nullable=False),
-)
+   id = Column('id', Integer, primary_key=True)
+   chrom = Column('chrom', SmallInteger, nullable=False)
+   pos = Column('pos', Integer, nullable=False)
+   pos19 = Column('pos19', Integer, nullable=True)
 
-coloc = Table(
-   'coloc', meta,
-   Column('id', Integer, primary_key = True),
-   Column('chrom', Integer, nullable=False),
-   Column('pos', Integer, nullable=False),
-   Column('rsid', String(50), ForeignKey("rsid2cytoband.rsid"), nullable=False),
-   Column('ref', String(50), nullable=False),
-   Column('alt', String(50), nullable=False),
-   Column('egene', String(50), ForeignKey("ensg2symbol.gene_id"), nullable=False),
-   Column('gwas_beta', Float, nullable=False),
-   Column('gwas_pval', Float, nullable=False),
-   Column('gwas_id', String(50), ForeignKey("gwas_annot.gwas_id"), nullable=False),
-   Column('eqtl_id', String(50), nullable=False),
-   Column('eqtl_beta', Float, nullable=False),
-   Column('eqtl_pval', Float, nullable=False),
-   Column('PP.H4.abf', Float, nullable=False),
[…]

### 2022-12-19 gwas2eqtl_pleiotropy/db2.py

@@ -1,90 +1,0 @@
-from sqlalchemy import Column, Integer, String, SmallInteger, Float, UniqueConstraint
-from sqlalchemy.orm import declarative_base
-from sqlalchemy.dialects.postgresql import INT4RANGE
-
-# declarative base class
-Base = declarative_base()
-
-
-class pos19(Base):
-   """scripts/insrt_pos19.py"""
-   __tablename__ = "pos19"
-   __table_args__ = (UniqueConstraint('chrom', 'pos', name='_pos19_uc'),)
-
-   id = Column('id', Integer, primary_key=True)
-   chrom = Column('chrom', SmallInteger, nullable=False)
-   pos = Column('pos', Integer, nullable=False)
-   pos19 = Column('pos19', Integer, nullable=True)
-
-
-class ensg2symbol(Base):
-   """scripts/insrt_gwas_annot.py"""
-   __tablename__ = "ensg2symbol"
-   gene_id = Column('gene_id', String(15), primary_key=True)
-   gene_symbol = Column('symbol', String(63), nullable=False, unique=False)
-
-
-class cytoband(Base):
-   """scripts/insrt_cytoband.py"""
-   __tablename__ = "cytoband"
-   __table_args__ = (UniqueConstraint('chrom', 'start_end38', name='_cytobad2_uc'),)
-
-   id = Column('id', String(15), primary_key=True)
-   chrom = Column('chrom', SmallInteger, nullable=False)
-   start_end38 = Column('start_end38', INT4RANGE, nullable=False)
-   cytoband = Column('cytoband', String(7), nullable=False)
-
-
-class gwas_annot(Base):
-   """scripts/insrt_gwas_annot.py"""
-   __tablename__ = "gwas_annot"
-
-   gwas_id = Column('gwas_id', String(63), primary_key=True)
-   gwas_trait = Column('gwas_trait', String(255), nullable=False)
-   gwas_class = Column('gwas_class', String(127), nullable=False)
-
-
-class eqtl_annot(Base):
-   """scripts/insrt_eqtl_annot.py"""
-   __tablename__ = "eqtl_annot"
-
-   eqtl_id = Column('eqtl_id', String(63), primary_key=True)
-   study = Column('study', String(63), nullable=False)
-   qtl_group = Column('qtl_group', String(63), nullable=False)
-   tissue_ontology_id = Column('tissue_ontology_id', String(63), nullable=False)
[…]

### 2022-12-19 scripts/plthtmp_disease_comorbidity_matrix.py

@@ -43,9 +43,9 @@ corr_df = corr_df.loc[corr_df.sum(axis=1) > 0, ]
 corr_df = corr_df[corr_df.columns[corr_df.sum(axis=0) > 0]]
 
 #%% filter row with at least a number of correlations >0.1 larger than 30
-corr_count_ser = (corr_df.abs() >= 0.001).sum(axis=1)
+corr_count_ser = (corr_df.abs() >= 0.05).sum(axis=1)
 # max number of correlation 80
-corr_count_ser = corr_count_ser.sort_values(ascending=False)[0:100]
+corr_count_ser = corr_count_ser.sort_values(ascending=False)[0:50]
 # mask = (corr_df >= 0.04).sum(axis=1) > 10
 mask = corr_df.index.isin(corr_count_ser.index)
 corr_df = corr_df.loc[mask]
@@ -70,7 +70,9 @@ annotation_df['pmid'] = annotation_df['pmid'].replace(math.nan, 0)
 annotation_df['pmid'] = annotation_df['pmid'].astype(int)
 annotation_df['trait'] = annotation_df['pmid'].astype(str) + '_' + annotation_df['trait']
 
-pmid_trait_mask = annotation_df['trait'].duplicated(keep='first')
+pmid_trait_dupli_mask = annotation_df['trait'].duplicated(keep='first')
+pmid_trait_uniq_mask = ~annotation_df['trait'].duplicated(keep=False)
+pmid_trait_mask = pmid_trait_dupli_mask + pmid_trait_uniq_mask
 dis_df = dis_df.loc[pmid_trait_mask, pmid_trait_mask]
 annotation_df = annotation_df.loc[pmid_trait_mask, ]
 
@@ -100,7 +102,7 @@ g = seaborn.clustermap(dis_df, **clustermap_args_dic)
 seaborn.set(font_scale=1)
 g.cax.set_visible(False)
 # import pdb; pdb.set_trace()
-g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize=10)
+g.ax_heatmap.set_yticklabels(g.ax_heatmap.get_ymajorticklabels(), fontsize=8)
 g.ax_heatmap.set(ylabel='Trait')
 #
 for label in class_labels.unique():

### 2022-12-22 ms/ms03_results.tex
MS

@@
 %
-% TODO Supplementary table 1 for newest pipeline
 127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many
@@
 %
-% TODO Supplementary table 2 for newest pipeline
 These filters resulted in 418 GWAS traits (Supplementary table 2).
 %
-Among these studies, there were 8,580 clumped leading variants with a p-value below 5e-8 from 284 GWAS (Supplementary file).
+These studies resulted in 8,580 clumped leading variants with a p-value below 5e-8 from 284 GWAS (Supplementary file).
 % select count(distinct gwas_id) from tophits; 284
@@
 My pipeline uncovered eQTL colocalizations for 5,484 or 64$\%$ of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
-% select count(*) from (select distinct tophits.rsid from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
+% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
 % select count(*) from (select distinct eqtl_gene_id from coloc where pp_h4_abf>=0.75) as sel;  7040
@@
 
-The largest classes of diseases by number of leading GWAS variants are autoimmune diseases (2107), cancer of breast (1976) and allergy (1321)
+The largest classes of diseases by number of leading GWAS variants are cancer of breast (1976), allergy (1288), and autoimmune diseases (1216).
 %
-In these diseases, the percentages of variants with a colocalized eQTL are 37\%, 21\% and 42\% in these classe, respectively (Supplementary table ST3).
-% TODO
+In these disease classes, the percentages of variants with a colocalized eQTL are 58\%, 61\% and 68\%, respectively (Supplementary table ST3).
+% Supplementary ST3
 % Number of tophits rsid per GWAS class
@@
 % select count(distinct tophits.rsid),gwas_class from gwas_annot, tophits, coloc where tophits.gwas_id=gwas_annot.gwas_id and tophits.rsid=coloc.rsid group by gwas_class order by count desc;
-% OLD SQLITE
-% select count(distinct tophits.rsid),gwas_class from gwas_annot join tophits on tophits.gwas_id=gwas_annot.gwas_id group by gwas_class order by count desc limit 5;
-%/* count tophits.rsid group by gwas_annot.gwas_class where coloc.`PP.H4.abf`>=0.8*/
-%select gwas_annot.gwas_class, count(distinct tophits.rsid) from tophits, coloc, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id and coloc.rsid=tophits.rsid and coloc.`PP.H4.abf`>=0.8 GROUP by gwas_annot.gwas_class;
-%/* count tophits.rsid group by gwas_annot.gwas_class*/
-%select gwas_annot.gwas_class, count(distinct tophits.rsid) from tophits, gwas_annot where gwas_annot.gwas_id=tophits.gwas_id GROUP by gwas_annot.gwas_class order by gwas_annot.gwas_class;
 
-These results are consistent with previous colocalization analysis that found explained percentages of work for autoimmune diseases \citep{2021.Li.Mu}.....
-% TODO
+These results is much higher than previous colocalization analyses for autoimmune diseases by previous authors that
+found 23-41\% depending on the study \citep{2016.Soranzo.Chen,2017.Cotsapas.Chun,2021.Li.Mu}.
 % Discuss 25% in Chun et al 2017, Nature Genetics
@@
 
-This pipeline resulted in 103,551 variants from 246 GWAS that colocalized (PP.H4.abf$\geq$0.8) with at least one eQTL.
+These data is a useful resource to annotate GWAS variants with genes and biosamples.
+Therefore I have developed a web application where I expose 138,301 variants from 293 GWAS that colocalize with at least
+one eQTL at a threshold value PP.H4.abf$\geq$0.75.
 % TODO
-%select count(coloc.rsid) from (select DISTINCT gwas_id FROM coloc where pp_h4_abf>=0.8);
-%select count(*) from (select distinct rsid from coloc where pp_h4_abf>=0.8) as sel;
-%
-This might be a useful resource to annotate GWAS variants with genes and biosamples.
-Therefore I have developed a web application where I expose 116,736??? variants from 246??? GWAS that colocalize with at least one eQTL at a threshold value PP.H4.abf$\geq$0.75.
-% TODO
-% select count(*) from (select DISTINCT chrom,pos38,alt FROM colocweb) as sel;
-% select count(*) from (select DISTINCT gwas_id FROM coloc) as sel;
+% select count(*) from (select DISTINCT chrom,pos38,alt FROM colocweb) as sel; 138301
+% select count(*) from (select DISTINCT gwas_id FROM colocweb) as sel; 293
 
@@
 
-% TODO Check if I aggregate different coloc windows or not.
-I also computed trait distances based on the spearman correlation of eQTL effect size for distinct eQTLs given by
-distinct chromosome, position, alternative allele, eQTL gene and eQTL biosample.
+Then I wanted to verify if different traits cluster coherently in different trait classes.
 %
-From this point on I defined colocalization with parameters PP.H4.abf>=0.75 and SNP.PP.H4>=0.5.
+As trait distance, I used the the spearman correlation of the eQTL splitted by
+chromosome, position, alternative allele, eQTL gene and eQTL biosample.
 %
-The argument for selecting a cutoff SNP.PP.H4>=0.5 is given below.
+To avoid non resolve eQTLs, I set a cutoff SNP.PP.H4>=0.5 in addition to the cutoff PP.H4.abf>=0.75.
+%
+The cutoff SNP.PP.H4>=0.5 was chosen, it allow to group pleiotropic eQTLs in a reasonable number of regions (See below).
 %
@@
 %
-I plotted only trait classes with more than a minimum number of traits (Fig. \ref{fig:gwas_distance}).
+Only trait classes with more than a minimum number of traits were plotted (Fig. \ref{fig:gwas_distance}).
 %
 Due to the large number of datasets, I have create a separate class for some cancers such as breast and skin.
+%
+In the OpenGWAS, some traits are redundant because the come from the same reference.
+%
+I plotted non-rendundant GWAS traits.
 %
@@
 
-% TODO create supplementary table 2
 I manually assigned the 418 GWAS to 112 classes to aggregate identical or similar traits (Supplementary table 2).
@@
 
+% TODO CONTINUE
 These trait class annotations were used to investigate pleiotropy of eQTLs.

### 2022-12-26 scripts/insrt_gwascatalog.py
ENH annotate GWAS based on OLS

@@ -1,0 +1,49 @@
+import os
+import pathlib
+
+import pandas
+import requests
+import sqlalchemy
+import sys
+
+from gwas2eqtl_pleiotropy.constants import public_data_dir
+from gwas2eqtl_pleiotropy.db import Base, gwascatalog
+from sqlalchemy import create_engine
+
+
+#%%
+help_cmd_str = "todo"
+try:
+    url = sys.argv[1]
+    if len(sys.argv) > 2:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+#%% Create all tables
+engine = create_engine(url)
+meta_data = sqlalchemy.MetaData(bind=engine)
+if sqlalchemy.inspect(engine).has_table("gwascatalog"):
+    gwascatalog.__table__.drop(engine)
+Base.metadata.create_all(engine)
+
+#%% Download
+url = "http://ftp.ebi.ac.uk/pub/databases/gwas/releases/2022/12/22/gwas-catalog-associations_ontology-annotated.tsv"
+tsv_path = os.path.join(public_data_dir, url.split('http://')[1])
+if not os.path.isfile(tsv_path):
+    pathlib.Path(os.path.dirname(tsv_path)).mkdir(exist_ok=True, parents=True)
+    resp = requests.get(url) # making requests to server
+    with open(tsv_path, "wb") as f: # opening a file handler to create new file
+        f.write(resp.content) # writing content to file
+
+df = pandas.read_csv(tsv_path, sep="\t", header=0, usecols=['PUBMEDID', 'STUDY', 'DISEASE/TRAIT', 'MAPPED_TRAIT', 'MAPPED_TRAIT_URI', 'STUDY ACCESSION'])
+df.drop_duplicates(inplace=True)
+df.rename(dict(zip(df.columns, [column.key for column in (gwascatalog.__table__.columns)][1:])), axis=1, inplace=True)
+df.to_sql('gwascatalog', con=engine, if_exists='append', index=False)
+
+
+

### 2023-01-06 ms/ms02_intro.tex
MS

@@
 %
-The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \citep{2007burton.barret,2018.Parkinson.Buniello}.
+The number of GWAS has rapidly increased since the first GWAS around 15 years ago and has resulted in comprehensive databases such as the NHGRI-EBI GWAS Catalog \cite{2007burton.barret,2018.Parkinson.Buniello}.
 %
-For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71,673 variant-trait associations from 3,567 publications \citep{2018.Parkinson.Buniello}.
+For instance, in 2019, NHGRI-EBI GWAS Catalog contained 5687 GWAS comprising 71,673 variant-trait associations from 3,567 publications \cite{2018.Parkinson.Buniello}.
 
 % Shortcomings of GWAS
-However, the interpretation of the GWAS variants is complex because of three reasons \citep{2020.Trynka.CanoGamez}.
+However, the interpretation of the GWAS variants is complex because of three reasons \cite{2020.Trynka.CanoGamez}.
 %
@@
 %
-Moreover, the distribution of eQTLs and GWAS variants in the genome are different because of negative selection \citep{2022.Pritchard.Mostafavi}.
+Moreover, the distribution of eQTLs and GWAS variants in the genome are different because of negative selection \cite{2022.Pritchard.Mostafavi}.
 %
-Nevertheless, eQTLs are still important data sets to interpret GWAS variants, because they provide gene regulatory information about the eQTL genes and cell types and tissues \citep{2021.Li.Mu}.
+Nevertheless, eQTLs are still important data sets to interpret GWAS variants, because they provide gene regulatory information about the eQTL genes and cell types and tissues \cite{2021.Li.Mu}.
 %
@@
 
-Many GWAS have allowed observing that many genetic variants are involved in different traits \citep{2019.Posthuma.Watanabe}.
+Many GWAS have allowed observing that many genetic variants are involved in different traits \cite{2019.Posthuma.Watanabe}.
 %
-In that work, the analyzed loci covered 60\% of the genome, and 90\% of these loci contained associations across different traits \citep{2019.Posthuma.Watanabe}.
+In that work, the analyzed loci covered 60\% of the genome, and 90\% of these loci contained associations across different traits \cite{2019.Posthuma.Watanabe}.
 %
-They also found that pleiotropic genes and variants were less tissue-specific in terms of gene expression and eQTLs \citep{2019.Posthuma.Watanabe}.
+They also found that pleiotropic genes and variants were less tissue-specific in terms of gene expression and eQTLs \cite{2019.Posthuma.Watanabe}.
 %
-However, the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \citep{2021.Li.Mu,2018.Vijayanand.Schmiedel,2017gtex.nature}.
+However, the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \cite{2021.Li.Mu,2018.Vijayanand.Schmiedel,2017gtex.nature}.
 %

### 2023-01-06 ms/ms03_results.tex
MS
ENH annotate GWAS based on OLS

@@
 127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many
-tissues and cell types \citep{2021.Alasoo.Kerimov} (Supplementary table 1).
+tissues and cell types \cite{2021.Alasoo.Kerimov} (Supplementary table 1).
 
-GWAS traits were selected from the IEU OpenGWAS database based on four criteria \citep{2018.Parkinson.Buniello}.
+GWAS traits were selected from the IEU OpenGWAS database based on four criteria \cite{2018.Parkinson.Buniello}.
 %
@@
 %
-These studies resulted in 8,580 clumped leading variants with a p-value below 5e-8 from 284 GWAS (Supplementary file).
-% select count(distinct gwas_id) from tophits; 284
-% select count(*) from (select DISTINCT chrom, pos, rsid, ea from tophits) as sel; 8,580
+These studies resulted in 10,621 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
+% select count(distinct gwas_id) from tophits; 335
+% select count(*) from (select DISTINCT chrom, pos, rsid, ea from tophits) as sel; 10621
 % tophits goes to OSF
@@
 My pipeline uncovered eQTL colocalizations for 5,484 or 64$\%$ of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
-% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
+% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid from tophits inner join coloc
+% on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
 % select count(*) from (select distinct eqtl_gene_id from coloc where pp_h4_abf>=0.75) as sel;  7040
@@
 These results is much higher than previous colocalization analyses for autoimmune diseases by previous authors that
-found 23-41\% depending on the study \citep{2016.Soranzo.Chen,2017.Cotsapas.Chun,2021.Li.Mu}.
+found 23-41\% depending on the study \cite{2016.Soranzo.Chen,2017.Cotsapas.Chun,2021.Li.Mu}.
 % Discuss 25% in Chun et al 2017, Nature Genetics
@@
 %
-I separated eQTLs according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence (\citep{2016.Cunningham.McLaren}).
+I separated eQTLs according to the count of GWAS classes and computed the EBI variant effect predictor (VEP) consequence (\cite{2016.Cunningham.McLaren}).
 %
@@
 
-Cis-regulatory modules (CRMs) are non-coding genomic regions with a higher density of bound cis-regulatory modules \citep{2021.Ballester.Hammal}.
+Cis-regulatory modules (CRMs) are non-coding genomic regions with a higher density of bound cis-regulatory modules \cite{2021.Ballester.Hammal}.
 %

### 2023-01-07 ms/ms06_floats.tex
ENH update

@@
 separator=tab,
-tabular=p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.45\textwidth},
+tabular=p{0.02\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.05\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.45\textwidth},
 head,
-table head=\bfseries Chrom & \bfseries Pleio RSID & \bfseries Pleio pos. & \bfseries Cytob. & \bfseries Start & \bfseries End & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Pleio RSID & \bfseries Pleio pos. & \bfseries GWAS Categories\\\hline,
 ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi\ & \csvcolvii}% specify your coloumns here
+{\csvcoli\ & \csvcolv\ & \csvcolvi\ & \csvcoliv\ & \csvcolii\ & \csvcoliii\ & \csvcolvii}% specify your coloumns here
 %\hline
@@
 \vspace{15pt}
-\caption{Pleiotropic regions involving 7 or more GWAS classes.
+\caption{Pleiotropic regions involving 6 or more GWAS classes.
 These regions were built with a sliding window of 1e5 nt.

### 2023-01-07 scripts/cmpt_pleiotropic_regions.py

@@ -41,7 +41,7 @@ df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
 
 pleio_df = pandas.DataFrame()
 df.sort_values(['chrom', 'pos38'], inplace=True)
-df['gwas_category_lst'] = df['gwas_category_lst'].str.split(',')
+df['gwas_category_lst'] = df['gwas_category_lst'].str.split(';')
 df['gwas_category_count'] = 1
 
 ########################
@@ -153,10 +153,10 @@ regions_pleio_ms_df = regions_pleio_ms_df.loc[regions_pleio_ms_df['gwas_category
 # format output
 # import pdb; pdb.set_trace()
 regions_pleio_ms_df.drop(['gwas_category_count'], inplace=True, axis=1)
-regions_pleio_ms_df['gwas_category_lst'] = regions_pleio_ms_df['gwas_category_lst'].apply(lambda x: ', '.join(x))
+regions_pleio_ms_df['gwas_category_lst'] = regions_pleio_ms_df['gwas_category_lst'].apply(lambda x: '; '.join(x))
 regions_pleio_ms_df['start'] = regions_pleio_ms_df['start'].apply(lambda x: '{0:,}'.format(x))
 regions_pleio_ms_df['end'] = regions_pleio_ms_df['end'].apply(lambda x: '{0:,}'.format(x))
-regions_pleio_ms_df['cytoband'] = regions_pleio_ms_df['cytoband'].apply(lambda x: ', '.join(x) )
+regions_pleio_ms_df['cytoband'] = regions_pleio_ms_df['cytoband'].apply(lambda x: '; '.join(x) )
 regions_pleio_ms_df = regions_pleio_ms_df[['chrom', 'pleio_rsid', 'pleio_pos38', 'cytoband', 'start', 'end', 'gwas_category_lst']]
 regions_pleio_ms_df.rename({'gwas_category_lst': 'GWAS categoryes'}, axis=1, inplace=True)
 

### 2023-01-07 scripts/cmpt_supplementary_files.py

@@ -102,16 +102,16 @@ for rowi, row in count_per_region_df.iterrows():
     start = row['start']
     end = row['end']
     eqtl_gene_symbol_ser = count_per_rsid_gwas_egene_etissue_df.query('chrom=={chrom} & pos38>={start} & pos38<={end}'.format(chrom=chrom, start=start, end=end))['eqtl_gene_symbol_lst']
-    eqtl_gene_symbol_lst = sorted(eqtl_gene_symbol_ser.str.split(', ').explode().unique())
+    eqtl_gene_symbol_lst = sorted(eqtl_gene_symbol_ser.str.split(';').explode().unique())
     count_per_region_df.loc[rowi, 'eqtl_gene_symbol'] = ', '.join(eqtl_gene_symbol_lst)
     #
     egene_ser = count_per_rsid_gwas_egene_etissue_df.query('chrom=={chrom} & pos38>={start} & pos38<={end}'.format(chrom=chrom, start=start, end=end))['egene_lst']
-    egene_lst = sorted(egene_ser.str.split(', ').explode().unique())
+    egene_lst = sorted(egene_ser.str.split(';').explode().unique())
     count_per_region_df.loc[rowi, 'egene'] = ', '.join(egene_lst)
     #
     # import pdb; pdb.set_trace()
     etissue_category_ser = count_per_rsid_gwas_egene_etissue_df.query('chrom=={chrom} & pos38>={start} & pos38<={end}'.format(chrom=chrom, start=start, end=end))['etissue_category_term_lst']
-    etissue_category_term_lst = sorted(etissue_category_ser.str.split(', ').explode().unique())
+    etissue_category_term_lst = sorted(etissue_category_ser.str.split(';').explode().unique())
     count_per_region_df.loc[rowi, 'etissue_category_term'] = ', '.join(etissue_category_term_lst)
 
 count_per_region_df.sort_values(by=['gwas_category_count', 'chrom', 'start'], ascending=[False, True, True], inplace=True)

### 2023-01-09 ms/ms04_discussion.tex
ENH PMIDs in disease matrix

@@
 More preciseley, in autoimmune diseases, we find colocalization for 37\% of leading GWAS variants, which is comparable
-to a previous study that found colocalization for 40.4\% of GWAS loci \citep{2021.Li.Mu}.
+to a previous study that found colocalization for 40.4\% of GWAS loci \cite{2021.Li.Mu}.
 %
@@
 We found 103,551 variants from 246 GWAS that colocalized with a posterior probability PP.H4.abf$\ge$0.8 with at least one eQTL.
-
-%TODO continue
 
@@
 
-I used these data sets here: eQTLs from the eQTL Catalogue \citep{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \citep{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \citep{2021.Ballester.Hammal}, and UCSC annotation data.
+I used these data sets here: eQTLs from the eQTL Catalogue \cite{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \cite{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \cite{2021.Ballester.Hammal}, and UCSC annotation data.
 %
-I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \citep{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
+I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \cite{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
 %

### 2023-01-10 ms/ms06_floats.tex
ENH added marker gene

@@
 \csvreader[separator=tab,
-tabular=ccrrp{0.4\textwidth},
+tabular=crclcp{0.4\textwidth},
 head,
-table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker & \bfseries GWAS Categories\\\hline,
 ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
 %\hline
@@
 %
-\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS classes. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
+\caption{eQTLs involved in four GWAS categories. The gene marker is the eQTL gene with highest PubMed publication count.}
+\label{tab:pleitropic_variants}
 \end{table}

### 2023-01-10 ms/ms06_floats.tex
MS

@@
 %
-\caption{eQTLs involved in four GWAS categories. The gene marker is the eQTL gene with highest PubMed publication count.}
+\caption{Example eQTLs involved in four GWAS categories in different cytobands.
+The whole list is given in the Supplementary files.
+The gene marker is the eQTL gene with the highest PubMed publication count.}
 \label{tab:pleitropic_variants}

### 2023-01-15 ms/ms06_floats.tex
ENH added marker genes

@@
 head,
-table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker eQTL gene& \bfseries GWAS Categories\\\hline,
 ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
@@
 separator=tab,
-tabular=p{0.02\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.05\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.45\textwidth},
+tabular=p{0.05\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.5\textwidth},
 head,
-table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Pleio RSID & \bfseries Pleio pos. & \bfseries GWAS Categories\\\hline,
+table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries GWAS Categories\\\hline,
 ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
-{\csvcoli\ & \csvcolv\ & \csvcolvi\ & \csvcoliv\ & \csvcolii\ & \csvcoliii\ & \csvcolvii}% specify your coloumns here
+{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
 %\hline

### 2023-01-15 scripts/cmpt_count_per_rsid.py

@@ -75,8 +75,17 @@ m_df = pandas.merge(m_df, egene_df, on=variant_def_lst)
 m_df = pandas.merge(m_df, etissue_df, on=variant_def_lst)
 m_df = pandas.merge(m_df, gene_pubmed_marker_df, on=variant_def_lst)
 m_df.sort_values(['gwas_category_count', 'gwas_ontology_term_count', 'chrom', 'pos38', 'rsid'], ascending=[False, False, True, True, True], inplace=True)
-columns = ['chrom', 'cytoband', 'pos38', 'rsid', 'alt', 'eqtl_gene_marker_symbol', 'pubmed_count', 'gwas_category_count', 'gwas_ontology_term_count', 'gwas_category_lst', 'gwas_ontology_term', 'egene_lst', 'eqtl_gene_symbol_lst', 'egene_count', 'etissue_category_term_lst', 'etissue_category_term_count', 'eqtl_gene_marker_id']
+columns = ['chrom', 'cytoband', 'pos38', 'rsid', 'alt', 'eqtl_gene_marker_symbol',
+           'pubmed_count', 'gwas_category_count', 'gwas_ontology_term_count',
+           'gwas_category_lst', 'gwas_ontology_term', 'egene_lst', 'eqtl_gene_symbol_lst', 'egene_count', 'etissue_category_term_lst', 'etissue_category_term_count', 'eqtl_gene_marker_id']
 m_df = m_df[columns]
+
+m_df['gwas_category_lst'] = m_df['gwas_category_lst'].apply(lambda x: ";".join(sorted([i for i in x if not i is None])))
+m_df['gwas_ontology_term'] = m_df['gwas_ontology_term'].apply(lambda x: ";".join(sorted([i for i in x if not i is None])))
+m_df['egene_lst'] = m_df['egene_lst'].apply(lambda x: ";".join(sorted([i for i in x if not i is None])))
+m_df['eqtl_gene_symbol_lst'] = m_df['eqtl_gene_symbol_lst'].apply(lambda x: ";".join(sorted([i for i in x if not i is None])))
+m_df['etissue_category_term_lst'] = m_df['etissue_category_term_lst'].apply(lambda x: ";".join(sorted([i for i in x if not i is None])))
+
 with pandas.ExcelWriter(count_per_rsid_gwas_egene_etissue_ods, engine="odf") as writer:
     m_df.to_excel(writer, index=False)
 
[…]

### 2023-01-15 scripts/pltsctr_x_per_rsid_y_gwas.py

@@ -1,16 +1,16 @@
-import sys
+"""Scatter of gwas count vs variant rsid"""
 
-import seaborn
-
-from gwas2eqtl_pleiotropy.PathManager import PathManager
 from matplotlib import pyplot as plt
 from gwas2eqtl_pleiotropy.constants import tick_fontsize, label_fontsize, scatter_dot_size, dpi
 
+import sys
+import seaborn
 import os
 import pandas
 import pathlib
 
 plt.rcParams["figure.figsize"] = (8, 6)
+
 from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 seaborn.set_theme(**seaborn_theme_dic)
 
@@ -18,8 +18,8 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     max_gwas_category_count = int(sys.argv[1])
-    region_window_100000_tsv_path = sys.argv[2]
-    count_per_rsid_gwas_tsv_path = sys.argv[3]
+    region_window_100000_ods_path = sys.argv[2]
+    count_per_rsid_gwas_ods_path = sys.argv[3]
     outdir_path = sys.argv[4]
     if len(sys.argv) > 5:
         print("""Two many arguments!
@@ -34,11 +34,13 @@ except IndexError:
 basename_str = "count_per_rsid_gwas.tsv"
 # cmpt_count_per_rsid_dir_path = os.path.join(PathManager.get_project_path(), "out", "cmpt_count_per_rsid.py")
 # tsv_path = os.path.join(cmpt_count_per_rsid_dir_path, basename_str)
-count_per_rsid_df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
+# count_per_rsid_df = pandas.read_csv(count_per_rsid_gwas_tsv_path, sep="\t")
+count_per_rsid_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine='odf')
 
 #%% input regions
 # region_window_100000_tsv_path = os.path.join(PathManager.get_outdir_path(), "cmpt_pleiotropic_regions.py", "region_window_100000.tsv")
-region_window_100000_df = pandas.read_csv(region_window_100000_tsv_path, sep="\t")
+# region_window_100000_df = pandas.read_csv(region_window_100000_tsv_path, sep="\t")
+region_window_100000_df = pandas.read_excel(region_window_100000_ods_path, engine='odf')
 
 # #%% Output
 # if not '__file__' in locals():

### 2023-01-15 presentation_230120_internal_seminar/intern_sem_230120.tex

\documentclass{beamer}
\usetheme{Boadilla}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{csvsimple}

% Set relative path to all figures
\newcommand*{\floatRelativePath}{../out/gwas417/pval_5e-08/r2_0.1/kb_1000/window_1000000/75_50}%
% Footnote without marker: https://tex.stackexchange.com/questions/30720/footnote-without-a-marker
\newcommand\blfootnote[1]{%
    \begingroup
    \renewcommand\thefootnote{}\footnote{#1}%
    \addtocounter{footnote}{-1}%
    \endgroup
}
% Change size of footnotes
\renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
% reset footnote counter for each frame in beamer class
\AtBeginEnvironment{frame}{\setcounter{footnote}{0}}

\title{Analysis of pleiotropy of eQTLs}
\subtitle{Internal Seminar}
\author{Aitor Gonz\'alez}
\institute{Aix Marseille Univ, INSERM, TAGC}
\date{Jan. 20, 2023}

% Add section slide
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}

        \titlepage

    \end{frame}

    \section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Genome-wide association studies (GWAS)}

        \begin{itemize}
            \item Genetic alleles with higher frequency in individuals with a given disease
            \item A very active field with many studies
            \item Many genetic variants are involved in several phenotypes
        \end{itemize}
%
        \vfill
%
        Questions:
%
        \begin{itemize}
            \item Correlation between variants. Which is the causal variant?
            \item Which cell type important for disease?
            \item Over 90\% GWAS variants are non-coding. What is their role?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Expression quantitative trait loci (eQTLs)}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}

            \end{column}
        \end{columns}

        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Annotation of GWAS variants with gene and tissue targets using colocalization analysis}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
                \end{center}
            \end{column}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4bc.png}
                \end{center}
            \end{column}
        \end{columns}

        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Objectives}

        \begin{enumerate}
            \item Large scale gene and tissue annotation of variants of complex traits using GWAS/eQTL colocalization analysis
            \item To identify pleiotropic eQTLs and eQTL-containing genomic regions
            \item To characterize the properties of pleiotropic eQTLs
            \item To share colocalizations with experimentalists
        \end{enumerate}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Data and Algorithms}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{itemize}
                    \item IEA OpenGWAS database\footnote{https://gwas.mrcieu.ac.uk}: 418 GWAS datasets
                    \item EBI eQTLs database\footnote{https://www.ebi.ac.uk/eqtl}: Processed eQTLs from 127 studies in human.
                    \item CRAN - Package coloc\footnote{https://cran.r-project.org/web/packages/coloc/index.html}: Colocalisation analysis of GWAS/eQTL with one causal variant
                \end{itemize}
            \end{column}
            \begin{column}{0.5\textwidth}
                \begin{center}
                    \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/pgen.1004383.g001.png}
                \end{center}
            \end{column}
        \end{columns}

%        \let\thefootnote\relax\footnotetext{$^1$https://gwas.mrcieu.ac.uk}
%        \let\thefootnote\relax\footnotetext{$^2$https://www.ebi.ac.uk/eqtl}
%        \let\thefootnote\relax\footnotetext{$^3$https://cran.r-project.org/web/packages/coloc/index.html}
    \end{frame}

    \section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Large scale GWAS/eQTL colocalization analysis}

        \begin{itemize}
            \item Colocalization between 293 GWAS and 127 eQTL studies
            \item 138e3 colocalized variants with probability $\geq$ 0.75
            \item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
        \end{itemize}

        \begin{figure}[!]
            \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/coloc_web.png}
        \end{figure}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Do GWAS traits cluster coherently?}

        GWAS distance based on correlation of beta of colocalized eQTL gene and tissue

        \begin{figure}[!]
            \includegraphics[width=0.5\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
        \end{figure}

        \begin{itemize}
            \item Categories are coherent
            \item Help categorize GWAS traits
        \end{itemize}

    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%\frametitle{Diseases?}
%
%\begin{table}[!tbp]
%\centering
%\scriptsize
%\hline
%\csvreader[separator=tab,
%tabular=ccrrp{0.4\textwidth},
%head,
%table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
%]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
%{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
%\hline
%%
%\vspace{15pt}
%%
%\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
%\end{table}
%
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{How are pleiotropic eQTLs defined?}

        \begin{itemize}
            \item Uniformly annotate GWAS traits based on clustering and EBI ontologies
            \item Aggregate colocalised eQTLs based on the number of trait categories
            \item Analyse eQTLs in groups based on the number of trait categories
        \end{itemize}

    \end{frame}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Winner (Most pleiotroic) eQTLs?}

        \begin{itemize}
            \item Example eQTLs involved in four GWAS categories in different cytobands
            \item The gene marker is the eQTL gene with the highest PubMed publication count
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
            \csvreader[separator=tab,
            tabular=crclcp{0.4\textwidth},
            head,
            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
            \vspace{15pt}
        \end{table}

    \end{frame}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Winner (Most pleiotroic) regions?}

        \begin{itemize}
            \item Pleiotropic regions involving 6 or more GWAS classes
            \item Regions were built with a sliding window of 1e5 nt
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
            \centering
            \footnotesize
%\hline
            \csvreader[
                separator=tab,
                tabular=p{0.02\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.05\textwidth}p{0.45\textwidth},
                head,
                table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries GWAS Categories\\\hline,
            ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
            {\csvcoli\ & \csvcolv\ & \csvcolvi\ & \csvcoliv\ & \csvcolvii}% specify your coloumns here
%\hline
%
            \vspace{15pt}
        \end{table}
    \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%    \frametitle{How are pleiotropic variants distributed?}
%
%    \begin{columns}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.32\textwidth}
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
%            \end{center}
%        \end{column}
%    \end{columns}
%
%    \begin{center}
%        \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/chr5_131912097_132802472.png}
%    \end{center}
%
%    \small
%    \begin{itemize}
%        \item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
%        \item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
%        \item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
%    \end{itemize}
%    \normalsize
%
%\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Where are the pleiotropic variants?\footnote{https://www.ensembl.org/vep}}

            \begin{center}
                \includegraphics[width=0.7\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
            \end{center}

            In the transcripts

%        \let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Do pleiotropic variants associate to more eQTL genes?}

            \begin{columns}
                \begin{column}{0.5\textwidth}
                    \begin{center}
                        \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_egenes.png}
                    \end{center}
                \end{column}
                \begin{column}{0.5\textwidth}  %%<--- here
                    \begin{center}
                        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
                    \end{center}
                \end{column}
            \end{columns}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Do pleiotropic variants associate to more eQTL tissues?}

            \begin{columns}
                \begin{column}{0.5\textwidth}
                    \begin{center}
                        \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_etissues.png}
                    \end{center}
                \end{column}
                \begin{column}{0.5\textwidth}  %%<--- here
                    \begin{center}
                        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
                    \end{center}
                \end{column}
            \end{columns}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Do pleiotropic variants bind more transcription factors?\footnote{https://remap.univ-amu.fr}}

            \begin{columns}
                \begin{column}{0.5\textwidth}
                    \begin{center}
                        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
                    \end{center}
                \end{column}
                \begin{column}{0.5\textwidth}  %%<--- here
                    \begin{center}
                        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
                    \end{center}
                \end{column}
            \end{columns}

%        \let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}

        \end{frame}

        \section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Concluding remarks}

            Outcomes interesting for the GOLD population genotypes
%
            \begin{itemize}
                \item A large set of causal GWAS variants annotated with target genes and tissues
                \item A list of pleiotropic variants and regions
            \end{itemize}
%
            \vfill
%
            What are the gene regulatory properties underlying pleiotropy between complex traits?
%
            \begin{itemize}
                \item Variant pleiotropy is correlated in regions
                \item Enrichment of variants upstream (incl. 5'UTR) and downstream of genes (incl. 3'UTR)
                \item Enrichment of transcription factor and CRM binding
                \item Increased number of associated egenes and etissues
            \end{itemize}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Perspectives}

            \begin{itemize}
                \item A web portail to share data
                \item Causality of eQTL genes on GWAS (Looking for experts ;-) )
            \end{itemize}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{Acknowledgements}

            \begin{itemize}
                \item L\'eopoldine Lecerf (M1)
                \item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
            \end{itemize}
%
            \vfill
%
            Funding
%
            \begin{itemize}
                \item Institut Cancer et Immunologie - Aix-Marseille Univ.
                \item Agence nationale de la recherche (ANR)
                \item GOLD
            \end{itemize}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \begin{frame}
            \frametitle{}

        \end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}
%    \frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
%
%    \begin{columns}
%        \begin{column}{0.5\textwidth}
%            \begin{center}
%                \includegraphics[width=0.7\textwidth]{../presentation_230120_gold2022_paris/fig/model_pleio_gwas.png}
%            \end{center}
%        \end{column}
%        \begin{column}{0.5\textwidth}  %%<--- here
%            \begin{center}
%                \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
%            \end{center}
%        \end{column}
%    \end{columns}
%
%\end{frame}

    \end{document}

### 2023-01-16 presentation_230120_internal_seminar/intern_sem_230120.tex
internal seminar

@@
             head,
-            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker & \bfseries GWAS Categories\\\hline,
+            table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
             ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
@@
 
-        \begin{itemize}
-            \item Pleiotropic regions involving 6 or more GWAS classes
-            \item Regions were built with a sliding window of 1e5 nt
-        \end{itemize}
-
 % full size table is table
@@
                 separator=tab,
-                tabular=p{0.02\textwidth}p{0.09\textwidth}p{0.09\textwidth}p{0.05\textwidth}p{0.45\textwidth},
+                tabular=p{0.005\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.05\textwidth}p{0.5\textwidth},
                 head,
-                table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries GWAS Categories\\\hline,
+                table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Known eQTL gene & \bfseries GWAS Categories\\\hline,
             ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
-            {\csvcoli\ & \csvcolv\ & \csvcolvi\ & \csvcoliv\ & \csvcolvii}% specify your coloumns here
+            {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
 %\hline
@@
             \vspace{15pt}
+            \caption{Pleiotropic regions involving 6 or more GWAS classes.
+These regions were built with a sliding window of 1e5 nt.
+Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
         \end{table}
+
     \end{frame}

### 2023-01-18 presentation_230120_internal_seminar/intern_sem_230120.tex
DOC

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%    The associations to diseases are heterougeneously distributed in the genome, Watanabe & Posthuma
+%    This is based on loci associated to multiple traits
     \begin{frame}
@@
         \begin{itemize}
-            \item Genetic alleles with higher frequency in individuals with a given disease
-            \item A very active field with many studies
-            \item Many genetic variants are involved in several phenotypes
+            \item Association test of effect allele on case individuals
+            \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
+            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories
         \end{itemize}
@@
         \vfill
-%
+%          However, how do we know if these variants are active in a tissue,
+%         or whether these variants false positives due to linkage disequilibrium,
+%         or how these variants act molecularly
         Questions:
@@
             \item Correlation between variants. Which is the causal variant?
-            \item Which cell type important for disease?
-            \item Over 90\% GWAS variants are non-coding. What is their role?
+            \item Which cell types important for disease?
+            \item In which tissues are these variants active?
         \end{itemize}
@@
 
+        \blfootnote{}
         \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}

### 2023-01-19 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

@@
         \vfill
+
 %          However, how do we know if these variants are active in a tissue,
@@
         \frametitle{Expression quantitative trait loci (eQTLs)}
-
+%eQTLs are variants that modify gene expression in a given tissue.
+%        eQTLs give a potential mechanism (the gene target) and a tissue.
         \begin{columns}
@@
         \end{columns}
+%
+        \vfill
+%
+%        To my knowledge, it is unknown where are the most pleiotropice Eqtls in the genome.
+%        This is important because this would give hints to putatitve pleiotropi genes (unknown or unknown) that are regulated by these pleiotropic eQTLs?
+        Pleiotropy of eQTLs in genome is currently unknown
 
         \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Objectives}
+
+        \begin{enumerate}
+            \item To look for pleiotropic eQTLs in the genome
+            \item To characterise their genomic regions, target genes and the cause of their pleiotropy
+        \end{enumerate}
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Strategy}
+
+        \begin{enumerate}
+            \item Annotate eQTLs with GWAS traits using colocalization analysis
+            \item Annotate GWAS traits in categories
+            \item Aggregate trait categories of eQTLs: Pleiotropic eQTLs
+            \item Aggregate pleiotropic eQTLs in genome regions
+            \item Analyse pleiotropic eQTLs and regions
+        \end{enumerate}
     \end{frame}
@@
         \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
-    \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \begin{frame}
-        \frametitle{Objectives}
-
-        \begin{enumerate}
-            \item Large scale gene and tissue annotation of variants of complex traits using GWAS/eQTL colocalization analysis
-            \item To identify pleiotropic eQTLs and eQTL-containing genomic regions
-            \item To characterize the properties of pleiotropic eQTLs
-            \item To share colocalizations with experimentalists
-        \end{enumerate}
     \end{frame}
@@
 
+%In this project, I have carried a large colocalization analysis between ~300 GWAS traits and 127 eQTL studies belonging to different tissues and mainly primary cell types.
+%In addition, by colocalizing statistically eQTLs and GWAS, we can to some extent, break, linkage disequilibrium
+
         \begin{itemize}
@@
     \begin{frame}
+        \frametitle{Web portail}
+
+%These data can be useful to interpret disease variants
+%        Therefore I have developed a small web portail
+
+        TODO
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
         \frametitle{Do GWAS traits cluster coherently?}
+
+%        To evaluate the coherence of the colocalization data, I have computed distances between traits based on the Spearman correlation of the betas of the eqtl-gene-tissue
 
@@
         \end{itemize}
+
+%        In this heatmap, we can see that similar traits cluster together
+%One difficulty for analysing pleiotropy is to distinguish similar traits for really different traits.
+%In the previous heatmap, we have seen that traits cluster around coherent categories
+%To look for real pleiotropic eQTLs I have manually annotated traits in different categories, such as cardiovascular disease, allergic, autoimmune, ....
 
@@
             \item Example eQTLs involved in four GWAS categories in different cytobands
-            \item The gene marker is the eQTL gene with the highest PubMed publication count
+            \item The representative gene is the eQTL gene with the highest PubMed publication count
         \end{itemize}
@@
                 {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-            \vspace{15pt}
+            \vspace{15pt}\label{tab:pleitropic_variants}
         \end{table}
@@
     \begin{frame}
-        \frametitle{Concluding remarks}
+        \frametitle{Conclusion}
+
+        \begin{itemize}
+            \item A list of pleiotrtA large set of causal GWAS variants annotated with target genes and tissues
+            \item A list of pleiotropic variants and regions
+        \end{itemize}
 
@@
         \begin{itemize}
-            \item A web portail to share data
-            \item Causality of eQTL genes on GWAS (Looking for experts ;-) )
+            \item Causality of eQTL genes on GWAS
+            \item Study loci in more
         \end{itemize}

### 2023-01-19 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

@@
 %         or how these variants act molecularly
-        Questions:
+        Questions:\footnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
 %
@@
 
-        \blfootnote{}
-        \blfootnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
     \end{frame}
@@
 %        eQTLs give a potential mechanism (the gene target) and a tissue.
-        \begin{columns}
-            \begin{column}{0.5\textwidth}
+
         \begin{center}
@@
         \end{center}
-            \end{column}
-            \begin{column}{0.5\textwidth}
-
-            \end{column}
-        \end{columns}
-%
-        \vfill
 %
@@
         \begin{enumerate}
-            \item To look for pleiotropic eQTLs in the genome
+            \item To identify pleiotropic eQTLs in the genome
             \item To characterise their genomic regions, target genes and the cause of their pleiotropy
@@
         \begin{enumerate}
-            \item Annotate eQTLs with GWAS traits using colocalization analysis
-            \item Annotate GWAS traits in categories
-            \item Aggregate trait categories of eQTLs: Pleiotropic eQTLs
+            \item Annotate eQTLs with GWAS traits (Colocalization analysis)
+            \item Define categories for GWAS traits
+            \item Aggregate categories of eQTLs: Pleiotropic eQTLs
             \item Aggregate pleiotropic eQTLs in genome regions
-            \item Analyse pleiotropic eQTLs and regions
+            \item Explore properties pleiotropic eQTLs and regions
         \end{enumerate}
@@
     \begin{frame}
-        \frametitle{Annotation of GWAS variants with gene and tissue targets using colocalization analysis}
+        \frametitle{Colocalization to annotate eQTLs with GWAS traits}
 
@@
     \begin{frame}
-        \frametitle{Data and Algorithms}
+        \frametitle{IEA OpenGWAS database\footnote{https://gwas.mrcieu.ac.uk}}
 
-        \begin{columns}
-            \begin{column}{0.5\textwidth}
         \begin{itemize}
-                    \item IEA OpenGWAS database\footnote{https://gwas.mrcieu.ac.uk}: 418 GWAS datasets
-                    \item EBI eQTLs database\footnote{https://www.ebi.ac.uk/eqtl}: Processed eQTLs from 127 studies in human.
-                    \item CRAN - Package coloc\footnote{https://cran.r-project.org/web/packages/coloc/index.html}: Colocalisation analysis of GWAS/eQTL with one causal variant
+            \item 418 GWAS
+            \item Physiological and disease traits
+            \item UK biobank, GWAS catalog, Other datasets
         \end{itemize}
-            \end{column}
-            \begin{column}{0.5\textwidth}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{EBI eQTLs database\footnote{https://www.ebi.ac.uk/eqtl}}
+
+        \begin{itemize}
+            \item 127 studies in human from different datasets
+            \item Tissues, cell types, stimulated conditions, ...
+            \item GTEX, DICE, ...
+        \end{itemize}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{CRAN - Package coloc\footnote{https://cran.r-project.org/web/packages/coloc/index.html}}
+
+        \begin{itemize}
+            \item Colocalisation analysis of GWAS/eQTL with one causal variant
+        \end{itemize}
+
         \begin{center}
@@
         \end{center}
-            \end{column}
-        \end{columns}
 
-%        \let\thefootnote\relax\footnotetext{$^1$https://gwas.mrcieu.ac.uk}
-%        \let\thefootnote\relax\footnotetext{$^2$https://www.ebi.ac.uk/eqtl}
-%        \let\thefootnote\relax\footnotetext{$^3$https://cran.r-project.org/web/packages/coloc/index.html}
     \end{frame}
@@
             \item Colocalization between 293 GWAS and 127 eQTL studies
+            \item 25\% of 10 k leading SNPs (SNP.PP.H4$\geq$0.5), colocalized with at least one eQTL
+            \item At SNP.PP.H4$\geq$0.5, 5 k eQTLs and 5 k target genes
             \item 138e3 colocalized variants with probability $\geq$ 0.75
@@
         \begin{figure}[!]
-            \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/coloc_web.png}
+            \includegraphics[width=0.5\textwidth]{fig/histogram_explained.png}\label{fig:perc_explained}
         \end{figure}
-
-    \end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \begin{frame}
-        \frametitle{Web portail}
-
-%These data can be useful to interpret disease variants
-%        Therefore I have developed a small web portail
-
-        TODO
 
@@
     \begin{frame}
-        \frametitle{Winner (Most pleiotroic) eQTLs?}
+        \frametitle{Most pleiotroic eQTLs}
 
@@
     \begin{frame}
-        \frametitle{Winner (Most pleiotroic) regions?}
+        \frametitle{Most pleiotroic regions}
 
@@
         \end{table}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{The IRF1/IL4 locus}
+
+        \begin{figure}[!]
+            \includegraphics[width=0.95\textwidth]{fig/il4}
+        \end{figure}
+
+        \begin{itemize}
+            \item Sevaral immune genes
+            \item GWAS categories: allergic disease; autoimmune disease; body height; breast cancer; cancer; cardiovascular disease; hypertensive disorder; respiratory system disease; skin disease
+            \item eQTL biosamples: adipose tissue;aorta;blood;brain;breast; ...
+        \end{itemize}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{The CDKN2A/P16 locus}
+
+        \begin{figure}[!]
+            \includegraphics[width=0.95\textwidth]{fig/cdk2na}
+        \end{figure}
+
+        \begin{itemize}
+            \item Cell cycle regulator
+            \item GWAS categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
+            \item eQTL biosamples: adipose tissue;blood;brain;colon;immune system; ...
+        \end{itemize}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{The TP53 locus}
+
+        \begin{figure}[!]
+            \includegraphics[width=0.95\textwidth]{fig/cdk2na}
+        \end{figure}
+
+        \begin{itemize}
+            \item Cell cycle regulator
+            \item GWAS categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
+            \item eQTL biosamples: adipose tissue;adrenal gland;aorta;blood;brain; ...
+        \end{itemize}
 
@@
 
-    \section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Web access}
+
+%These data can be useful to interpret disease variants
+%        Therefore I have developed a small web portail
+
+        \begin{figure}[!]
+            \includegraphics[width=\textwidth]{fig/web_portal.png}\label{fig:coloc_web}
+        \end{figure}
+
+    \end{frame}
+
+    \section{Concluding remarks} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
         \begin{itemize}
-            \item A list of pleiotrtA large set of causal GWAS variants annotated with target genes and tissues
-            \item A list of pleiotropic variants and regions
-        \end{itemize}
-
-        Outcomes interesting for the GOLD population genotypes
-%
-        \begin{itemize}
-            \item A large set of causal GWAS variants annotated with target genes and tissues
-            \item A list of pleiotropic variants and regions
-        \end{itemize}
-%
-        \vfill
-%
-        What are the gene regulatory properties underlying pleiotropy between complex traits?
-%
-        \begin{itemize}
-            \item Variant pleiotropy is correlated in regions
-            \item Enrichment of variants upstream (incl. 5'UTR) and downstream of genes (incl. 3'UTR)
-            \item Enrichment of transcription factor and CRM binding
+            \item A large set of GWAS variants annotated with putative target genes and tissues
+            \item A list of pleiotropic eQTLs and regions
+            \item Pleiotropic eQTL region include immune and cell cycle regulators
+            \item Enrichment of 5' UTR and transcript (missense) variants
             \item Increased number of associated egenes and etissues
+            \item A web site to access the data
         \end{itemize}
@@
         \begin{itemize}
-            \item Causality of eQTL genes on GWAS
-            \item Study loci in more
+            \item Verify causality of eQTL genes on GWAS
+            \item Analyze pleiotropic loci in more detail
         \end{itemize}

### 2023-01-19 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

@@
 \begin{frame}
+    \frametitle{How are pleiotropic variants distributed?}
+
+    \begin{columns}
+        \begin{column}{0.32\textwidth}
+            \begin{center}
+                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start132239645_end132497907_categories9.png}
+            \end{center}
+        \end{column}
+        \begin{column}{0.32\textwidth}
+            \begin{center}
+                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start132239645_end132497907_categories9.png}
+            \end{center}
+        \end{column}
+        \begin{column}{0.32\textwidth}
+            \begin{center}
+                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start132239645_end132497907_categories9.png}
+            \end{center}
+        \end{column}
+    \end{columns}
+
+    \begin{center}
+        \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/chr5_131912097_132802472.png}
+    \end{center}
+
+    \small
+    \begin{itemize}
+        \item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
+        \item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
+        \item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
+    \end{itemize}
+    \normalsize
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Does trait categories correlate with eQTL and biosample counts?}
+
+        \begin{figure}[!]
+            \includegraphics[width=0.5\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr}\label{fig:count_per_rsid_gwas_egene_etissue_corr}
+        \end{figure}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
         \frametitle{Most pleiotroic regions}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \begin{frame}
-        \frametitle{The IRF1/IL4 locus}
-
-        \begin{figure}[!]
-            \includegraphics[width=0.95\textwidth]{fig/il4}
-        \end{figure}
-
-        \begin{itemize}
-            \item Sevaral immune genes
-            \item GWAS categories: allergic disease; autoimmune disease; body height; breast cancer; cancer; cardiovascular disease; hypertensive disorder; respiratory system disease; skin disease
-            \item eQTL biosamples: adipose tissue;aorta;blood;brain;breast; ...
-        \end{itemize}
-
-    \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%    \begin{frame}
+%        \frametitle{The IRF1/IL4 locus}
+%
+%        \begin{figure}[!]
+%            \includegraphics[width=0.95\textwidth]{fig/il4}
+%        \end{figure}
+%
+%        \begin{itemize}
+%            \item Sevaral immune genes
+%            \item GWAS categories: allergic disease; autoimmune disease; body height; breast cancer; cancer; cardiovascular disease; hypertensive disorder; respiratory system disease; skin disease
+%            \item eQTL biosamples: adipose tissue;aorta;blood;brain;breast; ...
+%        \end{itemize}
+%
+%    \end{frame}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{frame}
-%    \frametitle{How are pleiotropic variants distributed?}
-%
-%    \begin{columns}
-%        \begin{column}{0.32\textwidth}
-%            \begin{center}
-%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-%            \end{center}
-%        \end{column}
-%        \begin{column}{0.32\textwidth}
-%            \begin{center}
-%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-%            \end{center}
-%        \end{column}
-%        \begin{column}{0.32\textwidth}
-%            \begin{center}
-%                \includegraphics[width=\textwidth]{\floatRelativePath/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-%            \end{center}
-%        \end{column}
-%    \end{columns}
-%
-%    \begin{center}
-%        \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/chr5_131912097_132802472.png}
-%    \end{center}
-%
-%    \small
-%    \begin{itemize}
-%        \item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
-%        \item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
-%        \item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
-%    \end{itemize}
-%    \normalsize
-%
-%\end{frame}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \begin{frame}
-        \frametitle{Do pleiotropic variants associate to more eQTL genes?}
+        \frametitle{Are pleiotropic variants associated to more eQTL genes?}
 
@@
     \begin{frame}
-        \frametitle{Do pleiotropic variants associate to more eQTL tissues?}
+        \frametitle{Are pleiotropic variants associated to more eQTL tissues?}
 

### 2023-01-20 presentation_230120_internal_seminar/intern_sem_230120.tex
ENH

@@
 
-\title{Analysis of pleiotropy of eQTLs}
+\title{Discovery and analysis of pleiotropic eQTLs}
 \subtitle{Internal Seminar}
@@
     \begin{frame}
-        \frametitle{Genome-wide association studies (GWAS)}
+        \frametitle{GWAS and pleiotropy}
 
         \begin{itemize}
-            \item Association test of effect allele on case individuals
-            \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
-            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories
+            \item Genome-wise association studies: Association test in the population of variants on common traits, eg. height or Crohn's disease
+            \item Used for frequent variants and common quantitative or qualitative traits
+            \item High-throughput genetics: 3000 unique loci for over thousands of disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
+            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories\footnote{Watanabe et al 2019. 10.1038/s41588-019-0481-0}
         \end{itemize}
 %
-        \vfill
+%         \vfill
 
@@
 %         or how these variants act molecularly
-        Questions:\footnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+%         Questions:\footnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+% %
+%         \begin{itemize}
+%             \item Correlation between variants. Which is the causal variant?
+%             \item Which genes are important for disease?
+%             \item In which tissues are these variants active?
+%         \end{itemize}
+% %
+%         \vfill
+% %
+%         Questions partially addressed by looking at eQTLs in different tissues
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%    The associations to diseases are heterougeneously distributed in the genome, Watanabe & Posthuma
+%    This is based on loci associated to multiple traits
+    \begin{frame}
+        \frametitle{Questions regarding pleiotropy analysis\footnote{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}}
+
+%         \begin{itemize}
+%             \item Association test of effect allele of individuals on traits
+%             \item High-throughput genetics: 3000 unique loci for over 250 disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
+%             \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories\footnote{Watanabe et al 2019. 10.1038/s41588-019-0481-0}
+%         \end{itemize}
+% %
+%         \vfill
+
+%          However, how do we know if these variants are active in a tissue,
+%         or whether these variants false positives due to linkage disequilibrium,
+%         or how these variants act molecularly
+
 %
@@
             \item Correlation between variants. Which is the causal variant?
-            \item Which cell types important for disease?
+            \item Which genes are important for disease?
             \item In which tissues are these variants active?
@@
             \item To identify pleiotropic eQTLs in the genome
-            \item To characterise their genomic regions, target genes and the cause of their pleiotropy
+            \item To characterise the causes of their pleiotropy
         \end{enumerate}
@@
         \begin{itemize}
-            \item 418 GWAS
+            \item UK biobank, GWAS catalog, Other datasets
             \item Physiological and disease traits
-            \item UK biobank, GWAS catalog, Other datasets
+            \item 418 GWAS, based on sample size and traits
         \end{itemize}
@@
         \begin{itemize}
+            \item GTEX, DICE, ...
+            \item Tissues, cell types, stimulated conditions, ...
             \item 127 studies in human from different datasets
-            \item Tissues, cell types, stimulated conditions, ...
-            \item GTEX, DICE, ...
         \end{itemize}
@@
         \begin{center}
-            \includegraphics[width=\textwidth]{../presentation_230120_gold2022_paris/fig/pgen.1004383.g001.png}
+            \includegraphics[width=0.5\textwidth]{../presentation_230120_gold2022_paris/fig/pgen.1004383.g001.png}
         \end{center}
@@
     \section{Results}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Colocalization of GWAS and eQTLs}
+
+%In this project, I have carried a large colocalization analysis between ~300 GWAS traits and 127 eQTL studies belonging to different tissues and mainly primary cell types.
+%In addition, by colocalizing statistically eQTLs and GWAS, we can to some extent, break, linkage disequilibrium
+
+        \begin{itemize}
+            \item Colocalization between 293 GWAS and 127 eQTL studies
+        \end{itemize}
+
+        \begin{figure}[!]
+            \includegraphics[width=\textwidth]{fig/web_portal.png}\label{fig:coloc_web}
+        \end{figure}
+
+    \end{frame}
 
@@
             \item Colocalization between 293 GWAS and 127 eQTL studies
-            \item 25\% of 10 k leading SNPs (SNP.PP.H4$\geq$0.5), colocalized with at least one eQTL
-            \item At SNP.PP.H4$\geq$0.5, 5 k eQTLs and 5 k target genes
+            \item 25\%\footnote{SNP.PP.H4$\geq$0.5} of 10,000 leading SNPs, colocalized with at least one eQTL
+            \item 5,000 eQTLs and 5,000 target genes\footnote{SNP.PP.H4$\geq$0.5}
             \item 138e3 colocalized variants with probability $\geq$ 0.75
-            \item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
+            \item Percentage of GWAS loci explained by eQTL: histogram
         \end{itemize}
@@
         \begin{itemize}
-            \item Uniformly annotate GWAS traits based on clustering and EBI ontologies
-            \item Aggregate colocalised eQTLs based on the number of trait categories
+            \item Annotate GWAS traits uniformly with ontologies
+            \item Annotate normalized trait name in higher-level categories, eg cardiovascular
+            \item Group categories of each eQTL
             \item Analyse eQTLs in groups based on the number of trait categories
@@
         \begin{itemize}
-            \item Example eQTLs involved in four GWAS categories in different cytobands
+            \item Example pleiotropic eQTLs in different cytobands
             \item The representative gene is the eQTL gene with the highest PubMed publication count
@@
             \item Cell cycle regulator
-            \item GWAS categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
+            \item Trait categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
             \item eQTL biosamples: adipose tissue;blood;brain;colon;immune system; ...
@@
         \begin{figure}[!]
-            \includegraphics[width=0.95\textwidth]{fig/cdk2na}
+            \includegraphics[width=0.95\textwidth]{fig/tp53}
         \end{figure}
@@
             \item Cell cycle regulator
-            \item GWAS categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
+            \item Trait categories: cancer;cardiovascular disease;hypertensive disorder;joint disease;respiratory system disease;skin cancer
             \item eQTL biosamples: adipose tissue;adrenal gland;aorta;blood;brain; ...
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \begin{frame}
-        \frametitle{Where are the pleiotropic variants?\footnote{https://www.ensembl.org/vep}}
-
-        \begin{center}
-            \includegraphics[width=0.7\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-        \end{center}
-
-        In the transcripts
-
-%        \let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}
-
-    \end{frame}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%     \begin{frame}
+%         \frametitle{Where are the pleiotropic variants?\footnote{https://www.ensembl.org/vep}}
+%
+%         \begin{center}
+%             \includegraphics[width=0.7\textwidth]{fig/vep.png}
+%         \end{center}
+%
+%         In the transcripts
+%
+% %        \let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}
+%
+%     \end{frame}
 
@@
     \begin{frame}
-        \frametitle{Are pleiotropic variants associated to more eQTL genes?}
+        \frametitle{Are pleiotropic eQTLs associated to more eQTL genes?}
 
@@
     \begin{frame}
-        \frametitle{Are pleiotropic variants associated to more eQTL tissues?}
+        \frametitle{Are pleiotropic eQTLs associated to more eQTL tissues?}
 
@@
     \begin{frame}
-        \frametitle{Do pleiotropic variants bind more transcription factors?\footnote{https://remap.univ-amu.fr}}
+        \frametitle{Do pleiotropic eQTLs bind more transcription factors?\footnote{https://remap.univ-amu.fr}}
 
@@
         \frametitle{Web access}
+
+        \url{https://gwas2eqtl.tagc.univ-amu.fr/gwas2eqtl/}
 

### 2023-02-05 ms/ms06_floats.tex
ENH
ENH improved ontology

@@
 %
-\caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different gene and samples.}
+    \caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different genes and samples.}
     \label{fig:gwas_distance}
@@
     head,
-table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Marker eQTL gene& \bfseries GWAS Categories\\\hline,
+    table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries GWAS Categories\\\hline,
     ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
@@
     \caption{Example eQTLs involved in four GWAS categories in different cytobands.
-The whole list is given in the Supplementary files.
-The gene marker is the eQTL gene with the highest PubMed publication count.}
+    The gene marker is the eQTL gene with the highest PubMed publication count.
+    The whole list is given in the Supplementary files.}
     \label{tab:pleitropic_variants}

### 2023-02-05 scripts/cmpt_supplementary_files.py

@@ -64,7 +64,8 @@ st_df.to_excel(st_writer, sheet_name=sheet_name, index=False, header=True)
 #%% ST2
 sheet_name = 'ST2'
 gwas_category_df = pandas.read_excel(gwas_category_ods_path)
-gwas_category_df.drop(['query', 'category'], axis=1, inplace=True)
+import pdb; pdb.set_trace()
+gwas_category_df.drop(['query_ontology', 'query_term', 'category'], axis=1, inplace=True)
 # import pdb; pdb.set_trace()
 # category_pleio_df = gwas_category_df[['icd10_code_level1.1', 'category_pleiotropy']]
 # category_pleio_df = category_pleio_df.loc[~category_pleio_df.isna().any(axis=1)]

### 2023-02-06 ms/ms06_floats.tex
ENH updated supplementary
ENH

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% Fig 1: Comorbidity matrix
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\begin{figure}[!tbp]
+    \centering
+%
+    \includegraphics[width=\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
+%
+    \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
+    I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
+    because this locus is not included in the colocalization pipeline.}
+    \label{fig:hist_perc_tophits_eqtl_excl_mhc}
+%
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     The gene marker is the eQTL gene with the highest PubMed publication count.
-    The whole list is given in the Supplementary files.}
+    The whole list is given in the Supplementary table ST2.}
     \label{tab:pleitropic_variants}

### 2023-02-06 scripts/cmpt_supplementary_files.py

@@ -10,10 +10,11 @@ try:
     gwas_trait_ods = sys.argv[1]
     gwas_category_ods = sys.argv[2]
     etissue_category_ods = sys.argv[3]
-    count_per_rsid_gwas_egene_etissue_ods = sys.argv[4]
-    region_window_100000_ods = sys.argv[5]
-    supp_tabl_xlsx_path = sys.argv[6]
-    if len(sys.argv) > 7:
+    perc_tophits_eqtl_tsv = sys.argv[4]
+    count_per_rsid_gwas_egene_etissue_ods = sys.argv[5]
+    region_window_100000_ods = sys.argv[6]
+    supp_tabl_xlsx_path = sys.argv[7]
+    if len(sys.argv) > 8:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -42,7 +43,8 @@ description_lst.append('Classification of eQTL tissues and cell types. The first
 sheet_name_lst.append('ST2')
 description_lst.append("Metadata and classification of GWAS")
 
-# ST3 ST3_gwas_trait_perc_explained_loci
+sheet_name_lst.append('ST3')
+description_lst.append("Percentage of tophits per GWAS that colocalized with at least one eQTL.")
 
 sheet_name_lst.append("ST4")
 description_lst.append("Count and list of GWAS phenotypes, egenes and etissues for each eQTL/GWAS variant")
@@ -84,6 +86,9 @@ st_df = gwas_trait_df.merge(gwas_category_df, on=['gwas_id', 'gwas_trait'])
 st_df.to_excel(st_writer, sheet_name=sheet_name, index=False, header=True)
 
 #%% ST3 ST3_gwas_trait_perc_explained_loci
+sheet_name = 'ST3'
+perc_tophits_eqtl_df = pandas.read_csv(perc_tophits_eqtl_tsv, sep="\t", index_col='gwas_id')
+perc_tophits_eqtl_df.to_excel(st_writer, sheet_name=sheet_name, index=True, header=True)
 
 #%% ST4
 sheet_name = 'ST4'

### 2023-02-06 scripts/plthst_perc_tophits_eqtl.py

@@ -1,0 +1,80 @@
+#%%
+import sys
+import matplotlib.pyplot as plt
+import os
+import pandas
+import pathlib
+import seaborn
+
+from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, dpi
+
+plt.rcParams["figure.figsize"] = (8, 6)
+from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
+seaborn.set_theme(**seaborn_theme_dic)
+
+#%%
+help_cmd_str = "todo"
+try:
+    perc_tophits_eqtl_tsv = sys.argv[1]
+    hist_perc_tophits_eqtl_png = sys.argv[2]
+    hist_perc_tophits_nomhc_eqtl_png = sys.argv[3]
+    if len(sys.argv) > 4:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+pathlib.Path(os.path.dirname(hist_perc_tophits_eqtl_png)).mkdir(parents=True, exist_ok=True)
+# import pdb; pdb.set_trace()
+# count_df = pandas.read_csv(perc_tophits_eqtl_tsv, sep="\t")
+perc_tophits_eqtl_df = pandas.read_csv(perc_tophits_eqtl_tsv, sep="\t", usecols=['gwas_id', 'loci_coloc_perc', 'loci_nomhc_coloc_perc'], index_col='gwas_id')
+# import pdb; pdb.set_trace()
+
+#%%
+ylabel = "Percentage of GWAS"
+# ylim=[0, 100]
+edgecolor='k'
+linewidth = 2
+grid_axis = 'y'
+hist_kwargs = {'density': 1, 'edgecolor': edgecolor, 'linewidth': linewidth}
+stat = 'percent'
+label_fontsize = 28
+
+#%% loci_coloc_perc
+title = "eQTL colocalized tophits (incl. MHC)"
+data_ser = perc_tophits_eqtl_df['loci_coloc_perc']
+shplt = seaborn.histplot(data_ser, stat=stat, discrete=True)
+
+plt.grid(visible=True, axis='y')
+plt.title(title, fontsize=label_fontsize)
+plt.xlabel("Percentage of tophits", fontsize=label_fontsize)
[…]

### 2023-02-06 ms/ms03_results.tex
ENH

@@
 These studies resulted in 10,621 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
-% select count(distinct gwas_id) from tophits; 335
+% select count(distinct gwas_id) from tophits
+% 335
 % select count(*) from (select DISTINCT chrom, pos, rsid, ea from tophits) as sel; 10621
@@
 
-My pipeline uncovered eQTL colocalizations for 5,484 or 64$\%$ of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
-% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid from tophits inner join coloc
-% on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75) as sel; 5484
+I excluded colocalization in the MHC locus in chromosome 6 between position 25,000,000 and 35,000,000.
+%
+My pipeline uncovered eQTL colocalizations for 5,484 of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
+% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid
+% from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75 and coloc.snp_pp_h4>=0.5) as sel
+% 5484
+% select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid
+% from tophits inner join coloc on coloc.rsid=tophits.rsid where coloc.pp_h4_abf>=0.75 and coloc.snp_pp_h4>=0.5) as sel
+% 2441
 % select count(*) from (select distinct eqtl_gene_id from coloc where pp_h4_abf>=0.75) as sel;  7040
+% select count(*) from (select distinct eqtl_gene_id from coloc where pp_h4_abf>=0.75 and coloc.snp_pp_h4>=0.5) as sel;  5620
 % select count(*) from (select distinct eqtl_id from coloc where pp_h4_abf>=0.75) as sel;  127
+%
+The histogram shows three group of GWAS (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}).
+%
+Around 25\% of GWAS had no tophits colocalized with eQTLs.
+%
+On the other extreme, 7.5\% of GWAS had all 100\% tophits colocalized with eQTLs.
+%
+The other GWAS had between 10\% and 50\% of tophits colocalized with eQTLs.
 
@@
 % TODO
-% select count(*) from (select DISTINCT chrom,pos38,alt FROM colocweb) as sel; 138301
+%select count(*) from (select DISTINCT chrom,pos,alt FROM coloc
+%where coloc.pp_h4_abf>=0.75 and coloc.snp_pp_h4>=0) as sel
+%138301
+%select count(*) from (select DISTINCT chrom,pos,alt FROM coloc
+%where coloc.pp_h4_abf>=0.75 and coloc.snp_pp_h4>=0.5) as sel
+%5343
 % select count(*) from (select DISTINCT gwas_id FROM colocweb) as sel; 293

### 2023-02-07 ms/ms04_discussion.tex
MS BMC

@@
 %This variants control the gene RERE, which is a nuclear receptor corregulator and is involved in neurodevelopmental disorders.\\
-
-\section*{Methods}\label{sec:methods}
-
-\subsection*{Data sets and data exploration}
-
-I used these data sets here: eQTLs from the eQTL Catalogue \cite{2021.Alasoo.Kerimov}, GWAS variants from the IEU OpenGWAS project \cite{2021.Marcora.Lyon}, chromatin immunoprecipitation (ChIP) transcription factors peaks from the ReMap database \cite{2021.Ballester.Hammal}, and UCSC annotation data.
-%
-I explore the data using the UCSC browser (\url{http://genome.ucsc.edu}) \cite{2021.Kent.Lee}, and OMIM database (\url{https://omim.org/}).
-%
-Colocalization and analysis pipelines were implemented with Snakemake (Supplementary Figure S1) .
-%
-The colocalization and analysis pipelines can be found here: \url{https://github.com/aitgon/gwas2eqtl} and \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
-
-\subsection*{Colocalization}
-
-My colocalization pipeline was inspired by another pipeline that is at the eQTL Catalogue Github repository ( \url{https://github.com/eQTL-Catalogue/colocalisation} ).
-%
-GWAS were downloaded from OpenGWAS and converted to hg38 coordinates.
-%
-Permuted eQTLs with FDR 0.05 in each sample were selected as lead variants.
-%
-eQTL and GWAS variants surrounded the lead variants with radius of 500,000 nt (Total window 1,000,000 nt) were retrieved.
-%
-These variants were kept if MAF was strictly greater than 0 and lower than 1, the variants were not duplicated and there was no missing data.
-%
-Colocalization between eQTL and GWAS variants was tested using the "coloc.abf" function of CRAN coloc for each window, each eQTL sample and each GWAS.
-
-\subsection*{Definition of pleiotropic regions}
-
-To define pleiotropic regions, I included "pleiotropic" variants defining as having more than 2 categories as long as the distance between the two "pleiotropic" variants is less than 100 000 bp.
-%
-Then the number of GWAS categories in Supplementary table TODO is given by the number of different variant categories in the region.
-
-\subsection*{Characterization of regulatory QTLs}
-
-\section*{Code availability}
-
-The eQTL/GWAS variant colocalization pipeline code is available at this repository: \url{https://github.com/aitgon/gwas2eqtl}.
-%
-The analysis pipeline of the eQTL/GWAS colocalization data and the source code of this manuscript is available at this repository: \url{https://github.com/aitgon/gwas2eqtl_pleiotropy}.
-
-\section*{Data availability}
-
-The raw colocalization data before filtering in a single file is available at this URL: \url{https://osf.io/hvmje}.
-%
-A website to access colocalization variants with PP.H4.abf $\geq$ 0.5 is available at this URL: TODO.
-
-\section*{Acknowledgements}
-
-Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
-%
-I thank Sandrine Marquet, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and L\'eopoldine Lecerf for a preliminary version of the pipeline developed during her internship.

### 2023-02-07 ms/ms06_figures.tex
MS Genome biology

--- 2023-02-06 ms/ms06_floats.tex
@@
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Tab 1: Winner eQTLs
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% full size table is table
-\begin{table}[!tbp]
-    \centering
-    \scriptsize
-%\hline
-    \csvreader[separator=tab,
-    tabular=crclcp{0.4\textwidth},
-    head,
-    table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries GWAS Categories\\\hline,
-    ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-        {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-%\hline
-%
-    \vspace{15pt}
-%
-    \caption{Example eQTLs involved in four GWAS categories in different cytobands.
-    The gene marker is the eQTL gene with the highest PubMed publication count.
-    The whole list is given in the Supplementary table ST2.}
-    \label{tab:pleitropic_variants}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Tab 2: Winner regions
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% full size table is table
-\begin{table}[!tbp]
-    \centering
-    \scriptsize
-%\hline
-    \csvreader[
-        separator=tab,
-        tabular=p{0.05\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.5\textwidth},
-        head,
-        table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries GWAS Categories\\\hline,
-    ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
-    {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-%\hline
-%
-    \vspace{15pt}
-    \caption{Pleiotropic regions involving 6 or more GWAS classes.
-These regions were built with a sliding window of 1e5 nt.
-Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
-\end{table}
 

### 2023-02-07 ms/ms07_tables.tex
MS Genome biology

--- 2023-02-05 ms/ms06_floats.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 1: Comorbidity matrix
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-    \centering
-%
-    \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
-%
-    \caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different genes and samples.}
-    \label{fig:gwas_distance}
-%
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 2: Length histogram of regions
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \begin{figure}[!tbp]
-% \centering
-% %
-% \begin{subfigure}[]{.33\textwidth}
-% \textbf{a}
-% \\
-% \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
-% \end{subfigure}
-% %
-% \begin{subfigure}[]{.33\textwidth}
-% \textbf{b}
-% \\
-% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
-% \end{subfigure}
-% %
-% \caption{Analysis of pleiotropic regions.
-\textbf{a}, Length distribution of pleiotropic regions.
-\textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
-% \end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 3: Histograms variants vs GWAS, eGene and eTissues
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-    \centering
-%
-    \begin{subfigure}[]{.32\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
-    \end{subfigure}
-%
-    \begin{subfigure}[]{.32\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
-    \end{subfigure}
-%
-    \begin{subfigure}[]{.32\textwidth}
-        \textbf{c}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
-    \end{subfigure}
-%
-    \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
-        (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 4: Correlation between GWAS, genes and tissue counts
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-    \centering
-%
-    \includegraphics[width=0.7\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
-%
-    \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL tissues.}
-    \label{fig:correlation_counts}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 5: VEP consequences
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!tbp]
-    \centering
-%
-    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-%
-    \caption{Odds ratio of the variants annotated with given variant effect
-    predictor (VEP) consequence and given GWAS class count compared with non
-    pleiotropic variants with GWAS class count one. The significance is given for the
-    q-value with False discovery rate 5\%.}
-    \label{fig:vep_consequence}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 6: Violin plots. eGenes, eTissues and gwas per variants
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-    \centering
-%
-    \begin{subfigure}[]{.32\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
-    \end{subfigure}
-%
-    \begin{subfigure}[]{.32\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
-    \end{subfigure}
-%
-%\begin{subfigure}[]{.32\textwidth}
-%\textbf{c}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-%\end{subfigure}
-%
-    \caption{Mean number of genes in eQTL-tissue pairs (\textbf{a}) and mean number of tissues in eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 7: TF count per GWAS class count
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!tbp]
-    \centering
-%
-    \begin{subfigure}[]{.33\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
-    \end{subfigure}
-%
-    \begin{subfigure}[]{.33\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
-    \end{subfigure}
-%
-    \caption{Analysis of transcription factor binding.
-    \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
-    \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.} \label{fig:freq_tf_per_variant}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 8: beta and log pval
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{figure}[!tbp]
-%\centering
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{b}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
-%\end{subfigure}
-%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{c}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{d}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
-%\end{subfigure}
-%
-%\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
-%%Correlation
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 9: graphical conclusions
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{figure}[!tbp]
-    \centering
-%
-    \begin{subfigure}[]{\textwidth}
-
-        \includegraphics[width=\textwidth]{fig/graphical_summary.png}
-    \end{subfigure}
-
-    \caption{\textbf{Model of regulatory variant pleiotropy.} I have investigated three possible mechanisms of pleiotropy.
-\textbf{Left}, Pleiotropic variants have more eGenes that result in more functions and more phenotypes.
-This might arise from an enrichment of pleiotropic variants in splicing or 3' UTR regions.
-\textbf{Center}, I have also found that eGenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
-This might be explained from variants being bound by more transcription factors.
-\textbf{Left} Triplets of variant-eGene-eTissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
-I have found that this might be explained by an enrichment of missense alleles.} \label{fig:beta}
-%
-\end{figure}

### 2023-02-16 scripts/pltbar_x_per_gwas_cat_y_allele_freq.py
ENH allele freq

@@ -1,0 +1,97 @@
+from sqlalchemy import create_engine, text
+
+from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, boxplot_kwargs, annotator_config_dic
+from statannotations.Annotator import Annotator
+
+import matplotlib.pyplot as plt
+import numpy
+import os
+import pandas
+import pathlib
+import seaborn
+import sys
+
+
+# Plot parameters
+plt.rcParams["figure.figsize"] = (8, 6)
+from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
+seaborn.set_theme(**seaborn_theme_dic)
+
+#%%
+help_cmd_str = "todo"
+try:
+    snp_pp_h4 = float(sys.argv[1])
+    url = sys.argv[2]
+    count_per_rsid_gwas_ods_path = sys.argv[3]
+    eur_af_png_path = sys.argv[4]
+    if len(sys.argv) > 5:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+#%% Input
+if not os.path.isfile(count_per_rsid_gwas_ods_path):
+    print("input file does not exit")
+    sys.exit(1)
+
+outdir_path = os.path.dirname(eur_af_png_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
+
+#%%
+sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
+engine = create_engine(url)
+# query = 'SELECT * FROM my_table'
+# import pdb; pdb.set_trace()
+with engine.begin() as conn:
+    h4_df = pandas.read_sql(text(sql), con=conn).drop_duplicates()
+
+#%%
+count_per_rsid_gwas_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine='odf')
+max_gwas_category_count = count_per_rsid_gwas_df['gwas_category_count'].max()
+
+#%%
+m_df = h4_df.merge(count_per_rsid_gwas_df[['chrom', 'pos38', 'rsid', 'ref', 'alt', 'gwas_category_count']], on=['chrom', 'pos38', 'rsid', 'ref', 'alt'])
[…]

### 2023-02-16 scripts/pltbar_x_per_variant_etissue_y_egene.py
ONGOING

@@ -1,3 +1,4 @@
+import sqlalchemy
 from statannotations.Annotator import Annotator
 from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, dpi, boxplot_kwargs, annotator_config_dic
@@ -20,11 +21,11 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    max_gwas_category_count = int(sys.argv[2])
-    url = sys.argv[3]
-    count_per_rsid_gwas_ods_path = sys.argv[4]
-    vlnplt_png_path = sys.argv[5]
-    if len(sys.argv) > 6:
+    # max_gwas_category_count = int(sys.argv[2])
+    sa_url = sys.argv[2]
+    count_per_rsid_gwas_ods_path = sys.argv[3]
+    vlnplt_png_path = sys.argv[4]
+    if len(sys.argv) > 5:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -43,7 +44,10 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 #%%
 sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
-h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
+# h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
+engine = sqlalchemy.create_engine(sa_url)
+with engine.begin() as conn:
+    h4_df = pandas.read_sql(sqlalchemy.text(sql), con=conn).drop_duplicates()
 
 #%%
 # count_per_rsid_gwas_df = pandas.read_csv(count_per_rsid_gwas_ods_path, sep="\t")
@@ -64,7 +68,7 @@ m2_df.to_csv(tsv_path, header=True, index=False, sep='\t')
 
 #%% set max_gwas_category_count
 m_df = m_df[sel_cols + ['gwas_category_count']]
-m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+# m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
 
[…]

### 2023-02-18 ms/ms02_intro.tex
MS BMC
MS

@@
-\section*{Introduction}\label{sec:introduction}
+\section*{Background}\label{sec:introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% I. FIRST HALF OF THE INTRODUCTION (State of the art)
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%1. Background key sentence (max < 1 ½ line):
+%Example: "Machupo virus infects 200,000 persons per year worldwide"
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % GWAS in general
-Genome-wide association studies (GWAS) have allowed the association of frequent genetic variants with common traits.
+Genome-wide association studies (GWAS) have allowed the association of many frequent genetic variants with common traits.
 %
@@
 
-% Shortcomings of GWAS
+%In addition, colocalization analysis can help prioritize causal variants.
+
+This wealth of GWAS has indicated that a large proportion of the genome is highly pleiotropic \cite{2019.Posthuma.Watanabe}.
+%
+More precisely, loci covering 60\% of the genome loci was analyzed, and 90\% of these loci contained associations across different traits \cite{2019.Posthuma.Watanabe}.
+%
+It was also found that pleiotropic genes and variants were less tissue-specific in terms of gene expression and eQTLs \cite{2019.Posthuma.Watanabe}.
+%
+%However, the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \cite{2021.Li.Mu,2018.Vijayanand.Schmiedel,2017gtex.nature}.
+%%
+%In addition, the mechanism of pleiotropy is not well understood yet.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 2. Put the hero of the study center stage1:
+% - Template: “X is a central component of/ one of the main parameters of/ a logical candidate for/an enigmatic/ etc
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Shortcomings of GWAS and advantage of eQTLs
 However, the interpretation of the GWAS variants is complex because of three reasons \cite{2020.Trynka.CanoGamez}.
@@
 % Interest in studying GWAS
-Expression QTLs (eQTLs) are variants that change the expression of a given gene in a cell type or tissue.
+Expression QTLs (eQTLs) are variants that change the expression of genes in particular cell types and tissue.
 %
-eQTLs share some of the same shortcomings as GWAS variants such as the inability to identify the causal variant in regions of high LD.
+Therefore, eQTLs can be used to annotate non-coding regions that often regulate gene expression.
 %
-Moreover, the distribution of eQTLs and GWAS variants in the genome are different because of negative selection \cite{2022.Pritchard.Mostafavi}.
+%In addition, colocalization analysis of eQTLs and GWAS variants is able to partially break linkage disequilibrium (LD).
+%%
+%eQTLs share some of the same shortcomings as GWAS variants such as the inability to identify the causal variant in regions of high LD.
+%%
+%%Moreover, the distribution of eQTLs and GWAS variants in the genome are different because of negative selection \cite{2022.Pritchard.Mostafavi}.
+%%
+%Nevertheless, eQTLs are still important data sets to interpret GWAS variants, because they provide gene regulatory information about the eQTL genes and cell types and tissues \cite{2021.Li.Mu}.
 %
-Nevertheless, eQTLs are still important data sets to interpret GWAS variants, because they provide gene regulatory information about the eQTL genes and cell types and tissues \cite{2021.Li.Mu}.
-%
-%In addition, colocalization analysis can help prioritize causal variants.
+There, eQTLs are important variants to interprete GWAS.
 
-Many GWAS have allowed observing that many genetic variants are involved in different traits \cite{2019.Posthuma.Watanabe}.
-%
-In that work, the analyzed loci covered 60\% of the genome, and 90\% of these loci contained associations across different traits \cite{2019.Posthuma.Watanabe}.
-%
-They also found that pleiotropic genes and variants were less tissue-specific in terms of gene expression and eQTLs \cite{2019.Posthuma.Watanabe}.
-%
-However, the question of tissue specificity of eQTLs has been analyzed in different works with conflicting conclusions \cite{2021.Li.Mu,2018.Vijayanand.Schmiedel,2017gtex.nature}.
-%
-In addition, the mechanism of pleiotropy is not well understood yet.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% I.3. Gap1:
+% - Template: We don't know / we don’t understand / we lack ..."
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+However, we don't know whether there are differences in pleiotropy among eQTLs and the distribution in the genome and properties of  pleiotropic eQTLs in the genome.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%4. Limitations of previous approaches1:
+%- Template: "Existing approaches couldn’t address the gap because they suffer from N main obstacles"
+%- Example: "2 obstacles prevented the determination of the structure of protein X: 1) the difficulty of producing it in sufficient quantity and 2) its inherent flexibility"
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Previous colocalization analysis between eQTLs and GWAS focused on particular sets of GWAS such as autoimmune traits \cite{2021.Li.Mu}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%5. How you overcame previous limitations:
+%- Template: We were able to address the gap thanks to approach X, which solves the N main limitations mentioned above
+%- Example: "We were able to determine the structure of X because 1) we managed to produce sufficient quantities in a cell-free system and 2) we reduced its flexibility by complexing it with an anti-X antibody”.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 In this work, I have taken advantage of two large eQTL and GWAS data sets from the EBI eQTL Catalogue and the IEU OpenGWAS, respectively,
-to develop a global colocalization analysis based on 127 eQTL studies and 418 GWAS across many cell types, tissues, and phenotypes.
+to develop a systematic colocalization analysis based on 127 eQTL studies and 417 GWAS across many cell types, tissues, and phenotypes.
 %
-I have used these eQTL/GWAS variants to investigate mechanisms of pleiotropic gene regulatory variants.
+I have annotated the 417 GWAS with trait categories and aggregated colocalized eQTLs into this trait categories.
 %
-These eQTL/GWAS variants might be a useful resource for the annotation and prioritization of frequent variants.
+This approach has allowed me to uncover pleiotropic eQTLs, locate the genomic regions with most pleiotropic eQTLs
+and identify some significant properties of these eQTLs.
 %
-Therefore, I have developed an easy-to-use web application to expose these colocalization data.
+%I have used these eQTL/GWAS variants to investigate mechanisms of pleiotropic gene regulatory variants.
+%%
+%These eQTL/GWAS variants might be a useful resource for the annotation and prioritization of frequent variants.
+%%
+%Therefore, I have developed an easy-to-use web application to expose these colocalization data.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% SECOND HALF OF THE INTRODUCTION
+% (Research questions and overview of experimental approach)
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-02-18 ms/ms03_results.tex
MS

@@
 
-To annotate traits from genome-wide associations studies (GWAS) with genes and tissues,
-I developed a pipeline that systematically colocalizes variants from GWAS and expression quantitative trait loci (eQTLs) (Section Methods).
+% Organizing words: Pleiotropic eQTLs
+
+To investigate the pleiotropy of eQTLs, I developed a pipeline that systematically colocalizes variants from GWAS and expression quantitative trait loci (eQTLs) (Section Methods).
 %
@@
 
+For the genome-wide association studies (GWAS), I used data from the IEU OpenGWAS database, that provides GWAS summary statistics
+from different batches such as the EBI database of complete GWAS summary data, manually collected GWAS and UK Biobank.
+%
 GWAS traits were selected from the IEU OpenGWAS database based on four criteria \cite{2018.Parkinson.Buniello}.
@@
 %
-These filters resulted in 418 GWAS traits (Supplementary table 2).
+These filters resulted in 417 GWAS traits (Supplementary table 2).
 %
 These studies resulted in 10,621 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
+% TODO: Add to supplementary table?
 % select count(distinct gwas_id) from tophits
@@
 
-I excluded colocalization in the MHC locus in chromosome 6 between position 25,000,000 and 35,000,000.
+The colocalization pipeline uncovered eQTL colocalizations for 5,484 of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples in the somatic chromosomes 1-22 \cite{2019.Posthuma.Watanabe,2021.Li.Mu}.
 %
-My pipeline uncovered eQTL colocalizations for 5,484 of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples.
 % select count(*) from (select distinct tophits.chrom,tophits.pos,tophits.ea,tophits.rsid
@@
 %
-The histogram shows three group of GWAS (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}).
+Variants in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000 were
+excluded from the colocalization pipeline, because of their complex linkage disequilibrium structure.
 %
-Around 25\% of GWAS had no tophits colocalized with eQTLs.
+The histogram shows three groups of GWAS (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}).
 %
-On the other extreme, 7.5\% of GWAS had all 100\% tophits colocalized with eQTLs.
+There were 100\% of leading variants explained with eQTLs for between 5\% and 12\% of the GWAS depending on whether
+we set PP.H4.SNP at 0.25, 0.5 or 0.75 (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc} and suppl. fig. S1).
+This parameter gives the posterior probability of a SNP in a window to be causal.
 %
-The other GWAS had between 10\% and 50\% of tophits colocalized with eQTLs.
+There were a middle group of maximally 5\% GWAS with maximum percentage of explained leading SNPs between 10\% and 50\%
+depending on the PP.H4.SNP parameter 0.25-0.75
+Around 25\% of GWAS had no tophits colocalized with eQTLs for a parameter PP.SNP.H4=0.5.
+%
+Finally, between 17\% and 27\% GWAS had 0\% of leading variants explained by eQTLs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc} and suppl. fig. S1).
 
@@
 % select count(distinct tophits.rsid),gwas_class from gwas_annot, tophits, coloc where tophits.gwas_id=gwas_annot.gwas_id and tophits.rsid=coloc.rsid group by gwas_class order by count desc;
-
+%
 These results is much higher than previous colocalization analyses for autoimmune diseases by previous authors that
@@
 
-These data is a useful resource to annotate GWAS variants with genes and biosamples.
+These variants annotated with GWAS association and gene expression associations in particular biosamples are useful functional informations.
+%
 Therefore I have developed a web application where I expose 138,301 variants from 293 GWAS that colocalize with at least
 one eQTL at a threshold value PP.H4.abf$\geq$0.75.
-% TODO
 %select count(*) from (select DISTINCT chrom,pos,alt FROM coloc
@@
 % Work with PP.H4.abf`>=0.75 and SNP.PP.H4>=0.5
-\subsection*{Traits classification based on the effect size of eQTLs}
+\subsection*{Trait clustering based on the beta of the eQTLs}
 %
@@
 
+% Organizing words: Pleiotropic eQTLs
+
+In the 417 GWAS, there are traits from different categories and belonging to independent studies.
+%
+To verify that the colocalization results are coherent between different GWAS datasets,
+I cluster the GWAS traits based on the Spearman correlation between beta values
+of colocalized eQTL variant, gene and biosample (Fig. \ref{fig:trait_clusters}).
+
 In the previous colocalization analysis, we found many eQTLs that are associated to very different traits.
+
+Many eQTLs were found to colocalize with GWAS traits from different categories.
 %
@@
 %
-Only trait classes with more than a minimum number of traits were plotted (Fig. \ref{fig:gwas_distance}).
+Only trait classes with more than a minimum number of traits were plotted (Fig. \ref{fig:trait_clusters}).
 %
@@
 %
-This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases, cancers and allergies (Fig. \ref{fig:gwas_distance}).
+This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases, cancers and allergies (Fig. \ref{fig:trait_clusters}).
 %

### 2023-02-18 ms/ms06_backmatter.tex
MS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

    \section*{Acknowledgements}%% if any
    Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
%
    I thank Sandrine Marquet, Marie Michel, Pascale Paul, Salvatore Spicuglia and Pascal Rihet for helpful discussions and Léopoldine Lecerf for a preliminary version of the colocalization pipeline developed during her internship.

%    \section*{Funding}%% if any
%    Text for this section\ldots

%    \section*{Abbreviations}%% if any
%    Text for this section\ldots

    \section*{Availability of data and materials}%% if any
    Text for this section\ldots

%    \section*{Ethics approval and consent to participate}%% if any
%    Text for this section\ldots

    \section*{Competing interests}
    The author declare that he has no competing interests.

%    \section*{Consent for publication}%% if any
%    Text for this section\ldots

    \section*{Authors' contributions}
    Aitor González: Conceptualization, Methodology, Software, Validation, Formal analysis, Resources, Data Curation, Visualization, Writing - original draft, Writing - review \& editing.

%    \section*{Authors' information}%% if any
%    Text for this section\ldots

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
    \bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
    \bibliography{ms_pleiotropy}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

    \section*{Figures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 1: Histogram of the percentage of the GWAS explained leading variants
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
        I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
            because this locus is not included in the colocalization pipeline.}
        \label{fig:hist_perc_tophits_eqtl_excl_mhc}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Trait distance matrix
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
        \label{fig:trait_clusters}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
    \begin{figure}[h!]
        %
        \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
            (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL biosamples.}
        \label{fig:correlation_counts}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 5: VEP consequences
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Odds ratio of the variants annotated with given variant effect
        predictor (VEP) consequence and given GWAS class count compared with non
        pleiotropic variants with GWAS class count one. The significance is given for the
        q-value with False discovery rate 5\%.}
        \label{fig:vep_consequence}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{TODO beta and p-val}
        \label{fig:beta_pval}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Mean number of eQTL genes per eQTL-biosample pair (\textbf{a}) and mean number of biosamples per eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 8: TF count per GWAS class count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{figure}[h!]
        %
        \caption{Analysis of transcription factor binding.
        \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
        \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.}
        \label{fig:freq_tf_per_variant}
        %
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
    \section*{Tables}

    \input{ms07_tables}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \section*{Additional Files}
    \subsection*{Additional file 1: Table S1.}
    Classification of eQTL tissues and cell types. The first six were downloaded from the EBI eQTL repository.
    The 7th column "etissue\_category\_term" is used here to compute tissue diversity.

    \subsection*{Additional file 2: Table S2.}
    Metadata and classification of GWAS.

    \subsection*{Additional file 3: Table S3.}
    Percentage of tophits per GWAS that colocalized with at least one eQTL.

    \subsection*{Additional file 4: Table S4.}
    Count and list of GWAS phenotypes, egenes and etissues for each eQTL/GWAS variant.

    \subsection*{Additional file 5: Table S5.}
    Count and list of GWAS phenotypes, egene symbols and ENSEMBL IDs and etissue classes for each pleiotropic region.

\end{backmatter}

### 2023-02-18 ms/ms07_tables.tex
MS

@@
 %
-% Tab 1: Winner eQTLs
+% Tab 1: Representative pleiotropic eQTLs of cytobands
 %
@@
 
-% full size table is table
-\begin{table}[!tbp]
+\begin{table}[h!]
     \centering
     \scriptsize
-%\hline
+
     \csvreader[separator=tab,
     tabular=crclcp{0.4\textwidth},
+    late after last line=\\\hline,
     head,
-    table head=\bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries GWAS Categories\\\hline,
+    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries GWAS Categories\\\hline,
     ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
         {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-%\hline
     %
@@
     %
-    \caption{Example eQTLs involved in four GWAS categories in different cytobands.
+    \caption{Represantive pleiotropic eQTLs involved in four GWAS categories in different cytobands.
     The gene marker is the eQTL gene with the highest PubMed publication count.
@@
 %
-% Tab 2: Winner regions
+% Tab 2: Region containing most pleiotropic eQTLs
 %
@@
 
-% full size table is table
-\begin{table}[!tbp]
+\begin{table}[h!]
     \centering
     \scriptsize
-%\hline
     \csvreader[
@@
         tabular=p{0.05\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.5\textwidth},
+        late after last line=\\\hline,
         head,
-        table head=\bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries GWAS Categories\\\hline,
+        table head=\\\hline \bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries GWAS Categories\\\hline,
     ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
     {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-%\hline
     %
     \vspace{15pt}
+    %
     \caption{Pleiotropic regions involving 6 or more GWAS classes.
@@
 Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
+    %
 \end{table}

### 2023-02-18 ms/ms08_figures.tex
MS

--- 2023-02-07 ms/ms06_figures.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 1: Comorbidity matrix
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{figure}[!tbp]
+% Fig 1: Histogram of percentage of explained of GWAS
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
     \centering
@@
     %
-    \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
-    I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
-    because this locus is not included in the colocalization pipeline.}
-    \label{fig:hist_perc_tophits_eqtl_excl_mhc}
+    \caption{}
     %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 1: Comorbidity matrix
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{figure}[!tbp]
+% Fig 2: Trait distance matrix
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+    %
     \centering
@@
     %
-    \caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different genes and samples.}
-    \label{fig:gwas_distance}
+    \caption{}
     %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 2: Length histogram of regions
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
+% Fig 2: Length histogram of regions
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \begin{figure}[!tbp]
@@
 % \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 3: Histograms variants vs GWAS, eGene and eTissues
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
 %
-\begin{figure}[!tbp]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+    %
     \centering
@@
     %
-    \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
-        (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
+    \caption{}
+    %
 \end{figure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 4: Correlation between GWAS, genes and tissue counts
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{figure}[!tbp]
+% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+    %
     \centering
     %
-    \includegraphics[width=0.7\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
+    \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
     %
-    \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL tissues.}
-    \label{fig:correlation_counts}
+    \caption{}
     %
@@
 
-\begin{figure}[!tbp]
+\begin{figure}[h!]
+    %
     \centering
@@
     %
-    \caption{Odds ratio of the variants annotated with given variant effect
-    predictor (VEP) consequence and given GWAS class count compared with non
-    pleiotropic variants with GWAS class count one. The significance is given for the
-    q-value with False discovery rate 5\%.}
-    \label{fig:vep_consequence}
+    \caption{}
     %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 6: Violin plots. eGenes, eTissues and gwas per variants
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{figure}[!tbp]
+% Fig 6: beta and pval
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+
+    \begin{subfigure}[]{.50\textwidth}
+        \textbf{a}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
+    \end{subfigure}
+    %
+    \begin{subfigure}[]{.50\textwidth}
+        \textbf{b}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{.50\textwidth}
+        \textbf{c}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_neglog10pval.png}
+    \end{subfigure}
+    %
+    \begin{subfigure}[]{.50\textwidth}
+        \textbf{d}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_neglog10pval.png}
+    \end{subfigure}
+    %
+    \caption{}
+    %
+
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+    %
     \centering
     %
-    \begin{subfigure}[]{.32\textwidth}
+    \begin{subfigure}[]{.45\textwidth}
         \textbf{a}
@@
         \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+        %
     \end{subfigure}
     %
-    \begin{subfigure}[]{.32\textwidth}
+    \begin{subfigure}[]{.45\textwidth}
         \textbf{b}
@@
     %
-%\begin{subfigure}[]{.32\textwidth}
-%\textbf{c}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-%\end{subfigure}
-%
-    \caption{Mean number of genes in eQTL-tissue pairs (\textbf{a}) and mean number of tissues in eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
+    \caption{} \label{fig:gwas_egene_etisue_per_variant}
     %
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 7: TF count per GWAS class count
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Fig 8: TF count per GWAS class count
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[!tbp]
+\begin{figure}[h!]
     \centering
     %
-    \begin{subfigure}[]{.33\textwidth}
+    \begin{subfigure}[]{.48\textwidth}
         \textbf{a}
@@
     %
-    \begin{subfigure}[]{.33\textwidth}
+    \begin{subfigure}[]{.48\textwidth}
         \textbf{b}
@@
     %
-    \caption{Analysis of transcription factor binding.
-    \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
-    \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.} \label{fig:freq_tf_per_variant}
+    \caption{}
     %
@@
 %\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
 %\end{subfigure}
@@
 %\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
+%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
 %\end{subfigure}
@@
 %
-\begin{figure}[!tbp]
+\begin{figure}[h!]
     \centering
 %
-    \begin{subfigure}[]{\textwidth}
-
-        \includegraphics[width=\textwidth]{fig/graphical_summary.png}
+    \begin{subfigure}[]{0.95\textwidth}
+        \textbf{a}
+        \\
+        \includegraphics[width=\textwidth]{fig/model1.png}
+    \end{subfigure}
+%
+    \begin{subfigure}[]{0.95\textwidth}
+        \textbf{b}
+        \\
+        \includegraphics[width=\textwidth]{fig/model2.png}
     \end{subfigure}

### 2023-02-24 ms/ms08_figures.tex
ENH egene distance

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/violin.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/violin.png}
     \end{subfigure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% Fig 8: beta and log pval
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-%\begin{figure}[!tbp]
-%\centering
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{a}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{b}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
-%\end{subfigure}
+% Fig 9: eQTL gene distance
 %
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{c}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
-%\end{subfigure}
-%%
-%\begin{subfigure}[]{.33\textwidth}
-%\textbf{d}
-%\\
-%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
-%\end{subfigure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h!]
+    \centering
     %
-%\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
-%%Correlation
-%\end{figure}
+    \begin{subfigure}[]{.48\textwidth}
+        \textbf{a}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+    \end{subfigure}
+    %
+    \caption{}
+    %
+\end{figure}
 

### 2023-02-24 ms/ms08_figures.tex
MS figurex

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/violin.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/violin.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
     \end{subfigure}
@@
     %
-    \begin{subfigure}[]{.48\textwidth}
+    \begin{subfigure}[]{.48\textwidth}vep.png
         \textbf{a}

### 2023-02-24 scripts/cmpt_ucsc_tracks_eqtl_inter_bed.py

@@ -10,7 +10,7 @@ import sqlalchemy
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    sa_url = sys.argv[2]
+    db_url = sys.argv[2]
     outdir_path = sys.argv[3]
     if len(sys.argv) > 4:
         print("""Two many arguments!
@@ -25,7 +25,7 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 
 #%%
 sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
-engine = sqlalchemy.create_engine(sa_url)
+engine = sqlalchemy.create_engine(db_url)
 with engine.begin() as conn:
     df = pandas.read_sql(sqlalchemy.text(sql), con=conn).drop_duplicates()
 

### 2023-02-25 ms/ms08_figures.tex
MS

@@
 % Fig 1: Histogram of percentage of explained of GWAS
+% Fig 2: ucsc screenshots
 %
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
-    %
+
+    \begin{subfigure}[]{0.49\textwidth}
+        \textbf{a}
+        \\
         \includegraphics[width=\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
-    %
+    \end{subfigure}
+
+    \begin{subfigure}[]{0.99\textwidth}
+        \textbf{b}
+        \\
+        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_bcell.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{0.99\textwidth}
+        \textbf{c}
+        \\
+        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_monocyte.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{0.99\textwidth}
+        \textbf{d}
+        \\
+        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_frontalcortex.png}
+    \end{subfigure}
+
     \caption{}
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     %
@@
 % Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
+% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
+% Fig 5: VEP consequences
 %
@@
     %
-    \begin{subfigure}[]{.32\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{a}
@@
     %
-    \begin{subfigure}[]{.32\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{b}
@@
     \end{subfigure}
-    %
-    \begin{subfigure}[]{.32\textwidth}
+
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{c}
@@
     %
-    \caption{}
-    %
-\end{figure}
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{d}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
+    \end{subfigure}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h!]
-    %
-    \centering
-    %
-    \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
-    %
-    \caption{}
-    %
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 5: VEP consequences
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h!]
-    %
-    \centering
-    %
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{c}
+        \\
         \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
+    \end{subfigure}
     %
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
 
-    \begin{subfigure}[]{.50\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{a}
@@
     %
-    \begin{subfigure}[]{.50\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{b}
@@
 
-    \begin{subfigure}[]{.50\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{c}
@@
     %
-    \begin{subfigure}[]{.50\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{d}
@@
     \end{subfigure}
-    %
+
+    \centering
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{e}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_allele_freq.py/eur_af.png}
+    \end{subfigure}
+
     \caption{}
-    %
 
@@
 % Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
+% Fig: TF count per GWAS class count
+% Fig : eQTL gene distance
 %
@@
 
-\begin{figure}[h!]
-    %
+\begin{figure}[!ht]
+
     \centering
-    %
-    \begin{subfigure}[]{.45\textwidth}
+
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{a}
@@
     %
-    \begin{subfigure}[]{.45\textwidth}
+    \begin{subfigure}[]{.49\textwidth}
         \textbf{b}
@@
     \end{subfigure}
+
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{c}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+    \end{subfigure}
     %
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{d}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+    \end{subfigure}
+
+    % Fig 9: eQTL gene distance
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{e}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+    \end{subfigure}
+
     \caption{} \label{fig:gwas_egene_etisue_per_variant}
@@
 %
-% Fig 8: TF count per GWAS class count
+% Fig: TF count per GWAS class count
+% Fig : eQTL gene distance
 %
@@
 
-\begin{figure}[h!]
-    \centering
+%\begin{figure}[!ht]
+%    \centering
 %
-    \begin{subfigure}[]{.48\textwidth}vep.png
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
-    \end{subfigure}
+%    \begin{subfigure}[]{.49\textwidth}vep.png
+%    \textbf{a}
+%    \\
+%    \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+%    \end{subfigure}
 %
-    \begin{subfigure}[]{.48\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
-    \end{subfigure}
+%    \begin{subfigure}[]{.49\textwidth}
+%        \textbf{b}
+%        \\
+%        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+%    \end{subfigure}
 %
-    \caption{}
+%    % Fig 9: eQTL gene distance
+%    \begin{subfigure}[]{.49\textwidth}
+%        \textbf{c}
+%        \\
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+%    \end{subfigure}
 %
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 9: eQTL gene distance
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[h!]
-    \centering
-    %
-    \begin{subfigure}[]{.48\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
-    \end{subfigure}
-    %
-    \caption{}
-    %
-\end{figure}
+%    \caption{}
+%    %
+%\end{figure}
 
@@
 %
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
 %
-    \begin{subfigure}[]{0.95\textwidth}
+    \begin{subfigure}[]{0.99\textwidth}
         \textbf{a}
@@
         \includegraphics[width=\textwidth]{fig/model1.png}
-    \end{subfigure}
-%
-    \begin{subfigure}[]{0.95\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{fig/model2.png}
     \end{subfigure}

### 2023-02-25 scripts/cmpt_vep.py

@@ -1,4 +1,6 @@
 """Variant effect predictor"""
+import sqlalchemy
+
 from gwas2eqtl_pleiotropy.Logger import Logger
 from gwas2eqtl_pleiotropy.PathManager import PathManager
 from gwas2eqtl_pleiotropy.constants import region_bin, label_fontsize, tick_fontsize
@@ -38,7 +40,10 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 #%%
 sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
 # columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
-h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
+# h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
+engine = sqlalchemy.create_engine(url)
+with engine.begin() as conn:
+    h4_df = pandas.read_sql(sqlalchemy.text(sql), con=conn).drop_duplicates()
 
 #%%
 # gwas_cat_df = pandas.read_csv(count_per_rsid_gwas_ods_path, sep="\t")

### 2023-02-25 ms/ms03_results.tex
ENH

@@
 %
-These studies resulted in 10,621 clumped leading variants with a p-value below 5e-8 from 335 GWAS (Supplementary file).
-% TODO: Add to supplementary table?
+These studies resulted in 10,621 clumped leading variants with a p-value below 5e-8 from 335 GWAS.
 % select count(distinct gwas_id) from tophits
@@
 
-The colocalization pipeline uncovered eQTL colocalizations for 5,484 of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples in the somatic chromosomes 1-22 \cite{2019.Posthuma.Watanabe,2021.Li.Mu}.
+The colocalization pipeline uncovered eQTL colocalizations for 5,484 of the leading GWAS variants with 7,040 eQTL genes and 127 eQTL biosamples in the somatic chromosomes \cite{2019.Posthuma.Watanabe,2021.Li.Mu}.
 %
@@
 %
-The histogram shows three groups of GWAS (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}).
+The histogram shows three groups of GWAS depending on the percentage of explained leading SNPs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}a).
 %
-There were 100\% of leading variants explained with eQTLs for between 5\% and 12\% of the GWAS depending on whether
-we set PP.H4.SNP at 0.25, 0.5 or 0.75 (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc} and suppl. fig. S1).
-This parameter gives the posterior probability of a SNP in a window to be causal.
+For around 25\% of GWAS, there are not colocalizations between the leading SNPs and eQTLs.
 %
-There were a middle group of maximally 5\% GWAS with maximum percentage of explained leading SNPs between 10\% and 50\%
-depending on the PP.H4.SNP parameter 0.25-0.75
-Around 25\% of GWAS had no tophits colocalized with eQTLs for a parameter PP.SNP.H4=0.5.
+On the other side, for around 7.5\% of GWAS, 100\% of the leading SNPs colocalize with eQTLs.
 %
-Finally, between 17\% and 27\% GWAS had 0\% of leading variants explained by eQTLs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc} and suppl. fig. S1).
+The remaining GWAS show between 10\% and 50\% colocalized leading SNPs with eQTLs.
+
+The number of colocalization depends on the PP.H4.SNP parameter that gives the probability of a SNP in the colocalized locus to be the causal SNP.
+%
+We have changed this parameter between 25\% and 75\%, which results in 5\% and 12\% of the GWAS
+showing colocalization with 100\% of their leading variants.
+%
+In this range, between 16\% and 25\% of the GWAS do not show any colocalization.
+%
+The remaining GWAS show between 5\% and 10\% of their leading variants colocalized with eQTLs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}a and Supplementary figure S1).
 
@@
 
-These variants annotated with GWAS association and gene expression associations in particular biosamples are useful functional informations.
+These variants annotated with GWAS associations and eQTLs are useful functional information.
 %
@@
 % select count(*) from (select DISTINCT gwas_id FROM colocweb) as sel; 293
+
+I provide a UCSC track hub with these GWAS colocalizations.
+%
+In Fig.
+\ref{fig:hist_perc_tophits_eqtl_excl_mhc}b-d, I show three screenshots of this UCSC track hub for three different biosamples, namely B-cells, naive monocytes and frontal cortex biosamples in the chromosome 5 around the gene IL4.
+%
+This screenshot shows that depending on the biosample, the regulated genes are very different.
+%
+In the
 
@@
 
-% TODO CONTINUE
 These trait class annotations were used to investigate pleiotropy of eQTLs.
@@
 allergy, autoimmune diseases, cancer of the breast, circulatory system diseases, hypertension, and respiratory system diseases
-(Table \ref{tab:pleitropic_variants} and Supplementary table 4).
+(Table \ref{tab:pleitropic_variants}a and Supplementary table 4).
 %

### 2023-02-26 scripts/cmpt_ucsc_tracks_pleiotropic_eqtls.py
ENH ucsc tracks

@@ -33,7 +33,6 @@ df['chrom'] = 'chr' + df['chrom'].astype(str)
 df['rsid'] = 'rs' + df['rsid'].astype(str)
 df.rename({'chrom': '#chrom', 'pos38': 'chromEnd', 'rsid': 'name', 'gwas_category_count': 'score'}, axis=1, inplace=True)
 df['score'] = df['score'] * 1000 / 4
-# import pdb; pdb.set_trace()
 
 track_config_str = """browser position chr5:132239646-132497907
 browser pack Pleiotropic_eQTLs

### 2023-02-26 ms/ms03_results.tex
ENH ucsc tracks
MS

@@
 %
-The histogram shows three groups of GWAS depending on the percentage of explained leading SNPs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}a).
+The histogram shows three groups of GWAS depending on the percentage of explained leading SNPs (Fig. \ref{fig:1}a).
 %
@@
 %
-The remaining GWAS show between 5\% and 10\% of their leading variants colocalized with eQTLs (Fig. \ref{fig:hist_perc_tophits_eqtl_excl_mhc}a and Supplementary figure S1).
+The remaining GWAS show between 5\% and 10\% of their leading variants colocalized with eQTLs (Fig. \ref{fig:1}a and Supplementary figure S1).
 
@@
 In Fig.
-\ref{fig:hist_perc_tophits_eqtl_excl_mhc}b-d, I show three screenshots of this UCSC track hub for three different biosamples, namely B-cells, naive monocytes and frontal cortex biosamples in the chromosome 5 around the gene IL4.
+\ref{fig:1}b-d, I show three screenshots of this UCSC track hub for three different biosamples, namely monocytes, and  frontal cortex biosamples in the chromosome 5 around the gene IL4 (Fig.
+\ref{fig:1}a).
 %
@@
 I cluster the GWAS traits based on the Spearman correlation between beta values
-of colocalized eQTL variant, gene and biosample (Fig. \ref{fig:trait_clusters}).
+of colocalized eQTL variant, gene and biosample (Fig. \ref{fig:2}).
 
@@
 %
-Only trait classes with more than a minimum number of traits were plotted (Fig. \ref{fig:trait_clusters}).
+Only trait classes with more than a minimum number of traits were plotted (Fig. \ref{fig:2}).
 %
@@
 %
-This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases, cancers and allergies (Fig. \ref{fig:trait_clusters}).
+This analysis shows a coherent clustering of traits belonging to autoimmune, circulatory system diseases, cancers and allergies (Fig. \ref{fig:2}).
 %
@@
 74\% of colocalized eQTL/GWAS variants are associated to one GWAS class, 21\% are associated to 2 classes and the remaining variants associated to
-3 or more trait classes make less that 4\% of colocalized eQTL/GWAS variants (Fig. \ref{fig:hist_gwas_egene_etissue}a).
+3 or more trait classes make less that 4\% of colocalized eQTL/GWAS variants (Fig. \ref{fig:3}a).
 %
 Regarding eQTL genes, 36\% of the colocalized eQTL/GWAS variants modulate a single gene, 24\% modulate two genes, and the
-remaining 40\% modulate three or more genes (Fig. \ref{fig:hist_gwas_egene_etissue}b).
+remaining 40\% modulate three or more genes (Fig. \ref{fig:3}b).
 %
 Concerning tissues, 27\% colocalized eQTL/GWAS variants are active in a single tissue, 16\% in two tissues, 11\% in three
-tissues, and the remaining 46\% in four or more tissues (Fig. \ref{fig:hist_gwas_egene_etissue}c).
+tissues, and the remaining 46\% in four or more tissues (Fig. \ref{fig:3}c).
 %
 In conclusion, trait pleiotropy is rather rare among colocalized eQTL/GWAS variants, while, regulation of at least two
-genes in at least two tissues is rather common (Fig. \ref{fig:hist_gwas_egene_etissue}).
+genes in at least two tissues is rather common (Fig. \ref{fig:3}).
 

### 2023-02-26 ms/ms06_backmatter.tex
MS

@@
 
-    \begin{figure}[h!]
+    \begin{figure}[!ht]
         %
@@
             because this locus is not included in the colocalization pipeline.}
-        \label{fig:hist_perc_tophits_eqtl_excl_mhc}
+        \label{fig:1}
         %
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[!ht]
         %
         \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
-        \label{fig:trait_clusters}
+        \label{fig:2}
         %
@@
 %
-    \begin{figure}[h!]
+    \begin{figure}[!ht]
         %
-        \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
-            (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
-        %
-    \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-    \begin{figure}[h!]
-        %
-        \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL biosamples.}
-        \label{fig:correlation_counts}
-        %
-    \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 5: VEP consequences
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-    \begin{figure}[h!]
-        %
-        \caption{Odds ratio of the variants annotated with given variant effect
-        predictor (VEP) consequence and given GWAS class count compared with non
-        pleiotropic variants with GWAS class count one. The significance is given for the
-        q-value with False discovery rate 5\%.}
-        \label{fig:vep_consequence}
+        \caption{Distribution of GWAS category counts, correlation and ENSEMBL variant effect prediction.
+        \textbf{a-c} Distribution of GWAS category, eQTL gene and eQTL biosample counts.
+        Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes (\textbf{b}) and eQTL tissues (\textbf{c}).
+        \textbf{d} Spearman correlation between the counts of GWAS categories, eQTL genes and eQTL biosamples.
+        \textbf{e} Odds ratio of the eQTLs annotated with the given ENSEMBL variant predictor
+        for eQTLs association withe the given GWAS category count versus the eQTLs with a single GWAS category count.
+        The significance is given for the q-value with False discovery rate 5\%.
+        }
+        \label{fig:3}
         %
@@
 
-    \begin{figure}[h!]
+    \begin{figure}[!ht]
         %
-        \caption{TODO beta and p-val}
-        \label{fig:beta_pval}
+        \caption{eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
+        \textbf{a,b} Average eQTL and GWAS beta values.
+        \textbf{c,d} Average eQTL and GWAS p-values.
+        \textbf{e} Average alternative allele frequency in the European population from the 1000 Genomes database.
+        }
+        \label{fig:4}
         %

### 2023-02-27 ms/ms03_results.tex
MS

@@
 
-I provide a UCSC track hub with these GWAS colocalizations.
+I provide a UCSC track hub with these GWAS/eQTL colocalizations.
 %
-In Fig.
-\ref{fig:1}b-d, I show three screenshots of this UCSC track hub for three different biosamples, namely monocytes, and  frontal cortex biosamples in the chromosome 5 around the gene IL4 (Fig.
-\ref{fig:1}a).
+In Fig. \ref{fig:1}b-d, I show three screenshots of these UCSC tracks for three different biosamples, namely monocytes, and  frontal cortex biosamples in the chromosome 5 around the gene IL4.
 %
-This screenshot shows that depending on the biosample, the regulated genes are very different.
+In these UCSC tracks, red represents a positive eQTL beta coefficient, which blue represents a negative eQTL beta coefficient.
 %
-In the
+This screenshot shows that depending of gene regulation based on the biosample.
+%
+In b-cells, there is mainly the SLC22A5 gene that is regulated (Fig. \ref{fig:1}b).
+%
+On the other hand, monocytes show a significant level changes of P4HA2, SLC22A4 and SLC22A5 (Fig. \ref{fig:1}c).
+%
+Finally, the cortex biosamples show significant eQTL effects for the PDLIM4, MIR3936HG, IRF-1S1 and RAD50 (Fig. \ref{fig:1}d).
+
+To visualize simultaneously, the GWAS and eQTL beta coefficients, I have create two tracks "beta equal" and "beta unequal".
+%
+In "beta equal", both GWAS and eQTL beta coefficients have the same sign while in "beta unequal", the coefficients show different sign.
+%
+This shows that in b-cells, an expression decrease of SLC22A5 gene correlates with a decrease of Asthma predisposition and an increase of extreme height and inflammatory bowel diseases.
+%
+Altogether, I provide two new resources, namely a web application and a UCSC track hub, that might be useful for the interpretation of genetic human variants.
 
@@
 %
-Different trait classes of eQTLs were aggregated and sorted according to the number of GWAS classes the eQTLs belong to (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
+Different trait classes of eQTLs were aggregated and sorted according to the number of GWAS classes the eQTLs belong to (Table \ref{tab:pleitropic_eqtls} and Supplementary table 4).
 %
-The most pleiotropic variants belong to 6 classes that include different traits such as autoimmune diseases, circulatory disases and cancer.
+The most pleiotropic variants belong to four classes that include different traits such as autoimmune diseases, circulatory disases and cancer.
 %
-The most pleiotropic eQTLs are found in cytobands 1p13.2, 12q24.12 and 19.q13.33.
+The most pleiotropic eQTLs are found in the loci around genes ALDH2 (12q24.12), ENO1 (1p36.23), ORMDL3 (17.q21.1), IL4 (5q31.1), BACH2 (6q15), FUT2 (19q13.33), SLC39A8 (4q24) and CYP1A1 (15q24.1)
+(Table \ref{tab:pleitropic_eqtls} and Supplementary table 4).
 %
-The eQTL rs2476601 is well known for rheumatoid arthritis, juvenile athritis, diabetes and cancer of skin REFs.
+We have generated this list based on different SNP.PP.H4 parameter values 0.25, 0.5 and 0.75 and we find the same variants except rs8067378 (Data not shown).
 %
-Our analysis annotates this variant with eQTL genes AP4B1,AP4B1-AS1,BCL2L15,DCLRE1B,MAGI3,PHTF1 and PTPN22 tissues Adipose,Brain,Colon,Fibroblast,ImmuneCell LCL and iPSC
+Most of these pleiotropic eQTLs are well known with Google Scholar citations (Accessed Feb. 27, 2023).
 
-% TODO add rs2522051
-This variant rs2522051 is involved in six classes of traits, namely,
-allergy, autoimmune diseases, cancer of the breast, circulatory system diseases, hypertension, and respiratory system diseases
-(Table \ref{tab:pleitropic_variants}a and Supplementary table 4).
+%By contrast, the variant rs13107325 has 596
+% between 596 for rs13107325
+%%
+%The eQTL rs2476601 is well known for rheumatoid arthritis, juvenile athritis, diabetes and cancer of skin REFs.
+%%
+%Our analysis annotates this variant with eQTL genes AP4B1,AP4B1-AS1,BCL2L15,DCLRE1B,MAGI3,PHTF1 and PTPN22 tissues Adipose,Brain,Colon,Fibroblast,ImmuneCell LCL and iPSC
 %
-The variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1
-%
-This variant rs2522051 is active in 23 classes of tissues including adipose, brain, digestive cells, immune cells, and sexual organs.
-%
-This variant rs2522051 regulates several genes such as IL13, PDLIM4, and RAD50.
+%% TODO add rs2522051
+%This variant rs2522051 is involved in six classes of traits, namely,
+%allergy, autoimmune diseases, cancer of the breast, circulatory system diseases, hypertension, and respiratory system diseases
+%(Table \ref{tab:pleitropic_eqtls}a and Supplementary table 4).
+%%
+%The variant rs2522051 is located downstream of the IRF1 gene in cytoband 5q31.1
+%%
+%This variant rs2522051 is active in 23 classes of tissues including adipose, brain, digestive cells, immune cells, and sexual organs.
+%%
+%This variant rs2522051 regulates several genes such as IL13, PDLIM4, and RAD50.
 
@@
 Pleiotropic eQTLs are concentrated in genomic regions such as in cytobands 3q23,
-5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Table \ref{tab:pleitropic_variants} and Supplementary table 4).
+5q31.1, 9p21.3, 15q24.1, and 19q13.33 for eQTLs involved on five or more classes (Table \ref{tab:pleitropic_eqtls} and Supplementary table 4).
 %
@@
 
-%
 % TODO Verify values
@@
 
-%\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
+\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
+
+Pleiotropic eQTLs affect simultaneously more trait classes.
 %
-%Pleiotropic eQTLs affect simultaneously more trait classes.
-%%
-%Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
-%%
-%I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
-%%
-%Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
 %
-%Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
-%%
-%I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
-%%
-%Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
 %
-%These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).
+Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+
+Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
+%
+I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
+%
+Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
+
+These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).

### 2023-02-27 ms/ms08_figures.tex
ENH ucsc tracks
MS

@@
 
-    \begin{subfigure}[]{0.49\textwidth}
+    \begin{subfigure}[]{0.32\textwidth}
         \textbf{a}
@@
     \end{subfigure}
-    %
+
     \caption{}
-    %
+
 \end{figure}
@@
 
-    \caption{} \label{fig:gwas_egene_etisue_per_variant}
-    %
+    \caption{}
+
 \end{figure}
@@
     \centering
-%
+
     \begin{subfigure}[]{0.99\textwidth}
@@
 
-    \caption{\textbf{Model of regulatory variant pleiotropy.} I have investigated three possible mechanisms of pleiotropy.
-\textbf{Left}, Pleiotropic variants have more eGenes that result in more functions and more phenotypes.
-This might arise from an enrichment of pleiotropic variants in splicing or 3' UTR regions.
-\textbf{Center}, I have also found that eGenes of pleiotropic variants are active in more etissues which result in more GWAS phenotypes.
-This might be explained from variants being bound by more transcription factors.
-\textbf{Left} Triplets of variant-eGene-eTissues are associated with more GWAS phenotypes, which directly affect the number of GWAS phenotypes.
-I have found that this might be explained by an enrichment of missense alleles.} \label{fig:beta}
-%
 \end{figure}

### 2023-03-11 ms/ms07_tables.tex
MS

@@
 
-\begin{table}[h!]
+\begin{table}[!ht]
     \centering
@@
     %
-    \caption{Represantive pleiotropic eQTLs involved in four GWAS categories in different cytobands.
-    The gene marker is the eQTL gene with the highest PubMed publication count.
-    The whole list is given in the Supplementary table ST2.}
-    \label{tab:pleitropic_variants}
+    \caption{
+%        Represantive pleiotropic eQTLs involved in four GWAS categories in different cytobands.
+%    The gene marker is the eQTL gene with the highest PubMed publication count.
+%    The whole list is given in the Supplementary table ST2.
+    }
+    \label{tab:pleitropic_eqtls}
 \end{table}
@@
     %
-    \caption{Pleiotropic regions involving 6 or more GWAS classes.
-These regions were built with a sliding window of 1e5 nt.
-Genomic coordinates are given for the hg38 assembly.}\label{tab:pleiotropic_regions}
+    \caption{
+%        Pleiotropic regions involving 6 or more GWAS classes. These regions were built with a sliding window of 1e5 nt. Genomic coordinates are given for the hg38 assembly.
+    }\label{tab:pleiotropic_regions}
     %

### 2023-03-16 ms/ms08_figures.tex
MS
ENH

@@
 % Fig 1: Histogram of percentage of explained of GWAS
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!ht]
+    \centering
+
+        \includegraphics[width=0.5\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
+
+    \caption{}
+%
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % Fig 2: ucsc screenshots
@@
 
-    \begin{subfigure}[]{0.32\textwidth}
+    \begin{subfigure}[]{0.99\textwidth}
         \textbf{a}
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
+        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_bcell_help.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_bcell.png}
+        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_monocyte.png}
     \end{subfigure}
@@
         \textbf{c}
-        \\
-        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_monocyte.png}
-    \end{subfigure}
-
-    \begin{subfigure}[]{0.99\textwidth}
-        \textbf{d}
         \\

### 2023-03-21 ms/ms08_figures.tex
ENH added watanabe analysis
ENH

@@
 
-% \begin{figure}[!tbp]
-% \centering
-% %
-% \begin{subfigure}[]{.33\textwidth}
-% \textbf{a}
-% \\
-% \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
-% \end{subfigure}
-% %
-% \begin{subfigure}[]{.33\textwidth}
-% \textbf{b}
-% \\
-% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
-% \end{subfigure}
-% %
-% \caption{Analysis of pleiotropic regions.
+ \begin{figure}[!tbp]
+ \centering
+ %
+ \begin{subfigure}[]{.33\textwidth}
+ \textbf{a}
+ \\
+ \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_hist_100000.png}
+ \end{subfigure}
+ %
+ \begin{subfigure}[]{.33\textwidth}
+ \textbf{b}
+ \\
+ \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
+ \end{subfigure}
+ %
+ \caption{Analysis of pleiotropic regions.
 \textbf{a}, Length distribution of pleiotropic regions.
 \textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
-% \end{figure}
+ \end{figure}
 

### 2023-03-24 ms/ms08_figures.tex
ENH

@@
 
-\begin{figure}[!ht]
-    \centering
+\begin{figure}[!tbp]
 
-        \includegraphics[width=0.5\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
+\begin{subfigure}[]{.24\textwidth}
+\textbf{a}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/autoimmune_disease.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.24\textwidth}
+\textbf{b}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/cardiovascular_disease.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.24\textwidth}
+\textbf{c}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/digestive_system_disease.png}
+\end{subfigure}
+%
+\begin{subfigure}[]{.24\textwidth}
+\textbf{d}
+\\
+\includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/mental_or_behavioural_disorder.png}
+\end{subfigure}
 
 \caption{}
+\label{fig:region_gwas_egenes_tissues}
 %
@@
 %
-% Fig 2: Trait distance matrix
+% Fig 3: Trait distance matrix
 %
@@
 % Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
-% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
-% Fig 5: VEP consequences
+% Fig: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
+% Fig: VEP consequences
 %
@@
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{c}
+        \textbf{e}
         \\
@@
     \end{subfigure}
+    %
+    % Fig 9: eQTL gene distance
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{e}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+    \end{subfigure}
 
@@
         \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
-    \end{subfigure}
-
-    % Fig 9: eQTL gene distance
-    \begin{subfigure}[]{.49\textwidth}
-        \textbf{e}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
     \end{subfigure}

### 2023-03-24 ms/ms08_figures.tex
ENH

@@
 %
-% Fig 1: Histogram of percentage of explained of GWAS
+% Fig 1a-d: Histogram of percentage of explained of GWAS
+% Fig 1e: Trait distance matrix
 %
@@
 
-\caption{}
-\label{fig:region_gwas_egenes_tissues}
-%
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Fig 2: ucsc screenshots
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!ht]
-    \centering
-
-    \begin{subfigure}[]{0.99\textwidth}
-        \textbf{a}
+    \begin{subfigure}[]{\textwidth}
+        \textbf{e}
         \\
-        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_bcell_help.png}
-    \end{subfigure}
-
-    \begin{subfigure}[]{0.99\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_monocyte.png}
-    \end{subfigure}
-
-    \begin{subfigure}[]{0.99\textwidth}
-        \textbf{c}
-        \\
-        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_frontalcortex.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
     \end{subfigure}
@@
 %
-% Fig 3: Trait distance matrix
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!ht]
-    %
-    \centering
-    %
-    \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
-    %
-    \caption{}
-    %
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 % Fig 2: Length histogram of regions
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
- \begin{figure}[!tbp]
- \centering
- %
- \begin{subfigure}[]{.33\textwidth}
- \textbf{a}
- \\
- \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_hist_100000.png}
- \end{subfigure}
- %
- \begin{subfigure}[]{.33\textwidth}
- \textbf{b}
- \\
- \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
- \end{subfigure}
- %
- \caption{Analysis of pleiotropic regions.
-\textbf{a}, Length distribution of pleiotropic regions.
-\textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
- \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 % Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
@@
     \centering
+
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{a}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_hist_100000.png}
+    \end{subfigure}
     %
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{a}
+        \textbf{b}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{c}
         \\
@@
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{b}
+        \textbf{d}
         \\
@@
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{d}
+        \textbf{f}
         \\
@@
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{e}
+        \textbf{g}
         \\
@@
     \begin{subfigure}[]{.49\textwidth}
-        \textbf{e}
+        \textbf{f}
         \\

### 2023-03-26 ms/ms08_figures.tex
ENH

@@
 \begin{figure}[!tbp]
-
-    \begin{subfigure}[]{.24\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/autoimmune_disease.png}
-    \end{subfigure}
-    %
-    \begin{subfigure}[]{.24\textwidth}
-        \textbf{b}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/cardiovascular_disease.png}
-    \end{subfigure}
-    %
-    \begin{subfigure}[]{.24\textwidth}
-        \textbf{c}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/digestive_system_disease.png}
-    \end{subfigure}
-    %
-    \begin{subfigure}[]{.24\textwidth}
-        \textbf{d}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/mental_or_behavioural_disorder.png}
-    \end{subfigure}
+    \centering
 
     \begin{subfigure}[]{\textwidth}
-        \textbf{e}
-        \\
         \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
@@
 
-    \begin{subfigure}[]{.49\textwidth}
-        \textbf{g}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
-    \end{subfigure}
-
     \caption{}
@@
         \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{g}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
     \end{subfigure}

### 2023-04-02 scripts/cmpt_count_per_rsid.py
ENH improved figures

@@ -7,7 +7,7 @@ import sys
 import sqlalchemy
 from matplotlib import pyplot as plt
 
-from gwas2eqtl_pleiotropy.constants import label_fontsize
+from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize
 
 #%%
 help_cmd_str = "todo"
@@ -112,13 +112,16 @@ ms_df.to_csv(tsv_path, sep="\t", index=False)
 
 #%%
 m2df = m_df[['gwas_category_count', 'etissue_category_term_count', 'egene_count']].copy()
-m2df.rename({'gwas_category_count': 'GWAS. cat. cnt.', 'etissue_category_term_count': 'eTissue cat. cnt.', 'egene_count': 'eGene cnt.', }, axis=1, inplace=True)
+m2df.rename({'gwas_category_count': 'Trait cat. cnt.', 'etissue_category_term_count': 'Tissue cnt.', 'egene_count': 'Gene cnt.', }, axis=1, inplace=True)
 corr = m2df.corr(method='spearman')
 plt.subplots_adjust(left=0.2, right=0.8, top=0.9, bottom=0.)
-ax = seaborn.heatmap(corr, annot=True, xticklabels=False)
-ax.set_yticklabels(ax.get_ymajorticklabels(), fontsize=12, rotation=45)
-plt.title("Spearman correlation", fontsize=16)
+ax = seaborn.heatmap(corr, annot=True, xticklabels=False, cmap="rocket_r", annot_kws={"size": 20})
+ax.set_yticklabels(ax.get_ymajorticklabels(), fontsize=20, rotation=45)
+#plt.yticks(fontsize=tick_fontsize, rotation=45)
+plt.title("Spearman correlation", fontsize=label_fontsize)
+plt.subplots_adjust(left=0.4, right=1)
 plt.savefig(count_per_rsid_gwas_egene_etissue_corr_png)
+plt.tight_layout()
 plt.clf()
 plt.close()
 

### 2023-04-02 scripts/plt_x_per_variant_y_egene_distance.py
ENH

@@ -106,6 +106,8 @@ plt.savefig(hist_png_path)
 plt.close()
 
 #%% Violin plot
+# import pdb; pdb.set_trace()
+m2_df['eqtl_gene_distance'] = m2_df['eqtl_gene_distance']/1000
 ax = seaborn.barplot(data=m2_df, x=x, y=y, estimator=numpy.mean, palette="rocket_r")
 
 annotator = Annotator(ax, pairs, data=m2_df, x=x, y=y, order=order)
@@ -114,7 +116,7 @@ annotator.apply_and_annotate()
 
 plt.title("Distance to closest gene", fontsize=label_fontsize)
 plt.xlabel("GWAS category count", fontsize=label_fontsize)
-plt.ylabel("Distance mean", fontsize=label_fontsize)
+plt.ylabel("Mean distance [kbp]", fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize, rotation=0)
 plt.yticks(fontsize=tick_fontsize)
 
@@ -122,6 +124,7 @@ plt.tight_layout()
 hist_png_path = os.path.join(outdir_path, "violin.png")
 plt.savefig(hist_png_path)
 plt.close()
+m2_df['eqtl_gene_distance'] = m2_df['eqtl_gene_distance']*1000
 
 ########################################################################################################################
 #

### 2023-04-02 scripts/pltbar_x_per_variant_etissue_y_egene.py

@@ -104,12 +104,13 @@ annotator = Annotator(ax, pairs, data=m_df, x=x, y=y, order=order)
 annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
 annotator.apply_and_annotate()
 
+ax.set_xticklabels(xticklabels)
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel(xlabel, fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize, rotation=0)
 plt.ylabel(ylabel, fontsize=label_fontsize)
+plt.ylim([1, 3])
 plt.yticks(fontsize=tick_fontsize)
-ax.set_xticklabels(xticklabels)
 
 plt.tight_layout()
 plt.savefig(vlnplt_png_path, dpi=dpi)

### 2023-04-02 scripts/pltbar_x_per_variant_pleiotropy_y_remapcrm.py

@@ -117,7 +117,7 @@ annotator.configure(**annotator_config_dic)
 annotator.annotate()
 
 ax.set_xticklabels(xticklabels)
-plt.grid(axis="y")
+# plt.grid(axis="y")
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel(xlabel, fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize, rotation=0)
@@ -137,7 +137,7 @@ annotator.configure(**annotator_config_dic)
 annotator.annotate()
 
 ax.set_xticklabels(xticklabels)
-plt.grid(axis="y")
+# plt.grid(axis="y")
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel(xlabel, fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize, rotation=0)

### 2023-04-02 scripts/pltbox_x_per_rsid_y_remapnr.py

@@ -1,6 +1,7 @@
 import os
 import pathlib
 
+import numpy
 import pandas
 import seaborn
 import sys
@@ -72,7 +73,7 @@ pairs = [(str(1), str(int(i))) for i in sorted(cat_df['gwas_category_count'].uni
 cat_df[x] = cat_df[x].astype(int).astype(str)
 
 #%%
-ax = seaborn.boxplot(x=x, y=y, data=cat_df, order=order, **boxplot_kwargs)
+ax = seaborn.barplot(x=x, y=y, data=cat_df, order=order, estimator=numpy.mean, palette="rocket_r")
 
 annotator = Annotator(ax, pairs, data=cat_df, x=x, y=y, order=order)
 annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)

### 2023-04-20 ms/ms07_tables.tex
ENH

@@
     head,
-    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries GWAS Categories\\\hline,
+    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries Trait categories\\\hline,
     ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
@@
         head,
-        table head=\\\hline \bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries GWAS Categories\\\hline,
+        table head=\\\hline \bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries Trait categories\\\hline,
     ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names

### 2023-04-20 scripts/cmpt_count_per_rsid.py

@@ -132,7 +132,7 @@ m2df = m_df[['rsid', 'ref', 'alt', 'gwas_category_count', 'domains_watanabe2019'
 order = [*range(1, m2df['gwas_category_count'].max()+1)]
 ax = seaborn.boxplot(x='gwas_category_count', y='domains_watanabe2019', data=m2df, order=order, palette="rocket_r")
 
-plt.xlabel("GWAS category count", fontsize=label_fontsize)
+plt.xlabel("trait category count", fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize)
 plt.yticks(fontsize=tick_fontsize)
 plt.ylabel("Watanabe cat. cnt.", fontsize=label_fontsize)
@@ -154,7 +154,7 @@ order = [*range(1, m2df['gwas_category_count'].max()+1)]
 ax = seaborn.barplot(x=m2df_watanabe2019_cat_count_df.index, y='watanabe_perc', data=m2df_watanabe2019_cat_count_df, order=order, palette="rocket_r")
 
 plt.title("Known in Watanabe 2019", fontsize=label_fontsize)
-plt.xlabel("GWAS category count", fontsize=label_fontsize)
+plt.xlabel("trait category count", fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize)
 plt.yticks(fontsize=tick_fontsize)
 plt.ylabel("Percentage", fontsize=label_fontsize)

### 2023-04-20 scripts/plt_violin_x_per_egene_y_interactors.py

@@ -101,7 +101,7 @@ order = [*range(1, max_gwas_class_count+1)]
 xticklabels = order.copy()
 xticklabels[-1] = '≥{}'.format(order[-1])
 title = "Interactors per egene"
-xlabel = "GWAS category count"
+xlabel = "Trait category count"
 ylabel = "Phys. interactor count"
 y = "interactor_count"
 

### 2023-04-20 scripts/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py

@@ -78,12 +78,12 @@ xticklabels = order.copy()
 # xticklabels[-1] = '≥{}'.format(order[-1])
 box_pairs = [(1, i) for i in range(1, max_gwas_category_count+1) ]
 x = 'gwas_category_count'
-xlabel = "GWAS category count"
+xlabel = "Trait category count"
 
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # boxplot and mann-whitney
 y = "gwas_neglog10pval"
-title = "GWAS neg. log10 p-val"
+title = "GWAS significance"
 ylabel = "Neg. log10 p-val mean"
 
 #%%
@@ -110,7 +110,7 @@ plt.close()
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # boxplot and mann-whitney
 y = "eqtl_neglog10pval"
-title = "eQTL neg. log10 p-val"
+title = "eQTL significance"
 ylabel = "Neg. log10 p-val mean"
 
 #%%
@@ -138,7 +138,7 @@ plt.close()
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # boxplot and mann-whitney
 y = "eqtl_beta_abs"
-title = "eQTL beta"
+title = "eQTL effect"
 ylabel = "Absolute beta mean"
 
 #%%
@@ -166,7 +166,7 @@ plt.close()
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # boxplot and mann-whitney
 y = "gwas_beta_abs"
-title = "GWAS beta"
+title = "GWAS effect"
 ylabel = "Absolute beta mean"
 
 #%%
@@ -193,7 +193,7 @@ plt.close()
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # violin and mann-whitney
[…]

### 2023-04-20 scripts/pltbar_x_per_variant_egene_y_etissue.py

@@ -83,7 +83,7 @@ order = [str(x) for x in range(1, max(m_df['gwas_category_count'].unique())+1)]
 xticklabels = order.copy()
 # xticklabels[-1] = '≥{}'.format(order[-1])
 title = "Tissues per eQTL-gene "
-xlabel = "GWAS category count"
+xlabel = "Trait category count"
 ylabel = "Tissue count mean"
 y = "etissue_category_term_count"
 x = "gwas_category_count"

### 2023-05-04 doc/poster_eccb22_barcelona/poster_eccb22_barcelona.tex
DOC presentation

--- 2022-08-31 poster_eccb22_barcelona/poster_eccb22_barcelona.tex
@@
 	%\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
-	\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
+	\documentclass[portrait,a0b,final]{a0poster}
 	%%% Option "a4resizeable" makes it possible ot resize the
@@
 	\item Here we use this GWAS database: The IEU OpenGWAS project\\ \url{https://gwas.mrcieu.ac.uk}
-	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (\citep{2019.Posthuma.Watanabe}).
+	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome \citep{2019.Posthuma.Watanabe}.
 	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is not well understood.
@@
 	\item Here we use this eQTL database: The EBI eQTL Catalogue\\ \url{https://www.ebi.ac.uk/eqtl}
+	\item Databases of GWAS and eQTL studies allow large colocalization analysis \citep{2021.Li.Mu}.
 	\item The R package coloc performs colocalization tests between genetic traits:\\ \url{https://CRAN.R-project.org/package=coloc}
@@
 	\begin{center}\pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightblue,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whiteblue,gradmidpoint=1.0,framesep=0.5em}{\begin{center}\fontheader{Results}\end{center}}\end{center}
-
-	- A pipeline of eQTL/GWAS variant colocalization (Screenshot pythonanywhere)
-	- Table with most pleiotropic variants (Table 1)
-	- Molecular mechanisms Fig 7 and fig 8a
-	- Model of mechanisms Fig 10
 
@@
 		    \end{minipage}
-		    \mycaptionfig{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.}
+		    \mycaptionfig{Gene ontology analysis of genes associated with eQTLs (eGenes) of variants associated to three (\textbf{Upper}) and two (\textbf{Bottom}) GWAS categories compared to eGenes of variants associated to one GWAS category.}
 	    \end{center}
@@
 		    \end{minipage}
-		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (Left), splice region variant (Center) and 3’ UTR variant (Right)}
+		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (\textbf{Left}), splice region variant (\textbf{Center}) and 3’ UTR variant (\textbf{Right})}
 	    \end{center}
@@
 
-	\vspace{0.5cm} % intersection space
+	\vspace{0.9cm} % intersection space
 
@@
 		    \end{minipage}
-		    \mycaptionfig{Count of eGenes per variant-tissue (\textbf{Left}), eTissues per variant-eGene (\textbf{Center}) and GWAS categories per variant-eGene-eTissue (\textbf{Right}).}
+		    \mycaptionfig{Count of eGenes per variant-tissue (\textbf{Left}), samples of eQTLs (eTissues) per variant-eGene (\textbf{Center}) and GWAS categories per variant-eGene-eTissue (\textbf{Right}).}
 	    \end{center}
@@
 		    \mycaptionfig{Model of regulatory variant pleiotropy. I have investigated
-	three possible mechanisms of pleiotropy. Left, Pleiotropic variants have more
+	three possible mechanisms of pleiotropy. (\textbf{Left}), Pleiotropic variants have more
 	eGenes that result in more functions and more phenotypes. This might arise
 	from an enrichment of pleiotropic variants in splicing or 3’ UTR regions.
-	Center, I have also found that eGenes of pleiotropic variants are active
+	(\textbf{Center}), I have also found that eGenes of pleiotropic variants are active
 	in more etissues which result in more GWAS phenotypes. This might be
-	explained from variants being bound by more transcription factors. Left
+	explained from variants being bound by more transcription factors. (\textbf{Right}),
 	Triplets of variant-eGene-eTissues are associated with more GWAS pheno-
@@
 	\begin{itemize}	
-	\item We have developed a pipeline of eQTL/GWAS colocalization (Fig. 1).
-	\item These colocalization results are in an easy to use web site (Fig. 1).
-	\item We have computed a list of pleiotropic eQTLs (Table 1).
-	\item Pleiotropic variants are enriched in missense and splice region VEP effect consequences (Fig. 2).
-	\item Pleiotropic eQTLs have more eGenes per variant-eTissue, are active in more eTissues per variant-eGene and are associated to more GWAS categories per variant-eGene-eTissue (Fig. 3).
-	\item Pleiotropic variants bind transcription factors (Fig. 4).
-	\item Model explaining the reason of pleiotropy for eQTLs (Fig. 5).
+	\item We have developed a pipeline of eQTL/GWAS colocalization based on 127 eQTL studies and 150 GWAS (\textbf{Fig. 1}).
+	\item These colocalization results are available in an easy to use web site (\textbf{Fig. 1}).
+	\item We have computed a list of pleiotropic eQTLs (\textbf{Table 1}).
+	\item eGenes of pleiotropic variants are enriched in immune related functions (\textbf{Fig. 2}).
+	\item Pleiotropic variants are enriched in missense and splice region VEP effect consequences (\textbf{Fig. 3}).
+	\item Pleiotropic eQTLs have more eGenes per variant-eTissue, are active in more eTissues per variant-eGene and are associated to more GWAS categories per variant-eGene-eTissue (\textbf{Fig. 4}).
+	\item Pleiotropic variants bind more transcription factors (\textbf{Fig. 5}).
+	\item We propose a model explaining the hypotheses underlying the pleiotropy for eQTLs (\textbf{Fig. 6}).
 	\end{itemize}

### 2023-05-04 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC presentation
MS presentation

--- 2022-10-18 presentation_gold2022_paris/presentation_gold2022_paris.tex
@@
 
+\newcommand*{\presentationgoldparispath}{../presentation_230120_gold2022_paris/fig/}%
+
 % Change size of footnotes
 \renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
-\title{Analysis of gene regulatory properties underlying trait pleiotropy}
-\subtitle{Meeting GOLD 2022}
+\title{Pleiotropic expression quantitative trait loci are located in cis-regulatory modules, closer to more target genes and with smaller effects on the genes}
+\subtitle{Centuri Day 2023}
 \author{Aitor Gonz\'alez}
 \institute{Aix Marseille Univ, INSERM, TAGC}
-\date{Oct. 19, 2022}
+\date{May 11, 2023}
 
@@
 \begin{frame}
-\frametitle{Genome-wide association studies (GWAS)}
+\frametitle{Genetic variants and diseases}
 
 \begin{itemize}
-\item Genetic alleles with higher frequency in individuals with a given disease
-\item A very active field with many studies
-\item Many genetic variants are involved in several phenotypes
+\item Phenotypic differences between human populations are mostly coded in frequent genetic variants
+\item These genetic variants also cause different disease susceptibility
+\item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
 \end{itemize}
@@
 %
-Limitations
-%
-\begin{itemize}
-\item Correlation between variants
-\item Which cell types are causal to the disease?
-\item Over 90\% GWAS variants fall in non-coding regions
-\end{itemize}
-%
-\vfill
-%
-These limitations can be partially addressed by looking at eQTLs in different tissues
-
-\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+What is the molecular mechanism and effect of tissue of these variants?
 
@@
 
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}
+\begin{itemize}
+\item Quantitative trait loci (QTL) are variants associated with changes of quantitative phenotypes
+\item Expression (eQTL) are variants associated with changes of gene expression in a given tissue
+\item eQTLs provide a putative molecular mechanism of a variant
+\end{itemize}
+%
+\vfill
+%
+What are the properties of pleiotropic eQTLs?
 
-\end{column}
-\end{columns}
+\includegraphics[width=\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
 
@@
 \begin{frame}
-\frametitle{Colocalisation analysis of GWAS and eQTL variants}
+\frametitle{Approach}
 
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4a.jpg}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/doi_10.3389_fgene.2020.00424_fig4bc.png}
-     \end{center}
-\end{column}
-\end{columns}
+\begin{enumerate}
+\item Compute colocalization of eQTLs and GWAS variants
+\item Split eQTLs by the number of associated disease categories
+\item Compare properties of pleiotropic eQTLs
+\end{enumerate}
+%
+\vfill
+%
+What are the properties of pleiotropic eQTLs?
+
+\includegraphics[width=\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
 
@@
 \begin{frame}
-\frametitle{Objectives}
-
-\begin{enumerate}
-\item Large scale colocalization analysis between GWAS and eQTL variants
-\item To identify pleiotropic genetic loci
-\item To characterize the regulatory causes of trait pleiotropy
-\end{enumerate}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Data and Algorithms}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-\begin{itemize}
-\item IEA OpenGWAS database$^1$: 244,881e6 associations from 42e3 GWAS datasets
-\item EBI eQTLs database$^2$: Processed gene expression and splicing QTLs from all public studies on human.
-\item CRAN - Package coloc$^3$: Genetic colocalisation analysis of two phenotypes to look for shared common genetic causal variants
-\end{itemize}
-\end{column}
-\begin{column}{0.5\textwidth}
-\begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/pgen.1004383.g001.png}
-\end{center}
-\end{column}
-\end{columns}
-
-\let\thefootnote\relax\footnotetext{$^1$https://gwas.mrcieu.ac.uk}
-\let\thefootnote\relax\footnotetext{$^2$https://www.ebi.ac.uk/eqtl}
-\let\thefootnote\relax\footnotetext{$^3$https://cran.r-project.org/web/packages/coloc/index.html}
-\end{frame}
-
-\section{Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Large scale GWAS/eQTL colocalization analysis}
-
-\begin{itemize}
-\item Colocalization analysis of 418 GWAS and 127 eQTL studies (53e3 pairs)
-\item 143e3 shared causal variants with probability $\geq$ 0.8
-\item Percentage of GWAS loci explained by eQTL: 0$\%$ - 77$\%$ (Average 33$\%$)
-\end{itemize}
-
-\begin{figure}[!]
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/coloc_web.png}
-\end{figure}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Do GWAS traits cluster coherently?}
-
-Distance between GWAS traits based on eQTL/egene beta
-
-\begin{figure}[!]
-\includegraphics[width=0.5\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/plthtmp_disease_comorbidity_matrix.py/corr.png}
-\end{figure}
-
-GWAS/eQTL colocalization analysis allows to cluster diseases coherently
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{frame}
-%\frametitle{Diseases?}
-
-%\begin{table}[!tbp]
-%\centering
-%\scriptsize
-%\hline
-%\csvreader[separator=tab,
-%tabular=ccrrp{0.4\textwidth},
-%head,
-%table head=\bfseries Chrom. & \bfseries Cytoband & \bfseries Pos (hg38) & \bfseries Variant & \bfseries GWAS Categories\\\hline,
-%]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-%{\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv & \csvcolv}% specify your coloumns here
-%\hline
-%%
-%\vspace{15pt}
-%%
-%\caption{Colocalized eQTL/GWAS variants involved in 5 or more GWAS categories. Genomic coordinates are given for the hg38 assembly. }\label{tab:pleitropic_variants}
-%\end{table}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{How are pleiotropic variants distributed?}
-
-\begin{columns}
-\begin{column}{0.32\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_gwas.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-     \end{center}
-\end{column}
-\begin{column}{0.32\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_egene.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-     \end{center}
-\end{column}
-\begin{column}{0.32\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltsctr_x_per_rsid_y_etissue.py/count_per_rsid_chr5_start131912097_end132802472_categories7.png}
-     \end{center}
-\end{column}
-\end{columns}
+\frametitle{Summary}
 
 \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/chr5_131912097_132802472.png}
+\includegraphics[width=0.5\textwidth]{fig/graphical_abstract.drawio.png}
 \end{center}
-
-\small
-\begin{itemize}
-\item The cytokine locus: chr5:131,912,097-132,802,472. Genes IRF1, IL3, IL4, IL5, ...
-\item Diseases: Allergy,Autoimmune dis.,Cancer of breast,Circulatory sys. dis.,Hypertension,Respiratory system dis.
-\item For instance: rs2522051 regulates 12 e-genes in 23 e-tissues
-\end{itemize}
-\normalsize
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Where are the pleiotropic variants?$^1$}
-
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_vep_consequence.py/vep.png}
-     \end{center}
-
-Upstream and downstream of genes and in the intron
-
-\let\thefootnote\relax\footnotetext{$^1$https://www.ensembl.org/vep}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Why are these variants pleiotropic? Do pleiotropic variants associate to more egenes?}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_egenes.png}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
-     \end{center}
-\end{column}
-\end{columns}
-
-More egenes per variant-etissue pair.
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Do pleiotropic variants associate to more etissues?}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_etissues.png}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
-     \end{center}
-\end{column}
-\end{columns}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Do pleiotropic variants associate to more traits (after fixing egene-etissue)?}
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/presentation_gold2022_paris/fig/model_pleio_gwas.png}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
-     \end{center}
-\end{column}
-\end{columns}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Do pleiotropic variants bind more transcription factors?$^1$}
-
-This could explain why pleiotropic variants active in more etissues and associated to more egenes
-
-\begin{columns}
-\begin{column}{0.5\textwidth}
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
-     \end{center}
-\end{column}
-\begin{column}{0.5\textwidth}  %%<--- here
-    \begin{center}
-\includegraphics[width=\textwidth]{/home/gonzalez/Repositories/gwas2eqtl_pleiotropy/out/gwas418/pval_5e-08/r2_0.1/kb_1000/window_1000000/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
-     \end{center}
-\end{column}
-\end{columns}
-
-\end{frame}
-
-\let\thefootnote\relax\footnotetext{$^1$https://remap.univ-amu.fr}
-
-\section{Conclusions, perspective and acknowledgements} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Conclusions}
-
-Regulatory characteristics of pleiotropic variants
-%
-\begin{itemize}
-\item Variant pleiotropy is correlated in regions
-\item Enrichment of variants upstream (incl. 5'UTR) and downstream of genes (incl. 3'UTR)
-\item Enrichment of transcription factor and CRM binding
-\item Increased number of associated egenes and etissues
-\end{itemize}
-%
-\vfill
-%
-Other outcomes of this study
-%
-\begin{itemize}
-\item A large analysis of GWAS/eQTL colocalisations
-\item A list of pleiotropic variants and regions
-\end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Perspectives}
-
-\begin{itemize}
-\item A web portail to share data
-\item Causality of egenes on gwas (Looking for experts ;-) )
-\end{itemize}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}
-\frametitle{Acknowledgements}
-
-\begin{itemize}
-\item L\'eopoldine Lecerf (M1)
-\item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
-\end{itemize}
-%
-\vfill
-%
-Funding
-%
-\begin{itemize}
-\item Institut Cancer et Immunologie (ICI)
-\item Agence nationale de la recherche (ANR)
-\item GOLD
-\end{itemize}
 

### 2023-05-04 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC

@@
 \newcommand*{\presentationgoldparispath}{../presentation_230120_gold2022_paris/fig/}%
+\newcommand*{\mspath}{../../out/gwas417/pval_5e-08/r2_0.1/kb_1000/window_1000000/75_50}%
 
@@
         \end{itemize}
-%
         \vfill
-%
         What are the properties of pleiotropic eQTLs?
 
-\includegraphics[width=\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
+        \includegraphics[width=0.7\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
 
@@
     \begin{frame}
-\frametitle{Approach}
+        \frametitle{Annotation of eQTLs traits using colocalization analysis}
+
+        \begin{center}
+            \includegraphics[width=0.6\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4bc.png}
+        \end{center}
+
+        \let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Strategy}
 
@@
             \item Split eQTLs by the number of associated disease categories
-\item Compare properties of pleiotropic eQTLs
+            \item Compare molecular properties of pleiotropic eQTLs
         \end{enumerate}
-%
-\vfill
-%
-What are the properties of pleiotropic eQTLs?
 
-\includegraphics[width=\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
+    \end{frame}
 
-\let\thefootnote\relax\footnotetext{Cano-Gamez et al. 2020. doi:10.3389/fgene.2020.00424}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Are pleiotropic variants associated to more eQTL genes?}
+
+        \begin{center}
+            \includegraphics[width=0.5\textwidth]{\mspath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+        \end{center}
+
     \end{frame}

### 2023-05-05 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC presentation

@@
 \documentclass{beamer}
-\usetheme{Boadilla}
+\usetheme{Warsaw}
+\usecolortheme{seahorse}
 
@@
 \usepackage{csvsimple}
+\usepackage{hyperref}
 
@@
 \renewcommand{\footnotesize}{\fontsize{5pt}{5pt}\selectfont}
-\title{Pleiotropic expression quantitative trait loci are located in cis-regulatory modules, closer to more target genes and with smaller effects on the genes}
+\title{Identification and analysis of pleiotropic expression quantitative trait loci}
 \subtitle{Centuri Day 2023}
@@
 
+    \section{Results} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \begin{frame}
-        \frametitle{Are pleiotropic variants associated to more eQTL genes?}
+        \frametitle{Is the effect size of pleiotropic eQTLs different?}
+
+        \begin{center}
+            \includegraphics[width=0.5\textwidth]{\mspath/plt_x_per_variant_y_egene_distance.py/violin.png}
+        \end{center}
+        \vfill
+        Yes, pleiotropic eQTLs have a lower effect on gene expression and traits
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Are pleiotropic eQTLs closer to genes?}
+
+        \begin{center}
+            \includegraphics[width=0.5\textwidth]{\mspath/plt_x_per_variant_y_egene_distance.py/violin.png}
+        \end{center}
+        \vfill
+        Yes, distance to genes decreases with the pleiotropy
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Are pleiotropic eQTLs associated to more genes?}
 
@@
             \includegraphics[width=0.5\textwidth]{\mspath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+        \end{center}
+        \vfill
+        Yes, number of target gene increaseas with the pleiotropy
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{A web portal to access the eQTL/trait annotations}
+
+        \url{https://gwas2eqtl.tagc.univ-amu.fr/gwas2eqtl}
+
+        \begin{center}
+            \includegraphics[width=0.9\textwidth]{fig/gwas2eqtl.png}
         \end{center}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Acknowledgements}
+
+        \begin{itemize}
+            \item P Paul
+            \item L Lecerf (M1), P Rihet, M Michel, S Marquet, S Spicuglia
+        \end{itemize}
+%
+        \vfill
+%
+        Funding
+%
+        \begin{itemize}
+            \item Institut Cancer et Immunologie - Aix-Marseille Univ.
+            \item Agence nationale de la recherche (ANR)
+        \end{itemize}
+
+    \end{frame}
+
 \end{document}

### 2023-05-08 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC centuri day

@@
         \begin{itemize}
-            \item Phenotypic differences between human populations are mostly coded in frequent genetic variants
-            \item These genetic variants also cause different disease susceptibility
-            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
+            \item Differences between humans coded in 80 x 10$^6$ frequent genetic variants
+            \item Genetic variants change susceptibilities for thousands of common traits and diseases
+            \item For example, breast cancer:
         \end{itemize}
         %
-        \vfill
-%
-        What is the molecular mechanism and effect of tissue of these variants?
+        \begin{figure}
+            \begin{center}
+                \includegraphics[width=0.5\textwidth]{\presentationgoldparispath/41586_2017_Article_BFnature24284_Fig1_HTML.jpg}
+            \end{center}
+            \caption{Genetic variant associations with breast cancer risk (DOI:10.1038/nature24284).}
+        \end{figure}
 
@@
     \begin{frame}
-        \frametitle{Expression quantitative trait loci (eQTL)}
+        \frametitle{Effect of genetic variants on gene expression: eQTLs}
 
         \begin{itemize}
-            \item Quantitative trait loci (QTL) are variants associated with changes of quantitative phenotypes
+            \item Quantitative trait loci (QTL) variants associated with changes of quantitative phenotypes
             \item Expression (eQTL) are variants associated with changes of gene expression in a given tissue
@@
         \end{itemize}
-        \vfill
-        What are the properties of pleiotropic eQTLs?
 
@@
     \begin{frame}
+        \frametitle{Pleiotropic genetic variants}
+
+        \begin{itemize}
+            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
+        \end{itemize}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Summary of previous and my work here}
+
+        \begin{center}
+            \includegraphics[width=0.6\textwidth]{fig/graphical_intro_approach.drawio.png}
+        \end{center}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
         \frametitle{Strategy}
@@
     \section{Results} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
+        \frametitle{Most pleiotroic eQTLs}
+
+        \begin{itemize}
+            \item Example pleiotropic eQTLs in different cytobands
+            \item The representative gene is the eQTL gene with the highest PubMed publication count
+        \end{itemize}
+
+% full size table is table
+        \begin{table}[!tbp]
+    \centering
+    \scriptsize
+
+    \csvreader[separator=tab,
+    tabular=crcp{0.4\textwidth},
+    late after last line=\\\hline,
+    head,
+    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Gene marker& \bfseries Trait categories\\\hline,
+    ]{\mspath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+        {\csvcoli\ & \csvcolii\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
+            \vspace{15pt}\label{tab:pleitropic_variants}
+        \end{table}
+
+    \end{frame}
 
@@
     \begin{frame}
-        \frametitle{Summary}
+        \frametitle{Summary of my work}
 

### 2023-05-09 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
ENH

@@
             \item Differences between humans coded in 80 x 10$^6$ frequent genetic variants
-            \item Genetic variants change susceptibilities for thousands of common traits and diseases
+            \item Genetic variants change susceptibilities to thousands of human traits
             \item For example, breast cancer:
@@
         \begin{itemize}
-            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
+            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, for instance:
         \end{itemize}
+
+        % full size table is table
+        \begin{table}[!tbp]
+            \centering
+            \scriptsize
+
+            \csvreader[separator=tab,
+            tabular=crcp{0.4\textwidth},
+            late after last line=\\\hline,
+            head,
+            table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Gene marker& \bfseries Trait categories\\\hline,
+            ]{\mspath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+                {\csvcoli\ & \csvcolii\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
+            \vspace{15pt}\label{tab:pleitropic_variants}
+        \end{table}
 
@@
 
-        \begin{center}
-            \includegraphics[width=0.5\textwidth]{\mspath/plt_x_per_variant_y_egene_distance.py/violin.png}
-        \end{center}
+        \begin{figure}[!ht]
+
+            \begin{subfigure}[]{.49\textwidth}
+                \textbf{a}
+                \\
+                \includegraphics[width=\textwidth]{\mspath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
+            \end{subfigure}
+            %
+            \begin{subfigure}[]{.49\textwidth}
+                \textbf{b}
+                \\
+                \includegraphics[width=\textwidth]{\mspath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
+            \end{subfigure}
+
+        \end{figure}
+        %
         \vfill
@@
             \item Agence nationale de la recherche (ANR)
+            \item Centre de Calcul Intensif d'Aix-Marseille
         \end{itemize}

### 2023-05-10 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC centuri

@@
     \begin{frame}
-        \frametitle{Annotation of eQTLs traits using colocalization analysis}
+        \frametitle{Genetic variants with effect on trait and gene expression}
+
+        Colocalization of eQTLs and GWAS variants
 
         \begin{center}
-            \includegraphics[width=0.6\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4bc.png}
+            \includegraphics[width=0.5\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4bc.png}
         \end{center}
@@
         \vfill
-        Yes, number of target gene increaseas with the pleiotropy
+        Yes, number of target gene increases with the pleiotropy
 
@@
     \begin{frame}
+        \frametitle{Applications / Collaborations}
+
+        \begin{itemize}
+            \item Dataset of traits, categories, tissues and cell types
+            \item Annotation of genomic regions
+            \item Annotation of genes
+        \end{itemize}
+
+    \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \begin{frame}
         \frametitle{Acknowledgements}

### 2023-05-10 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
ENH
DOC centuri

@@
 \newcommand*{\mspath}{../../out/gwas417/pval_5e-08/r2_0.1/kb_1000/window_1000000/75_50}%
+\newcounter{frame}[frame]
+\setbeamertemplate{footline}[frame number]
 
@@
 \title{Identification and analysis of pleiotropic expression quantitative trait loci}
-\subtitle{Centuri Day 2023}
+\subtitle{CENTURI Scientific Day 2023}
+\date{May 11, 2023 - Hexagone auditorium, Luminy campus}
 \author{Aitor Gonz\'alez}
 \institute{Aix Marseille Univ, INSERM, TAGC}
-\date{May 11, 2023}
 
@@
         \begin{itemize}
-            \item Differences between humans coded in 80 x 10$^6$ frequent genetic variants
+            \item Trait differences in human due to 80 x 10$^6$ genetic variants
             \item Genetic variants change susceptibilities to thousands of human traits
@@
 
+        $\rightarrow$ What is the molecular mechanism of the trait association?
+
     \end{frame}
@@
     \begin{frame}
-        \frametitle{Effect of genetic variants on gene expression: eQTLs}
+        \frametitle{eQTLs: Genetic variants that change gene expression}
+
+        \includegraphics[width=0.7\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
 
         \begin{itemize}
-            \item Quantitative trait loci (QTL) variants associated with changes of quantitative phenotypes
-            \item Expression (eQTL) are variants associated with changes of gene expression in a given tissue
-            \item eQTLs provide a putative molecular mechanism of a variant
+            \item eQTLs useful to indicate putative molecular mechanism of a trait variants
         \end{itemize}
 
-        \includegraphics[width=0.7\textwidth]{\presentationgoldparispath/doi_10.3389_fgene.2020.00424_fig4a.jpg}
+        $\rightarrow$ How do we annotate trait variants with eQTLs?
 
@@
     \begin{frame}
-        \frametitle{Genetic variants with effect on trait and gene expression}
+        \frametitle{Annotation of trait variants with eQTLs}
 
-        Colocalization of eQTLs and GWAS variants
+        $\rightarrow$ Colocalization of eQTLs and trait variants
 
@@
         \begin{itemize}
-            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, for instance:
+            \item Many genetic variants change simultaneously susceptibility to several diseases, for instance:
         \end{itemize}
 
-        % full size table is table
         \begin{table}[!tbp]
@@
             \scriptsize
-
             \csvreader[separator=tab,
-            tabular=crcp{0.4\textwidth},
+            tabular=crrcp{0.4\textwidth},
             late after last line=\\\hline,
             head,
-            table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Gene marker& \bfseries Trait categories\\\hline,
-            ]{\mspath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-                {\csvcoli\ & \csvcolii\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-            \vspace{15pt}\label{tab:pleitropic_variants}
+            table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Ref. & \bfseries Alt. & \bfseries Trait categories\\\hline,
+            ]{fig/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolvi}% specify your coloumns here
+            \vspace{15pt}\label{tab:2}
         \end{table}
@@
     \begin{frame}
-        \frametitle{Summary of previous and my work here}
+        \frametitle{Positioning of our work here in the state of the art}
 
         \begin{center}
-            \includegraphics[width=0.6\textwidth]{fig/graphical_intro_approach.drawio.png}
+            \includegraphics[width=0.6\textwidth]{fig/graphical_intro_approach.v2.drawio.png}
         \end{center}
@@
 
-    \section{Results} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \section{Results and discussion} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \begin{frame}
-        \frametitle{Most pleiotroic eQTLs}
+        \frametitle{Pleiotroic eQTLs}
 
         \begin{itemize}
-            \item Example pleiotropic eQTLs in different cytobands
-            \item The representative gene is the eQTL gene with the highest PubMed publication count
+            \item Analyzed: 138,274 variants from 293 traits and the 127 cell types/tissues
+            \item Found: 476 pleiotropic eQTLs associated to $\geq$ two traits categories
+            \item Important genomic regions: ABO, ALDH2, IL4, TP53, ...
         \end{itemize}
 
-% full size table is table
         \begin{table}[!tbp]
@@
             \scriptsize
-
             \csvreader[separator=tab,
-            tabular=crcp{0.4\textwidth},
+            tabular=crrrcp{0.4\textwidth},
             late after last line=\\\hline,
             head,
-            table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Gene marker& \bfseries Trait categories\\\hline,
-            ]{\mspath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
-                {\csvcoli\ & \csvcolii\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
-            \vspace{15pt}\label{tab:pleitropic_variants}
+            table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Ref. & \bfseries Alt. & \bfseries Gene marker& \bfseries Trait categories\\\hline,
+            ]{fig/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
+                {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
+            \vspace{15pt}\label{tab:1}
         \end{table}
@@
     \begin{frame}
-        \frametitle{Summary of my work}
+        \frametitle{Graphical summary}
 
         \begin{center}
-            \includegraphics[width=0.5\textwidth]{fig/graphical_abstract.drawio.png}
+            \includegraphics[width=0.6\textwidth]{fig/graphical_abstract.drawio.png}
         \end{center}
@@
     \begin{frame}
-        \frametitle{Applications / Collaborations}
+        \frametitle{\includegraphics[width=0.1\textwidth]{fig/opentonetwork.png} Applications of these datasets}
 
         \begin{itemize}
-            \item Dataset of traits, categories, tissues and cell types
-            \item Annotation of genomic regions
-            \item Annotation of genes
+            \item Large dataset of genetic variants ...
+            \item ... associated to traits and gene expression changes in 127 tissues and cell types
+            \item Useful to annotate genomic regions and genes
         \end{itemize}
@@
 
+%
+        Collaborators and colleagues at TAGC, specially:
+        %
         \begin{itemize}
-            \item P Paul
-            \item L Lecerf (M1), P Rihet, M Michel, S Marquet, S Spicuglia
+            \item P. Paul, L. Lecerf (M1), P. Rihet, M. Michel, S. Marquet, S. Spicuglia.
         \end{itemize}
@@
 %
-        Funding
+        Funding and resources:
 %
@@
         \end{itemize}
+%
+        \vfill
+        %
+        Thank you for your attention!!
 

### 2023-09-01 doc/presentation_230120_internal_seminar/intern_sem_230120.tex

--- 2023-01-20 presentation_230120_internal_seminar/intern_sem_230120.tex
@@
         \tableofcontents[currentsection]
-    \end{frame}
+    \end{frame}Large scale GWAS/eQTL colocalization analysis
 }
@@
         \begin{itemize}
-            \item Genome-wise association studies: Association test in the population of variants on common traits, eg. height or Crohn's disease
+            \item Genome-wise association studies: Test of association of variants to traits in a population, eg. height or Crohn's disease
             \item Used for frequent variants and common quantitative or qualitative traits
@@
         \begin{itemize}
-            \item Correlation between variants. Which is the causal variant?
+            \item Which is the causal variant?
             \item Which genes are important for disease?
@@
             \item 25\%\footnote{SNP.PP.H4$\geq$0.5} of 10,000 leading SNPs, colocalized with at least one eQTL
-            \item 5,000 eQTLs and 5,000 target genes\footnote{SNP.PP.H4$\geq$0.5}
+            \item 5,000 target genes\footnote{SNP.PP.H4$\geq$0.5}
             \item 138e3 colocalized variants with probability $\geq$ 0.75
@@
             \item Pleiotropic eQTL region include immune and cell cycle regulators
-            \item Enrichment of 5' UTR and transcript (missense) variants
             \item Increased number of associated egenes and etissues
@@
             \item Analyze pleiotropic loci in more detail
+            \item Compute colocalization of new GWAS datasets
         \end{itemize}
@@
             \item Agence nationale de la recherche (ANR)
-            \item GOLD
         \end{itemize}

### 2023-09-01 ms/ms08_figures.tex
ENH improved figures
ENH
ENH plots

@@
 
-    \begin{subfigure}[]{\textwidth}
+    \begin{subfigure}[]{0.65\textwidth}
+        \textbf{a}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/loci_explained_perc.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{0.85\textwidth}
+        \textbf{b}
+        \\
         \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_hist_100000.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_hist_100000.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/histplot.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/histplot.png}
     \end{subfigure}
@@
 %
-\begin{figure}[!ht]
-    \centering
-
-    \begin{subfigure}[]{0.99\textwidth}
-        \textbf{a}
-        \\
-        \includegraphics[width=\textwidth]{fig/model1.png}
-    \end{subfigure}
-
-\end{figure}
+%\begin{figure}[!ht]
+%    \centering
+%
+%    \begin{subfigure}[]{0.99\textwidth}
+%        \textbf{a}
+%        \\
+%        \includegraphics[width=\textwidth]{fig/model1.png}
+%    \end{subfigure}
+%
+%\end{figure}

### 2023-09-01 scripts/cmpt_count_per_rsid.py

@@ -129,7 +129,9 @@ plt.close()
 #%########################################### watanabe 2019 category count
 m2df = m_df[['rsid', 'ref', 'alt', 'gwas_category_count', 'domains_watanabe2019']].copy()
 
-order = [*range(1, m2df['gwas_category_count'].max()+1)]
+m2df.loc[m2df['gwas_category_count']>=manuscript_pleio_cutoff, 'gwas_category_count'] = '≥' + str(manuscript_pleio_cutoff)
+
+order = [*range(1, manuscript_pleio_cutoff)] + ['≥' + str(manuscript_pleio_cutoff)]
 ax = seaborn.boxplot(x='gwas_category_count', y='domains_watanabe2019', data=m2df, order=order, palette="rocket_r")
 
 plt.xlabel("trait category count", fontsize=label_fontsize)
@@ -150,7 +152,7 @@ m2df_watanabe2019_cat_count_df = m2df_watanabe2019_cat_count_df.merge(m2df_cat_c
 m2df_watanabe2019_cat_count_df.columns = ['watanabe_count', 'count']
 m2df_watanabe2019_cat_count_df['watanabe_perc'] = (m2df_watanabe2019_cat_count_df['watanabe_count'] / m2df_watanabe2019_cat_count_df['count'] * 100).astype(int)
 
-order = [*range(1, m2df['gwas_category_count'].max()+1)]
+order = [*range(1, manuscript_pleio_cutoff)] + ['≥' + str(manuscript_pleio_cutoff)]
 ax = seaborn.barplot(x=m2df_watanabe2019_cat_count_df.index, y='watanabe_perc', data=m2df_watanabe2019_cat_count_df, order=order, palette="rocket_r")
 
 plt.title("Known in Watanabe 2019", fontsize=label_fontsize)

### 2023-09-02 ms/ms08_figures.tex
ENH

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/histplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/plt.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/histplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/plt.png}
+    \end{subfigure}
+
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{aa}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
+        %
+    \end{subfigure}
+    %
+    \begin{subfigure}[]{.49\textwidth}
+        \textbf{bb}
+        \\
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
     \end{subfigure}
@@
 %    \\
-%    \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+%    \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
 %    \end{subfigure}
@@
 %        \\
-%        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
 %    \end{subfigure}

### 2023-09-02 scripts/intrsct_remapnr_gwas_categories.py

@@ -1,4 +1,3 @@
-import multiprocessing
 import os
 import pathlib
 import shlex
@@ -7,17 +6,16 @@ import sys
 from multiprocessing import Pool
 
 from gwas2eqtl_pleiotropy.Logger import Logger
-from gwas2eqtl_pleiotropy.constants import public_data_dir
 
 #%%
 help_cmd_str = "todo"
 try:
-    # max_gwas_category_count = int(sys.argv[1])
     threads = int(sys.argv[1])
-    remap_nr_path = sys.argv[2]
-    eqtl_pleio_1_flank_10_hg38_bed = sys.argv[3]
-    remap_nr_pleio_1_flank_10_hg38_bed = sys.argv[4]
-    if len(sys.argv) > 5:
+    max_gwas_category_count = int(sys.argv[2])
+    remap_nr_path = sys.argv[3]
+    eqtl_pleio_1_flank_10_hg38_bed = sys.argv[4]
+    remap_nr_pleio_1_flank_10_hg38_bed = sys.argv[5]
+    if len(sys.argv) > 6:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -35,15 +33,14 @@ indir_path = os.path.dirname(eqtl_pleio_1_flank_10_hg38_bed)
 #%% bedtools intersect
 flank = 10
 
-for max_gwas_category_count in range(1, 99):
-    bed_path = os.path.join(indir_path, "eqtl_pleio_{}_flank_{}_hg38.bed".format(max_gwas_category_count, flank))
+for gwas_category_count_i in range(max_gwas_category_count + 1):
+    bed_path = os.path.join(indir_path, "eqtl_pleio_{}_flank_{}_hg38.bed".format(gwas_category_count_i, flank))
     if not os.path.isfile(bed_path):
         break
 
-max_gwas_category_count = max_gwas_category_count - 1
-print(max_gwas_category_count)
+# max_gwas_category_count = max_gwas_category_count - 1
+# print(max_gwas_category_count)
 def intrsct_remapnr(count_pleio):
-    # print(count_pleio)
     bed_path = os.path.join(indir_path, "eqtl_pleio_{}_flank_{}_hg38.bed".format(count_pleio, flank))
     intersect_basename = os.path.basename(remap_nr_pleio_1_flank_10_hg38_bed).replace('pleio_1', 'pleio_' + str(count_pleio))
[…]

### 2023-09-02 scripts/plt_x_per_rsid_y_remapnr.py

@@ -18,10 +18,10 @@ seaborn.set_theme(**seaborn_theme_dic)
 #%%
 help_cmd_str = "todo"
 try:
-    # max_gwas_category_count = int(sys.argv[1])
-    remap_nr_pleio_1_flank_10_hg38_bed = sys.argv[1]
-    tf_flank_10_png = sys.argv[2]
-    if len(sys.argv) > 3:
+    max_gwas_category_count = int(sys.argv[1])
+    remap_nr_pleio_1_flank_10_hg38_bed = sys.argv[2]
+    tf_flank_10_png = sys.argv[3]
+    if len(sys.argv) > 4:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -41,7 +41,7 @@ flank = 10
 
 cat_df = pandas.DataFrame({'gwas_category_count': [], 'rsid': [], 'tf': []})
 # for pleio in range(1, max_gwas_category_count+1):
-for pleio in range(1, 99):
+for pleio in range(1, max_gwas_category_count + 1):
     pleio_path = remap_nr_pleio_1_flank_10_hg38_bed.replace('pleio_1', 'pleio_{}'.format(pleio))
     if not os.path.isfile(pleio_path):
         break
@@ -55,13 +55,15 @@ for pleio in range(1, 99):
     df['gwas_category_count'] = pleio
     cat_df = pandas.concat([cat_df, df], axis=0)
 
+cat_df.loc[cat_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+
 describe_tsv_path = os.path.join(os.path.dirname(tf_flank_10_png), "describe.tsv")
 cat_df.groupby('gwas_category_count')['tf'].apply(lambda x: x.describe()).to_csv(describe_tsv_path, sep="\t")
+
 
 #%%
 order = [str(int(x)) for x in cat_df['gwas_category_count'].unique()]
 xticklabels = order.copy()
-# xticklabels[-1] = '≥{}'.format(order[-1])
 title = "TFs per variant"
 xlabel = "Trait category count"
 ylabel = "TF count"
@@ -81,16 +83,18 @@ annotator.apply_and_annotate()
 
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel(xlabel, fontsize=label_fontsize)
[…]

### 2023-09-02 scripts/pltbar_x_per_gwas_cat_y_allele_freq.py

@@ -21,10 +21,11 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    url = sys.argv[2]
-    count_per_rsid_gwas_ods_path = sys.argv[3]
-    eur_af_png_path = sys.argv[4]
-    if len(sys.argv) > 5:
+    max_gwas_category_count = int(sys.argv[2])
+    url = sys.argv[3]
+    count_per_rsid_gwas_ods_path = sys.argv[4]
+    eur_af_png_path = sys.argv[5]
+    if len(sys.argv) > 6:
         print("""Two many arguments!
         {}""".format(help_cmd_str))
         sys.exit(1)
@@ -49,11 +50,12 @@ with engine.begin() as conn:
 
 #%%
 count_per_rsid_gwas_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine='odf')
-max_gwas_category_count = count_per_rsid_gwas_df['gwas_category_count'].max()
+# max_gwas_category_count = count_per_rsid_gwas_df['gwas_category_count'].max()
 
 #%%
 m_df = h4_df.merge(count_per_rsid_gwas_df[['chrom', 'pos38', 'rsid', 'alt', 'gwas_category_count']], on=['chrom', 'pos38', 'rsid', 'alt'])
 m_df = m_df[['chrom', 'pos38', 'rsid', 'ref', 'alt', 'afr_af', 'amr_af', 'eas_af', 'eur_af', 'sas_af', 'gwas_category_count']].drop_duplicates()
+m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
 
 #%%
 order = [str(x) for x in range(1, max(m_df['gwas_category_count'].unique()) + 1)]
@@ -84,13 +86,24 @@ for y,ytitle in zip(y_labels, y_titles):
     ax.set_xticklabels(xticklabels)
     plt.title(ytitle, fontsize=label_fontsize)
     plt.xlabel(xlabel, fontsize=label_fontsize)
-    plt.xticks(fontsize=tick_fontsize, rotation=0)
+    # plt.xticks(fontsize=tick_fontsize, rotation=0)
+    xticks_labels = [str(x) for x in (plt.xticks()[0] + 1)]
+    xticks_labels[-1] = '≥' + str(xticks_labels[-1])
[…]

### 2023-09-02 scripts/plthst_gwas_egene_etissue.py

@@ -9,6 +9,7 @@ from matplotlib.ticker import FuncFormatter
 
 from gwas2eqtl_pleiotropy.constants import tick_fontsize, dpi
 from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
+from matplotlib.ticker import MaxNLocator
 
 plt.rcParams["figure.figsize"] = (8, 6)
 seaborn.set_theme(**seaborn_theme_dic)
@@ -36,59 +37,64 @@ pathlib.Path(os.path.dirname(hist_rsid_etissue_path)).mkdir(parents=True, exist_
 count_df = pandas.read_excel(count_per_rsid_gwas_egene_etissue_ods, engine='odf')
 
 #%%
-ylabel = "Percentage"
-ylim = [0, 100]
+ylabel = "Proportion"
+ylim = [0.0001, 1]
 edgecolor = 'k'
 linewidth = 2
 grid_axis = 'y'
 hist_kwargs = {'density': 1, 'edgecolor': edgecolor, 'linewidth': linewidth}
-stat = 'percent'
+stat = 'proportion'
 label_fontsize = 28
 
 #%% gwas
 data_ser = count_df['gwas_category_count']
-shplt = seaborn.histplot(data_ser, stat=stat, discrete=True)
+ax = seaborn.histplot(data_ser, stat=stat, discrete=True)
 
+ax.xaxis.set_major_locator(MaxNLocator(integer=True))
 plt.grid(visible=True, axis='y')
 title = "Trait specificity"
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel("Trait category count", fontsize=label_fontsize)
 plt.xticks(fontsize=tick_fontsize)
 plt.ylabel(ylabel, fontsize=label_fontsize)
-plt.ylim(ylim)
 plt.yticks(fontsize=tick_fontsize)
+plt.yscale('log')
+plt.ylim(ylim)
 
 plt.tight_layout()
 plt.savefig(hist_rsid_gwas_path, dpi=dpi)
 plt.close()
 
 with open(hist_rsid_gwas_path + ".txt", 'w') as fout:
-    fout.write('\n'.join([str(h.get_height()) for h in shplt.patches]))
+    fout.write('\n'.join([str(h.get_height()) for h in ax.patches]))
 
 #%% eQTL gene
 data_ser = count_df['egene_count']
-shplt = seaborn.histplot(data_ser, stat=stat, discrete=True)
[…]

### 2023-09-03 scripts/cmpt_vep.py
MS

@@ -79,7 +79,7 @@ vep_df.to_csv(vep_input_path, sep="\t", index=False, header=False)
 
 #%% run vep command
 dir_cache = "/".join(vep_cache_info.split('/')[:-3])
-cmd_str = "vep --offline --dir_cache {dir_cache} --plugin TSSDistance --force_overwrite -i {vep_input} " \
+cmd_str = "vep --offline --dir_cache {dir_cache} --cache_version 108 --plugin TSSDistance --force_overwrite -i {vep_input} " \
           "--output_file {vep_output}".format(vep_input=vep_input_path, dir_cache=dir_cache, vep_output=vep_output_path)
 Logger.info(cmd_str)
 output = subprocess.run(shlex.split(cmd_str), capture_output=True)

### 2023-09-06 ms/ms07_tables.tex
MS

@@
         separator=tab,
-        tabular=p{0.05\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.5\textwidth},
+        tabular=p{0.03\textwidth}p{0.1\textwidth}p{0.1\textwidth}p{0.06\textwidth}p{0.08\textwidth}p{0.45\textwidth}p{0.08\textwidth}p{0.08\textwidth},
         late after last line=\\\hline,
         head,
-        table head=\\\hline \bfseries Chr. & \bfseries Start & \bfseries End & \bfseries Cytob. & \bfseries Marker eQTL gene & \bfseries Trait categories\\\hline,
+        table head=\\\hline \bfseries Chr.
+& \bfseries Start & \bfseries End & \bfseries Cytob.
+& \bfseries Marker eQTL gene & \bfseries Trait categories & \bfseries Length & \bfseries eQTLs cnt.\\\hline,
     ]{\floatRelativePath/cmpt_pleiotropic_regions.py/100000/region_window_ms.tsv}{}% use head of csv as column names
-    {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
+    {\csvcoli\ & \csvcolii\ & \csvcoliii\ & \csvcoliv\ & \csvcolv\ & \csvcolvi\ & \csvcolvii\ & \csvcolviii}% specify your coloumns here
     %

### 2023-09-06 scripts/plt_x_per_variant_egene_y_etissue.py

@@ -21,7 +21,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    max_gwas_category_count = int(sys.argv[2])
+    pleio_high_cutoff = int(sys.argv[2])
     sa_url = sys.argv[3]
     count_per_rsid_gwas_ods_path = sys.argv[4]
     png_path = sys.argv[5]
@@ -64,7 +64,7 @@ sel_cols = ['rsid', 'eqtl_gene_id', 'etissue_category_term']  # tissue per varia
 
 #%% set max_gwas_category_count
 m_df = m_df[sel_cols + ['gwas_category_count']]
-m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+m_df.loc[m_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 #%% keep unique rsid-etissue_category_term pairs with max. gwas category
 m_df.sort_values('gwas_category_count', ascending=False, inplace=True)
@@ -80,10 +80,10 @@ describe_tsv_path = os.path.join(outdir_path, "describe.tsv")
 m_df.groupby('gwas_category_count')['etissue_category_term_count'].apply(lambda x: x.describe()).to_csv(describe_tsv_path, sep="\t")
 
 #%%
-m_df.loc[m_df['gwas_category_count']>=max_gwas_category_count, 'gwas_category_count'] = '≥' + str(max_gwas_category_count)
+m_df.loc[m_df['gwas_category_count'] >= pleio_high_cutoff, 'gwas_category_count'] = '≥' + str(pleio_high_cutoff)
 
 # order = [str(x) for x in range(1, max(m_df['gwas_category_count'].unique())+1)]
-order = [str(x) for x in [*range(1, max_gwas_category_count)] + ['≥' + str(max_gwas_category_count)]]
+order = [str(x) for x in [*range(1, pleio_high_cutoff)] + ['≥' + str(pleio_high_cutoff)]]
 
 # xticklabels = order.copy()
 # xticklabels[-1] = '≥{}'.format(order[-1])

### 2023-09-06 scripts/plt_x_per_variant_y_egene_distance.py

@@ -20,7 +20,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    max_gwas_category_count = int(sys.argv[2])
+    pleio_high_cutoff = int(sys.argv[2])
     sa_url = sys.argv[3]
     count_per_rsid_gwas_ods_path = sys.argv[4]
     hist_png_path = sys.argv[5]
@@ -60,7 +60,7 @@ gwas_category_count_max_int = count_per_rsid_gwas_df['gwas_category_count'].max(
 
 #%%
 m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos38', 'rsid', 'ref', 'alt'])
-m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+m_df.loc[m_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 # distance to tss
 m_df['eqtl_gene_distance'] = math.nan

### 2023-09-06 scripts/pltbar_x_per_gwas_cat_y_allele_freq.py

@@ -21,7 +21,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    max_gwas_category_count = int(sys.argv[2])
+    pleio_high_cutoff = int(sys.argv[2])
     url = sys.argv[3]
     count_per_rsid_gwas_ods_path = sys.argv[4]
     eur_af_png_path = sys.argv[5]
@@ -55,12 +55,12 @@ count_per_rsid_gwas_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine=
 #%%
 m_df = h4_df.merge(count_per_rsid_gwas_df[['chrom', 'pos38', 'rsid', 'alt', 'gwas_category_count']], on=['chrom', 'pos38', 'rsid', 'alt'])
 m_df = m_df[['chrom', 'pos38', 'rsid', 'ref', 'alt', 'afr_af', 'amr_af', 'eas_af', 'eur_af', 'sas_af', 'gwas_category_count']].drop_duplicates()
-m_df.loc[m_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+m_df.loc[m_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 #%%
 order = [str(x) for x in range(1, max(m_df['gwas_category_count'].unique()) + 1)]
 xticklabels = order.copy()
-box_pairs = [(1, i) for i in range(1, max_gwas_category_count+1) ]
+box_pairs = [(1, i) for i in range(1, pleio_high_cutoff + 1)]
 x = 'gwas_category_count'
 xlabel = "Trait category count"
 
@@ -69,7 +69,7 @@ xlabel = "Trait category count"
 ylabel = "Alternative allele freq."
 
 #%%
-pairs = [(str(1), str(i)) for i in range(2, max_gwas_category_count+1)]
+pairs = [(str(1), str(i)) for i in range(2, pleio_high_cutoff + 1)]
 m_df[x] = m_df[x].astype(str)
 
 y_labels = ['afr_af', 'amr_af', 'eas_af', 'eur_af', 'sas_af']

### 2023-09-06 scripts/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py

@@ -22,7 +22,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 help_cmd_str = "todo"
 try:
     snp_pp_h4 = float(sys.argv[1])
-    max_gwas_category_count = int(sys.argv[2])
+    pleio_high_cutoff = int(sys.argv[2])
     sa_url = sys.argv[3]
     count_per_rsid_gwas_ods_path = sys.argv[4]
     eqtl_beta_png_path = sys.argv[5]
@@ -68,19 +68,19 @@ m_gwas_df = m_gwas_df.loc[m_gwas_df['gwas_pval']!=0]  # remove pval with zeros
 m_gwas_df[['gwas_beta_abs']] = m_gwas_df[['gwas_beta']].abs()
 m_gwas_df[['gwas_neglog10pval']] = -numpy.log10(m_gwas_df[['gwas_pval']])
 m_gwas_df = m_gwas_df[['gwas_category_count', 'gwas_beta_abs', 'gwas_neglog10pval']]
-m_eqtl_df.loc[m_eqtl_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+m_eqtl_df.loc[m_eqtl_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 m_eqtl_df[['eqtl_beta_abs']] = m_eqtl_df[['eqtl_beta']].abs()
 m_eqtl_df[['eqtl_neglog10pval']] = -numpy.log10(m_eqtl_df[['eqtl_pval']])
 m_eqtl_df = m_eqtl_df[['gwas_category_count', 'eqtl_beta_abs', 'eqtl_neglog10pval']]
-m_eqtl_df.loc[m_eqtl_df['gwas_category_count'] >= max_gwas_category_count, "gwas_category_count"] = max_gwas_category_count
+m_eqtl_df.loc[m_eqtl_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 
 #%%
-order = [str(x) for x in range(1, max_gwas_category_count+1)]
+order = [str(x) for x in range(1, pleio_high_cutoff + 1)]
 xticklabels = order.copy()
 # xticklabels[-1] = '≥{}'.format(order[-1])
-box_pairs = [(1, i) for i in range(1, max_gwas_category_count+1) ]
[…]

### 2023-09-06 scripts/plthst_perc_tophits_eqtl.py

@@ -17,7 +17,6 @@ help_cmd_str = "todo"
 try:
     perc_tophits_eqtl_tsv = sys.argv[1]
     hist_perc_tophits_eqtl_png = sys.argv[2]
-    # hist_perc_tophits_nomhc_eqtl_png = sys.argv[3]
     if len(sys.argv) > 4:
         print("""Two many arguments!
         {}""".format(help_cmd_str))

### 2023-09-06 scripts/plthtmp_disease_comorbidity_matrix.py

@@ -7,10 +7,10 @@ import seaborn
 import sqlalchemy
 
 from gwas2eqtl_pleiotropy.Logger import Logger
-from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic, dpi
+from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 from matplotlib import pyplot as plt
 from matplotlib.pyplot import gcf
-from sqlalchemy import create_engine
+
 
 #%%
 seaborn.set_theme(**seaborn_theme_dic)

### 2023-09-06 scripts/pltsctr_x_per_rsid_y_etissue.py

@@ -16,7 +16,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 #%%
 help_cmd_str = "todo"
 try:
-    max_gwas_category_count = int(sys.argv[1])
+    pleio_high_cutoff = int(sys.argv[1])
     region_window_100000_ods_path = sys.argv[2]
     count_per_rsid_gwas_ods_path = sys.argv[3]
     outdir_path = sys.argv[4]
@@ -57,7 +57,7 @@ count_per_rsid_df['pos38'] = count_per_rsid_df['pos38'].astype('int')
 
 #%% Loop over regions
 pleiotropic_regions_df = region_window_100000_df.loc[
-    region_window_100000_df['gwas_category_count'] >= max_gwas_category_count, ['chrom', 'start', 'end', 'gwas_category_count']]
+    region_window_100000_df['gwas_category_count'] >= pleio_high_cutoff, ['chrom', 'start', 'end', 'gwas_category_count']]
 for rowi, row in pleiotropic_regions_df.iterrows():
     chrom = row['chrom']
     start = row['start']

### 2023-09-06 scripts/pltsctr_x_per_rsid_y_gwas.py

@@ -17,7 +17,7 @@ seaborn.set_theme(**seaborn_theme_dic)
 #%%
 help_cmd_str = "todo"
 try:
-    max_gwas_category_count = int(sys.argv[1])
+    pleio_high_cutoff = int(sys.argv[1])
     region_window_100000_ods_path = sys.argv[2]
     count_per_rsid_gwas_ods_path = sys.argv[3]
     outdir_path = sys.argv[4]
@@ -58,7 +58,7 @@ ylabel = "GWAS cat. count"
 count_per_rsid_df['pos38'] = count_per_rsid_df['pos38'].astype('int')
 
 #%% Loop over regions
-pleiotropic_regions_df = region_window_100000_df.loc[region_window_100000_df['gwas_category_count'] >= max_gwas_category_count, ['chrom', 'start', 'end', 'gwas_category_count']]
+pleiotropic_regions_df = region_window_100000_df.loc[region_window_100000_df['gwas_category_count'] >= pleio_high_cutoff, ['chrom', 'start', 'end', 'gwas_category_count']]
 for rowi, row in pleiotropic_regions_df.iterrows():
     chrom = row['chrom']
     start = row['start']

### 2023-09-08 ms/ms08_figures_boxenplot.tex
MS

--- 2023-09-02 ms/ms08_figures.tex
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_effect_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_effect_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_neglog10pval.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_signif_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_neglog10pval.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_signif_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_allele_freq.py/eur_af.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_allele_freq.py/eur_af_boxenplot.png}
     \end{subfigure}
@@
 
-    \begin{subfigure}[]{.49\textwidth}
-        \textbf{aa}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
-        %
-    \end{subfigure}
-    %
-    \begin{subfigure}[]{.49\textwidth}
-        \textbf{bb}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
-    \end{subfigure}
+%    \begin{subfigure}[]{.49\textwidth}
+%        \textbf{aa}
+%        \\
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
+%        %
+%    \end{subfigure}
+%    %
+%    \begin{subfigure}[]{.49\textwidth}
+%        \textbf{bb}
+%        \\
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
+%    \end{subfigure}
 

### 2023-09-08 scripts/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py

@@ -68,7 +68,7 @@ m_gwas_df = m_gwas_df.loc[m_gwas_df['gwas_pval']!=0]  # remove pval with zeros
 m_gwas_df[['gwas_beta_abs']] = m_gwas_df[['gwas_beta']].abs()
 m_gwas_df[['gwas_neglog10pval']] = -numpy.log10(m_gwas_df[['gwas_pval']])
 m_gwas_df = m_gwas_df[['gwas_category_count', 'gwas_beta_abs', 'gwas_neglog10pval']]
-m_eqtl_df.loc[m_eqtl_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
+m_gwas_df.loc[m_gwas_df['gwas_category_count'] >= pleio_high_cutoff, "gwas_category_count"] = pleio_high_cutoff
 
 m_eqtl_df[['eqtl_beta_abs']] = m_eqtl_df[['eqtl_beta']].abs()
 m_eqtl_df[['eqtl_neglog10pval']] = -numpy.log10(m_eqtl_df[['eqtl_pval']])
@@ -115,7 +115,19 @@ plt.close()
 #%% boxenplot
 pairs = [(str(1), str(i)) for i in range(2, pleio_high_cutoff + 1)]
 m_gwas_df[x] = m_gwas_df[x].astype(str)
-ax = seaborn.boxenplot(x=x, y=y, data=m_gwas_df, showfliers=False)
+ax = seaborn.boxenplot(x=x, y=y, data=m_gwas_df, showfliers=True, palette="rocket_r")
+
+annotator = Annotator(ax, pairs, data=m_gwas_df, x=x, y=y, order=order)
+annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
+annotator.apply_and_annotate()
+
+plt.title(title, fontsize=label_fontsize)
+plt.xlabel(xlabel, fontsize=label_fontsize)
+xticks_labels = [str(x) for x in (plt.xticks()[0] + 1)]
+xticks_labels[-1] = '≥' + str(xticks_labels[-1])
+plt.xticks(ticks=(plt.xticks()[0]), labels=xticks_labels, fontsize=tick_fontsize, rotation=0)
+plt.ylabel(ylabel, fontsize=label_fontsize)
+plt.yticks(fontsize=tick_fontsize)
 
 plt.tight_layout()
 png_path = os.path.join(outdir_path, "gwas_signif_boxenplot.png")
[…]

### 2023-09-10 ms/ms08_figures_boxenplot.tex
MS

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_effect_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_effect_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_signif_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_signif_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/plt.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot.png}
     \end{subfigure}
@@
 %    \\
-%    \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+%    \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
 %    \end{subfigure}
@@
 %        \\
-%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapcrm.py/remapcrm_flank10.png}
 %    \end{subfigure}

### 2023-09-10 scripts/plt_x_per_variant_etissue_y_egene.py

@@ -182,11 +182,11 @@ plt.savefig(hist_png_path)
 plt.close()
 
 #%% boxenplot
-ax = seaborn.boxenplot(data=m_df, x=x, y=y, order=order, palette="rocket_r", showfliers = False, scale='area')
+ax = seaborn.boxenplot(data=m_df, x=x, y=y, order=order, palette="rocket_r", showfliers=True, scale='area')
 
-# annotator = Annotator(ax, pairs, data=m_df, x=x, y=y, order=order)
-# annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
-# annotator.apply_and_annotate()
+annotator = Annotator(ax, pairs, data=m_df, x=x, y=y, order=order)
+annotator.configure(test='Mann-Whitney', text_format='star', loc='inside', **annotator_config_dic)
+annotator.apply_and_annotate()
 
 plt.title(title, fontsize=label_fontsize)
 plt.xlabel(xlabel, fontsize=label_fontsize)

### 2023-09-11 ms/ms08_figures_boxenplot.tex
MS

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_effect_ms_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_effect_ms_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_signif_ms_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_signif_ms_boxenplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/boxenplot_ms_closest_distance.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/histplot.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/histplot.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot_ms.png}
     \end{subfigure}

### 2023-09-11 scripts/plt_x_per_gwas_count_y_beta_neglog10pval.py

@@ -1,4 +1,5 @@
 import scipy
+from gwas2eqtl_pleiotropy import boxenplot_with_mannwhitneyu
 
 from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, annotator_config_dic
 from statannotations.Annotator import Annotator
@@ -126,23 +127,9 @@ plt.close()
 #%% boxenplot
 ax = seaborn.boxenplot(x=x, y=y, data=m_eqtl_df, showfliers=False, palette="rocket_r")
 
-# annotator = Annotator(ax, pairs, data=m_eqtl_df, x=x, y=y, order=order)
-# annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
-# annotator.apply_and_annotate()
-
-x1=0.1; x2=1.1; y=1.2; h=0.05; col='k'
-group1 = m_eqtl_df.where(m_eqtl_df.gwas_category_count=='1').dropna()['eqtl_beta_abs']
-group2 = m_eqtl_df.where(m_eqtl_df.gwas_category_count=='2').dropna()['eqtl_beta_abs']
-scipy.stats.mannwhitneyu(group1, group2)
-plt.plot([x1, x1, x2, x2], [y, y+h, y+h, y], lw=1.5, c=col)
-plt.text((x1+x2)*.5, y+h, "****", ha='center', va='bottom', color=col)
-
-x1=0.1; x2=2.1; y=1.3; h=0.1; col='k'
-group1 = m_eqtl_df.where(m_eqtl_df.gwas_category_count=='1').dropna()['eqtl_beta_abs']
-group2 = m_eqtl_df.where(m_eqtl_df.gwas_category_count=='3').dropna()['eqtl_beta_abs']
-scipy.stats.mannwhitneyu(group1, group2)
-plt.plot([x1, x1, x2, x2], [y, y+h, y+h, y], lw=1.5, c=col)
-plt.text((x1+x2)*.7, y+h, "****", ha='center', va='bottom', color=col)
+annotator = Annotator(ax, pairs, data=m_eqtl_df, x=x, y=y, order=order)
+annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
+annotator.apply_and_annotate()
 
 plt.title(title, fontsize=label_fontsize)
[…]

### 2023-09-11 scripts/pltbar_x_per_gwas_cat_y_allele_freq.py

@@ -1,3 +1,4 @@
+import scipy
 import sqlalchemy
 
 from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, boxplot_kwargs, annotator_config_dic
@@ -101,7 +102,8 @@ for y,ytitle in zip(y_labels, y_titles):
 
     # %% boxenplot
     ax = seaborn.boxenplot(x=x, y=y, data=m_df, order=order, showfliers=True, palette="rocket_r")
-
+    import pdb; pdb.set_trace()
+    scipy.stats.shapiro(m_df[y])
     annotator = Annotator(ax, pairs, data=m_df, x=x, y=y, order=order)
     annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
     annotator.apply_and_annotate()

### 2023-09-12 ms/ms09_suppl_boxenplot.tex
MS

%  CONFIGURE NEWcitep SINGLE-PAGE FORMAT

\onecolumn % go back to one column
%\fancyhead{} % make sure we get no headers
\renewcommand{\floatpagefraction}{0.1}
%\lfoot[\bSupInf]{\dAuthor}
%\rfoot[\dAuthor]{\cSupInf}
\newpage

%\captionsetup*{format=largeformat} % make figure legend slightly larger than in the paper
\setcounter{figure}{0} % reset figure counter for Supp. Figures
\setcounter{equation}{0} % reset equation counter for Supp. Equations
%\setcounter{page}{1} % reset page count
%\makeatletter
%\renewcommand{\thefigure}{S\@arabic\c@figure} % make Figure legend start with Figure S
%\makeatother
%\def\theequation{S\arabic{equation}}
\renewcommand{\figurename}{Supplementary Figure}

%  MAIN TEXT

\newpage
\section*{Supplementary Information}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Supp fig 1: Percent of explained loci
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{figure}[!tbp]
%    \centering
%    \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_perc_tophits_eqtl.py/subplots.png}
%
%    \caption{}
%%
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: ucsc screenshots
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    \centering

    \begin{subfigure}[]{0.99\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_bcell_help.png}
    \end{subfigure}

    \begin{subfigure}[]{0.99\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_monocyte.png}
    \end{subfigure}

    \begin{subfigure}[]{0.99\textwidth}
        \textbf{c}
        \\
        \includegraphics[width=\textwidth]{fig/ucsc_gwas2eqtl_il4_frontalcortex.png}
    \end{subfigure}

    \caption{}

\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 5: Comparison with Watanabe 2019
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
    %
    \centering
    %
    \begin{subfigure}[]{.49\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/watanabe_cat_count.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.49\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/watanabe_percentage.png}
    \end{subfigure}

    \caption{}

\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Supp fig: allele frequencies
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!tbp]

    \begin{subfigure}[]{.32\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/afr_af_ms_boxenplot.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/amr_af_ms_boxenplot.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{c}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eas_af_ms_boxenplot.png}
    \end{subfigure}

    \centering
    \begin{subfigure}[]{.32\textwidth}
        \textbf{d}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eur_af_ms_boxenplot.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{e}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/sas_af_ms_boxenplot.png}
    \end{subfigure}

    \caption{}

\end{figure}

### 2023-09-12 gwas2eqtl_pleiotropy/__init__.py

@@ -16,4 +16,4 @@ def boxenplot_with_mannwhitneyu(group1, group2, x1, x2, y, h):
         pvalue_asterisk = "*"
     else:
         pvalue_asterisk = "ns"
-    plt.text((x1 + x2) * .5, y + h, pvalue_asterisk, ha='center', va='bottom', color=col)
+    plt.text((x1 + x2) * .5, y + h, pvalue_asterisk, ha='center', va='bottom', color=col, fontsize=14)

### 2023-09-13 ms/ms08_figures_boxenplot.tex
MS

--- 2023-09-10 ms/ms08_figures_boxenplot.tex
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/eqtl_effect_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_effect_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/gwas_effect_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/eqtl_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/eqtl_signif_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_gwas_count_y_beta_neglog10pval.py/gwas_signif_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/gwas_signif_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_allele_freq.py/eur_af_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eur_af_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/boxenplot_custom_closest_distance.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/plt.png}
         %
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_egene_y_etissue.py/plt.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapcrm.py/remapcrm_flank10.png}
     \end{subfigure}

### 2023-09-13 ms/ms09_suppl_boxenplot.tex
MS

@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/afr_af_ms_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/amr_af_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/amr_af_ms_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/amr_af_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eas_af_ms_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eas_af_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eur_af_ms_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eur_af_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/sas_af_ms_boxenplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/sas_af_custom.png}
     \end{subfigure}

### 2023-09-13 scripts/plt_x_per_variant_y_allele_freq.py

@@ -2,7 +2,8 @@ import scipy
 import sqlalchemy
 
 from gwas2eqtl_pleiotropy import boxenplot_with_mannwhitneyu
-from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, boxplot_kwargs, annotator_config_dic
+from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, boxplot_kwargs, annotator_config_dic, \
+    boxenplot_kws
 from statannotations.Annotator import Annotator
 
 import matplotlib.pyplot as plt
@@ -77,9 +78,10 @@ m_df[x] = m_df[x].astype(str)
 y_labels = ['afr_af', 'amr_af', 'eas_af', 'eur_af', 'sas_af']
 y_titles = ['African population', 'American population', 'East Asian population', 'European population', 'South Asian population']
 
-for y,ytitle in zip(y_labels, y_titles):
-    print(y)
-    ax = seaborn.barplot(x=x, y=y, data=m_df, order=order, estimator=numpy.mean, palette="rocket_r")
+for y, ytitle in zip(y_labels, y_titles):
+
+    # %% boxenplot
+    ax = seaborn.boxenplot(x=x, y=y, data=m_df, order=order, **boxenplot_kws)
 
     annotator = Annotator(ax, pairs, data=m_df, x=x, y=y, order=order)
     annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
@@ -88,12 +90,10 @@ for y,ytitle in zip(y_labels, y_titles):
     ax.set_xticklabels(xticklabels)
     plt.title(ytitle, fontsize=label_fontsize)
     plt.xlabel(xlabel, fontsize=label_fontsize)
-    # plt.xticks(fontsize=tick_fontsize, rotation=0)
     xticks_labels = [str(x) for x in (plt.xticks()[0] + 1)]
     xticks_labels[-1] = '≥' + str(xticks_labels[-1])
     plt.xticks(ticks=(plt.xticks()[0]), labels=xticks_labels, fontsize=tick_fontsize, rotation=0)
     plt.ylabel(ylabel, fontsize=label_fontsize)
-    plt.ylim([0.2, 0.7])
     plt.yticks(fontsize=tick_fontsize)
 
     plt.tight_layout()
[…]

### 2023-09-13 scripts/plt_x_per_variant_y_beta_neglog10pval.py

@@ -1,8 +1,8 @@
-import scipy
 from gwas2eqtl_pleiotropy import boxenplot_with_mannwhitneyu
-
-from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, annotator_config_dic
+from gwas2eqtl_pleiotropy.constants import label_fontsize, tick_fontsize, annotator_config_dic, boxenplot_kws, \
+    boxenplot_line_kws
 from statannotations.Annotator import Annotator
+from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
 
 import sys
 import matplotlib.pyplot as plt
@@ -16,7 +16,7 @@ import sqlalchemy
 
 # Plot parameters
 plt.rcParams["figure.figsize"] = (8, 6)
-from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic
+
 seaborn.set_theme(**seaborn_theme_dic)
 
 #%%
@@ -95,7 +95,7 @@ box_pairs = [(1, i) for i in range(1, pleio_high_cutoff + 1)]
 x = 'gwas_category_count'
 xlabel = "Trait category count"
 
-#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+#%% eqtl effect %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # boxplot and mann-whitney
 y = "eqtl_beta_abs"
 title = "eQTL effect"
@@ -105,27 +105,8 @@ ylabel = "Absolute beta mean"
 pairs = [(str(1), str(i)) for i in range(2, pleio_high_cutoff + 1)]
 m_eqtl_df[x] = m_eqtl_df[x].astype(str)
 
-ax = seaborn.barplot(x=x, y=y, data=m_eqtl_df, order=order, estimator=numpy.mean, palette="rocket_r")
-
-annotator = Annotator(ax, pairs, data=m_eqtl_df, x=x, y=y, order=order)
-annotator.configure(test='Mann-Whitney', text_format='star', **annotator_config_dic)
-annotator.apply_and_annotate()
-
-plt.title(title, fontsize=label_fontsize)
[…]

### 2023-09-13 scripts/plt_x_per_variant_y_egene_distance.py

@@ -10,11 +10,11 @@ import matplotlib.pyplot as plt
 from gwas2eqtl_pleiotropy import boxenplot_with_mannwhitneyu
 
 from statannotations.Annotator import Annotator
-
+from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic, annotator_config_dic, label_fontsize, tick_fontsize, \
+    palette, boxenplot_line_kws, boxenplot_kws
 
 #%%
 plt.rcParams["figure.figsize"] = (8, 6)
-from gwas2eqtl_pleiotropy.constants import seaborn_theme_dic, annotator_config_dic, label_fontsize, tick_fontsize
 seaborn.set_theme(**seaborn_theme_dic)
 
 #%%
@@ -46,15 +46,12 @@ pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
 #%%
 sql = 'select distinct * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
 columns = ['chrom', 'pos38', 'rsid', 'ref', 'alt', 'eqtl_refseq_transcript_id', 'eqtl_refseq_transcript_start38', 'eqtl_refseq_transcript_end38', 'eqtl_refseq_transcript_strand']
-# columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
-# h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
 engine = sqlalchemy.create_engine(sa_url)
 with engine.begin() as conn:
     h4_df = pandas.read_sql(sqlalchemy.text(sql), con=conn)
 h4_df = h4_df[columns].drop_duplicates()
 
 #%%
-# count_per_rsid_gwas_df = pandas.read_csv(count_per_rsid_gwas_ods_path, sep="\t")
 count_per_rsid_gwas_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine='odf', usecols=['chrom', 'pos38', 'rsid', 'ref', 'alt', 'gwas_category_count'])
 count_per_rsid_gwas_df = count_per_rsid_gwas_df.drop_duplicates()
 gwas_category_count_max_int = count_per_rsid_gwas_df['gwas_category_count'].max()
@@ -68,6 +65,7 @@ m_df['eqtl_gene_distance'] = math.nan
 
[…]

### 2023-09-13 scripts/plt_x_per_variant_y_remapcrm.py

@@ -2,7 +2,8 @@
 odds ratio and p-value of CRM vs. Non-CRM variants"""
 
 from gwas2eqtl_pleiotropy.Logger import Logger
-from gwas2eqtl_pleiotropy.constants import annotator_config_dic, label_fontsize, tick_fontsize, dpi, seaborn_theme_dic
+from gwas2eqtl_pleiotropy.constants import annotator_config_dic, label_fontsize, tick_fontsize, dpi, seaborn_theme_dic, \
+    palette
 from matplotlib import pyplot as plt
 from scipy.stats import fisher_exact
 from statannotations.Annotator import Annotator
@@ -93,9 +94,6 @@ out_df.loc[out_df['p'] <= 1.00e-04, 'signif'] = '****'
 out_df.rename({'pleio_count': 'gwas_category_count'}, axis=1, inplace=1)
 
 #%%
-# out_df = in_df.loc[in_df['consequence'] == consequence, ['gwas_category_count', 'oddsr', 'p', 'signif']]
-
-#%%
 out_df['gwas_category_count'] = [str(i) for i in out_df['gwas_category_count']]
 order = out_df['gwas_category_count'].tolist()
 xticklabels = order.copy()
@@ -110,7 +108,7 @@ pairs = [('1', x) for x in out_df['gwas_category_count'] if x != "1"]
 formatted_pvalues = out_df['signif'].tolist()[1:]
 
 #%% barplot
-ax = seaborn.barplot(x=x, y=y, data=out_df, order=order, palette="rocket_r")
+ax = seaborn.barplot(x=x, y=y, data=out_df, order=order, palette=palette)
 
 annotator = Annotator(ax, pairs, data=out_df, x=x, y=y, order=order, size=label_fontsize)
 annotator.set_custom_annotations(formatted_pvalues)

### 2023-09-13 scripts/pltbar_vep_consequence.py

@@ -4,7 +4,7 @@ import sys
 import pandas
 import seaborn
 
-from gwas2eqtl_pleiotropy.constants import dpi, seaborn_theme_dic, label_fontsize, tick_fontsize
+from gwas2eqtl_pleiotropy.constants import dpi, seaborn_theme_dic, label_fontsize, tick_fontsize, palette
 from matplotlib import pyplot as plt
 
 seaborn.set_theme(**seaborn_theme_dic)
@@ -54,11 +54,11 @@ in_df.loc[in_df['gwas_category_count'] == max(in_df['gwas_category_count']), 'gw
 
 order = in_df['consequence'].unique()
 hue_order = in_df['gwas_category_count'].unique()
-g = seaborn.barplot(data=in_df, x="consequence", y="oddsr", hue="gwas_category_count", palette="rocket_r", orient='v', order=order, hue_order=hue_order)
+g = seaborn.barplot(data=in_df, x="consequence", y="oddsr", hue="gwas_category_count", palette=palette, orient='v', order=order, hue_order=hue_order)
 for c in g.containers:
     gwas_category_count = c.get_label()
     labels = in_df.loc[in_df['gwas_category_count'] == gwas_category_count, 'signif'].tolist()
-    g.bar_label(c, labels=labels, label_type='edge', padding=5)
+    g.bar_label(c, labels=labels, label_type='edge', padding=5, fontsize=label_fontsize)
 
 # legend_labels = hue_order
 # legend_labels[-1] = '≥' + hue_order[-1]

### 2023-09-14 ms/ms08_figures.tex
MS

--- 2023-09-01 ms/ms08_figures.tex
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/eqtl_effect_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/gwas_effect_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_neglog10pval.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/eqtl_signif_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_neglog10pval.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_beta_neglog10pval.py/gwas_signif_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_allele_freq.py/eur_af.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_allele_freq.py/eur_af_custom.png}
     \end{subfigure}
     %
-    % Fig 9: eQTL gene distance
-    \begin{subfigure}[]{.49\textwidth}
-        \textbf{f}
-        \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/violin.png}
-    \end{subfigure}
-
     \begin{subfigure}[]{.49\textwidth}
@@
 
+    % Fig 9: eQTL gene distance
     \begin{subfigure}[]{.49\textwidth}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/histplot.png}
-        %
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_egene_distance.py/boxenplot_custom_closest_distance.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/histplot.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_etissue_y_egene.py/plt.png}
+        %
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/boxenplot_custom.png}
     \end{subfigure}
@@
         \\
-        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapcrm.py/remapcrm_flank10.png}
     \end{subfigure}
@@
 %    \\
-%    \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
+%    \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
 %    \end{subfigure}
@@
 %        \\
-%        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
+%        \includegraphics[width=\textwidth]{\floatRelativePath/plt_x_per_variant_y_remapcrm.py/remapcrm_flank10.png}
 %    \end{subfigure}

### 2023-09-14 setup.py

@@ -9,6 +9,6 @@ setup(
     description='Description of my package',
     packages=find_packages(),
     install_requires=['crossmap', 'matplotlib', 'mysql-client', 'mysql-connector-python', 'odfpy', 'pandas',
-                      'requests', 'seaborn', 'snakemake', 'sqlalchemy', 'statannot', 'statsmodels',
+                      'requests', 'seaborn>=0.12.2', 'snakemake', 'sqlalchemy', 'statannot', 'statsmodels',
                       'termcolor'],
 )

### 2023-09-15 gwas2eqtl_pleiotropy/__init__.py
MS

@@ -7,7 +7,7 @@ from gwas2eqtl_pleiotropy.constants import label_fontsize
 def boxenplot_with_mannwhitneyu(group1, group2, x1, x2, y, h):
     col = 'k'
     statistic, pvalue = scipy.stats.mannwhitneyu(group1, group2)
-    plt.plot([x1, x1, x2, x2], [y, y + h, y + h, y], lw=1.5, c=col)
+    plt.plot([x1, x1, x2, x2], [y, y + h, y + h, y], lw=1.5, c=col, clip_on = False)
     if pvalue < 1.00e-04:
         pvalue_asterisk = "****"
     elif pvalue < 1.00e-03:

### 2023-09-15 scripts/plt_x_per_variant_y_remapnr.py

@@ -99,9 +99,10 @@ plt.savefig(tf_flank_10_png)
 plt.close()
 
 #%% boxenplot ms
-ax = seaborn.boxenplot(x=x, y=y, data=cat_df, order=order, **boxenplot_kws, line_kws=boxenplot_line_kws)
+# ax = seaborn.boxenplot(x=x, y=y, data=cat_df, order=order, **boxenplot_kws, line_kws=boxenplot_line_kws)
+ax = seaborn.boxplot(x=x, y=y, data=cat_df, order=order, **boxenplot_kws)
 
-ylim = [0, 100]
+ylim = [0, 130]
 x1_annot1 = 0.
 delta_h = 0.03
 

### 2023-10-18 ms/ms07_tables.tex
ENH

@@
     head,
-    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries RSID & \bfseries Gene marker& \bfseries Trait categories\\\hline,
+    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Cytoband & \bfseries rsid & \bfseries Gene marker& \bfseries Trait categories\\\hline,
     ]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names

### 2024-11-10 ms/ms01_abstract.tex
First commit after acceptation

--- 2022-11-12 ms/ms01_titlepage.tex
@@
-%\leadauthor{Scientist}
+\begin{abstract} % abstract
 
-\title{Pleiotropy of gene regulatory variants increases with the number of regulated genes, active cell types and tissues and bound transcription factors}
-%\shorttitle{Pleiotropy and gene regulatory variants}
-
-\author[1,*]{Aitor González\,\orcidlink{0000-0002-8402-1655}}
-
-%\author[1,\Letter]{Aitor González}
-\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
-\affil[*]{Corresponding author: aitor.gonzalez@univ-amu.fr}
-% \orcidlink{0000-0002-8402-1655}
-%\affil[1]{Aix Marseille Univ, INSERM, TAGC, 13288 Marseille, France}
-\date{}
-
-\maketitle
-
-\begin{abstract}
-
+    \parttitle{Background} %if any
+    %
     Pleiotropic genetic variants affecting several traits are relatively common in the genome as shown by the increasing number of genome-wide association studies (GWAS).
 %
-Frequent variants are often located in non-coding regions and are likely gene regulatory variants.
+    Complex disease variants are often located in non-coding regions and are likely gene regulatory variants.
 %
-In this work, I investigate the link between frequent gene regulatory variants and pleiotropy.
+    In this work, I investigate the link between frequent expression quantitative trait loci and pleiotropy.
 
+    \parttitle{Results} %if any
 % Verify numbers
-First I have computed the colocalization probability of 418 GWAS and 127 eQTL studies and I have observed 10,334 GWAS/eQTL pairs with at least one colocalized variants with a probability above 0.8.
+    First I have computed the colocalization probability of 417 GWAS and 127 eQTL studies and I have observed 10,334 GWAS/eQTL pairs with at least one colocalized variants with a probability above 0.8.
 %SELECT count(DISTINCT concat(`gwas_id`, `eqtl_id`)) FROM `gwas2eqtl_modelgwas2eqtl` where `PP_H4_abf`>=0.8; 10334
@@
 
+    \parttitle{Conclusions} %if any
     In conclusion, our work suggest that pleiotropy of gene regulatory variants arise from an increase of bound transcription factors, target egenes and etissues and severity of molecular effect consequences.
-
+    %
 \end{abstract}
-
-% Keywords
-% eGenes; eQTLs

### 2024-11-10 ms/ms05_methods.tex
First commit after acceptation

@@
 
-Full association data were downloaded from OpenGWAS and converted to hg38 coordinates using Picard and Crossmap \citep{2021.Marcora.Lyon,Picard2019toolkit,2013.Wang.Zhao}.
+Full association data were downloaded from OpenGWAS and converted to hg38 coordinates using Picard and Crossmap \cite{2021.Marcora.Lyon,Picard2019toolkit,2013.Wang.Zhao}.
 %
@@
 %
-Missing minor allele frequency variants (MAF) were retrieved from the European population in the 1000 genome database \citep{2015.Abecasis.Auton}.
+Missing minor allele frequency variants (MAF) were retrieved from the European population in the 1000 genome database \cite{2015.Abecasis.Auton}.
 %

### 2024-11-10 src/cmpt_comorbidites.py

@@ -1,0 +1,48 @@
+import os
+import pandas
+import pathlib
+import sys
+import seaborn as sns;
+import sqlalchemy
+
+from gwas2eqtl_pleiotropy_logger import Logger
+
+sns.set_theme(color_codes=True)
+
+
+#%%
+help_cmd_str = "todo"
+try:
+    snp_pp_h4 = float(sys.argv[1])
+    sa_url = sys.argv[2]
+    disease_corr_tsv_path = sys.argv[3]
+    if len(sys.argv) > 4:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+outdir_path = os.path.dirname(disease_corr_tsv_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
+
+# sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
+# columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
+# d_df = pandas.read_sql(sql, con=sa_url, columns=columns).drop_duplicates()
+
+#%%
+columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
+sql = 'select * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
+engine = sqlalchemy.create_engine(sa_url)
+with engine.begin() as conn:
+    c_df = pandas.read_sql(sqlalchemy.text(sql), con=conn, columns=columns).drop_duplicates()
+
+#%%
+c_df = c_df.pivot_table(values='eqtl_beta', index=['rsid', 'eqtl_gene_id', 'eqtl_id'], columns='gwas_id', fill_value=0)
+
+Logger.info("Spearman correlation")
+c_df = c_df.corr(method='spearman')
+c_df.to_csv(disease_corr_tsv_path, sep='\t', index=True, header=True)
+

### 2024-11-10 src/insrt_cytoband.py

@@ -1,0 +1,46 @@
+import sqlalchemy
+
+from gwas2eqtl_pleiotropy_logger import Logger
+from psycopg2.extras import NumericRange
+from gwas2eqtl_pleiotropy_db import Base, cytoband
+from sqlalchemy import create_engine
+
+import pandas
+import sys
+
+#%%
+
+help_cmd_str = "todo"
+try:
+    sa_url = sys.argv[1]
+    if len(sys.argv) > 2:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+#%% Create all tables
+engine = create_engine(sa_url)
+if sqlalchemy.inspect(engine).has_table(cytoband.__tablename__):
+    cytoband.__table__.drop(engine)
+Base.metadata.tables[cytoband.__tablename__].create(bind=engine)
+
+#%% cytoband
+Logger.info("Insert cytobands")
+cytoband_url = "http://hgdownload.cse.ucsc.edu/goldenpath/hg38/database/cytoBand.txt.gz"
+cyto_df = pandas.read_csv(cytoband_url, sep="\t", header=None, usecols=[0, 1, 2, 3])
+cyto_df.columns = ['chrom', 'start', 'end', 'cytoband']
+cyto_df['chrom'] = cyto_df['chrom'].str.replace('chr', '')
+cyto_df = cyto_df.loc[cyto_df['chrom'].isin([str(chrom) for chrom in range(1, 23)])]
+cyto_df['chrom'] = cyto_df['chrom'].astype(int)
+cyto_df.set_index(cyto_df['chrom'].astype(str) + cyto_df['cytoband'], inplace=True, verify_integrity=True)
+
+with engine.connect() as connection:
+    connection.execute((Base.metadata.tables['cytoband']).delete())
+
+cyto_df['start_end38'] = cyto_df.apply(lambda x: NumericRange(x['start']+1, x['end'], '[]'), axis=1)
+cyto_df.drop(['start', 'end'], axis=1, inplace=True)
+cyto_df.to_sql('cytoband', con=engine, if_exists='append', index=True, index_label='id')

### 2024-11-10 src/plt_x_per_variant_egene_y_etissue.py

@@ -1,0 +1,120 @@
+import sqlalchemy
+from constants import label_fontsize, tick_fontsize
+from matplotlib.ticker import MaxNLocator
+
+import os
+import pandas
+import pathlib
+import seaborn
+import sys
+import matplotlib.pyplot as plt
+
+
+#%%
+plt.rcParams["figure.figsize"] = (8, 6)
+from constants import seaborn_theme_dic
+seaborn.set_theme(**seaborn_theme_dic)
+
+#%%
+help_cmd_str = "todo"
+try:
+    snp_pp_h4 = float(sys.argv[1])
+    pleio_high_cutoff = int(sys.argv[2])
+    sa_url = sys.argv[3]
+    count_per_rsid_gwas_ods_path = sys.argv[4]
+    png_path = sys.argv[5]
+    if len(sys.argv) > 6:
+        print("""Two many arguments!
+        {}""".format(help_cmd_str))
+        sys.exit(1)
+except IndexError:
+    print("""Argument missing!
+    {}""".format(help_cmd_str))
+    sys.exit(1)
+
+
+#%%
+if not os.path.isfile(count_per_rsid_gwas_ods_path):
+    print("input file does not exit")
+    sys.exit(1)
+
+outdir_path = os.path.dirname(png_path)
+pathlib.Path(outdir_path).mkdir(parents=True, exist_ok=True)
+
+#%%
+sql = 'select distinct * from colocpleio where snp_pp_h4>={}'.format(snp_pp_h4)
+# columns = ['rsid', 'eqtl_beta', 'eqtl_gene_id', 'gwas_id', 'eqtl_id']
+# h4_df = pandas.read_sql(sql, con=url).drop_duplicates()
+engine = sqlalchemy.create_engine(sa_url)
+with engine.begin() as conn:
+    h4_df = pandas.read_sql(sqlalchemy.text(sql), con=conn).drop_duplicates()
+
+#%%
+# count_per_rsid_gwas_df = pandas.read_csv(count_per_rsid_gwas_ods_path, sep="\t")
+count_per_rsid_gwas_df = pandas.read_excel(count_per_rsid_gwas_ods_path, engine='odf')
+gwas_category_count_max_int = count_per_rsid_gwas_df['gwas_category_count'].max()
+
+#%%
+m_df = h4_df.merge(count_per_rsid_gwas_df, on=['chrom', 'pos38', 'rsid'])
+
+# %%
[…]

### 2024-11-10 ms/ms06_backmatter.tex

@@
     %
-        \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
+    \caption{Distribution of GWAS based on eQTL colocalization and UCSC screenshots with colocalized GWAS/eQTLs.
+    \textbf{a} Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
     I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
-            because this locus is not included in the colocalization pipeline.}
+    because this locus is not included in the colocalization pipeline.
+    \textbf{b-d} UCSC screenshots of eQTLs colocalized with GWAS variants in the chr5:132,137,990-132,728,110 (hg38)
+    for the b-cells (\textbf{b}), classical monocytes (\textbf{c}) and cortex cells (\textbf{d}).
+    There are two tracks for each biosample: "beta equal" and "beta unequal".
+    "beta equal" corresponds to colocalized GWAS/eQTLs where both the GWAS beta and the eQTL beta coefficient have the same sign.
+    "beta unequal" corresponds to colocalized GWAS/eQTLs where the GWAS beta and the eQTL beta coefficients have different signs.
+    This allows to infer the beta coefficient of the GWAS variant eventhough the sign is shown with the column for the eQTL.
+    }
     \label{fig:1}
@@
     %
-        \caption{Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.}
+    \caption{
+    Trait clustering based on the Spearman correlation between beta values of colocalized eQTL variants, genes and biosamples.
+    }
     \label{fig:2}
@@
 % Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
-% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
-% Fig 5: VEP consequences
+% Fig : Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
+% Fig : VEP consequences
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
+
 \begin{figure}[!ht]
@@
 %
-% Fig 6: beta and pval
+% Fig 4: beta and pval
 %
@@
     %
-        \caption{eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
+    \caption{
+    eQTL and GWAS beta and p-values and allele frequencies for different GWAS category counts.
     \textbf{a,b} Average eQTL and GWAS beta values.
@@
 %
-% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
+% Fig 5: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
+% Fig: TF count per GWAS class count
+% Fig: eQTL gene distribution
 %
@@
     %
-        \caption{Mean number of eQTL genes per eQTL-biosample pair (\textbf{a}) and mean number of biosamples per eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
+    \caption{
+    Analysis of eQTL gene and biosample counts, transcription factor binding and gene distance as a function of the GWAS category count.
+    \textbf{a} Mean number of eQTL genes per eQTL-biosample pair.
+    \textbf{b} Mean number of biosamples per eQTL-gene pair.
+    \textbf{c} Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
+    \textbf{d} Odds ratio of eQTLs annotated as belonging to a cis regulatory modules versus non-annotated.
+    \textbf{e} Violin plots of the eQTL gene distances.
+    }
+    \label{fig:5}
     %
@@
 %
-% Fig 8: TF count per GWAS class count
+% Graphical conclusions
 %
@@
     %
-        \caption{Analysis of transcription factor binding.
-        \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
-        \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.}
-        \label{fig:freq_tf_per_variant}
+        \caption{
+        Graphical conclusions. I have analysed the differences underlying pleiotropy of eQTLs.
+        I have found that pleiotropic eQTLs have more target genes and lower GWAS and expression beta coefficients.
+        There is some evidence that pleitropic eQTLs overlap with more cis-regulatory modules.
+        }
+    \label{fig:6}
     %
